### 2015-01-12 Papeleo/ProyectoParaBecaProyecto01.tex
Papelo anadido

\documentclass[letterpaper, 10pt]{article}

\usepackage[spanish]{babel}

\title{Proyecto de Investigacion}
\author{Wilhelm P. K. Zapfe Zaldivar}

\begin{document}

\maketitle

El estudio de las estructuras neuronales tiene que enfocarse en varios niveles
de complejidad. La evidencia apunta a a que existen arreglos de neuronas
altamente conectadas que operan como microcircuitos con funciones bien
definidas y que estos a su vez se encuentran ordenados modularmente. 
Entre pares de neuronas se conocian, hasta hace poco, dos formas de
comunicacion: puramente electrica o a travez de un neurotransmisor quimico. 
Las investigaciones del Dr. Rafael Gutierrez han 
demostrado que existe al menos otro tipo de comunicacion entre ellas,
una forma mixta, y esto
abre la posibilidad de que los circuitos sean mucho mas complejos de lo 
previamente esperado. Un arreglo de neuronas podria comunicarse entre ellas 
de las tres formas y formar un circuito operacional bien definido.

Por otro lado tenemos modelos que simulan la actividad de grupos completamente
conectados de neuronas y ralamente conectados entre ellos que parecen
indicar la existencia de ciertas estructuras generales en los  niveles mas alto
de la red. Un estudio parece indicar que ese modelo es adecuado para
explicar fenomenos de sincronizacion (y posiblemente ataques epilepticos)
a travez de la topologia de esta red superior.

Eso ultimo aun se encuentra lejos de poder resolver el problema
de entender los posibles circuitos que podrian estar relacionados
con funciones basicas del sistema nervioso. 
El laboratorio dirigido por el Dr. Gutierrez cuenta con el equipo y personal para 
poder llevar a cabo un estudio a ese nivel. El ejemplo particular
que se trabaja actualmente es la actividad electrica de pequeñas rebanadas
del hipocampo de la rata comun. El equipo nos permite obtener bastante mas finos
espacialmente y con una resolucion temporal del orden de varios kilohertz. 
Esto ha permitito establecer una serie de objetivos con miras a entender la relacion
funcional en las conexiones del Giro Dentado (GA) y la estructura llamada
Cuerno de Amon (CA). Dado que la estructura esta organizada ahi de forma 
escencialmente laminar, el estudio profundo de la actividad en rebanadas
podra dilucidar la topologia de la red, y eventualmente, 
la funcionalidad de las conexiones en el hipocampo.

\end{document}

### 2015-01-12 Papeleo/ProyectoParaBecaProyecto02.tex
Papelo anadido

\documentclass[letterpaper, 10pt]{article}

\usepackage[spanish]{babel}
\usepackage[utf8]{inputenc}

\title{Proyecto de Investigacion}
\author{Wilhelm P. K. Zapfe Zaldivar}

\begin{document}

\maketitle

La investigación de las estructuras neuronales actualmente
ha establecido varios niveles funcionales que parecen ser
comunes. Hay evidencia muy clara de que las neuronas se agrupan
en módulos relativamente pequeños y altamente conectados que funcionarían
como microcircuitos y éstos a su vez estarían ordenados
en redes poco conectadas. La topología de una y otra
estructura jerárquica daría función al circuito así establecido.
Los trabajos enfocados en descubrir esas estructuras son
en este momento investigación de punta.

El Dr. Rafael Gutiérrez ha establecido que existe, al menos, otro
tipo de comunicación frente a los dos ya conocidos, de corriente iónica
intercelular y de neurotransmisores químicos. Uno podría describirse como
una señal analógica o de valores continuos, mientras que el otro
es una señal binaria discreta. El tipo mixto muestra características
de ambas en las mediciones del potencial eléctrico intracelular. También
encontró    que los tres tipos de conexiones se dan entre el giro dentado
(GA) y la zona CA3 en el hipotálamo de las ratas comunes. Esta
información previa es de fundamental importancia para
poder descubrir las estructuras funcionales que podríamos
describir como microcircuitos en el hipotálamo.

Actualmente se dispone de la capacidad de recopilar datos de pequeñas
rebanadas \emph{in vitro} del hipotálamo de las ratas. El equipo y personal
experimental nos permite observar la actividad eléctrica de pequeñas 
láminas del cerebro y grabarlas con una densidad muy alta en el espacio y de
muy alta resolución en el tiempo. Con estos datos, y el conocimiento
de las huellas de los diferentes tipos de conexiones podemos, a través
de un análisis estadístico y causal, proponer posibles redes que
conecten ambas estructuras.

Para poder establecer finalmente un buen modelo explicativo de red para
los circuitos así observados sería bueno conocer, al menos cualitativamente,
las propiedades de las redes posibles, entre ellas, cualidades topológicas
y de las posibles dinámicas que podrían generar. El trabajo que yo aportaría
al grupo sería en esta línea, aparte de apoyo para el procesamiento
computacional de los datos. Mis conocimientos previos
de sistemas dinámicos no lineales y caóticos y capacidad de seguir
argumentos matemáticos abstractos en teoría de redes me permitiría
establecer cualidades generales en las propuestas de redes que obtendríamos
del análisis de los datos experimentales.

\end{document}

### 2015-03-18 CenterOfMass/PlotActiandCenterMass01.sh
Reporte anhadido

@@ -5,8 +5,9 @@
 datosfondobasename="LaplacianFilteredSpaceTime-"
 numero=$1
 dataname=$datosfondobasename$numero".dat"
-usandonegativo='2:1:($3/4.)'
-usandopositivo='2:1:($3/4.)'
+usandonegativo='(-$2/4.):(-$1/4):($3/500.)'
+usandopositivo='($2/4.):($1/4):($3/500.)'
+usandomatrix='($1/4.):($2/4):($3)'
 
 outname=`basename $dataname .dat`.png
 index="::::$numero"
@@ -21,7 +22,7 @@ gnuplot <<EOF
 
 set xr[-0.50:63.5]
 set yr[-0.5:63.5]
-set cbr[-120:120]
+set cbr[-10:10]
 set size ratio -1
 set pointsize 0.35
 unset key
@@ -32,7 +33,7 @@ set title titulo
 set palette defined (0 "#000077", 1 "#0000FF", 2 "white", 3 "#FF0000", 4 "#990000")
 #set palette defined (0 "white", 1 "#0000FF", 2 "black")
 set out "$outname"
-plot "$dataname" matrix  w image, "DatosCMNegativo01.dat" usi $usandonegativo every $index w lp ls 7 palette ,"DatosCMPositivo01.dat" usi $usandopositivo every $index w lp ls 7 palette 
+plot "$dataname" using $usandomatrix matrix  w image, "DatosCMNegativo01.dat" usi $usandonegativo every $index w lp ls 7 palette ,"DatosCMPositivo01.dat" usi $usandopositivo every $index w lp ls 7 palette 
  
 set out
 

### 2015-03-20 Reportes/Reporte01.tex
CentroDeMasa mejor mas chido

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}
\usepackage{graphicx}

\title{Reporte Centro de Masa, Densidad de Fuentes I}

\author{W. P. Karel Zapfe}

\begin{document}

\maketitle

\section{La Señal Cruda}

El aparato, el BioCAM 4096S, es un arreglo de 4096 electrodos en un 
cuadrángulo con 64 renglones y 64 columnas en un cuadrado de
$(2.67 mm)^2$. Los electrodos permiten medir variaciones del orden
de unos micro-volts hasta aproximadamente 16 mili-volts 
en el potencial eléctrico de tejido colocado sobre el arreglo. Cuando
todos los electrodos están registrando actividad, tenemos una frecuencia de muestreo
de 7022 tomas por segundo. Cada electrodo tiene $21 \mu m$ de lado, lo
que hace que sus dimensiones sean del mismo orden de magnitud que el 
soma de una neurona piramidal típica. El aparato tiene, de acuerdo
con sus fabricantes, una señal ruidosa de $26 \mu V$rms. Dicha señal
está presente en todos los electrodos excepto en un par que se
muestran saturados todo el tiempo en varios registros.

Cuando una célula piramidal dispara se puede medir un cambio de potencial
en el medio extracelular. Si tomamos medidas \emph{muy} cercanas al soma
(menos de $25 \mu m$)
y no hay mas influencias, este cambio puede ser de aproximadamente
$-250\mu V$ \cite{Obien2015}. Poco después ocurre una repolarización
y se observa una señal contraria de aproximadamente $+100\mu V$.
Dado que el potencial decrece como el inverso de la distancia
y es aditivo, nuestras espigas discernibles en la señal tienen tamaños mucho
menores, pero distintamente arriba del nivel de ruido.

La actividad neuronal producirá una distribución diferente en las medidas
de los electrodos cercanos, mientras que aquellos que se encuentren
suficientemente alejados de las fuentes y sumideros de corriente
percibirán mayoritariamente ruido y tendrán otra estadística.
Esto nos permite separar los electrodos sobre los cuales
se encuentren cuerpos densamente llenos de células piramidales
de aquellas regiones vacías, y podremos identificar el CA.

\section{Limpiando la señal}

Para obtener una señal \emph{limpia} tenemos que tomar en cuenta
cual es nuestro propósito. Para realizar el estudio que está llevando a cabo
Franco basta con poder identificar las espigas del ruido.
Pero para poder obtener una densidad de fuentes de corriente esto no basta.
El procedimiento formal para obtener esta densidad, que denotaré
 por CSD, es calcular el \emph{Laplaciano} del potencial de
campo. Esto es un operador diferencial (la suma de las dobles derivadas 
cartesianas), y por ende requiere una \emph{función suave} para actuar. 
El Laplaciano es especialmente sensible al ruido, tanto así, que su
función en procesamiento de señales es detectar \emph{bordes duros}.
Esto es especialmente de interés para nosotros: su implementación
más común  se usa
justamente en la detección de bordes en imágenes  digitales.
El ruido es básicamente un montón de bordes duros en cada instante temporal.
Estamos obligados, por ende, a filtrar todo el ruido que sea posible.

\subsection{Identificando CA}

Una gran cantidad de electrodos se encuentran lejos de las células
piramidales y sólo registran ruido. Este ruido se puede caracterizar
por sus propiedades estadísticas. Un método común de separar secciones
en una imagen bidimensional que resulta fácil de implementar y que
da resultados sorprendentemente buenos es el llamado método de Otsu
\cite{Otsu79}.

A grandes rasgos, lo que hace la llamada ``umbralización de Otsu'' es
separar la señal bidimensional en dos conjuntos tales que su 
estadística sea \emph{lo más distante posible}, sin hacer ninguna
suposición al respecto de las características de cada señal.

Decidí confrontar una implementación de este método con
el método que usa Franco.
Rescaté algunas ideas de Franco para obtener una máscara que
revelara la parte del arreglo que nos interesa.
Los detalles de la implementación específica 
se encuentran en un \emph{notebook}. El resultado se muestra en la figura
\ref{OtsuMask}.  Toda señal fuera de esa máscara la trataremos como cero.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{MascaraOtsu01.png}
  \caption{Mascara obtenida mediante la suma sucesiva de varias
    umbralizaciones de Otsu. La figura amarilla cubre los electrodos
    que se encuentran cerca del cuerpo celular CA.} \label{OtsuMask}
\end{figure}

\subsection{Quitando ruido}

La señal ruidosa se superpone a la señal de actividad neuronal en
los electrodos seleccionados. Una posibilidad para separarlas
sería utilizar las características de frecuencia de alguna de ambas 
y hacer pasar la señal de cada electrodo por un filtro pasa-bajos
o pasa-altos según sea el caso (Mario Treviño dixit).
Esto tiene un problema intrínseco: la señal ruidosa está acoplada
con la frecuencia de alternancia de la corriente, que en México es
de entre 55 a 65 Hz. Una espiga putativa en nuestra señal
(ver figura \ref{espiga}) tiene una estructura con duración
temporal de aproximadamente $5 ms$. Esto es aproximadamente 
un tercio del tamaño de la estructura asociada a los 60Hz,
asi que es demasiado cercana como para usarse confiadamente.
Preferí usar un simple criterio de umbral: todo aquello
que estuviera por debajo de $60\mu V$ en valor absoluto 
sería considerado ruido y transformado a ceros. La
comparación entre una señal original y la señal umbralizada
se muestra en la figura \ref{umbralc}.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.75\textwidth]{ComparaUmbral.png}
  \caption{Mascara obtenida mediante la suma sucesiva de varias
    umbralizaciones de Otsu. La figura amarilla cubre los electrodos
    que se encuentran cerca del cuerpo celular CA.} \label{umbralc}
\end{figure}

\subsection{Suavizando la señal}

La implementación directo del operador
Laplaciano discreto en una retícula bidimensional produce
efectos de borde muy notorios que ofuscarían la visualización
de las áreas con verdadera actividad. Es por ello que se
desarrollo un método de suavización de bordes en la  implementación
del operador llamado ``filtro Gauss-Laplace''. Consiste en pasar
un filtro Gaussiano por la imágen o señal primero, y luego el 
operador Laplaciano discreto. Esta fue la ``segunda derivada'' que
decidí implementar para buscar la densidad de fuentes de corrientes.
Esto producirá un efecto de señal de potencial similar a 
la de un campo apantallado, que es, hasta donde yo se,
 adecuada para modelar iones disueltos en un  medio
 complicado. Un ejemplo del efecto de cruz en esta resolución
 se puede ver en la figura \ref{LaplacianEfect}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{EjemploLaplacian01.png}
  \caption{Efecto de ``cruz'' por la implementación del Laplaciano discreto
    en una malla de 64 por 64 píxeles. La escala es muy cercana al
  $\epsilon$ de la derivada. }\label{LaplacianEfect}
\end{figure}

La señal ya así obtenida tiene todavía el inconveniente
de ser muy ``dura'' en el dominio espacial. Aunque $64\times 64$
puntos de muestreo en tan poco espacio tenga una resolución
alta en comparación con las estructuras fisiológicas, es baja
para obtener derivadas numéricas descentes.
Las gráficas obtenidas en Matlab por Treviño et all y Luzdas et all
se muestran mucho más suaves al calcular la CSD. Esto es escencialmente
debido a que Matlab interpola valores entre puntos de la malla
para crear un gradiante suave de colores. Si vamos a usar una
interpolación, es conveniente que usemos una que no aporte información
espuria a los datos y que podamos controlar los parámetros directamente.
Para ello usaremos una rutina de interpolación cuadrática
(que garantiza la existencia de segundas derivadas) y condiciones
en la frontera que extiendan el último renglón o fila.

\section{Centros de Masa}

\bibliographystyle{plain}

\bibliography{BiblioReportes01}

\end{document}

### 2015-03-20 CenterOfMass/PlotActiandCenterMass01.sh

@@ -5,9 +5,14 @@
 datosfondobasename="LaplacianFilteredSpaceTime-"
 numero=$1
 dataname=$datosfondobasename$numero".dat"
-usandonegativo='(-$2/4.):(-$1/4):($3/500.)'
-usandopositivo='($2/4.):($1/4):($3/500.)'
-usandomatrix='($1/4.):($2/4):($3)'
+#usandonegativo='(-$2/4.):(-$1/4):($3/500.)'
+#usandopositivo='($2/4.):($1/4):($3/500.)'
+#usandomatrix='($1/4.):($2/4):($3)'
+
+usandonegativo='(-$2):(-$1):($3/1500.)'
+usandopositivo='($2):($1):($3/1500.)'
+usandomatrix='($1):($2):($3)'
+
 
 outname=`basename $dataname .dat`.png
 index="::::$numero"

### 2015-03-23 Reportes/Reporte01.tex
CentroDeMasaMasDatos ya mejor con ContLapl

@@
 
-El aparato, el BioCAM 4096S, es un arreglo de 4096 electrodos en un 
+El aparato, el BioCAM 4096S, es un arreglo de 4096 electródos en un 
 cuadrángulo con 64 renglones y 64 columnas en un cuadrado de
-$(2.67 mm)^2$. Los electrodos permiten medir variaciones del orden
-de unos micro-volts hasta aproximadamente 16 mili-volts 
-en el potencial eléctrico de tejido colocado sobre el arreglo. Cuando
-todos los electrodos están registrando actividad, tenemos una frecuencia de muestreo
+$(2.67 mm)^2$. Los electrodos permíten medir variaciones del orden
+de unos microvolts hasta aproximadamente 16 milivolts 
+en el potencial eléctrico de muestras colocadas sobre el arreglo. Cuando
+todos están registrando actividad, tenemos una frecuencia de muestreo
 de 7022 tomas por segundo. Cada electrodo tiene $21 \mu m$ de lado, lo
@@
 suficientemente alejados de las fuentes y sumideros de corriente
-percibirán mayoritariamente ruido y tendrán otra estadística.
+percibiran mayoritariamente ruido y tendrán otra estadística.
 Esto nos permite separar los electrodos sobre los cuales
 se encuentren cuerpos densamente llenos de células piramidales
-de aquellas regiones vacías, y podremos identificar el CA.
+de aquellas regiones vacias, y podremos identificar el CA.
 
@@
 Pero para poder obtener una densidad de fuentes de corriente esto no basta.
-El procedimiento formal para obtener esta densidad, que denotaré
- por CSD, es calcular el \emph{Laplaciano} del potencial de
-campo. Esto es un operador diferencial (la suma de las dobles derivadas 
+El procedimiento formal para obtener esta densidad, que denotaré siguiendo
+la tradición, por $\rho$, es calcular el \emph{Laplaciano} del potencial de
+campo. Esto es un operador diferencial (la suma de las dobles derivada 
 cartesianas), y por ende requiere una \emph{función suave} para actuar. 
@@
 función en procesamiento de señales es detectar \emph{bordes duros}.
-Esto es especialmente de interés para nosotros: su implementación
-más común  se usa
-justamente en la detección de bordes en imágenes  digitales.
-El ruido es básicamente un montón de bordes duros en cada instante temporal.
+Esto es especialmente de interés para nosotros, ya que se usa
+justamente en la detección de bordes en imágenes bidimensionales digitales.
+El ruido es basicamente un monton de bordes duros en cada instante temporal.
 Estamos obligados, por ende, a filtrar todo el ruido que sea posible.
@@
 Una gran cantidad de electrodos se encuentran lejos de las células
-piramidales y sólo registran ruido. Este ruido se puede caracterizar
+piramidales y sólo registran ruido. Este ruido se puede caracterízar
 por sus propiedades estadísticas. Un método común de separar secciones
 en una imagen bidimensional que resulta fácil de implementar y que
-da resultados sorprendentemente buenos es el llamado método de Otsu
-\cite{Otsu79}.
-
+da resultados sorprendentemente buenos es el llamado método de Otsu.
 A grandes rasgos, lo que hace la llamada ``umbralización de Otsu'' es
@@
 estadística sea \emph{lo más distante posible}, sin hacer ninguna
-suposición al respecto de las características de cada señal.
+suposición al respecto de las caractéristicas de cada señal.
 
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.45\textwidth]{MascaraOtsu01.png}
-  \caption{Mascara obtenida mediante la suma sucesiva de varias
-    umbralizaciones de Otsu. La figura amarilla cubre los electrodos
-    que se encuentran cerca del cuerpo celular CA.} \label{OtsuMask}
-\end{figure}
-
 \subsection{Quitando ruido}
@@
 La señal ruidosa se superpone a la señal de actividad neuronal en
-los electrodos seleccionados. Una posibilidad para separarlas
+los electrodos seleccionados. Una posibilidad para separlas
 sería utilizar las características de frecuencia de alguna de ambas 
-y hacer pasar la señal de cada electrodo por un filtro pasa-bajos
-o pasa-altos según sea el caso (Mario Treviño dixit).
+y hacer pasar la señal de cada electródo por un filtro pasabajos
+o pasaltos segun sea el caso (Mario Treviño dixit).
 Esto tiene un problema intrínseco: la señal ruidosa está acoplada
@@
 comparación entre una señal original y la señal umbralizada
-se muestra en la figura \ref{umbralc}.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.75\textwidth]{ComparaUmbral.png}
-  \caption{Mascara obtenida mediante la suma sucesiva de varias
-    umbralizaciones de Otsu. La figura amarilla cubre los electrodos
-    que se encuentran cerca del cuerpo celular CA.} \label{umbralc}
-\end{figure}
+se muestra en la figura \ref{umbral}.
 
@@
 
-La implementación directo del operador
+La señal ya así obtenida tiene todavía el inconveniente
+de ser muy ``dura'' en el dominio espacial. Aunque $64\times 64$
+puntos de muestreo en tan poco espacio tenga una resolución
+alta en comparación con las estructuras fisiológicas, es baja
+para ``hacer cálculo''. La implementación directo del operador
 Laplaciano discreto en una retícula bidimensional produce
 efectos de borde muy notorios que ofuscarían la visualización
-de las áreas con verdadera actividad. Es por ello que se
-desarrollo un método de suavización de bordes en la  implementación
+de las areas con verdadera actividad. Es por ello que se
+desarollo un método de suavización de bordes en la  implementación
 del operador llamado ``filtro Gauss-Laplace''. Consiste en pasar
@@
  adecuada para modelar iones disueltos en un  medio
- complicado. Un ejemplo del efecto de cruz en esta resolución
- se puede ver en la figura \ref{LaplacianEfect}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.45\textwidth]{EjemploLaplacian01.png}
-  \caption{Efecto de ``cruz'' por la implementación del Laplaciano discreto
-    en una malla de 64 por 64 píxeles. La escala es muy cercana al
-  $\epsilon$ de la derivada. }\label{LaplacianEfect}
-\end{figure}
-
-La señal ya así obtenida tiene todavía el inconveniente
-de ser muy ``dura'' en el dominio espacial. Aunque $64\times 64$
-puntos de muestreo en tan poco espacio tenga una resolución
-alta en comparación con las estructuras fisiológicas, es baja
-para obtener derivadas numéricas descentes.
-Las gráficas obtenidas en Matlab por Treviño et all y Luzdas et all
-se muestran mucho más suaves al calcular la CSD. Esto es escencialmente
-debido a que Matlab interpola valores entre puntos de la malla
-para crear un gradiante suave de colores. Si vamos a usar una
-interpolación, es conveniente que usemos una que no aporte información
-espuria a los datos y que podamos controlar los parámetros directamente.
-Para ello usaremos una rutina de interpolación cuadrática
-(que garantiza la existencia de segundas derivadas) y condiciones
-en la frontera que extiendan el último renglón o fila.
+complicado.
 
@@
 
-\bibliographystyle{plain}
-
-\bibliography{BiblioReportes01}
-
 \end{document}

### 2015-04-08 CenterOfMass/PloteaCMdisjoint01.sh
Testing Denoiz

@@ -9,8 +9,8 @@ numero=$1
 dataname=$datosfondobasename$numero".dat"
 datosCMNeg=$datosCMNegname$numero".dat"
 datosCMPos=$datosCMPosname$numero".dat"
-#usandonegativo='(-$2/4.):(-$1/4):($3/500.)'
-#usandopositivo='($2/4.):($1/4):($3/500.)'
+#usandonegativo='(-$2/4.):(-$1/4):($3/1000.)'
+#usandopositivo='($2/4.):($1/4):($3/1000.)'
 #usandomatrix='($1/4.):($2/4):($3)'
 
 usandonegativo='($1):($2):(abs($3*0.002))'

### 2015-04-10 Reportes/DiapositivasAbril01.tex
la la la la

\documentclass[xetex,mathserif,serif]{beamer}

\begin{document}
  \begin{frame}
    \frametitle{CSDA}
    Our measurements would allow us to obtain the \emph{Current Source Density}
    of the electric field potentials  recorded in the experiments.
    The fundamental equation which relates this two quantities is
    Gauss Law for an ohmic, resistive and uniform medium:
    \begin{equation}
      \sigma_e \nabla^2 \Phi = -I_m.
    \end{equation}
    Here $\sigma_e$ is the electric conductivity of the extracellular medium
    and $I_m$ is the density of current sources. A more general approach
    can be seen at \cite{Bedard11}.

  \end{frame}

  \begin{frame}
     \frametitle{Implementation}
     The more direct approach for the CSD in 2D would be to apply
     a discrete Laplacian Filter directly on the data, if
     the spacing of the data were of the order of the dipolar structures
     that we wish to see. Our main problem is that such a filter is
     \emph{extremely} sensitive to noisse. This noice can be clearly seen
     on the temporal domain as random ls spikes ranging  $-40\mu V$ to 
     $+40\mu V$ 
  \end{frame}
% etc
\end{document}

### 2015-04-13 Reportes/DiapositivasAbril01.tex
notas

@@
 \documentclass[xetex,mathserif,serif]{beamer}
+
+%\usepackage[numberedbib]{apacite}
+%\usepackage[english]{babel}
+\usepackage{natbib}
+\usepackage{bibentry}
+\bibliographystyle{apalike}
 
@@
     Here $\sigma_e$ is the electric conductivity of the extracellular medium
-    and $I_m$ is the density of current sources. A more general approach
+    and $I_m$ is the density of current sources. A deeper approach
     can be seen at \cite{Bedard11}.
@@
    \begin{frame}
-     \frametitle{Implementation}
-     The more direct approach for the CSD in 2D would be to apply
-     a discrete Laplacian Filter directly on the data, if
-     the spacing of the data were of the order of the dipolar structures
-     that we wish to see. Our main problem is that such a filter is
-     \emph{extremely} sensitive to noisse. This noice can be clearly seen
-     on the temporal domain as random ls spikes ranging  $-40\mu V$ to 
-     $+40\mu V$ 
+     \frametitle{Implementation: de-noising the Data}
+     A direct application of a Laplacian Filter is not viable, since
+     this operator is extremely sensitive to noise. The noise is coupled
+     to the line power (\~ 60Hz). A (temporal) Gaussian convolution takes care
+     of this and its harmonics.
   \end{frame}
   % etc
+
+  \begin{frame}
+     \frametitle{Implementation: de-noising the Data}
+     The Laplacian implementation is, in this discrete form, a convolution
+     with the following matrix. This is in accordance with the
+     scales that we are interested in. Compare with \cite{Vaknin88}.
+
+  \end{frame}
+
+  \begin{frame}
+     \frametitle{Enter of Mass Dynamics}
+     Following \cite{Manjarrez07}, we desire to follow a putative CoM,
+     to discover how activity propagates along the CA.
+
+  \end{frame}
+
+  \begin{frame}
+     \frametitle{Implementation: naïve search}
+     A direct approach proves useless, as the activity occurs at separated
+     sites along CA1 and CA3. The concave geometry of our slice
+     produces a centre of mass outside the CA, and without physiological
+     interpretation.
+  \end{frame}
+
+  \begin{frame}
+     \frametitle{Implementation: disjoint set detection.}
+     It is better to focus on the \emph{disjoint sets} revealed by
+     the CSDA. We have a clear separation between sources and sinks,
+     and every set would have a centre of mass inside the CA, thus
+     allowing us to follow the propagation of activity on different
+     components \emph{simultaneously}.
+
+  \end{frame}
+
+  \begin{frame}{Bibliography}
+   % \bibliographystyle{apacite}
+    \bibliography{BiblioReportes01}
+  \end{frame}
+
 \end{document}

### 2015-05-06 Reportes/ExplicacionDetallada01.tex
detalles

--- 2015-03-23 Reportes/Reporte01.tex
@@
 
-\title{Reporte Centro de Masa, Densidad de Fuentes I}
+\title{Densidad de Fuentes de Corriente}
 
@@
 y es aditivo, nuestras espigas discernibles en la señal tienen tamaños mucho
-menores, pero distintamente arriba del nivel de ruido.
+menores, pero claramente sobre el nivel de ruido.
 
@@
 Esto nos permite separar los electrodos sobre los cuales
-se encuentren cuerpos densamente llenos de células piramidales
-de aquellas regiones vacias, y podremos identificar el CA.
+se encuentren cuerpos densamente llenos de células piramidales y dendritas
+de aquellas regiones relativamente vacias.
 
@@
 Franco basta con poder identificar las espigas del ruido.
-Pero para poder obtener una densidad de fuentes de corriente esto no basta.
+Pero para poder obtener una densidad de fuentes de corriente 
+çesto no es suficiente.
 El procedimiento formal para obtener esta densidad, que denotaré siguiendo
@@
 
-\subsection{Identificando CA}
-
-Una gran cantidad de electrodos se encuentran lejos de las células
-piramidales y sólo registran ruido. Este ruido se puede caracterízar
-por sus propiedades estadísticas. Un método común de separar secciones
-en una imagen bidimensional que resulta fácil de implementar y que
-da resultados sorprendentemente buenos es el llamado método de Otsu.
-A grandes rasgos, lo que hace la llamada ``umbralización de Otsu'' es
-separar la señal bidimensional en dos conjuntos tales que su 
-estadística sea \emph{lo más distante posible}, sin hacer ninguna
-suposición al respecto de las caractéristicas de cada señal.
-
-Decidí confrontar una implementación de este método con
-el método que usa Franco.
-Rescaté algunas ideas de Franco para obtener una máscara que
-revelara la parte del arreglo que nos interesa.
-Los detalles de la implementación específica 
-se encuentran en un \emph{notebook}. El resultado se muestra en la figura
-\ref{OtsuMask}.  Toda señal fuera de esa máscara la trataremos como cero.
-
 \subsection{Quitando ruido}
 
-La señal ruidosa se superpone a la señal de actividad neuronal en
-los electrodos seleccionados. Una posibilidad para separlas
-sería utilizar las características de frecuencia de alguna de ambas 
-y hacer pasar la señal de cada electródo por un filtro pasabajos
-o pasaltos segun sea el caso (Mario Treviño dixit).
-Esto tiene un problema intrínseco: la señal ruidosa está acoplada
-con la frecuencia de alternancia de la corriente, que en México es
-de entre 55 a 65 Hz. Una espiga putativa en nuestra señal
-(ver figura \ref{espiga}) tiene una estructura con duración
-temporal de aproximadamente $5 ms$. Esto es aproximadamente 
-un tercio del tamaño de la estructura asociada a los 60Hz,
-asi que es demasiado cercana como para usarse confiadamente.
-Preferí usar un simple criterio de umbral: todo aquello
-que estuviera por debajo de $60\mu V$ en valor absoluto 
-sería considerado ruido y transformado a ceros. La
-comparación entre una señal original y la señal umbralizada
-se muestra en la figura \ref{umbral}.
+Una técnica de \emph{suavización} de una señal en función del tiempo
+ la convolución con una Gaussiana relativamente estrecha. Esto es en realidad
+un filtro pasabajos. El teorema de la convolución nos garantiza que esta
+convolución es equivalente a multiplicar la señal por una Gaussiana 
+en el dominio de las frecuencias, efectivamente aplastando todo aquello
+que se encuentre lejos del origen. El ancho efectivo (la tolerancia a 
+frecuencias altas) está dado por el inverso del parámetro $\sigma$ de
+la Gaussiana en el ámbito temporal. El ruido está acoplado con 
+las capacidades de máxima resolución de la máquina, así que una 
+Gaussiana relativamente delgada eliminará estás estructuras (véase fig. 
+\ref{Compara01}).
+
+\begin{figure}[h]
+\includegraphics[width=0.9\textwidth]{ComparaSuaveyCruda01.pdf}
+\caption{La señal de un electrodo partícular cruda y después de
+la convolución con una Gaussiana con ancho $\sigma=$. En la gráfica
+menor se aprecia el efecto a escala aumentada.
+}\label{Compara01}
+\end{figure}
 

### 2015-05-14 Reportes/CSDAyOrthCoord01.tex
Reportes mas

\documentclass{article}

%usepackage[utf8]{inputenc}
 \usepackage{fontspec}
\usepackage{amsmath}
\usepackage[spanish]{babel}
\usepackage{graphicx}

\usepackage{caption}
\usepackage{subcaption}

 \setmainfont[Ligatures=TeX]{Oldstyle HPLHS}

\newcommand{\Jd}{\mathbf{J}}
\newcommand{\EF}{\mathbf{E}}
\newcommand{\cond}{\boldsymbol{\sigma}}
\DeclareMathOperator{\diver}{div}

\title{Densidad de Fuentes de Corriente y Coordenadas generalizadas}
\author{W. P. Karel Zapfe}

\begin{document}

\maketitle

\section{Cálculo Diferencial y Electrodinámica sobre el Hipocampo}

En principio, la densidad de fuentes de corriente se puede obtener
de la densidad de corriente por continuidad:
\begin{equation}
\nabla \cdot \Jd +\frac{\partial \rho}{\partial t}=0.
\end{equation}
Es usual llamarle a $\partial \rho /\partial t =-I_m$, la densidad de 
fuentes de corriente.
Ahora es momento para recordar que la notación ``nabla'' resulta especialmente
inconveniente para cambios de coordenadas, pues nos hace creer (y meter la pata)
que sus componentes cambian siempre igual, y eso no es el caso. Por
ello voy a reescribir de una forma más algorítmica la ecuación de arriba
de la siguiente forma:
\begin{equation}
\diver \Jd =I_m.
\end{equation}
Obsérvese que la cantidad $\Jd$ es una cantidad \emph{vectorial}, cuyos componentes
cambian de acuerdo a las coordenadas usadas localmente, sin embargo $I_m$ es un
escalar, es decir, es invariante frente a cambios de coordenadas. Por
ende, la operación $\diver ()$ tiene que contrarrestar el cambio de coordenadas,
es decir, es un operador \emph{contravariante}, a diferencia del operador
gradiente, que denotamos como $\nabla$ sin el puntido de producto interno\ldots

Si el material es ohmico y descente esperamos esta relación:
\begin{equation}
\Jd=\cond \EF,
\end{equation}
despreciando la contribución del campo de desplazamiento eléctrico (ver
las eqs. 18 y 28 los argumentos entre ellas en \cite{Bedard11}).
Por el momento supongamos que estamos en un régimen linear, donde
$\cond$ no depende de $\EF$. Sin embargo no hay ninguna razón
de entrada para asumir $\cond$ como isotrópico o homogéneo. Así que en
general consideremos a $\cond$ como un tensor, cuyas componentes generales
son:
\begin{equation}
\cond=\begin{pmatrix}
\sigma_{xx} & \sigma_{xy} \\
\sigma_{yx} & \sigma_{yy}
\end{pmatrix}
\end{equation}
Nótese que ya estoy restringiendome a un universo bidimensional.
Este tensor es real y simétrico, así que básicamente esto quiere decir
que podemos encontrar un sistema de coordendadas \emph{ortonormales}
donde el tensor sólo tenga componentes diagonales (el sistema estaría
alineado con los ``componentes principales'', véase por ejemplo la
ecuación 6 en Mitzdorf,
\cite{Mitzdorf85}). El cambio de coordenadas del sistema $(x,y)$ al
sistema de componentes principales, que denotaré por $(t',a')$ es localmente
\emph{ortogonal}, pues preserva norma y producto interno de nuestra
base cartesiana usual. Este cambio de coordenadas es local, ya que como
no hemos supuesto que $\cond$ sea constante, los ejes principales
pueden variar de punto a punto. Vamos a considerar que son 
bastante suaves y que por lo tanto dicha transformación local es la
diferencial de un cambio de cordenadas ortogonal adecuado.
Dado que la conductividad, a pesar de estar representada por este
objeto matemático abstracto (un tensor de rango dos), es algo muy concreto,
la sabiduría convencional indica que los ejes principales en cada punto
deben de estar alineados con las estructuras notorias de nuestro material.

En nuestro caso particular, las estructuras de interés son los cuerpos laminados
donde se encuentran los somas de las neuronas, el ``estrato piramidal'',
y las capas dendríticas cerca de éste. Podríamos imaginar entonces
que los ejes principales siguen, punto por punto, a lo largo del
CA, un vector tangente al centro de este estrato piramidal y otro vector
normal a éste, que llamare ``apical'' por analogía, véase la fig. 
\cite{esquemas01}.

\begin{figure}[h]
\centering
\begin{subfigure}[b]{0.45\textwidth}
\includegraphics[width=\textwidth]{DiagramaCAetDG02.pdf}
\caption{Esquema}
\label{diagCA}
\end{subfigure}
\quad
\begin{subfigure}[b]{0.45\textwidth}
\includegraphics[width=\textwidth]{PseudoCoordinates01.pdf}
\caption{Coordenadas}
\label{pseudocor}
\end{subfigure}
\caption{ (\ref{diagCA}) Esquema del hipocámpo en nuestras rebanadas y su
colocación aproximada en el BioCAM. Se muestran los estratos somáticos.
(\ref{pseudocor}) Coordenadas ortogonales putativas, a lo largo del eje
``tangente al estrato piramidal'', $t$ y ``apical'', $a$. } 
\label{esquemas01}
\end{figure}

Obsérvese que para el argumento no es central el hecho de que estas coordenadas
estén realmente alineadas con la dirección apical y el estrato piramidal, 
lo único importante es que existen ejes
principales para la conductividad en cada punto, y que la transformación
localmente puede ser vista como una rotación (ya que los eigenvectores pueden
escogerse normales). Voy  a denotar por un tilde a los objetos representados
en las coordenadas de los ejes principales, y por $T$ 
a la transformación diferencial:
\begin{equation}
\tilde{\cond}=
\begin{pmatrix}
\sigma_{t} & 0 \\
0 & \sigma_{a}
\end{pmatrix} = T \cond T^{-1}
\end{equation}

\begin{equation}
  \tilde{\Jd} = \tilde{\cond}\tilde{\EF}= T\Jd= (T \cond T^{-1}) T \EF = T\cond \EF
\end{equation}

\bibliographystyle{plain}
\bibliography{BiblioReportes01}

\end{document}

### 2015-05-15 Reportes/CSDAyOrthCoord01.tex
bbuu

@@
 \DeclareMathOperator{\diver}{div}
+\DeclareMathOperator{\grad}{grad}
 
@@
 normal a éste, que llamare ``apical'' por analogía, véase la fig. 
-\cite{esquemas01}.
+\ref{esquemas01}.
 
@@
 \end{equation}
-
 \begin{equation}
@@
 \end{equation}
+Entonces,
+\begin{equation}\label{imcambio}
+\tilde{I_m} = \widetilde{\diver} \tilde{\Jd} = \widetilde{\diver} T\cond \EF
+\end{equation}
+Como podemos escoger los eigenvectores de $\cond$ normalizados, $T$ es 
+una localmente una rotación, digamos que por el ángulo $\theta$. Esto quiere
+decir que en las coordenadas $(x,y)$ la conductividad tiene la 
+representación
+\begin{equation}
+\cond=
+\begin{pmatrix}
+\cos^2 \theta \sigma_t+\sen^2 \theta \sigma_a & \cos\theta \sen\theta (\sigma_t-\sigma_a) \\
+\cos\theta \sen\theta (\sigma_t-\sigma_a) &\sen^2 \theta \sigma_t+\cos^2 \theta \sigma_a
+\end{pmatrix} 
+\end{equation}
+usando
+\begin{equation}
+T^{-1}=
+\begin{pmatrix}
+\cos \theta  & - \sen\theta \\
+\sen\theta  &\cos \theta 
+\end{pmatrix} 
+\end{equation}
+Obtenemos $\EF$ aplicando el gradiente del potencial electroestático local. 
+En el sistema cartesiano usual,
+\begin{equation}
+\EF=-\grad \phi=-(\partial_x \phi, \partial_y \phi)
+\end{equation}
+La ecuación \ref{imcambio} queda entonces asì:
+\begin{equation}\label{imcambio}
+\tilde{I_m} = \widetilde{\diver} 
+(E_x\sigma_t \cos\theta - E_y \sigma_a \sen \theta,
+E_x\sigma_t \sen\theta + E_y \sigma_a \cos \theta)
+\end{equation}
+... y ha hacer talaaacha!
 

### 2015-05-29 Reportes/CSDADoneRight01.tex
EmpezandoPaper

\documentclass{article}

%usepackage[utf8]{inputenc}
 \usepackage{fontspec}
\usepackage{amsmath}
%\usepackage[spanish]{babel}
\usepackage{graphicx}

\usepackage{caption}
\usepackage{subcaption}

\setmainfont[Ligatures=TeX]{Oldstyle HPLHS}

\newcommand{\Jd}{\mathbf{J}}
\newcommand{\EF}{\mathbf{E}}
\newcommand{\cond}{\boldsymbol{\sigma}}
\DeclareMathOperator{\diver}{div}
\DeclareMathOperator{\grad}{grad}

\title{CSDA in 2D slices of the hippocampus \emph{done right}}
\author{W. P. K. Zapfe, Whoever Helps}

\begin{document}

\maketitle

\begin{abstract}
As high density electrode arrays become cheaper and more common, 
electrophysiological measurements of activity in neural tissue approach
the resolution of the studied structures. The electrodes measure
directly the Local Field Potential, but it is aknoledged that the
Current Source Density has more resolving power, as it points
more accurately to the actual sources and sinks of ionic current,
and thus, to the more active places in the tissue. Nevertheless,
the role of the conductivity tensor is rarely taken into account
in the derivation of the CSD, which could potentialy lead to
spurious distributions of the sources and sinks and deformation
of their structure, obfuscating the real activity underneath.
Here we show a possible qualitative treatment of the conductivity
across the stratum piramidale of the hippocampus, an expemplification
with two toy models and a possible numerical implementation with
high density real data.
\end{abstract}

\section{Introduction}

As Bédard and Destexhe pointed out in \cite{Bedard11}, the CSDA, as it
is performed usually, rests on
assumptions which could leave important features out. One of these 
assumptions is that the extracellular medium has very homogeneous 
electrical propierties, but this has been shown to not be
the case for dense neuronal layers such as the piramidal
strata in the hippocampus \cite{Holsheimer87, Lopez01, MarioPersonal}.
The evidence mostly points out that resistivity grows 
considerably around the somatic layers. Here we shall investigate
posible influence of this change of resistivity on current density
($\Jd$) and current source density ($I_m$).

\section{Analytical  derivation of the CSD}

For a carefull and complete derivation and discution of the CSD, 
see \cite{Bedard11}. We start with the charge conservation equation,
\begin{equation}
\nabla \cdot \Jd + \frac{d \rho}{d t} =0.
\end{equation}

\end{document}

### 2015-06-02 Reportes/CSDADoneRight01.tex
huevadas de eq

@@
 
-%usepackage[utf8]{inputenc}
+%\usepackage[utf8]{inputenc}
 \usepackage{fontspec}
 \usepackage{amsmath}
-%\usepackage[spanish]{babel}
+\usepackage[spanish]{babel}
 \usepackage{graphicx}
@@
 the resolution of the studied structures. The electrodes measure
-directly the Local Field Potential, but it is aknoledged that the
+directly the Local Field Potential, but it is aknowledged that the
 Current Source Density has more resolving power, as it points
-more accurately to the actual sources and sinks of ionic current,
+more accurately to the actual sources and sinks themselves,
 and thus, to the more active places in the tissue. Nevertheless,
@@
 spurious distributions of the sources and sinks and deformation
-of their structure, obfuscating the real activity underneath.
+of their structure, obfuscating the precise place of the
+poles underneath.
 Here we show a possible qualitative treatment of the conductivity
-across the stratum piramidale of the hippocampus, an expemplification
+across the stratum piramidale of the hippocampus, an examplification
 with two toy models and a possible numerical implementation with
-high density real data.
+high density experimental data.
 \end{abstract}
@@
 The evidence mostly points out that resistivity grows 
-considerably around the somatic layers. Here we shall investigate
-posible influence of this change of resistivity on current density
+considerably around thick somatic layers, or conversely, that
+the conductivity drops considerably there. In ohmic (linear) media,
+we expect the following relation to hold:
+\begin{equation}
+\Jd=\cond \EF,
+\end{equation}
+that is, the current density $\Jd$ is the result of the
+local electric field $\EF$ and the local conductivity, $\cond$.
+We usually take that last quantity as an homogeneous, isotropic,
+factor, at least over the scales that we are interested in.
+The conductivity in inhomogeneous media can be rarely
+simplified to a constant, as it can depend both from the direction
+of the fields and currents, and vary from point to point. 
+Then the above expression is just shorthand notation for the
+explicit  \emph{tensorial} relationship:
+\begin{equation}
+  \begin{pmatrix}
+    J_x \\
+    J_y \\
+    J_z
+  \end{pmatrix}
+  =
+   \begin{pmatrix}
+     \sigma_{xx}E_x+ \sigma_{xy}E_y+\sigma_{xz}E_z \\
+     \sigma_{xy}E_x+ \sigma_{yy}E_y+\sigma_{yz}E_z \\
+     \sigma_{xz}E_x+ \sigma_{yz}E_y+\sigma_{zz}E_z \\
+  \end{pmatrix},
+\end{equation}
+where we allready used the fact that $\cond$ is simmetric when
+expressed on cartessian coordinates.
+We assume that the electric field can be obtained as the
+gradient of the LFP, $\phi$:
+\begin{equation}
+  -\grad \phi=\EF
+\end{equation}
+This LFP is the quantity that the microelectrode arrays (MEAs) permits us
+to measure on the extracelular medium.
+To obtain the sources and sink densities, we use Gauss Divergence Theorem,
+stated in differencial form:
+\begin{equation}
+\nabla \cdot \Jd =I_m
+\end{equation}
+We perform the divergence operation on the current density, not
+on the electric field. Only when the medium is isotropic and homogeneous
+we can be sure that both operations coincide within a positive factor.
+
+Here we shall investigate
+posible influence of changing conductivity on current density
 ($\Jd$) and current source density ($I_m$).
@@
 \section{Analytical  derivation of the CSD}
+
+In experiments of this sort performed on hippocampus slices, the
+MEAs permit us to measure a sample of LFP over discrete points on a
+two dimensional lattice, so we do not have possibility of
+calculating currents or electric fields on the third dimension of space.
+The slices of the hippocampus are very thin, and we do not consider
+the posibility of current going outside the neural tissue sample (in one
+direction there is air and on the other the MEA), so we
+can reduce our formalism to two dimensional space. We then take
+$J_z=0$ and use $(x,y)$ as a cartesian coordinate system aligned
+with the MEA. Then the \emph{flat} current density  is:
+\begin{equation}
+  \begin{pmatrix}
+    J_x\\
+    J_y
+  \end{pmatrix}  =
+  \begin{pmatrix}
+    \sigma_{xx}E_x+\sigma_{xy}E_y \\
+    \sigma_{xy}E_x+\sigma_{yy}E_y 
+  \end{pmatrix}=
+  \begin{pmatrix}
+    -\sigma_{xx}\partial_x \phi -\sigma_{xy}\partial_y \phi \\
+    -\sigma_{xy}\partial_x \phi -\sigma_{yy}\partial_y \phi 
+  \end{pmatrix}.
+\end{equation}
+The divergence, in cartesian coordinates, yields:
+\begin{equation}\label{csd2dexp}
+  I_m=\partial_x J_x+\partial_y J_y=
+  -\partial_x( \sigma_{xx}\partial_x \phi +\sigma_{xy}\partial_y \phi)
+  -\partial_y( \sigma_{xy}\partial_x \phi +\sigma_{yy}\partial_y \phi). 
+\end{equation}
+Each one of the terms of the right hand side of eq. \ref{csd2dexp}
+has the form
+\begin{multline}
+  \partial_j (\sigma_{jj}\partial_j\phi) +\partial_j (\sigma_{jk}\partial_k \phi)
+  =  \\
+  (\partial_j \sigma_{jj})(\partial_j\phi) +\sigma_{jj}\partial_j^2\phi +
+  (\partial_j \sigma_{jk})(\partial_k \phi)+\sigma _{jk}\partial_j\partial_k \phi.
+\end{multline}
+The complete expression is cumbersome to write, but it is important to
+have \emph{full} expansion in order to make the numerical considerations
+in what follows.
+\begin{equation}
+  \begin{split}
+  -I_m= & (\partial_x \sigma_{xx}) (\partial_x \phi) +
+  (\partial_y \sigma_{yy}) (\partial_y \phi)\\
+& +(\partial_x \sigma_{xy}) (\partial_y \phi) +(\partial_y \sigma_{xy}) (\partial_x \phi)\\
+& + \sigma_{xx} \partial_x^2 \phi + \sigma_{yy} \partial_y^2 \phi\\
+& + 2\sigma_{xy} \partial_x\partial_y \phi. 
+\end{split}
+ \end{equation}
 

### 2015-06-02 Reportes/CSDADoneRight01.tex
Abstract Epileptic Soc.

@@
 the case for dense neuronal layers such as the piramidal
-strata in the hippocampus \cite{Holsheimer87, Lopez01, MarioPersonal}.
+strata in the hippocampus \cite{Holsheimer87, Lopez01, TrevinoPersonal}.
 The evidence mostly points out that resistivity grows 
@@
 
+\bibliographystyle{plain}
+\bibliography{BiblioReportes01}
+
 \end{document}

### 2015-06-03 Reportes/CSDADoneRight01.tex
Tengo un desmadre en las derivadas

@@
 \usepackage{amsmath}
-\usepackage[spanish]{babel}
+
 \usepackage{graphicx}
@@
 in what follows.
-\begin{equation}
+\begin{equation}\label{Imcart}
   \begin{split}
@@
  \end{equation}
+
+\begin{figure}[h]
+\centering
+\begin{subfigure}[t]{0.20\textwidth}
+\includegraphics[width=\textwidth]{DiagramaCAetDG02.pdf}
+\caption{}
+\label{diagCA}
+\end{subfigure}
+\quad
+\begin{subfigure}[t]{0.33\textwidth}
+\includegraphics[width=0.67\textwidth]{DiferentialCoordinates01.pdf}
+\caption{}
+\label{loceig}
+\end{subfigure}
+\begin{subfigure}[t]{0.20\textwidth}
+\includegraphics[width=\textwidth]{PseudoCoordinates01.pdf}
+\caption{}
+\label{pseudocor}
+\end{subfigure}
+\caption{\ref{diagCA}) A diagram showing the position of
+the hippocampus slice over the MEA. \ref{loceig}) 
+A tangent space basis along the
+eigenvectors of $\cond$. The red vectors indicate the
+Cartesian Coordinates aligned with the MEA latice. 
+\ref{pseudocor}) A putative local coordinate system spawned by the 
+eigenvectors $(t,a)$ at every point of intereset. 
+ }\label{diagdif}
+\label{esquemas01}
+\end{figure}
+
+This expressions may be regardes as inescesarily long, as the 
+$\cond$ tensor, being a real symmetric tensor, can be represented
+in local coordinates which diagonalize it. Those coordinates have
+physical meaning, as they are the resolving directions in which
+conductivity behaves differently and denote some underlying structure
+in the studied material. The usual assumption is that 
+the conductivity tensor is isotropic but not homogeneous, and it 
+changes along the apical direction, and not along the 
+differentiated strata (e.g. stratum pyramidale, stratum granolosum).
+Let us suppose that this is approximately correct, and that very close
+to the stratum pyramidale we represent $\cond$ into the coordinates
+defined by its eigenvectors (see fig. \ref{esquemas01}).
+ We shall call them $(t,a)$ coordinates
+and asume that the first is approximately tangent to the
+strata and the other to the apical direction. The $\cond$ tensor would 
+have the expression in such coordinates
+\begin{equation}
+\tilde{\cond}=
+\begin{pmatrix}
+\sigma_t & 0 \\
+0 & \sigma_a
+\end{pmatrix}
+\end{equation} 
+Then, the equation \ref{Imcart} becomes
+\begin{equation}\label{Imapic}
+\begin{split}
+-I_m= & (\partial_t \sigma_{t}) (\partial_t \phi) +
+(\partial_a \sigma_{a}) (\partial_a \phi)  \\
+& + \sigma_{a} \partial_a^2 \phi + \sigma_{t} \partial_t^2 \phi, 
+\end{split}
+\end{equation}
+as $(t,a)$ are also local ortogonal coordinates. 
+One may wonder why we are explicitly stating the non-diagonal 
+version of the equations. The data as it is delivered by
+the measuring apparatus is on a strictly cartessian grid, and
+we have to perform the numerical derivatives over this data, so
+we have to know how these assumptions appear in both sets
+of coordinates. Besides, the \emph{tangent-apical} coordinates
+are just a teoretical tool, a real implementation of such a coordinate
+set would be extremely complicated and prone to numerical errors.
+
+If the medium where the ions diffuse is liquid, and the structures
+therein are randomly distributed, the assumption of isotropy holds, 
+$\sigma_t=\sigma_a=\sigma$. Then, as we stated, there is no
+change along the strata, but there is change along the appical
+direction:
+\begin{align}
+\partial_t \sigma & =0 \\
+\partial_a \sigma & \neq 0. 
+\end{align}
+We are then left with
+\begin{equation}
+I_m=-\partial_t \sigma \partial_t \phi - \sigma \nabla^2 \phi.
+\end{equation}
+Here we must point out that the Laplacian is an invariant operator,
+so the last summand of the equation could be in principle be implemented
+in \emph{any} convenient set of coordinates, but in numerical aplications,
+the \emph{Discrete Laplacian Operator} is \emph{not} invariant (further
+discution in the following section). 
+The term $\partial_t \sigma \partial_t \phi$ is the hindrance to 
+ionic displacement due to dense layers of soma and dendrites near the
+CA structure. Most autors have neglected its efects, and in one
+dimensional analysis the structure of sources and sinks is not affected
+by this. Now, as the aquisition of data in two dimensions becomes
+finer, this term could lead to inexactitude in the location of the
+sinks and sources, depending on the magnitude of the
+derivative of the conductance. Locally both $(x,y)$ and
+$(t,a)$ are orthonormal systems, so we can change from one to another 
+with a rotation. If $theta$ is the angle between the $x$ vector
+and $t$, the partial derivatives follow the next set of relations:
+\begin{align}
+\partial_t \sigma &= \cos\theta \partial_x \sigma +\sin\theta \partial_y \sigma=0\\
+\partial_a \sigma &= -\sin\theta \partial_x \sigma +\cos\theta \partial_y \sigma
+\end{align}
+
+After a bit of algebraic juggling we end up with
+\begin{equation}
+-I_m=\sigma\nabla^2 \phi + 
+\partial_x\sigma (\partial_x \phi - \tan \theta \partial_y \phi)
+\end{equation}
+
+In order to estimate the influence
+of this summand on the result, we shall make some toy models which
+represent the variation of $\sigma$ along the apical direction.
 

### 2015-06-05 Reportes/CSDADoneRight01.tex
cambios pecuecos

@@
 of coordinates. Besides, the \emph{tangent-apical} coordinates
-are just a teoretical tool, a real implementation of such a coordinate
-set would be extremely complicated and prone to numerical errors.
+are just a teoretical tool, a numerical implementation of such a coordinate
+set would be  complicated and prone to numerical errors.
 
 If the medium where the ions diffuse is liquid, and the structures
-therein are randomly distributed, the assumption of isotropy holds, 
-$\sigma_t=\sigma_a=\sigma$. Then, as we stated, there is no
+therein are randomly distributed, the assumption of isotropy may hold 
+on average in sufficiently big scales \cite{Bedard11}. 
+In order to not make the notation more cumbersome, let us aggree
+that we are dealing with averages over volumes of the aproximate size
+of the electrodes. Then, 
+$\sigma_t=\sigma_a=\sigma$. As we stated, there is no
 change along the strata, but there is change along the appical

### 2015-06-30 Reportes/ConexionesyCriteriosdeCorrelacion01.tex
Reporte Networks

\documentclass{article}

%\usepackage[utf8]{inputenc}
\usepackage{fontspec}
\usepackage{amsmath}
\usepackage[spanish]{babel}

\usepackage{graphicx}

\usepackage{caption}
\usepackage{subcaption}

\setmainfont[Ligatures=TeX]{Oldstyle HPLHS}

\newcommand{\Jd}{\mathbf{J}}
\newcommand{\EF}{\mathbf{E}}
\newcommand{\cond}{\boldsymbol{\sigma}}
\DeclareMathOperator{\diver}{div}
\DeclareMathOperator{\grad}{grad}

\title{Criterios de Conexión causal entre GD y CA}
\author{F. Ortiz,  WPK.Zapfe,  MA. Aviño }

\begin{document}

\maketitle

\section{Introduction}

En este momento, contamos con datos de alta densidad temporal
y espacial de actividad eléctrica de rebanadas ventrales del 
hipocampo de las ratas. Nuestras medidas son potenciales eléctricos
medidos en el medio extracelular inmediato a las estructuras de nuestr
interés, es decir, el giro dentado (GD) y el \emph{cuerno de Amón} (CA). 
En ese medio extracelular es posible detectar, de forma indirecta,
los potenciales de acción, cambios abruptos en el potencial
eléctrico, acción del disparo de las neuronas. Un criterio \emph{ad hoc}
nos permite establecer un momento temporal asociado al pico
del potencial de acción. La pregunta fundamental es la siguiente:
¿Qué partes del GD hacen disparar a qué partes del CA? 
Pragmáticamente esperariamos que un disparo detectado en GD fuera seguido
por uno en CA después de un intervalo de tiempo, entre 2ms  a 5 ms.  
Si con nuestro criterio de detección de picos encontramos dichos eventos
consecutivos, ¿qué tan significativo son como criterio de relación causal?
Aquí mostramos tres gráficas con criterios de significancia distinta y 
discutimos brevemente sus virtudes y defectos.

\section{Los disparos \emph{eco}}

Para no empantanar ciertos términos, vamos a describir un putativo
potencial de acción en CA como \emph{eco} si ocurre entre 
2 y 5 milisegundos después que un disparo de GD. Este término
no implicará ninguna causalidad, simplemente que ocurre dentro
del intervalo de tiempo esperado para una conexión sináptica.
Para darle un significado de conectiviad plausible, necesitamos
un criterio que establezca un nivel de causalidad entre los disparos.
Los disparos eco como medida absoluta pueden ya indicar una cierta
relación de este tipo entre GD y CA, así que proponemos, antes
que nada, observar la gráfica de ecos meramente y ver si tiene
algun sentido dentro de la fisiología ya conocida. En la figura 
\ref{EcosTotales} presentamos dicha gráfica.

\begin{figure}[h]
\centering
\includegraphics[width=0.85\textwidth]{ConexionesEcos01.png}
\caption{ Los disparos ecos representados como lineas entre
los electrodos involucrados. Los puntos violeta representan 
electrodos debajo del GD, mientras que los negros
debajo de CA. El grosor y nivel de rojo de las lineas indican
el número de ecos totales encontrados. La linea mas gruesa
representa 210 ecos, mientras que la más delgada uno solo. 
Nótese lo poco homogeneo de la distribución. Los electrodos
que no registraron ecos no aparecen. 
}\label{EcosTotales}
\end{figure}

Notemos de esta primera figura lo poco homogeneo de la distribución.
Mientras que hay 6 electródos del GD fuertemenente conectados
a diversas regiones de CA, la mayoría de los puntos registran muy pocos
ecos, o activiad de algun tipo. Particularmente hay un electrodo en
la región infrapiramidal de GD que muestra altísima conectividad con
la región media y distal de CA3c, con una buena parte de CA3b y el
centro de CA3a. De igual forma podemos apreciar un grupo de vecinos
en la frontera entre la cresta y la región infrapiramidal que
muestra también una conexión similarmente fuerte a 
las mismas regiones. El último electrodo de la región infrapiramidal
Debajo de la cresta también encontramos un gripo de tres electrdoos
fuertemente conectados con las mismas regiones. En esta gráfica
puede haber muchos ecos casuales, es decir, falsos positivos,
dada la actividad notoria de todo CA, aún cuando GD parece 
estar silente. En cierta forma, esta imagen no está adecuadamente
normalizada a la actividad ``promedio'' de una región u otra y
de ahí la gran variedad de pesos que encontramos. Aun así, esta
gráfica tiene información que puede ser analisada formalmente y
dar resultados significativos. En particular convendría 
hacer un análisis de la distribución de las conexiones así establecidas.
Es conocimiento común que las redes que ocurren en fenómenos biológicos
deben de presentar ciertas características generales, como son 
robustez (estabilidad frente a perturbaciones o fallas), redundancia
(nodos que duplican o refuerzan conexiones), y ser \emph{libres de escala}.
Esto hace que sus conexiones sigan cierta distribución estadística
expresable como una ley de potencias con exponentes entre 1.4 y 3.
Deberíamos de revisar si es el caso en esta primera aproximación.

La siguiente figura (fig. \ref{Jacard} muestra los ecos 
después de  una normalización que propuso
la profesora Aviño, cuyo resultante es conocido como 
``coeficiente de similiaridad de Jaccard'', y escencialmente 
intenta medir que tan iguales son dos conjuntos a su intersección. 
En este caso estamos contando los ecos como una intersección de eventos
entre GD y CA. A pesar de que este coeficiente ha sido usado en la litaretura
con propósitos similares (), admitimos que encontramos su aplición un 
tanto dudosa para nuestro caso por lo siguente. Dada la gran actividad
que presenta CA en relación a GD, estamos dividiendo el número
de ecos por números muy  grandes que representan actividad que no esta
relacionada de forma alguna con la conexión que estamos buscando. Dicho de 
otra forma: interpretar los disparos ecos como una intersección de
eventos entre dos posibles conjuntos de otros eventos no parece
tener mucho sentido. De todas formas  hay que ver que posibles
detalles revela la figura y ver su posible implicación fisiológica.

\begin{figure}[h]
\centering
\includegraphics[width=0.85\textwidth]{ConexionesJacard01.png}
\caption{ Los ecos divididos entre el la suma de disparos
\emph{sin eco} entre los electrodos relacionados. De igual forma
lineas más gruesas y más rojas indican valor más alto, sin embargo,
la escala es diferente, dado que aquí el mayor valor posible sería uno.
En esta gráfica los grosores están exagerados, dado que el el 
coeficiente de Jaccard da valores muy bajos (del orden de milésimas).
}\label{Jacard}
\end{figure}

Es notorio en esta segunda figura que los mismos electrodos muestran
la misma conectividad alta que en los ecos crudos, sin embargo el 
peso relativo que tienen es mucho más homogeneo. A su vez otros pares
de electrodos cobran un peso mucho mayor, debido a la baja cantidad
de picos registrados en ellos, en particular está el par que une
a un elemento de la hoja suprapidamidal con el centro de CA3a. 
Sin embargo no debemos de concederle aún excesivo valor a esto
sin corroborar con otra serie de datos, pues puede ser un 
artefacto debido a la baja actividad detectada en ambos sitios. 
Por otro lado el tercer electrodo de derecha a izquierda de la región
infrapiramidal muestra aún una conexión fuerte con el centro de 
CA3a. Esta en particular puede ser ya una pista importante
sobre cierto grupo de conexiones operando en estas regiones. Otra
cosa que podría decirse de esta figura es que, siendo la normalización
tan baja, permite una resolución visiual más clara en la 
estructura de abaníco formada por los electrodos más conectados.

Finalmente Karel propuso normalizar el número de ecos entre el número
de \emph{ecos posibles}, esto es, el número menor entre los disparos
de un electrodo en GD y otro en CA. Típicamente este resulta ser
el correspondiente a GD, pero no siempre. Dado que no puede haber más 
``ecos'' que el número menor de disparos en un par de electrodos, 
esto nos normaliza los ecos a un valor entre 0 y 1 donde 0 es
``ningun eco'' y 1 es ``todos los disparos tienen un eco o son un eco''.
La figura \ref{MinimalCrit} muestra este criterio de peso en la
gráfica.

\begin{figure}[h]
\centering
\includegraphics[width=0.85\textwidth]{ConexionesPseudoCausales01.png}
\caption{ Los ecos divididos entre el la suma de disparos
\emph{sin eco} entre los electrodos relacionados. De igual forma
lineas más gruesas y más rojas indican valor más alto, sin embargo,
la escala es diferente, dado que aquí el mayor valor posible sería uno.
En esta gráfica los grosores están exagerados, dado que el el 
coeficiente de Jacard da valores muy bajos (del orden de milésimas).
}\label{MinimalCrit}
\end{figure}

La figura \ref{MinimalCrit}
es notoriamente diferente de las otras dos. En particular
muestra una conexión mucho más homogenea de todo el GD a la 
región CA3c, y muestra conexiones muy fuertes en la cresta del GD 
hacia varios electrodos de CA3c y la frontera con CA3b, conexiones
que aparecen de muy bajo peso en las dos figuras anteriores.
De la misma forma que la linea más pesada de la figura \ref{Jacard}
puede deberse a un número muy bajo de registros, lo mismo puede
suceder aquí, y no debemos de confiar en estos valores hasta no 
corroborar con mayores datos. Este criterio podría, tal vez, 
revelarnos conexiones discretas, sin inferencias de peso sobre 
las diferentes regiones, pero para ello necesitariamos muestras
con mayores datos del GD. Afortunadamente el problema del procesamiento
de los datos para buscar pares de ecos ya se encuentra bastante 
más avanzado, y podemos analizar juegos de datos mucho mayores en menor
tiempo, así que el problema consiste ahora en encontrar mejores
métodos para detectar actividad espontanea en el giro dentado. Por
otra parte, aquí no aparecen considerados otro tipo de respuestas
diferentes al potencial de acción como eco.

\end{document}

### 2015-06-30 CenterOfMass/PloteaCMdisjoint02.sh

@@ -2,7 +2,7 @@
 
 ##ploteando un chingo con varios archivos
 
-datosfondobasename="CSDPure-"
+datosfondobasename="CSD-"
 datosCMNegname="CMNegLapla-"
 datosCMPosname="CMPosLapla-"
 numero=$1
@@ -45,8 +45,8 @@ set label "Sink" at graph 1.1, 0.02
 set style fill transparent solid 0.95 noborder
 #set palette defined (0 "white", 1 "#0000FF", 2 "black")
 set out "$outname"
-#plot "$dataname" using $usandomatrix matrix  w image, "$datosCMNeg" usi $usandonegativo  w circles lw 3 lc rgb "#000075"
-plot  "$datosCMNeg" usi $usandonegativo  w circles lw 3 lc rgb "#000075", "$datosCMPos" usi $usandopositivo  w circles lw 3 lc rgb "#750000"
+plot "$dataname" using $usandomatrix matrix  w image #, "$datosCMNeg" usi $usandonegativo  w circles lw 3 lc rgb "#000075"
+#plot  "$datosCMNeg" usi $usandonegativo  w circles lw 3 lc rgb "#000075", "$datosCMPos" usi $usandopositivo  w circles lw 3 lc rgb "#750000"
 set out
 
 EOF

### 2015-07-01 Reportes/CSDAattheSomaticLayersScale.tex
Pum pum CSDA

--- 2015-06-05 Reportes/CSDADoneRight01.tex
@@
 
-\title{CSDA in 2D slices of the hippocampus \emph{done right}}
+\title{Two Dimensional CSDA at the scale of neuronal tissue.}
 \author{W. P. K. Zapfe, Whoever Helps}
@@
 electrophysiological measurements of activity in neural tissue approach
-the resolution of the studied structures. The electrodes measure
+the scale of the underlying structures. The electrodes measure
 directly the Local Field Potential, but it is aknowledged that the
-Current Source Density has more resolving power, as it points
-more accurately to the actual sources and sinks themselves,
-and thus, to the more active places in the tissue. Nevertheless,
-the role of the conductivity tensor is rarely taken into account
-in the derivation of the CSD, which could potentialy lead to
-spurious distributions of the sources and sinks and deformation
-of their structure, obfuscating the precise place of the
-poles underneath.
+Current Source Density has more resolving power and distinguishes
+clearly between the loci of outgoing current and ingoing currents.
+ The role of the conductivity tensor is rarely taken into account
+in the derivation of the CSD, and this could potentialy lead to
+spurious distributions of the sources and sinks- 
+In dense neural tissue, such as the
+stratum piramidale layer of the hippocampus, clear
+deviations from average conductivity have been measured. 
 Here we show a possible qualitative treatment of the conductivity
-across the stratum piramidale of the hippocampus, an examplification
-with two toy models and a possible numerical implementation with
-high density experimental data.
+across the layers of the hippocampus, and compare the results
+when this inhomogenity is taken into account and when it not.
 \end{abstract}
@@
 assumptions is that the extracellular medium has very homogeneous 
-electrical propierties, but this has been shown to not be
+electrical propierties. This has been shown to not be
 the case for dense neuronal layers such as the piramidal
@@
 where we allready used the fact that $\cond$ is simmetric when
-expressed on cartessian coordinates.
+expressed on Cartessian coordinates.
 We assume that the electric field can be obtained as the
@@
 \begin{equation}
-  -\grad \phi=\EF
+  -\nabla \phi=\EF
 \end{equation}
@@
 To obtain the sources and sink densities, we use Gauss Divergence Theorem,
-stated in differencial form:
 \begin{equation}
@@
 \end{equation}
-We perform the divergence operation on the current density, not
-on the electric field. Only when the medium is isotropic and homogeneous
-we can be sure that both operations coincide within a positive factor.
-
-Here we shall investigate
-posible influence of changing conductivity on current density
-($\Jd$) and current source density ($I_m$).
+The usual definition of $-\partial_t \rho =I_m$ has been made here. 
+We perform the divergence operation on the current density, \emph{not}
+on the electric field. When the medium is isotropic and homogeneous
+both derivations coincide within a positive factor. 
+In the next section we investigate the role of the inhomogenities 
+of the conductivity on current density
+($\Jd$) and current source density ($I_m$) at scales around
+the $20\mu m$.
 
@@
 The divergence, in cartesian coordinates, yields:
-\begin{equation}\label{csd2dexp}
-  I_m=\partial_x J_x+\partial_y J_y=
-  -\partial_x( \sigma_{xx}\partial_x \phi +\sigma_{xy}\partial_y \phi)
-  -\partial_y( \sigma_{xy}\partial_x \phi +\sigma_{yy}\partial_y \phi). 
-\end{equation}
-Each one of the terms of the right hand side of eq. \ref{csd2dexp}
-has the form
-\begin{multline}
-  \partial_j (\sigma_{jj}\partial_j\phi) +\partial_j (\sigma_{jk}\partial_k \phi)
-  =  \\
-  (\partial_j \sigma_{jj})(\partial_j\phi) +\sigma_{jj}\partial_j^2\phi +
-  (\partial_j \sigma_{jk})(\partial_k \phi)+\sigma _{jk}\partial_j\partial_k \phi.
-\end{multline}
-The complete expression is cumbersome to write, but it is important to
-have \emph{full} expansion in order to make the numerical considerations
-in what follows.
 \begin{equation}\label{Imcart}
@@
 & +(\partial_x \sigma_{xy}) (\partial_y \phi) +(\partial_y \sigma_{xy}) (\partial_x \phi)\\
-& + \sigma_{xx} \partial_x^2 \phi + \sigma_{yy} \partial_y^2 \phi\\
-& + 2\sigma_{xy} \partial_x\partial_y \phi. 
+& + 2\sigma_{xy} \partial_x\partial_y \phi \\
+& + \sigma_{xx} \partial_x^2 \phi + \sigma_{yy} \partial_y^2 \phi.
+\end{split}
+ \end{equation}
+The \emph{tacit} assumption that can be noted in most of the CSDA previously
+done is that the derivatives of $\cond$ are negligible and only
+the terms on the last line are  taken into account. 
+It can be argued that for scales that
+encompas many neurons and interneuronal space, the averaged conductivity is
+constant and has then its derivatives vanish. For MEAS that take measuraments
+more than $100 \mu m$ appart, this may be a reasonable assumption (see
+for example, eq. 12 in \cite{Bedard11} and the arguments leading to it).
+Examples of such derivations are those obtained by the MED64 probe in the
+work of Wang et all \cite{Wang10}, etc etc. Also sofisticated
+numerical methods, such as iCSD \cite{Pettersen2006, Leski11}
+and kCSD \cite{Potworowsky2011} use this approximation, albeit
+explicitly.
+
+If the interelectrode spacing
+is less than $100 \mu m$ then the neuronal tissue would begin
+to play a role in the inhomogenities of $\cond$, although the isotropic
+assumption may still be valid at scales larger than $20 \mu m$. At such 
+scales the laminar structure of the Cornu Amonis or the Gyrus Dentate is
+the main feature, and we should expect variations on the electrical 
+propierties of the media along the apical direction. Measurements made
+by Holsheimer showed that measurements in the CA1 region of a mouse
+hipocampus show variations of  $40\% $ in the resistivity around the mean
+value at distances of about $60 \mu m$.
+\cite{Holsheimer87}. This measurements have been confirmed
+recently by a collaborator \cite{TrevinoPersonal}. On the other hand
+we expect very little variation of the conductivity tensor along
+each structural layer. The symmetry of the conductivity tensor represents
+the structural underlying propierties of the media: as it can be 
+represented on a diagonal form on a suitable set of \emph{orthogonal}
+coordinates, then such coordinates must be associated with the 
+structures that give $\cond$ its values. Then the representation
+of $\cond$ in diagonal form should represent the conductivity
+along the axonal-apical direction, with high variation, and another
+axis transverse to it, in which $\cond$ showing very little variation.
+We can imagine that we can set a locally orthonormal set of coordinates
+alignate with such structures (see fig. \ref{esquemas01}).
+ Then the expression in eq. \ref{Imcart}
+can be reduced substantialy.  The $\cond$ tensor would appear as
+\begin{equation}
+\tilde{\cond}=
+\begin{pmatrix}
+\sigma_t & 0 \\
+0 & \sigma_a
+\end{pmatrix}
+\end{equation} 
+where $(t,a)$ are the transveral and axonal-apical local coordinates.
+The CSD would be obtained as
+\begin{equation}\label{Imapic}
+\begin{split}
+-I_m= & (\partial_t \sigma_{t}) (\partial_t \phi) +
+(\partial_a \sigma_{a}) (\partial_a \phi)  \\
+& + \sigma_{a} \partial_a^2 \phi + \sigma_{t} \partial_t^2 \phi. 
 \end{split}
@@
 \end{figure}
-
-This expressions may be regardes as inescesarily long, as the 
-$\cond$ tensor, being a real symmetric tensor, can be represented
-in local coordinates which diagonalize it. Those coordinates have
-physical meaning, as they are the resolving directions in which
-conductivity behaves differently and denote some underlying structure
-in the studied material. The usual assumption is that 
-the conductivity tensor is isotropic but not homogeneous, and it 
-changes along the apical direction, and not along the 
-differentiated strata (e.g. stratum pyramidale, stratum granolosum).
-Let us suppose that this is approximately correct, and that very close
-to the stratum pyramidale we represent $\cond$ into the coordinates
-defined by its eigenvectors (see fig. \ref{esquemas01}).
- We shall call them $(t,a)$ coordinates
-and asume that the first is approximately tangent to the
-strata and the other to the apical direction. The $\cond$ tensor would 
-have the expression in such coordinates
-\begin{equation}
-\tilde{\cond}=
-\begin{pmatrix}
-\sigma_t & 0 \\
-0 & \sigma_a
-\end{pmatrix}
-\end{equation} 
-Then, the equation \ref{Imcart} becomes
-\begin{equation}\label{Imapic}
-\begin{split}
--I_m= & (\partial_t \sigma_{t}) (\partial_t \phi) +
-(\partial_a \sigma_{a}) (\partial_a \phi)  \\
-& + \sigma_{a} \partial_a^2 \phi + \sigma_{t} \partial_t^2 \phi, 
-\end{split}
-\end{equation}
-as $(t,a)$ are also local ortogonal coordinates. 
-One may wonder why we are explicitly stating the non-diagonal 
-version of the equations. The data as it is delivered by
-the measuring apparatus is on a strictly cartessian grid, and
-we have to perform the numerical derivatives over this data, so
-we have to know how these assumptions appear in both sets
-of coordinates. Besides, the \emph{tangent-apical} coordinates
-are just a teoretical tool, a numerical implementation of such a coordinate
-set would be  complicated and prone to numerical errors.
 

### 2015-07-02 Reportes/CSDAattheSomaticLayersScale.tex
Graficas mandadas

@@
 across the layers of the hippocampus, and compare the results
-when this inhomogenity is taken into account and when it not.
+when this inhomogenity is taken into account and when its not.
 \end{abstract}
@@
 assumptions is that the extracellular medium has very homogeneous 
-electrical propierties. This has been shown to not be
+electrical propierties. It has been shown not to be
 the case for dense neuronal layers such as the piramidal
@@
 The evidence mostly points out that resistivity grows 
-considerably around thick somatic layers, or conversely, that
-the conductivity drops considerably there. In ohmic (linear) media,
-we expect the following relation to hold:
+considerably around thick somatic layers.
+The following linear relationship is a good aproximation for
+most media at convencional regimes: 
 \begin{equation}
@@
 local electric field $\EF$ and the local conductivity, $\cond$.
-We usually take that last quantity as an homogeneous, isotropic,
-factor, at least over the scales that we are interested in.
+We usually take this quantity as constant, 
+at least over macroscopic scales.
 The conductivity in inhomogeneous media can be rarely
-simplified to a constant, as it can depend both from the direction
+simplified this much, as it can depend both from the direction
 of the fields and currents, and vary from point to point. 
@@
 both derivations coincide within a positive factor.
+This makes sense when we are dealing with macroscopic quantities,
+that is, that the are at last one order of magnitude larger
+than the underlying physical structures which could cause 
+inhomogenities, but as our measurements get finner and finner,
+this treatment begins to be faulty. 
 In the next section we investigate the role of the inhomogenities 

### 2015-07-03 Reportes/CSDAattheSomaticLayersScale.tex
corrigiendo textos

@@
 Here we show a possible qualitative treatment of the conductivity
-across the layers of the hippocampus, and compare the results
-when this inhomogenity is taken into account and when its not.
+across the layers of the hippocampus based on coarse measurements
+and general smoothness arguments.
 \end{abstract}
@@
 assumptions which could leave important features out. One of these 
-assumptions is that the extracellular medium has very homogeneous 
-electrical propierties. It has been shown not to be
-the case for dense neuronal layers such as the piramidal
-strata in the hippocampus \cite{Holsheimer87, Lopez01, TrevinoPersonal}.
-The evidence mostly points out that resistivity grows 
-considerably around thick somatic layers.
+is that the extracellular medium has very homogeneous 
+electrical propierties. This may depend on the scale in which measurements
+take place and the size of the neuronal structures involved. We shall
+focus our atention on the hippocampus of the rat and mouse
+ and its typical sizes in what follows.
+
+The CSDA is the application of charge conservation theorem
+for ionic dilutions in the extracelular medium for neuronal
+tissue. CSD are obtained either directly through application
+of a numeric Laplacian operator directly to LFP data \cite{Nicholson75}, 
+or through  more refinated
+techniques, called inverse CSDA \cite{Pettersen2006, Leski2011} or kernel CSDA 
+\cite{Potworowsky2011}. There the conductivity $\cond$ is taken as 
+ a numerical constant, that is:
+\begin{equation}
+-\nabla^2 \Phi = \frac{\Jd}{\cond}. 
+\end{equation}
+Here $\Phi$ is the LFP and $\Jd$ the current density. 
+This is indeed a reasonable treatment at scales larger than $100\mu m$,
+where each electrode measures averaged contributions from many different
+densities of activity and structures. Even more, data are 
+generaly smoothed out for presentation, and this contributes to the
+averaging process.
+
+On the other hand, it has been shown that at 
+dense neuronal layers such as the piramidal
+strata in the hippocampus the conductivity (or equivalently,
+the resistivity) varies greatly around its mean value along
+an axis which transverses said structure
+\cite{Holsheimer87, Lopez01, TrevinoPersonal}. 
+The evidence points out that resistivity grows 
+considerably around the thick somatic layers. These layers 
+have thickness between $40\mu m$ to $100 \mu m$.
 The following linear relationship is a good aproximation for

### 2015-07-03 CenterOfMass/DiferentialOperators01.jl
trabajando gradiantes

@@ -1,0 +1,16 @@
+function Gradiente2D(z::Array)
+    (alto,ancho)=size(z)
+    result=Array(Any, (alto,ancho))
+    aux=copy(z)
+    aux=vcat(aux[1,:],aux,aux[end,:])
+    aux=hcat(aux[:,1],aux,aux[:,end])
+    for j=2:alto+1, k=2:ancho+1
+        dfdx=aux[j,k+1]-aux[j,k-1]
+        #los arreys tienen un eje y invertido. Su y es la -y cartesiana
+        dfdy=-(aux[j+1,k]-aux[j-1,k]) 
+        result[j-1,k-1]=(dfdx. dfdy)
+    end
+    return result
+end
+
+        

### 2015-07-06 CenterOfMass/DiferentialOperators01.jl
Derivads jl

@@ -5,6 +5,7 @@ function Gradiente2D(z::Array)
     aux=vcat(aux[1,:],aux,aux[end,:])
     aux=hcat(aux[:,1],aux,aux[:,end])
     for j=2:alto+1, k=2:ancho+1
+        #Derivada de la secante
         dfdx=aux[j,k+1]-aux[j,k-1]
         #los arreys tienen un eje y invertido. Su y es la -y cartesiana
         dfdy=-(aux[j+1,k]-aux[j-1,k]) 
@@ -14,3 +15,21 @@ function Gradiente2D(z::Array)
 end
 
         
+function Divergencia2D(z::Array)
+    (alto,ancho)=size(z)
+    result=Array(Real, (alto,ancho))
+    aux=vcat(aux[1,:],aux,aux[end,:])
+    aux=hcat(aux[:,1],aux,aux[:,end])
+    for j=2:alto+1, k=2:ancho+1
+        #Derivada de la secante
+        vx2=aux[j,k+1]
+        vx1=aux[j,k-1]
+        vy2=aux[j+1,k]
+        vy1=aux[j-1,k] 
+        dvxdx=vx2-vx1
+        #los arreys tienen un eje y invertido. Su y es la -y cartesiana
+        dvydy=-(vy2-vy1)
+        result[j-1,k-1]=dvxdx+dvydy
+    end
+    return result
+end

### 2015-09-29 Reportes/PlaticaChicosITAM092015.tex
Platica Chicos Itam

\documentclass{beamer}

\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}
\usepackage{color}

\decimalpoint

\usefonttheme{serif}

\newcommand{\dd}{\, \mathrm{d}}
\newcommand{\xq}{\, \mathbf{x}}

\begin{document}

\begin{frame}
  \frametitle{De las medidas eléctricas a lo subyacente:} 
  \begin{equation}
    \varphi(\xq)
    =\int \frac{1}{4\pi\epsilon}
    \frac{\rho(\xq')}{|\xq-\xq'|} \dd^3 \xq'  
  \end{equation}
  \begin{center}
  \begin{tabular}{ccc}
   \textcolor{red}{LFP} & 
   \includegraphics[width=0.4\textwidth]{Decaimiento01.pdf} &
   \textcolor{blue}{CSD}
  \end{tabular}
  \end{center}
\end{frame}

\begin{frame}
\frametitle{Del Potencial Local a la Densidad de Fuentes de Corriente} 
El operador Laplaciano formalmente invierte la expresión para
el potencial eléctrico (medio uniforme e isotrópico).
\begin{equation}
  -\nabla^2\phi \propto \rho
\end{equation}
\begin{equation}
I=\partial_t \rho= - \rho \cdot \sigma/\epsilon 
  \end{equation}
\ldots pero el op. Laplaciano es una idealización:
\begin{equation}
  \nabla^2 f (\xq) =\lim_{\varepsilon \to 0}
  \frac{\sum_k f(\xq+\varepsilon_k)+ f(\xq-\varepsilon_k)-2 f(\xq)}
       {\varepsilon^2}  
\end{equation}
\begin{flushright}
  {\tiny Nicholson \emph{et} Freeman , J.   Neurophys. 38, 356 (1975)\\
         Bédard \emph{et} Destexhe, PRE 84, 041909 (2011) }
\end{flushright}

\end{frame}

\begin{frame}
  \frametitle{Un Laplaciano Numérico}
  \begin{equation}
    \nabla^2_{n}=2/3
    \begin{pmatrix}
      0 & 1 & 0 \\
      1 & -4 & 1 \\
      0 & -4 & 0    
    \end{pmatrix}
    +1/3
    \begin{pmatrix}
      0.5 & 0 & 0.5 \\
      0 & -2 & 0 \\
      0.5 & 0 & 0.5    
    \end{pmatrix}
  \end{equation}
  \begin{center}
  \begin{tabular}{cc}
    \includegraphics[width=0.4\textwidth]{LFP-cb-716.png} &
    \includegraphics[width=0.45\textwidth]{CSD_SeminarioJunio-712.png}    
  \end{tabular}
  \end{center}
\begin{flushright}
  {\tiny Lindberg, PAMI, 90, 234 (1990)}   
\end{flushright}
\end{frame}

\begin{frame}
  \frametitle{Una aproximación \emph{inversa}: iCSDA, kCSDA}
    \begin{align}
    \varphi(\xq)
    & =\int \frac{1}{4\pi\sigma}
    \frac{I(\xq')}{|\xq-\xq'|} \dd^3 \xq' \\
    & \approx \sum_j \frac{1}{4\pi\sigma}
    \frac{I(\xq_j)}{|\xq-\xq_j|} |\Delta \xq|     
  \end{align}
    Esto puede ponerse como una transformación lineal:
\begin{equation}
    \phi_k=\sum_j F_{jk} I_j
\end{equation}
Donde $F$ contiene suposiciónes sobre la distribución, e.g.
\begin{equation}
  F_{jk}=\iiint_{\xq_k} \dd x \dd y \dd z
\frac{H(z)}{4\pi\sigma \sqrt{(x-x_j)^2+(y-y_j)^2+z^2}} 
\end{equation}
\end{frame}

\begin{frame}
  \frametitle{iCSDA, kCSDA II}
  \ldots y sólo tendriamos que invertir $F$ para obtener $I$
  \begin{equation}
    I_m=F^{-1}\phi 
  \end{equation}
  Esto requiere resolver $n\times n$ integrales, cada una con
  $n\times n$ contribuciones, y luego invertir la matriz, que es una operación
  que crece como $n^3$ \ldots
  Si nos va bien, esta operación crece como $O(n^7)$.

  Resuelta \emph{en principio}.

 \begin{flushright}
  {\tiny  Łesky \emph{ et all}, Neuroinform, 9, 401 (2011) \\
      Pettersen \emph{et all} , JNM  154, 116 (2011) \\
      Potworowsky \emph{et all}, Neur. Comp. 24, 541 (2012). }
  \end{flushright}
\end{frame}

\begin{frame}
\frametitle{Un cero bien localizado.} 
Tanto en la densidad de cargas, como en la densidad de fuentes de corriente
hay una distinción clara entre \emph{pozos} y \emph{fuentes}.
No es el caso para el LFP, que mide las cosas a partir de una ``tierra''
arbitraria. 
\begin{center}
    \includegraphics[width=0.75\textwidth]{CSD_SeminarioJunio-cb-732.png}
\end{center}
\end{frame}

\begin{frame}
  \frametitle{Promedios Vectoriales o \emph{Centros de Masa}}
  \begin{equation}
    \langle \xq(t) \rangle=\sum_k m(\xq_k)\xq/\sum_k m(\xq_k)
  \end{equation}

  \begin{center}
    \includegraphics[width=0.75\textwidth]{ImagenManjarrez01.jpg}
  \end{center}

 \begin{flushright}
  {\tiny  Manjarrez \emph{ et all}, Brain Research, 1145, 239 (2007) \\
      Chao \emph{ et all}, Neuroinformatics, 3, 263 (2005) }
  \end{flushright}

\end{frame}

\begin{frame}
\frametitle{Promedios vectoriales o \emph{Centro de Masa}} 
Tanto en la densidad de cargas, como en la densidad de fuentes de corriente
hay una distinción clara entre \emph{pozos} y \emph{fuentes}.
\begin{center}
  \begin{tabular}{cc}
    \includegraphics[width=0.35\textwidth]{DiagramaCAetDGCSDEspanol01.png}   &
    \includegraphics[width=0.55\textwidth]{CSD_CM_SeminarioJunio-732.png}
  \end{tabular}
\end{center}  
\end{frame}

\begin{frame}
\frametitle{Centros de Masa en el espacio de las CSD.} 
Tanto en la densidad de cargas, como en la densidad de fuentes de corriente
hay una distinción clara entre \emph{pozos} y \emph{fuentes}.
\begin{center}
  \begin{tabular}{cc}
    \includegraphics[width=0.35\textwidth]{DiagramaCAetDGCSDEspanol01.png}   &
    \includegraphics[width=0.55\textwidth]{CSD_CM_SeminarioJunio-732.png}
  \end{tabular}
\end{center}  
\end{frame}

\begin{frame}
\frametitle{Centros de Masa en el espacio de las CSD.} 
Tanto en la densidad de cargas, como en la densidad de fuentes de corriente
hay una distinción clara entre \emph{pozos} y \emph{fuentes}.
\begin{center}
  \begin{tabular}{cc}
    \includegraphics[width=0.35\textwidth]{DiagramaCAetDGCSDEspanol01.png}   &
    \includegraphics[width=0.55\textwidth]{CSD_CM_SeminarioJunio-732.png}
  \end{tabular}
\end{center}  
\end{frame}

\begin{frame}
\frametitle{Trayectorias de Actividad} 
El seguimiento continuo de los CM.
\begin{center}
    \includegraphics[width=0.7\textwidth]{TrayectoriasLimpiasNumeros01.png} 
\end{center}  
\end{frame}

\end{document}

### 2015-10-09 Reportes/ReporteExpress01.tex
ReporteOctubrePrincipio

\documentclass{beamer}

\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}
\usepackage{color}

\decimalpoint

\usefonttheme{serif}

\newcommand{\dd}{\, \mathrm{d}}
\newcommand{\xq}{\, \mathbf{x}}

\begin{document}

\begin{frame}
  \frametitle{De la actividad evocada a las trayectorias de los CM}
  Dado que la actividad evocada produce respuestas
  \emph{perfectamente} acompazadas, podemos
  promediar sobre varios eventos experimentales
  a fin de tener una señal más limpia. 
  \begin{center}
   \includegraphics[width=0.8\textwidth]{PotencialesEvocados01.pdf} 
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{¿Se vale promediar sobre CSD?}
    Téoricamente si, el CSDA es una operación lineal
    (formas más sofisticadas pueden no serlo). Esto es:

\begin{equation}
 \langle CSD(\phi) \rangle = CSD (\langle \phi \rangle)
\end{equation}

\begin{center}
  \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{CSDEvocado-Trancazo_1-0177.png} &
    \includegraphics[width=0.5\textwidth]{CSDEvocado-Promedio-0177.png} 
  \end{tabular}
\end{center}

\end{frame}

\begin{frame}
  \frametitle{No podemos \emph{``cerar''} los canales saturados}
  Un criterio simple nos dice cuando un canal no registro nada,
  el ataque usual era poner ese canal en ceros, pero al derivar
  nos produce un efecto de borde muy grande.

\begin{center}
  \begin{tabular}{cc}
    \includegraphics[width=0.4\textwidth]{LFPEvocado_1R4-Promedio-177.png} &
    \includegraphics[width=0.4\textwidth]{CSDEvocado-Promedio-0177.png} 
  \end{tabular}
\end{center}
Solución: tenemos que hacer un promedio sobre la vecindad local
para cada canal saturado, en cada \emph{frame}.
\end{frame}

\begin{frame}
\frametitle{¿Promedio de Trayectorias?}
Usando normalización sobre la longitud de arco, el
promedio de varias curvas esta bien definido.
\begin{center}
\includegraphics[width=0.5\textwidth]{PromedioSobreUnaTrayectoria01.pdf}
\end{center}
\end{frame}

\begin{frame}
  \frametitle{¿Promedio sobre los Centros de Masa?}
  Si el centro de masa fuera sobre todo el espacio, una
  vez más sería válido por linealidad. ¡La
  detección de componentes disjuntos nos puede echar todo a
  perder!
\begin{center}
 \includegraphics[width=0.5\textwidth]{promedioComponentesDisjuntos01.pdf}
\end{center}
\end{frame}

\begin{frame}
  \frametitle{¿Promedio sobre los Centros de Masa?}
  Si tenemos una familia de funciones regulares, podemos garantizar
  que eso no pasa. Una \emph{familia normal}\ldots
\begin{center}
 \includegraphics[width=0.7\textwidth]{FuncionesRegulares01.png}
\end{center}
\end{frame}

\begin{frame}
  \frametitle{Ceros sobre funciones regulares.}
  Toca recordar como se usaban esas cosas.
\begin{center}
 \includegraphics[width=0.9\textwidth]{FuncionesRegularesCeros01.png}
\end{center}
\end{frame}

\end{document}

### 2015-10-09 CenterOfMass/RenameAndEncode01.sh

@@ -23,6 +23,6 @@ rename $argumento $argumento$tresceros $argumento$unnum".png"
 
 echo $paraencoderbien
 
-mencoder "$paraencoderbien"  -o $outputbien -ovc lavc -mf fps=24
+mencoder "$paraencoderbien"  -o $outputbien -ovc lavc -mf fps=12
 
 

### 2015-10-18 Reportes/Cocoyoc2015.tex
Cococyoc

\documentclass{beamer}

\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}
\usepackage{color}

\decimalpoint

\usefonttheme{serif}

\newcommand{\dd}{\, \mathrm{d}}
\newcommand{\xq}{\, \mathbf{x}}

\begin{document}

\begin{frame}
  \frametitle{De las medidas eléctricas a lo subyacente:} 
  \begin{equation}
    \varphi(\xq)
    =\int \frac{1}{4\pi\epsilon}
    \frac{\rho(\xq')}{|\xq-\xq'|} \dd^3 \xq'  
  \end{equation}
    \begin{equation}
      \frac{\partial \rho}\partial{t}(\xq)=
      -\frac{\sigma}\partial{\epsilon}\rho(\xq)
    \end{equation}
  \begin{center}
  \begin{tabular}{ccc}
   \textcolor{red}{LFP} & 
   \includegraphics[width=0.3\textwidth]{Decaimiento01.pdf} &
   \textcolor{blue}{CSD}
  \end{tabular}
  \end{center}
  {\tiny Nicholson \emph{et} Freeman , J.   Neurophys. 38, 356 (1975)\\
         Bédard \emph{et} Destexhe, PRE 84, 041909 (2011) }
\end{frame}

\begin{frame}
\frametitle{Un cero bien localizado.} 
Tanto en la densidad de cargas, como en la densidad de fuentes de corriente
hay una distinción clara entre \emph{pozos} y \emph{fuentes}.
No es el caso para el LFP, que mide las cosas a partir de una ``tierra''
arbitraria. 
\begin{tabular}{cc}
      \includegraphics[width=0.4\textwidth]{VectorFields.png} &
    \includegraphics[width=0.55\textwidth]{CSD_SeminarioJunio-cb-732.png} 
\end{tabular}
\end{frame}

\begin{frame}
  \frametitle{Promedios Vectoriales o \emph{Centros de Masa}}
  \begin{equation}
    \langle \xq(t) \rangle=\sum_k m(\xq_k)\xq/\sum_k m(\xq_k)
  \end{equation}

  \begin{center}
    \includegraphics[width=0.75\textwidth]{ImagenManjarrez01.jpg}
  \end{center}

 \begin{flushright}
  {\tiny  Manjarrez \emph{ et all}, Brain Research, 1145, 239 (2007) \\
      Chao \emph{ et all}, Neuroinformatics, 3, 263 (2005) }
  \end{flushright}

\end{frame}

\begin{frame}
\frametitle{Centros de Masa en Geometrias Cóncavas} 
\begin{center}
  \begin{tabular}{cc}
    \includegraphics[width=0.45\textwidth]{ImagenGizMultiColorChica01.png} &  
    \includegraphics[width=0.45\textwidth]{DiagramaCAetDGCSDEspanol01.png}   
   % \includegraphics[width=0.55\textwidth]{CSD_CM_SeminarioJunio-732.png}
  \end{tabular}
\end{center}
\end{frame}

\begin{frame}
  \frametitle{Centros de Masa en Densidades de Fuentes}
  Podemos dividir cada conjunto (pozos o fuentes) en sus componentes
  disjuntas.
\begin{center}
  \begin{tabular}{cc}
    \includegraphics[width=0.55\textwidth]{CSD_SeminarioJunio-732.png} &  
    \includegraphics[width=0.35\textwidth]{ComponentesDisjuntosCSD01.png}   
   % \includegraphics[width=0.55\textwidth]{CSD_CM_SeminarioJunio-732.png}
  \end{tabular}
\end{center}
\end{frame}

\begin{frame}
  \frametitle{Un conjunto instantaneo de CM}
  El CM de cada \emph{Componente Disjunto} indica un putativo ``centro de actividad''
  rastreable, localizado mucho más cerca de donde
  la señal es más fuerte. Esto es interpretable como una serie de ``focos''
  activos localizados de mayor actividad.
  \begin{center}
   \includegraphics[width=0.65\textwidth]{CSD_SeminarioJunio-900.png}   
   \end{center}
\end{frame}

\begin{frame}
  \frametitle{Trayectorias de los CM}
  Usando un algorítmo que conecte los CM cuadro por cuadro comienzan a
  formarse \emph{``trayectorias''} 
  con precisión de 1/7022 de segundo.
\begin{center}
   \includegraphics[width=0.67\textwidth]{TrayectoriasLimpiasNumeros01.png}   
   \end{center}
\end{frame}

\begin{frame}
  \frametitle{Posibles implicaciones para el conectoma funcional}
  De esta serie de datos de alta densidad, ordenados por aparición,
  intensidad, y criterio anatómico, debe ser posible descubrir conexiones
  funcionales dentro de CA3. El formalismo es general, por ende debe de ser aplicable
  también a otras estructuras.

  (Estos ejemplos fueron creados usando actividad epileptiforme inducida por
  4AP.)
  \end{frame}

\end{document}

### 2015-10-19 Reportes/Cocoyoc2015.tex
Tururu

@@
 \usepackage{color}
+\usepackage{ragged2e}
+
+\justifying
 
@@
     \begin{equation}
-      \frac{\partial \rho}\partial{t}(\xq)=
-      -\frac{\sigma}\partial{\epsilon}\rho(\xq)
+      \frac{\partial \rho}{\partial{t}}(\xq)=
+      -\frac{\sigma}{\epsilon}\rho(\xq)
     \end{equation}
@@
 arbitraria.
+
 \begin{tabular}{cc}
-      \includegraphics[width=0.4\textwidth]{VectorFields.png} &
+      \includegraphics[width=0.4\textwidth]{VectoresRojos_732.png} &
     \includegraphics[width=0.55\textwidth]{CSD_SeminarioJunio-cb-732.png} 
@@
   \frametitle{Un conjunto instantaneo de CM}
-  El CM de cada \emph{Componente Disjunto} indica un putativo ``centro de actividad''
-  rastreable, localizado mucho más cerca de donde
+  El CM de cada \emph{Componente Disjunto} indica un putativo ``centro de actividad'' rastreable,
+  localizado mucho más cerca de donde
   la señal es más fuerte. Esto es interpretable como una serie de ``focos''
@@
   \frametitle{Posibles implicaciones para el conectoma funcional}
+
   De esta serie de datos de alta densidad, ordenados por aparición,
   intensidad, y criterio anatómico, debe ser posible descubrir conexiones
-  funcionales dentro de CA3. El formalismo es general, por ende debe de ser aplicable
-  también a otras estructuras.
+  funcionales dentro de CA3. El formalismo es general, es aplicable
+  a otras estructuras.
 

### 2015-11-06 Reportes/CuadrosPoster.tex
cuadros poster 01

\documentclass{article}

\usepackage[utf8]{inputenc}

\begin{document}

\section{CSD and centers of mass}

For certain purposes the representation of the measured data in CSD
representation has advantages over the recorded LFP. The principal one is
that it minimizes the volume-conducted activity, pointing more acurattely
to the sites in which electrical-measurable events are taking place. We have
another more important reason to resort to this representation: namely,
as it names implies, it represents a \emph{density}, so, it indicates
quantity of something per unit volume, which permits us to apply certain
concepts in a rigourous, physicaly interpretable manner.

Densities are instrumental for taking vector averages, also known as centers
of mass. A charged or flux density is in this regard no different that a
mass density and can be used to obtain significant ``centers'' of
physical objects. In our case, our physical objects are active neurons and
their inmediate extracellular space, and is our interest to track a
putative ``center of activity'' in order to trace routes in which this
activity flows across the tissues that we are studing. Also, in this regard,
it delimitates (instantaneous) active areas from their surroundings by
a (aproximate) zero Set, which is the locus of all points in the measurement
area that register a null CSD. This set is typically a regular set of
curves which separate two sets, the ``sources'' and ``sinks'', i.e. the
set of all places where there is outflux of charge and the set of all places
where there is influx of charge. Each of these sets can be decomposed in
its disjoint components using the zero set as the border between them.

Each disjoint component is an instantaneous locus where we localize one pole of
an active unit. These poles will appear to move as the activity grows in neighbouring
units or vanishes. The high frecuency of sampling at our disposal permits us to
track very precicely the apparent track that this apparition, grows, wanning and
dissapearence of poles produce. Thus, by calculating the center of mass of
each one of these poles at succesive recording times
we can obtain a trayectory of activity at local units.

The HD MEAS of the kind of the BioCAM 4096  an electrode density comparable
to the sizes of of typical neurons at the hipocampal tissue, which is around
the limit in which mean size approximations are still applicable \cite{Bedard2011}.
At such ranges a simple convex numeric Laplacian yields adecuate results without
the need to restore to computer costly alternatives such as the iCSDA or kCSDA
\cite{Leski2011, Potworowsky2011}. A nice implementation of a 2D numeric Laplacian
which reduces the typical cross effect is given by a convolution of
the 2D array of data with the following matrix:
\begin{equation}

\end{equation}

\end{document}

### 2015-11-06 Reportes/CuadrosPoster.tex
mas texto poster

@@
 For certain purposes the representation of the measured data in CSD
-representation has advantages over the recorded LFP. The principal one is
+representation has advantages over the recorded LFP. One is
 that it minimizes the volume-conducted activity, pointing more acurattely
-to the sites in which electrical-measurable events are taking place. We have
-another more important reason to resort to this representation: namely,
+to the sites in which electrical-measurable events are taking place \cite{Mitzdorf85}.
+We have another important reason to resort to this representation:
 as it names implies, it represents a \emph{density}, so, it indicates
-quantity of something per unit volume, which permits us to apply certain
+quantity of something per unit volume. Thus we are able to apply certain
 concepts in a rigourous, physicaly interpretable manner.
@@
 Densities are instrumental for taking vector averages, also known as centers
-of mass. A charged or flux density is in this regard no different that a
+of mass. A charge or flux density is in this regard no different that a
 mass density and can be used to obtain significant ``centers'' of
-physical objects. In our case, our physical objects are active neurons and
-their inmediate extracellular space, and is our interest to track a
-putative ``center of activity'' in order to trace routes in which this
-activity flows across the tissues that we are studing. Also, in this regard,
-it delimitates (instantaneous) active areas from their surroundings by
-a (aproximate) zero Set, which is the locus of all points in the measurement
-area that register a null CSD. This set is typically a regular set of
-curves which separate two sets, the ``sources'' and ``sinks'', i.e. the
-set of all places where there is outflux of charge and the set of all places
-where there is influx of charge. Each of these sets can be decomposed in
-its disjoint components using the zero set as the border between them.
+physical quantities. In our case, this are the ionic flows across
+the membrane, which, looked macroscopically from the extracellular
+space, appear as sinks and sources. These are evidence of neuronal
+activity and is and is our interest to locate a
+putative ``instantaneous center'' of said activity and then track
+trajectories which could sintetize the sequence of diverse units
+incorporating themselves to the action.
+
+Sinks and sources are separated by a zero set. This is also an advantage
+over the LFP representation, in which any particular value has no
+interpretative significance, and only \emph{differences} between values
+have meaning. 
+Our zero  set is typically a union of regular 
+curves which separate sources from sinks. As we choose a numerical error
+tolerance, these sets appear in our measurements conformed by disjoint
+components. This permits us two confirm measurements that reveal
+the expected polar activity of firing neurons, and to locate centers
+of  activity by calculating the center of mass in each disjoint component
+at each frame from our data. We then have an instantaneous ``center of mass''
+for each disjoint component, which we shall call it the ``instantaneous local
+center of activity'' or ILCA for short. These ILCAs have an apparent
+movement as sinks and sources appear and fade, so we can track their
+position and intensity frame by frame, producing a putative trayectory,
+which could reveal displacement of activity or the succesion of recruiting
+neurons. 
+Thanks to the high frecuency sampling of our experiments, we can observe
+very small displacements of this centers of activities. Care must be taken
+to interpret this displacements below the electrode scale.
+
+Each trayectory thus obtained encodes and synthetizes the following information:
+\begin{description}
+\item[Start] location and time of the apparition of the source/sink.
+\item[Path] succesive locus and timestamp of the putative centers.
+\item[Intensity] The time dependant integrated density of the source/sink.  
+\item[End] location and time of the disapearence of the source/sink.
+\end{description}
+
+Visual comparition with the sink/source spatiotemporal structure coincides
+with known measuraments of CSD.
 
@@
 \begin{equation}
-
+f(x)
 \end{equation}

### 2015-11-09 Reportes/CuadrosPoster.tex
tralala

@@
 representation has advantages over the recorded LFP. One is
-that it minimizes the volume-conducted activity, pointing more acurattely
-to the sites in which electrical-measurable events are taking place \cite{Mitzdorf85}.
+that it minimizes the volume-conducted activity, pointing more accurately
+to the sites in which electrical-measurable events are taking place %\cite{Mitzdorf85}.
 We have another important reason to resort to this representation:
@@
 quantity of something per unit volume. Thus we are able to apply certain
-concepts in a rigourous, physicaly interpretable manner.
+concepts in a rigorous, physically interpretable manner.
 
-Densities are instrumental for taking vector averages, also known as centers
-of mass. A charge or flux density is in this regard no different that a
+Densities are instrumental for taking vector averages (``centers
+of mass''). A charge or flux density is in this regard no different to a
 mass density and can be used to obtain significant ``centers'' of
@@
 space, appear as sinks and sources. These are evidence of neuronal
-activity and is and is our interest to locate a
+activity. It is our interest to locate a
 putative ``instantaneous center'' of said activity and then track
 trajectories which could sintetize the sequence of diverse units
-incorporating themselves to the action.
+incorporating themselves to the action or spaciotemporal variations of
+said action across units.
 
-Sinks and sources are separated by a zero set. This is also an advantage
+Sinks and sources are separated by a zero set. This is an advantage
 over the LFP representation, in which any particular value has no
@@
 have meaning. 
-Our zero  set is typically a union of regular 
+The zero  set is typically a union of regular 
 curves which separate sources from sinks. As we choose a numerical error
@@
 components. This permits us two confirm measurements that reveal
-the expected polar activity of firing neurons, and to locate centers
-of  activity by calculating the center of mass in each disjoint component
-at each frame from our data. We then have an instantaneous ``center of mass''
-for each disjoint component, which we shall call it the ``instantaneous local
-center of activity'' or ILCA for short. These ILCAs have an apparent
+the expected polar activity of neurons
+at each frame from the data. We have observed that each disjoint component
+of each sources and sinks corresponds roughly to the poles of firing neurons
+or superpositions of them (although sinks/sources of lower intensity
+can appear below the firing threshold also).
+
+Each of these components has its own  ``center of mass'', which can be
+calculated in the usuall maner. We shall call them the ``instantaneous local
+centers of activity'' or ILCAs for short. These ILCAs have an apparent
 movement as sinks and sources appear and fade, so we can track their
@@
 \item[Start] location and time of the apparition of the source/sink.
-\item[Path] succesive locus and timestamp of the putative centers.
-\item[Intensity] The time dependant integrated density of the source/sink.  
-\item[End] location and time of the disapearence of the source/sink.
+\item[Path] successive locus and timestamps of the putative centers.
+\item[Intensity] the time dependent integrated density of the source/sink.  
+\item[End] location and time of the disappearance of the source/sink.
 \end{description}
 
-Visual comparition with the sink/source spatiotemporal structure coincides
-with known measuraments of CSD.
-
-Each disjoint component is an instantaneous locus where we localize one pole of
-an active unit. These poles will appear to move as the activity grows in neighbouring
-units or vanishes. The high frecuency of sampling at our disposal permits us to
-track very precicely the apparent track that this apparition, grows, wanning and
-dissapearence of poles produce. Thus, by calculating the center of mass of
-each one of these poles at succesive recording times
-we can obtain a trayectory of activity at local units.
-
-The HD MEAS of the kind of the BioCAM 4096  an electrode density comparable
-to the sizes of of typical neurons at the hipocampal tissue, which is around
-the limit in which mean size approximations are still applicable \cite{Bedard2011}.
-At such ranges a simple convex numeric Laplacian yields adecuate results without
-the need to restore to computer costly alternatives such as the iCSDA or kCSDA
-\cite{Leski2011, Potworowsky2011}. A nice implementation of a 2D numeric Laplacian
-which reduces the typical cross effect is given by a convolution of
-the 2D array of data with the following matrix:
-\begin{equation}
-f(x)
-\end{equation}
-
 \end{document}

### 2015-11-11 Reportes/CuadrosPoster.tex
Muchos Registros

@@
 
+\begin{figure}
+  \includegraphics[width=0.9\textwidth]{LFP_700_more_selected_electrodes_01.svg}
+  \caption{LFP recorded from experiment at the onset of the epileptic activity.
+Selected electrodes from which electrophysiological recording is shown are marked  along various transveral lines to CA.
+The data is presented as recorded, without any smoothing or denoizing.
+}
+
 \end{document}

### 2015-11-18 Reportes/PosterPhiladelphia/IntroPoster01.tex
Intro Poster

Epileptic or epileptic-like activity is characterized by excessive synchronous action 
in the brain.  High density microelectronic arrays (HDMEAs) can be used to record
the generated LFP associated to this activity at hundreds or thousands of near
locations and at a  high frequency. These last years said devices are
approximating resolutions comparable to the neuron scales involved (of the
order 50 µm). However, comparing thousands of electro-physiological traces with dense
patterns can obstacle the discovery of the location, succession, and mode of
propagation of the activity.

An usual technique to discover such information is to plot ``heat maps'' of
the absolute diference in voltage of the LFP at each recording time,
and then animating the
color plots in succesion to observe the start, propagation, and end 
of the seizure.  This raw treatment can serve in general to ilustrate and to learn
the general characteristics of the seizure, but it is poor in resolution and
numerically vague. Here we propose a refination of the  methods used for analysing
propagation of activity in electrophysiological recordings. Our contribution permits
to separate various paths of symultaneous activity and track fine variations
between various almost synchronic units.

The method in question consists on extacting rigorous separable data of the CSD
representation and then applying a ``Center of Mass analysis'' on each set. 
This enables us to sepparate precisely units of
activity by their succesive activities. Our analyisis consist on four steps.

On the first step we transform the LFP data into CSD by a simple numerical
Laplacian operator. This representation of the data has some advantages over the LFP.
The most cited one is that the spread of the signal over area is reduced, pointing
more accurately to sites of activity. Also it represents, as it name suggest,
a density. That means that it can be used for obtaining weighted averages,
in particular, we use it to obtain vector averages of sinks and sources. Lastly, it
has a zero value with a precise meaning, separating sinks from sources (influx and
eflux sites). The zero set is a natural border that allows us to separate
the various poles that are formed by active units. This is important for the
second step of the analysis, in which we use this border to segment the sets of
sources and sinks into disjoint components, which correspond roughly to poles
of firing neurons. The third step is to calculate the ``center of mass'' of each
one of these components, frame by frame. As ``mass'' we use the CSD, which is, by
definition, a density, and then the center thus obtained is precisely the center of
mass for a negative/positive pole. At this step we have a collection of centers
of mass for both sinks and sources, each one encompasing its ``intensity'' (total
mass) and spatiotemporal coordinates. In the fourth step we try to concatenate the
centers of each frame with those of the next, trying to find the most probable
succesor to each one by distance and intensity. Every chain thus obtained
is a path, which reveals very fine and rapid displacements of the neuronal activity
on each site of the tissue.

In the following figures we illustrate this ideas with data obtained from a
rat's hippocampus wich has beeen prepared with 4AP... bla bla Franco, check this.

### 2015-12-11 ArticuloCSDyCM/Paper01.tex
papar

\documentclass{article}

\usepackage{amsmath}
\usepackage[utf8]{inputenc}

\begin{document}

\section{CSD for High density MEAs}

The concepts presented here require first obtaining the CSD of the recorded LFP. Any method would conceptualy present the same utility, but thanks to the density and scales of the BioCAM 4096, in our case a ``classical'' numerical derivation of the CSD is the most adecuate route to take. A numerical Laplacian Operator which reduces the cross efect of rectangular grids is the convolution of the data with the following matrix \cite{Lindberg90}:
\begin{equation}
\nabla^2_{1/3}=(2/3)
\begin{pmatrix}
  0 & 1 & 0 \\
  1 & -4 & 1 \\
  0 & 1 & 0
\end{pmatrix}
+1/3
\begin{pmatrix}
  0.5 & 0 & 0.5 \\
  0 & -2 & 0 \\
  0.5 & 0 & 0.5
\end{pmatrix}  
\end{equation}
At the border we duplicate the last rows of data and take the results knowingly of the border effects. As this sort of operators are extremelly sensitive to noice, a Gaussian blur is applied before presenting the graphical data. This softening of the edges in the CSD image also represents our ignorance of the finer details of the components of the distribution (see section bla bla).

More sophisticated methods such as iCSDA \cite{Leski2011} and kCSDA \cite{Potworowski2011} prove little advantage in estimating the sources under the ohmic, isotropic and homogeneous assumptios at the scales involved in these data, and turn out to be computer-costly for high density MEAs and dense temporal sampling.

\section{CSD and disjoint components}

One of the most obvious uses of the CSD in place of the LFP has been overlocked, and this might have to do that samplings at lesser spatial density did not permit extending the implications of the concept of density, namely, that one of ``masses'', or, in this case, charges. Center of Mass Analysis (CMA) has been previously been used in order to track putative ``trayectories'' of the activity or information across neural tissue or even the whole brain \cite{Chao05, Manjarrez07}. In such works the choosen ``mass'' was an heuristic measure of activity adecuate for representing the corse-grain picture of displacement of activity or some unit of ``active information''. But these present the usual problem of justifing the choice maide on practical grounds. For calculating a vectorial average, the adecuate choice of wheights per vector is a density, a positive quantity that measures the amount of something per unit volume. Current Source Density, as it name implies, is the density of sinks and sources of charged current per unit volume. CSD sepparates naturally the measured activity into two distinct sets, namely, the set of all sources and the set of all sinks, separated by the zero set. This is in contrast to the LFP, where the zero value of the function has no solid meaning as electrical potentials can be measured to arbitrary ground values.

A visual inspection of the CSD colormap at a given instant in time shows that both sources and sinks appear in rather large and significant patches. These appear to correspond to the expected double and triple poles of fyring neurons in structured tissue \cite{} and inspired us to sepparate the CSD subsets into their \emph{disjoint components}. This concept is used in mathematics to denote subsets of a geometric defined set that do not touch each other. Their rough correspondence to poles of fyring neurons would be most usefull in tracking the succesive locus of the activity.

To clarify the concept and how we are we using it an example is in order. In figure \ref{ejemplodisjuntos} we show geometrical sets enclosed by simple curves as color patches, where every color indicates a Set. The set A consists of a single component, that means that one can make  path between any two points in the set that consist entirely of points of the same set. In contrast, the set B is made of two disjoint components, labeled B1 and B2. If one chooses a point in each of the components, one cannot draw a path joining them without leaving the set. But each one of those components is connected, meaning that if we choose the pair of points in one componet, then again we can trace a path joining them in the set. So the set B consists of two disjoint components, but each component can be regarded as a connected set on their own. Likewise the set C is made of three disjoint components.

CSD divides the set of measuring points into two subsets: the set of all sources and the set of all sinks, i.e, the set of all measuring points that have a CSD positive value, and the set of those who have a negative CSD value. The zero set is a border between them, but due to the finite precision of the measurement (and the validity of the mean field approach!) these sets can make contact with each other. Still, noice between patches corresponding to a single set could make a ``bridge'' between components, making identification of poles inexact. In order to avoid this we create a ``fat zero'' set of all the recording points whose CSD values are inside a small error interval around zero. This sepperates even more the components of the set, and helps us to simplify the following analysis.

\section{CSD and Center of Mass}

Densities provide the adecuate conceptual tool to obtain ``Centers of Mass'' or vector averages. A density is a measure of a quantity per spatial spread, at precise locus in space. CSD has, in contast to mass density, negative values. But these signs are a matter of convention, it is more important to say that it has two different cualitative propiertes and that sepparates the domain into two sets. Each set has then a total ``mass'', the integral of the absolute value of the CSD. Then the concept of ``Center of Mass'' applies naturally. Once that we sepparate the two sets and apply the concept in each one independently, even the sign convention of the densities seases to be a problem, as it cancels in the formula for the CM:
\begin{equation}
\langle x(t) \rangle =\frac{\sum_j x_{j} (t) I_{j} (x,t)} {\sum_j I_{j}(x,t)}
\end{equation}
An instantaneous Center of Mass, thus calculated, would present little if none physiological interpretation. The activity could be spread over non-convex sets (which would include disjoint and concave sets) and so the vector average, the Center of Mass, would be located in points where there is no activity. In our example the use of the CA complex of the hippocampus proves an ideal example. This structure is concave, and during epileptic burst, can have activity spread over many disjoint sites simultaneously. We must not take out of sight the phenomena which is producing the CSD distribution: neurons have a finite size, and the effect of their action a finite extent. Every patch that we can delimit as a disjoint component of the Sources/Sink sets indicates a coherent, joint active group of units. A center of mass for each disjoint component would then have much

\end{document}

### 2016-01-06 ArticuloCSDyCM/Paper01.tex
uura papaer

@@
 The concepts presented here require first obtaining the CSD of the recorded LFP.
-Any method would conceptualy present the same utility, but thanks to the density and scales of the BioCAM 4096, in our case a ``classical'' numerical derivation of the CSD is the most adecuate route to take.
+Any method would present the same utility.
+Thanks to the density and scales of the BioCAM 4096, in our case a ``classical'' numerical derivation of the CSD is the most adecuate route to take.
 A numerical Laplacian Operator which reduces the cross efect of rectangular grids is the convolution of the data with the following matrix \cite{Lindberg90}:
@@
 At the border we duplicate the last rows of data and take the results knowingly of the border effects.
-As this sort of operators are extremelly sensitive to noice, a Gaussian blur is applied before presenting the graphical data.
+As this sort of operators are extremely sensitive to noise, a Gaussian blur is applied before presenting the graphical data.
 This softening of the edges in the CSD image also represents our ignorance of the finer details of the components of the distribution (see section bla bla).
 
-More sophisticated methods such as iCSDA \cite{Leski2011} and kCSDA \cite{Potworowski2011} prove little advantage in estimating the sources under the ohmic, isotropic and homogeneous assumptios at the scales involved in these data, and turn out to be computer-costly for high density MEAs and dense temporal sampling.
+More sophisticated methods such as iCSDA \cite{Leski2011} and kCSDA \cite{Potworowski2011} prove little advantage in estimating the sources under the ohmic, isotropic and homogeneous assumptions at the scales involved in our data, and turn out to be computer-costly for high density MEAs and dense temporal sampling.
 
@@
 
-One of the most obvious uses of the CSD in place of the LFP has been overlocked, and this might have to do that samplings at lesser spatial density did not permit extending the implications of the concept of density, namely, that one of ``masses'', or, in this case, charges.
-Center of Mass Analysis (CMA) has been previously been used in order to track putative ``trayectories'' of the activity or information across neural tissue or even the whole brain \cite{Chao05, Manjarrez07}.
-In such works the choosen ``mass'' was an heuristic measure of activity adecuate for representing the corse-grain picture of displacement of activity or some unit of ``active information''.
+One of the most obvious uses of the CSD in place of the LFP has been overlooked.
+This might have occurred because  samplings at lesser spatial density did not permit extending the implications of the concept of density.
+Density is \emph{quantifiable entity per unit of space}, or, in other words, concentration of something distributed in space.
+In this case, the ``something'' is the sources and sinks of electric current, which can be identified with the interchange of charged ions from the inside to the outside of cellular membranes.
+As these are indicators of activity in the neurons, the CSD is also, indirectly, a \emph{``density of local activity''}.
+The advantages of this representation of the activity in place of the LFP have been stated many times (better spatial resolution, mostly), but is use as \emph{density} has been overlooked.
+
+Center of Mass Analysis (CMA) has previously been used in order to track putative  ``trajectories'' of the activity or information across neural tissue or even the whole brain \cite{Chao05, Manjarrez07}.
+In such works the chosen ``mass'' was an heuristic measure of activity adecuate for representing the corse-grain picture of displacement of activity or some unit of ``active information''.
 But these present the usual problem of justifing the choice maide on practical grounds.
 For calculating a vectorial average, the adecuate choice of wheights per vector is a density, a positive quantity that measures the amount of something per unit volume.
-Current Source Density, as it name implies, is the density of sinks and sources of charged current per unit volume.
 CSD sepparates naturally the measured activity into two distinct sets, namely, the set of all sources and the set of all sinks, separated by the zero set.
-This is in contrast to the LFP, where the zero value of the function has no solid meaning as electrical potentials can be measured to arbitrary ground values.
+The sepparation is not quantitative but qualitative: even when the signs of electric charge are arbitrary, the distintion between them is not.
+This is also in contrast to LFP, where the zero of the function has no precise meaning as electrical potentials can be measured from arbitrary ground values.
 
 A visual inspection of the CSD colormap at a given instant in time shows that both sources and sinks appear in rather large and significant patches.
-These appear to correspond to the expected double and triple poles of fyring neurons in structured tissue \cite{} and inspired us to sepparate the CSD subsets into their \emph{disjoint components}.
-This concept is used in mathematics to denote subsets of a geometric defined set that do not touch each other.
-Their rough correspondence to poles of fyring neurons would be most usefull in tracking the succesive locus of the activity.
+These appear to correspond to the expected double and triple poles of fyring neurons in structured tissue \cite{} and inspired us to separate the CSD subsets into their \emph{disjoint components}.
+This concept is used in mathematics to denote subsets of a given geometric set that do not touch each other.
+Their rough correspondence to poles of firing neurons would be most useful in tracking the successive locus of the activity.
 
@@
 If one chooses a point in each of the components, one cannot draw a path joining them without leaving the set.
-But each one of those components is connected, meaning that if we choose the pair of points in one componet, then again we can trace a path joining them in the set.
+But each one of those components is connected, meaning that if we choose the pair of points in one component, then again we can trace a path joining them in the set.
 So the set B consists of two disjoint components, but each component can be regarded as a connected set on their own.
@@
 The zero set is a border between them, but due to the finite precision of the measurement (and the validity of the mean field approach!) these sets can make contact with each other.
-Still, noice between patches corresponding to a single set could make a ``bridge'' between components, making identification of poles inexact.
-In order to avoid this we create a ``fat zero'' set of all the recording points whose CSD values are inside a small error interval around zero.
+Still, noise between patches corresponding to a single set could make a ``bridge'' between components, making identification of poles inexact.
+In order to avoid this we create a ``thick zero'' set of all the recording points whose CSD values are inside a small error interval around zero.
 This sepperates even more the components of the set, and helps us to simplify the following analysis.
@@
 
-Densities provide the adecuate conceptual tool to obtain ``Centers of Mass'' or vector averages.
+Densities provide the adequate conceptual tool to obtain ``Centers of Mass'' or vector averages.
 A density is a measure of a quantity per spatial spread, at precise locus in space.
@@
 \end{equation}
-An instantaneous Center of Mass, thus calculated, would present little if none physiological interpretation.
+An instantaneous Center of Mass, thus calculated, would present little or none physiological interpretation.
 The activity could be spread over non-convex sets (which would include disjoint and concave sets) and so the vector average, the Center of Mass, would be located in points where there is no activity.
-In our example the use of the CA complex of the hippocampus proves an ideal example.
+In our example the use of the CA complex of the hippo campus proves an ideal example.
 This structure is concave, and during epileptic burst, can have activity spread over many disjoint sites simultaneously.
@@
 Every patch that we can delimit as a disjoint component of the Sources/Sink sets indicates a coherent, joint active group of units.
-A center of mass for each disjoint component would then have much
+A center of mass for each disjoint component would then indicate an average locus for this active group, a putative center of activity to which we can give a position, intensity and time.
+An instantaneous snapshot of such information would depict the centers of activity of an ordered structure of neurons and their relative intensities.
+But it is in the successive analysis of a series of such snapshots that this analysis can show its true power.
+The apparent , movement, and disappearance of such centers could reveal very subtle details of the dynamics of neural activity.
+Very small, fast displacements of the Centers of mass, could reveal slight asynchronism between neurons which could belong to the same active group.
+Large, slower than a few firing periods, could indicate physiological connections between various groups.
+
+\section{The Recipe}
+
+In order to perform the analysis described here, we summarize the steps.
+\begin{enumerate}
+\item Acquire high density electrophysiological data.
+The data must be two or three dimensional in space.
+Distance between the recording sites must be of the order of the size of the typical neurons involved, and such sites must be on a dense enough grid to perform differential operation over them.
+The sampling frequency should be enough to detect subthreshold activity, of at least 5kHz.
+Also, the data should come from structured tissue.
+Unordered, highly homogeneous tissue, will not yield significant results.
+\item Obtain, frame by frame, the CSD from the data. Any method may be used, but computer expensive techniques are discouraged if the data is dense enough. Common sense and tests are guidelines here.
+\item Separate the CSD data into two three sets: sources, sinks, and border or indiscernible (below error) data.
+\item Perform disjoint component analysis in the sources and sinks sets, frame by frame.
+\item  Obtain the CM for each disjoint component.
+\item For each frame and its successor, and for each CM, perform a ``most probable successor'' detection.
+Here we must put numerical criteria according to the data avaible, i.e.
+how far awat and how different in intensity should be allowed a CM to be at time $t+1$ to be considered succesor in time of another at time $t$.
+  \item plot the data and try to figure what it means.
+\end{enumerate}
+
+A series of experiments done in the laboratory on rat's hippocampus provided us with the data that we shall use to illustrate the CSD and CM analysis proposed above.
+
+\section{Application to epileptic-like activity in the hippocampus}
+Chronologically speaking, we performed first the analysis to the most complicated data (numerically and qualitatively), which where a series of epileptic burst that occured after 15 minutes of application of 4AP to living slices of hippocampus.
+The burst occured with activity spreading over many sites simultaneously.
+This turned out to be a test of the applicability and robustness of our method.
+The necessity of having good error value for the ``thick zero'' set was more apparent here than anywhere else.
+Also the posibilty to create new structures of numerical data was seen here.
+But also the most strickig features of the propagation of activity, coded now as trajectories of the CM, could be seen here, as the epileptic burst covered the whole CA structure.
+
+\section{Application to evocated activity in idem.}
+
+The evocated activity seem to yield much cleaner results, as we can now compare the trajectories of successive, very similar events.
+Our CSD and CM analysis can be used to detect consistent features in these experiments, such as the direction the waxing and wanning of activity.
+Small deviations from the trajectory in successive experiments that are above the numerical error could indicate inconsistencies in the network below.
+
+\section{Application to spontaneous activity in idem.}
+
+We had one experiment in which we could trace very nicely the data on DG and CA simultaneously without adding excitatory drugs.
 

### 2016-01-18 ArticuloCSDyCM/Paper03.tex
version 03 a Rafa

--- 2016-01-06 ArticuloCSDyCM/Paper01.tex
@@
 \usepackage[utf8]{inputenc}
+\usepackage[margin=1in]{geometry}
+\usepackage{lscape}
+\usepackage{mathptmx}
+\usepackage[scaled=.90]{helvet}
+\usepackage{courier}
+%\usepackage{lipsum}% just to generate filler text
+\usepackage{setspace}
+\usepackage{graphicx}
+\usepackage{longtable}
 
@@
 Thanks to the density and scales of the BioCAM 4096, in our case a ``classical'' numerical derivation of the CSD is the most adecuate route to take.
-A numerical Laplacian Operator which reduces the cross efect of rectangular grids is the convolution of the data with the following matrix \cite{Lindberg90}:
+A numerical Laplacian Operator which reduces the cross effect of rectangular grids is the convolution of the data with the following matrix \cite{Lindberg90}:
 \begin{equation}
@@
 \end{equation}
-At the border we duplicate the last rows of data and take the results knowingly of the border effects.
-As this sort of operators are extremely sensitive to noise, a Gaussian blur is applied before presenting the graphical data.
-This softening of the edges in the CSD image also represents our ignorance of the finer details of the components of the distribution (see section bla bla).
+This sort of operators are extremely sensitive to edges.
+The noise of the data can be perceived in the spatial domain as rapidly varying edges.
+A Gaussian spatial denoising can be performed before or after the Laplacian operator to the data in order to reduce excessive borders.
+Our Gaussian blur filter has a $\sigma$ value of three electrodes or $126 \mu m$.
+Intuitively this means that it smooths over each structure smaller than the soma of a typical pyramidal cell.
+This is  in accordance to our mean field approach, where we expect smooth differentiable fields over the scale of the neurons, but not in finer scales (for details see \cite{Bedard11}, further work on this subject will be presented elsewhere \cite{IsabelYo}).
 
+On previous works it has been argued that numerical difference operators sacrifice all data on the borders of the array.
+For electrode arrays of lesser density, this could posse a problem, but in our case we only renounce to 27 out of 4095 channels (one is grounded).
 More sophisticated methods such as iCSDA \cite{Leski2011} and kCSDA \cite{Potworowski2011} prove little advantage in estimating the sources under the ohmic, isotropic and homogeneous assumptions at the scales involved in our data, and turn out to be computer-costly for high density MEAs and dense temporal sampling.
@@
 As these are indicators of activity in the neurons, the CSD is also, indirectly, a \emph{``density of local activity''}.
-The advantages of this representation of the activity in place of the LFP have been stated many times (better spatial resolution, mostly), but is use as \emph{density} has been overlooked.
+The advantages of this representation of the activity in place of the LFP have been stated many times (better spatial resolution, mostly).
+To the best of our knowledge, the conceptual advantages of using this density as such has never been done.
 
-Center of Mass Analysis (CMA) has previously been used in order to track putative  ``trajectories'' of the activity or information across neural tissue or even the whole brain \cite{Chao05, Manjarrez07}.
-In such works the chosen ``mass'' was an heuristic measure of activity adecuate for representing the corse-grain picture of displacement of activity or some unit of ``active information''.
-But these present the usual problem of justifing the choice maide on practical grounds.
-For calculating a vectorial average, the adecuate choice of wheights per vector is a density, a positive quantity that measures the amount of something per unit volume.
+On the other hand, Center of Mass Analysis (CMA) has been used previously in order to track putative  ``trajectories'' of the activity or information across neural tissue or even the whole brain \cite{Chao05, Chao07, Manjarrez07, Manjarrez09}.
+In such works the chosen ``mass'' was an heuristic measure adecuate for representing the corse-grain picture of displacement of activity.
+These present the usual problem of justifing the choice made on practical grounds.
+For calculating a vectorial average, the adecuate wheight per vector is a density, a positive quantity that measures the amount of something per unit volume.
 CSD sepparates naturally the measured activity into two distinct sets, namely, the set of all sources and the set of all sinks, separated by the zero set.
 The sepparation is not quantitative but qualitative: even when the signs of electric charge are arbitrary, the distintion between them is not.
-This is also in contrast to LFP, where the zero of the function has no precise meaning as electrical potentials can be measured from arbitrary ground values.
+Therefore, CSD represents on each of the two sets (sinks and sources) exactly what we need, a unit of concentration per unit space, in this case, per area.
+Also, these separation by the zero value set of the CSD is in contrast to LFP or EEG measurements, where the zero set has no precise meaning as it only represents a practical ground value.
+This is aknowledged in most of the literature, where only the scales of the LFP recording are shown, without reference to the sign of the values.
 
-A visual inspection of the CSD colormap at a given instant in time shows that both sources and sinks appear in rather large and significant patches.
+A first visual inspection of the CSD colormap at a given instant in time shows that both sources and sinks appear in rather large and significant patches.
 These appear to correspond to the expected double and triple poles of fyring neurons in structured tissue \cite{} and inspired us to separate the CSD subsets into their \emph{disjoint components}.
@@
 
-CSD divides the set of measuring points into two subsets: the set of all sources and the set of all sinks, i.e, the set of all measuring points that have a CSD positive value, and the set of those who have a negative CSD value.
-The zero set is a border between them, but due to the finite precision of the measurement (and the validity of the mean field approach!) these sets can make contact with each other.
-Still, noise between patches corresponding to a single set could make a ``bridge'' between components, making identification of poles inexact.
+Theoretically, CSD would divide the set of measuring points into three subsets: the set of all sources, the set of all sinks, and the border between them i.e, the set of all measuring points that have a CSD positive value, the set of those who have a negative CSD value, and the set of those points having exactly the zero value in CSD space.
+Due to the finite precision of the measurement, and of the validity of the mean field approach the sinks and sources set can make contact with each other.
+This is because our measurements produce a function over a discrete set, a continuous function corresponding CSD is only a theoretical tool.
+Even more, noise between patches corresponding to a single set could make a ``bridges'' between components, making identification of poles inexact.
 In order to avoid this we create a ``thick zero'' set of all the recording points whose CSD values are inside a small error interval around zero.
-This sepperates even more the components of the set, and helps us to simplify the following analysis.
+This separates even more the components of the set, represents our ignorance of precise borders and helps us to simplify the following analysis.
 
@@
  An instantaneous Center of Mass, thus calculated, would present little or none physiological interpretation.
-The activity could be spread over non-convex sets (which would include disjoint and concave sets) and so the vector average, the Center of Mass, would be located in points where there is no activity.
-In our example the use of the CA complex of the hippo campus proves an ideal example.
+The activity could be spread over non-convex sets (which would include disjoint and concave sets) and so the vector average would be located in points where there is no measurement of activity.
+In our example the use of the CA complex of the hippocampus proves an ideal example.
 This structure is concave, and during epileptic burst, can have activity spread over many disjoint sites simultaneously.
+If our method is reliable, then, even under such constrains we may be able to track the activity and provide some physiological interpretation.
+That is why we have used data that was obtained from an experiment in which 4AP (bla bla for Franco).
+
  We must not take out of sight the phenomena which is producing the CSD distribution: neurons have a finite size, and the effect of their action a finite extent.
-Every patch that we can delimit as a disjoint component of the Sources/Sink sets indicates a coherent, joint active group of units.
+Every patch that we can delimit as a disjoint component of the Sources/Sink set indicates a coherent, joint active group of units.
 A center of mass for each disjoint component would then indicate an average locus for this active group, a putative center of activity to which we can give a position, intensity and time.
@@
 But it is in the successive analysis of a series of such snapshots that this analysis can show its true power.
-The apparent , movement, and disappearance of such centers could reveal very subtle details of the dynamics of neural activity.
+The appearance , displacement, and disappearance of such centers could reveal very subtle details of the dynamics of neural activity.
 Very small, fast displacements of the Centers of mass, could reveal slight asynchronism between neurons which could belong to the same active group.
-Large, slower than a few firing periods, could indicate physiological connections between various groups.
+Larger and  slower than a few firing periods displacements could indicate physiological connections between various groups.
 
@@
  \begin{enumerate}
-\item Acquire high density electrophysiological data.
+ \item Acquire high density electro-physiological data.
 The data must be two or three dimensional in space.
-Distance between the recording sites must be of the order of the size of the typical neurons involved, and such sites must be on a dense enough grid to perform differential operation over them.
-The sampling frequency should be enough to detect subthreshold activity, of at least 5kHz.
+Distance between the recording sites must be of the order of the size of the typical neurons involved, and such sites must be on a dense enough grid to perform numerical difference operations over them.
+The sampling frequency should be enough to detect sub-threshold activity, of at least 5kHz.
 Also, the data should come from structured tissue.
 Unordered, highly homogeneous tissue, will not yield significant results.
-\item Obtain, frame by frame, the CSD from the data. Any method may be used, but computer expensive techniques are discouraged if the data is dense enough. Common sense and tests are guidelines here.
-\item Separate the CSD data into two three sets: sources, sinks, and border or indiscernible (below error) data.
+\item Obtain, frame by frame, the CSD from the data.
+Any method may be used, but computer expensive techniques are discouraged if the data is dense enough.
+Common sense and tests are guidelines here.
+We recommend simpler, difference based approaches.
+\item Separate the CSD data into three sets: sources, sinks, and border or indiscernible (below error) data.
 \item Perform disjoint component analysis in the sources and sinks sets, frame by frame.
+  A single pass algorithm could be used here \cite{Vincent91, Abubaker07}.
 \item  Obtain the CM for each disjoint component.
@@
 \section{Application to epileptic-like activity in the hippocampus}
-Chronologically speaking, we performed first the analysis to the most complicated data (numerically and qualitatively), which where a series of epileptic burst that occured after 15 minutes of application of 4AP to living slices of hippocampus.
-The burst occured with activity spreading over many sites simultaneously.
+Chronologically speaking, we performed first the analysis to the most complicated data (numerically and qualitatively), which where a series of epileptic burst that occurred after 15 minutes of application of 4AP to living slices of hippocampus.
+The burst occurred with activity spreading over many sites simultaneously.
 This turned out to be a test of the applicability and robustness of our method.
@@
 
-\section{Application to evocated activity in idem.}
+\section{Application to evoked activity in idem.}
 
-The evocated activity seem to yield much cleaner results, as we can now compare the trajectories of successive, very similar events.
+The evoked activity seems to yield much cleaner results, as we can now compare the trajectories of successive, very similar events.
 Our CSD and CM analysis can be used to detect consistent features in these experiments, such as the direction the waxing and wanning of activity.
@@
 
+\bibliographystyle{plain}
+\bibliography{../Reportes/BiblioReportes01}
+
 \end{document}

### 2016-01-28 ArticuloCSDyCM/Paper03.tex
paper03 mas y mas

@@
 
-\section{CSD for High density MEAs}
+\section{Introduction}
 
-The concepts presented here require first obtaining the CSD of the recorded LFP.
+Spatialy tracking the activity of electrophysiological data in a clear, concise, and interpretable manner is a challenging problem.
+
+\section{CSD, CM and Disjoint Components}
+
+\subsection{CSD as weight density for vectors}
+
+One of the most obvious uses of the Current Source Density (CSD) in place of the Local Field Potentials (LFP) has been overlooked.
+This might have occurred because  samplings at lesser spatial resolution did not permit extending the implications of the concept of density.
+Density is \emph{quantifiable entity per unit of space}, or, in other words, concentration of something which is distributed in space.
+In this case, the ``something'' are the sources and sinks of electric current.
+Those can be interpreted as the interchange of charged ions from the inside to the outside of cellular membranes.
+This is an indicator of activity in the neurons, therefore the CSD is also, indirectly, a \emph{``density of local activity''}.
+Some of the advantages of this representation of the activity in place of the LFP have been stated many times (better spatial resolution, mostly).
+To the best of our knowledge, the conceptual advantages of using this density as such has never been done.
+A density is the adequate concept for calculating integrated measures, such as \emph{weights} of units or entities, \emph{averages} and others.
+We are particularly interested in its use for obtaining \emph{vector averages}, also known informally as Centers of Mass (CM).
+
+Center of Mass Analysis (CMA) has been used previously in order to track putative  ``trajectories'' of the activity or information across neural tissue or even the whole brain \cite{Chao05, Chao07, Manjarrez07, Manjarrez09}.
+In such works the chosen ``mass'' was an heuristic measure adecuate for representing the corse-grain picture of displacement of activity.
+These present the problem of justifing the choice made on practical grounds.
+For calculating a vectorial average, the adecuate wheight per vector is a density, a positive quantity that measures the amount of something per unit volume, so it is only adequate that we search for a natural density in the electrophysiological experiments, and the mos obvious one is the CSD which can be obtained through various methods from the LFP recordings if the spatial sampling is rich enough.
+
+Apparently, CSD would not fit the requirements of non-negativity for its use as a mathematical density.
+CSD sepparates the measured activity into three distinct sets, namely, the set of all sources, the set of all sinks and the zero set.
+The sepparation is not quantitative but qualitative: even when the signs of electric charge are arbitrary, the distintion between them is not.
+Therefore, CSD represents on each of the two sets (sinks and sources) exactly what we need, a unit of concentration per unit space, in this case, per area, and the sign can be ignored.
+The separation by the zero value set of the CSD is in contrast to LFP or EEG measurements, where the zero set has no precise meaning as it only represents a practical ground value.
+This is aknowledged in most of the literature, where only the scales of the LFP recording are shown, without reference to the sign of the values.
+Here the zero set has a precise meaning as instantaneous lack of activity and sepparation between sinks and sources.
+This shall be exploited when we explore a further refinement of the analysis.
+
+Once the sepparation has been made, we would have two densities, or more preciselly, two different density functions defined over two sepparate sets.
+Trying to use this for obtaining vector averages would still be to crude.
+If these sets spread out over non-convex geometries, the vector averages could lie outside the sets, having little or none physiological interpretation.
+Therefore we decided to introduce another concept from geometricall set theory in order to make the analysis more local in character.
+
+\subsection{CSD and Disjoint Components}
+
+A first visual inspection of the CSD color map at a given instant in time shows that both sources and sinks appear in rather large and significant patches.
+These appear to correspond to the expected double and triple poles of firing neurons in structured tissue \cite{} and inspired us to separate the sink and source  sets into their \emph{disjoint components}.
+This concept is used in mathematics to denote subsets of a given geometric set that do not touch each other.
+Their rough correspondence to poles of firing neurons would be most useful in tracking the successive locus of the activity.
+
+To clarify the concept and how we are we using it an example is in order.
+In figure \ref{ejemplodisjuntos} we show geometrical sets enclosed by simple curves as color patches, where every color indicates a Set.
+The set A consists of a single component, that means that one can make  path between any two points in the set that consist entirely of points of the same set.
+In contrast, the set B is made of two disjoint components, labeled B1 and B2.
+If one chooses a point in each of the components, one cannot draw a path joining them without leaving the set.
+But each one of those components is connected, meaning that if we choose the pair of points in one component, then again we can trace a path joining them in the set.
+So the set B consists of two disjoint components, but each component can be regarded as a connected set on their own.
+Likewise the set C is made of three disjoint components.
+
+Theoretically, CSD would divide the set of measuring points into three subsets: the set of all sources, the set of all sinks, and the border between them i.e, the set of all measuring points that have a CSD positive value, the set of those who have a negative CSD value, and the set of those points having exactly the zero value in CSD space.
+Due to the finite precision of the measurement, the validity of the mean field approach, and the noise in the measurements, the sinks and sources sets can make contact with each other.
+Said measurements produce a function over a discrete point set, but we tend to think of them as continuous functions.
+Also, noise between patches corresponding to a single set could make a ``bridge'' between components, making identification of poles inexact.
+In order to avoid this we create a ``thick zero'' set of all the recording points whose CSD values are inside a small error interval around zero.
+This separates even more the components of the set, represents our ignorance of precise borders and helps us to simplify the following analysis.
+
+The ``thick zero'' would help us to sepparate the sinks and sources sets into its disjoint components, which would correspond then to clear patches of definite activity.
+If each one of these patches could be interpreted as an active unit, we may asign to it a ``center of activity'' in order to follow their displacements.
+A Center of Mass for each nstantaneous disjoint sink/source would provide us a very sensitive measurement of ``active locus''.
+We can then calculate at each time frame the CM for each disjoint component before tracking displacements of activity.
+
+At each disjoint component at a precise time $t$, indexed $k$, we have a number of points with CSD of the same sign, indexed $j$. Then, the CM of the $k$ component would have the usual definition:
+\begin{equation}\label{cmparadisj}
+   \langle q(t) \rangle_k =\frac{\sum_j q_{j,k} (t) I_{j,k} (q_{j,k},t)}
+           {\sum_j I(q_{j,k},t)},
+\end{equation}
+where $I(q_{j,k},t)$ is the CSD at point $q_{j,k}=(x_{j,k}, y_{j,k})$ at time $t$.
+These disjoint components, while not rigorously convex, are less spread over the space of the measurements and tend to have their CM over the actual active points, or at the border of the active units.
+
+\subsection{CM and trajectories}
+
+We must not take out of sight the phenomena which is producing the CSD distribution: neurons have a finite size, and the effect of their action a delimited extent.
+Every patch that we can identify as a disjoint component of the Sources/Sink sets indicates a coherent, joint active group of units.
+A center of mass for each disjoint component would then indicate an average locus for this active group, a putative center of activity to which we can give a position, intensity and time.
+An instantaneous snapshot of such information would depict the centers of activity of an ordered structure of neurons and their relative intensities.
+But it is in the successive depiction of a series of such snapshots that this analysis can show power.
+The appearance , displacement, and disappearance of such centers could reveal very subtle details of the dynamics of neural activity.
+Very small, fast displacements of the Centers of mass, may be occur due to slight asynchronism between neurons which could belong to the same active group.
+Larger and  slower than a few firing periods displacements could indicate physiological connections between various groups.
+
+The next step in our analysis would then a procedure to systematically associate to each CM at time $t$ a ``successor'' at time $t+\Delta t$, where $\Delta t$ is the sampling time step.
+We have used a pseudo-distance function which takes into account the geometrical distance from the putativa succesors for a CM, and also the intensity.
+For two CM inside our measurement space we define our pseudo-distance as follows.
+Let $q_1=(x_1, y_1)$ be the coordinates of the first point and $q_2=(x_2,y_2)$, the corresponding coordinates of the second, and let $I_1, I_2$ be their corresponding integrated intensities.
+Then the ``distance'' that separates them shall be
+\begin{equation}\label{pseudodist}
+d(q_1,q_2):=\sqrt{(x_1-x_2)^2+(y_1+y_2)^2+C(I_1-I_2)},
+\end{equation}
+where $C$ is a conveniently chosen constant, so as to give a similar importance to the intensities as to the positions.
+The integrated intensity of a CM is simply the denominator in the eq. \ref{cmparadisj}, that is, the total ``mass'' obtained by integrating the CSD in a specific disjoint component:
+\begin{equation}
+I_k(t)=\sum_j I (q_{j,k}, t)
+\end{equation}
+We specify then a tolerance $\delta$.
+If a certain $t$ a CM is below $\delta$ distance from another at $t+\Delta t$, we consider the latter the successor in time of the former.
+By continuously applying such procedure, we can begin to trace trajectories of the CMs, as they appear, wander, and fade.
+
+ \section{The Procedure}
+\subsection{Recipe in a nutshell}
+
+ In order to perform the analysis described here, we summarize the steps.
+ \begin{enumerate}
+ \item Acquire high density electro-physiological data.
+The data must be two or three dimensional in space.
+Distance between the recording sites must be of the order of the size of the typical neurons involved, and such sites must be on a dense enough grid to perform numerical difference operations over them.
+The sampling frequency should be enough to detect sub-threshold activity, of at least 5kHz.
+Also, the data should come from structured tissue.
+Unordered, highly homogeneous tissue, will not yield significant results.
+\item Obtain, frame by frame, the CSD from the data.
+Any method may be used, but computer expensive techniques are discouraged if the data is dense enough.
+Common sense and tests are guidelines here.
+We recommend simpler, difference based approaches.
+Our approach is detailed below.
+\item Separate the CSD data into three sets: sources, sinks, and border or indiscernible (below error) data.
+\item Perform disjoint component analysis in the sources and sinks sets, frame by frame.
+  A single pass algorithm could be used here \cite{Vincent91, Abubaker07}.
+\item  Obtain the CM for each disjoint component, using eq. \ref{cmparadisj}.
+\item For each frame and its successor, and for each CM, perform a ``most probable successor'' detection.
+Here we must put numerical criteria according to the data avaible, i.e.
+how far away and how different in intensity should be allowed a CM to be at time $t+1$ to be considered succesor in time of another at time $t$.
+We use the distance defined in eq.
+\ref{pseudodist}, a sort of Euclidian-like distance.
+\item ``Connect the dots'', that is, apply succesively the step above for each CM, until it stops having a succesor or its intensity falls below the error criteria.
+Then plot the succesive CMs as one trayectory.
+  \item Plot all the trayectories and sort visually those that may have physiological interpretation from those who may artifacts of the method. 
+\end{enumerate}
+
+\subsection{CSD for high density MEAS}
+
+The concepts presented here require first obtaining the CSD from the recorded LFP.
 Any method would present the same utility.
@@
 
-\section{CSD and disjoint components}
+\section{Experimental examples}
 
-One of the most obvious uses of the CSD in place of the LFP has been overlooked.
-This might have occurred because  samplings at lesser spatial density did not permit extending the implications of the concept of density.
-Density is \emph{quantifiable entity per unit of space}, or, in other words, concentration of something distributed in space.
-In this case, the ``something'' is the sources and sinks of electric current, which can be identified with the interchange of charged ions from the inside to the outside of cellular membranes.
-As these are indicators of activity in the neurons, the CSD is also, indirectly, a \emph{``density of local activity''}.
-The advantages of this representation of the activity in place of the LFP have been stated many times (better spatial resolution, mostly).
-To the best of our knowledge, the conceptual advantages of using this density as such has never been done.
-
-On the other hand, Center of Mass Analysis (CMA) has been used previously in order to track putative  ``trajectories'' of the activity or information across neural tissue or even the whole brain \cite{Chao05, Chao07, Manjarrez07, Manjarrez09}.
-In such works the chosen ``mass'' was an heuristic measure adecuate for representing the corse-grain picture of displacement of activity.
-These present the usual problem of justifing the choice made on practical grounds.
-For calculating a vectorial average, the adecuate wheight per vector is a density, a positive quantity that measures the amount of something per unit volume.
-CSD sepparates naturally the measured activity into two distinct sets, namely, the set of all sources and the set of all sinks, separated by the zero set.
-The sepparation is not quantitative but qualitative: even when the signs of electric charge are arbitrary, the distintion between them is not.
-Therefore, CSD represents on each of the two sets (sinks and sources) exactly what we need, a unit of concentration per unit space, in this case, per area.
-Also, these separation by the zero value set of the CSD is in contrast to LFP or EEG measurements, where the zero set has no precise meaning as it only represents a practical ground value.
-This is aknowledged in most of the literature, where only the scales of the LFP recording are shown, without reference to the sign of the values.
-
-A first visual inspection of the CSD colormap at a given instant in time shows that both sources and sinks appear in rather large and significant patches.
-These appear to correspond to the expected double and triple poles of fyring neurons in structured tissue \cite{} and inspired us to separate the CSD subsets into their \emph{disjoint components}.
-This concept is used in mathematics to denote subsets of a given geometric set that do not touch each other.
-Their rough correspondence to poles of firing neurons would be most useful in tracking the successive locus of the activity.
-
-To clarify the concept and how we are we using it an example is in order.
-In figure \ref{ejemplodisjuntos} we show geometrical sets enclosed by simple curves as color patches, where every color indicates a Set.
-The set A consists of a single component, that means that one can make  path between any two points in the set that consist entirely of points of the same set.
-In contrast, the set B is made of two disjoint components, labeled B1 and B2.
-If one chooses a point in each of the components, one cannot draw a path joining them without leaving the set.
-But each one of those components is connected, meaning that if we choose the pair of points in one component, then again we can trace a path joining them in the set.
-So the set B consists of two disjoint components, but each component can be regarded as a connected set on their own.
-Likewise the set C is made of three disjoint components.
-
-Theoretically, CSD would divide the set of measuring points into three subsets: the set of all sources, the set of all sinks, and the border between them i.e, the set of all measuring points that have a CSD positive value, the set of those who have a negative CSD value, and the set of those points having exactly the zero value in CSD space.
-Due to the finite precision of the measurement, and of the validity of the mean field approach the sinks and sources set can make contact with each other.
-This is because our measurements produce a function over a discrete set, a continuous function corresponding CSD is only a theoretical tool.
-Even more, noise between patches corresponding to a single set could make a ``bridges'' between components, making identification of poles inexact.
-In order to avoid this we create a ``thick zero'' set of all the recording points whose CSD values are inside a small error interval around zero.
-This separates even more the components of the set, represents our ignorance of precise borders and helps us to simplify the following analysis.
-
- \section{CSD and Center of Mass}
-
- Densities provide the adequate conceptual tool to obtain ``Centers of Mass'' or vector averages.
-A density is a measure of a quantity per spatial spread, at precise locus in space.
-CSD has, in contast to mass density, negative values.
-But these signs are a matter of convention, it is more important to say that it has two different cualitative propiertes and that sepparates the domain into two sets.
-Each set has then a total ``mass'', the integral of the absolute value of the CSD.
-Then the concept of ``Center of Mass'' applies naturally.
-Once that we sepparate the two sets and apply the concept in each one independently, even the sign convention of the densities seases to be a problem, as it cancels in the formula for the CM:
- \begin{equation}
- \langle x(t) \rangle =\frac{\sum_j x_{j} (t) I_{j} (x,t)} {\sum_j I_{j}(x,t)}
- \end{equation}
- An instantaneous Center of Mass, thus calculated, would present little or none physiological interpretation.
-The activity could be spread over non-convex sets (which would include disjoint and concave sets) and so the vector average would be located in points where there is no measurement of activity.
-In our example the use of the CA complex of the hippocampus proves an ideal example.
-This structure is concave, and during epileptic burst, can have activity spread over many disjoint sites simultaneously.
-If our method is reliable, then, even under such constrains we may be able to track the activity and provide some physiological interpretation.
-That is why we have used data that was obtained from an experiment in which 4AP (bla bla for Franco).
-
- We must not take out of sight the phenomena which is producing the CSD distribution: neurons have a finite size, and the effect of their action a finite extent.
-Every patch that we can delimit as a disjoint component of the Sources/Sink set indicates a coherent, joint active group of units.
-A center of mass for each disjoint component would then indicate an average locus for this active group, a putative center of activity to which we can give a position, intensity and time.
-An instantaneous snapshot of such information would depict the centers of activity of an ordered structure of neurons and their relative intensities.
-But it is in the successive analysis of a series of such snapshots that this analysis can show its true power.
-The appearance , displacement, and disappearance of such centers could reveal very subtle details of the dynamics of neural activity.
-Very small, fast displacements of the Centers of mass, could reveal slight asynchronism between neurons which could belong to the same active group.
-Larger and  slower than a few firing periods displacements could indicate physiological connections between various groups.
-
- \section{The Recipe}
-
- In order to perform the analysis described here, we summarize the steps.
- \begin{enumerate}
- \item Acquire high density electro-physiological data.
-The data must be two or three dimensional in space.
-Distance between the recording sites must be of the order of the size of the typical neurons involved, and such sites must be on a dense enough grid to perform numerical difference operations over them.
-The sampling frequency should be enough to detect sub-threshold activity, of at least 5kHz.
-Also, the data should come from structured tissue.
-Unordered, highly homogeneous tissue, will not yield significant results.
-\item Obtain, frame by frame, the CSD from the data.
-Any method may be used, but computer expensive techniques are discouraged if the data is dense enough.
-Common sense and tests are guidelines here.
-We recommend simpler, difference based approaches.
-\item Separate the CSD data into three sets: sources, sinks, and border or indiscernible (below error) data.
-\item Perform disjoint component analysis in the sources and sinks sets, frame by frame.
-  A single pass algorithm could be used here \cite{Vincent91, Abubaker07}.
-\item  Obtain the CM for each disjoint component.
-\item For each frame and its successor, and for each CM, perform a ``most probable successor'' detection.
-Here we must put numerical criteria according to the data avaible, i.e.
-how far awat and how different in intensity should be allowed a CM to be at time $t+1$ to be considered succesor in time of another at time $t$.
-  \item plot the data and try to figure what it means.
-\end{enumerate}
-
-A series of experiments done in the laboratory on rat's hippocampus provided us with the data that we shall use to illustrate the CSD and CM analysis proposed above.
+The necessity to create such analysis came originally from the study of electro-physiological data obtained from the rat's hippocampus.
+The highly structured CA region of the hippocampus provides strong, discernible patterns of Sources and Sinks.
+Following the pattern of activity using Center of Mass analysis directly from the data, though, proves unfruitful, due to the non-convex nature of the layered tissue.
+Using CSD and disjoint component tracking over them provided the right conceptual framework to follow the displacement of active areas during highly active events, such as epileptic-like activity that was pharmacologically induced.
+The proposed analysis would not produce vector averages over different simultaneously active areas, on the contrary, it would separate them and track them independently.
+So the most violent epileptic burst provided us with the data necessary to prove the robustness of the method.
 
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{abstract}
 \bibliography{../Reportes/BiblioReportes01}

### 2016-01-29 ArticuloCSDyCM/Paper03.tex
mas intro paper03

@@
 Spatialy tracking the activity of electrophysiological data in a clear, concise, and interpretable manner is a challenging problem.
+The data from electrophysiological recordings are usually presented first, as a series of traces that correspond each to a sampling point in the studied structure.
+As Micro Electrode Arrays (MEAs) become denser, this representation of the data becomes saturated and the discovery of patterns turns out a difficult task, and novel information gets cluttered or obfuscated by the density of the data.
+
+The succesive electrophysiological action of distinct units on structured tissue would trace a ``path'' or ``trajectory'' of the activity.
+A tracing of such trajectories would help to indicate causal or correlated ``efective conections'' in the tissue.
+How to define and acquire such traces is an open question.
+In previous works, some form of vector averaging using heuristic wheights obtained from the Local Field Potential or the number of action potentials have been presented \cite{Chao05, Manjarrez07}.
+Here we advance on such methdos by incorporating mathematically sound definitions and techniques which allow us to sepparate spatialy distinct traces.
+
+An important piece of information that can be aquired from a dense enough MEA would be putative trajectories traced out by active units, as they fire in succesion by causal relations or as a side effect of correlation in their workings.
+This is part of the ``efective  conectome problem'', that is, the search for effective causality paths that convey activity (and ultimately, information) inside the nervous structures.
 

### 2016-02-01 ArticuloCSDyCM/Paper03.tex
paper03 again

--- 2016-01-28 ArticuloCSDyCM/Paper03.tex
@@
 
+\author{Laboratorio 19 Crew}
+\title{CSD, Disjoint components and trayectory detection}
+
 \begin{document}
+
+\maketitle
 
@@
 
-Spatialy tracking the activity of electrophysiological data in a clear, concise, and interpretable manner is a challenging problem.
+Spatially tracking the activity of electro-physiological data in a concise, rigorous and interpretable manner poses a challenging problem.
+Measurement recordings are usually presented first as a series of traces that correspond each to a sampling point in the studied structure.
+As Micro Electrode Arrays (MEAs) become denser, this representation of the data becomes saturated and the discovery of patterns turns into a difficult task.
+Novel information is also cluttered or obfuscated by the same high density acquisition that allows us to discover it.
+Therefore, representing  the recorded data in ways which can unveil previously unobtainable information.
+Our task in this paper is to present a method for tracing out ``functional connectivity trajectories'' of activity in structured neural tissue, and we shall exemplify the method by application to the data obtained from recordings of the rats hippocampus.
+
+Succesive  of distinct units on structured tissue would trace a ``trajectory'' of the activity.
+A tracing of such trajectories would help to indicate causal or correlated ``functional conections'' over the tissue.
+Neighbouring neurons that fire in succesion, integrated subthreshold activity and ephaptic coupling could all be interpreted parts of the paths where the activity gets transmitted.
+Nevertheless, how to define and acquire the corresponging traces is an open question.
+
+In previous works, some form of vector averaging has been proposed and shown to reveal some trajectrories traced out by the activity.
+We shall call such methods Center of Mass Analyses (CMA).
+The \emph{wheights} that have been chosen for obtaining the average have been selected on heuristic grounds.
+In the works of Chao \emph{et al.} \cite{Chao05, Chao07}, the density of spikes in a certain interval was used, a sort of ``density of action'', while Manjarrez \emph{et al.} \cite{Manjarrez07, Manjarrez09} used the positive half of the LFP recording as their weight.
+Here we advance on such methods by incorporating mathematically sound definitions and concepts which allow us to rigorously separate spatially distinct traces.
+Our weight is actually the natural density function provided by the data: the Current Source Density.
+This representation not only provides a mathematical fit measure for the weights in the vector averaging process, but helps to separate the data into disjoint components.
+Based on this, a further refinement of the technique, given by the polar structure of firing neurons, allowed us to obtain accurate, well defined trajectories of the activity in the studied tissue.
+
+The application of the method requires electro-physiological data acquired through bi-dimensional MEAs, although the extension to three dimensional arrays is possible in principle.
+The MEA should have dense enough spatial and temporal samplings, so as to obtain good estimates of the CSD and their changes in time.
+
+Our method permits us to map possible trajectories of activity inside structured neuronal tissue, without requiring ad hoc geometric constraints.
+As a prime example, non-convex structures which may show simultaneous activity in separate parts could test the robustness of our method.
+We have tested then the method on slices of rat's hippocampus, concentrating in the CA region.
+We exemplify the usage of the method with data that has been obtained by an experiment \textbf{bla bla bla describr experimentos}.
 
@@
 
-One of the most obvious uses of the Current Source Density (CSD) in place of the Local Field Potentials (LFP) has been overlooked.
-This might have occurred because  samplings at lesser spatial resolution did not permit extending the implications of the concept of density.
-Density is \emph{quantifiable entity per unit of space}, or, in other words, concentration of something which is distributed in space.
-In this case, the ``something'' are the sources and sinks of electric current.
-Those can be interpreted as the interchange of charged ions from the inside to the outside of cellular membranes.
-This is an indicator of activity in the neurons, therefore the CSD is also, indirectly, a \emph{``density of local activity''}.
-Some of the advantages of this representation of the activity in place of the LFP have been stated many times (better spatial resolution, mostly).
-To the best of our knowledge, the conceptual advantages of using this density as such has never been done.
-A density is the adequate concept for calculating integrated measures, such as \emph{weights} of units or entities, \emph{averages} and others.
-We are particularly interested in its use for obtaining \emph{vector averages}, also known informally as Centers of Mass (CM).
-
 Center of Mass Analysis (CMA) has been used previously in order to track putative  ``trajectories'' of the activity or information across neural tissue or even the whole brain \cite{Chao05, Chao07, Manjarrez07, Manjarrez09}.
@@
 These present the problem of justifing the choice made on practical grounds.
-For calculating a vectorial average, the adecuate wheight per vector is a density, a positive quantity that measures the amount of something per unit volume, so it is only adequate that we search for a natural density in the electrophysiological experiments, and the mos obvious one is the CSD which can be obtained through various methods from the LFP recordings if the spatial sampling is rich enough.
+For calculating a vectorial average, the adecuate wheight per vector is a density, a positive quantity that measures the concentration of something per space unit, so it is only adequate that we search for a natural density in the electrophysiological experiments.
+The Current Source Density fits the mathematical and interpretative requirements.
+We shall denote it in the following as $I$.
+This measure is usually presented in arbitrary units, which permits us to interpret it two-fold.
+One, directly from its definition, as the \emph{differential divergence} of the vectorial electric current density:
+\begin{equation}
+  I:=\nabla \cdot J
+\end{equation}
+Due to conservation of charges, this is the same as as the temporal variation of density of charge:
+\begin{equation}
+  I=\frac{\partial \rho}{\partial t}
+\end{equation}
+Current Source Density (CSD) as a mathematical density (a quasi-probability density function)  allows  to rigorously apply the concepts of vector averaging.
+This use has been overlooked.
+This might have occurred because  samplings at lesser spatial resolution did not permit estimations of this quantity with enough precision to make reliable spatial implications, even with the use of sophisticated CSD estimation methods.
+In our case, we want to use the density as a mathematician would: a function which assigns weights to points in the space of interest, and permits us to carry linear operations over them.
+The space here is the sampling points of the electrodes at a given instant in time, and the weight, the CSD, is amount interchange of charged ions from the inside to the outside of cellular membranes, per unit space.
+This is an indicator of activity in the neurons, therefore the CSD is also, indirectly, a \emph{``density of local activity''}.
 
-Apparently, CSD would not fit the requirements of non-negativity for its use as a mathematical density.
-CSD sepparates the measured activity into three distinct sets, namely, the set of all sources, the set of all sinks and the zero set.
-The sepparation is not quantitative but qualitative: even when the signs of electric charge are arbitrary, the distintion between them is not.
-Therefore, CSD represents on each of the two sets (sinks and sources) exactly what we need, a unit of concentration per unit space, in this case, per area, and the sign can be ignored.
-The separation by the zero value set of the CSD is in contrast to LFP or EEG measurements, where the zero set has no precise meaning as it only represents a practical ground value.
-This is aknowledged in most of the literature, where only the scales of the LFP recording are shown, without reference to the sign of the values.
-Here the zero set has a precise meaning as instantaneous lack of activity and sepparation between sinks and sources.
-This shall be exploited when we explore a further refinement of the analysis.
+Apparently, CSD would not fit the requirements of non-negativity for its use as a mathematical density, but this turns out to be not a problem.
+CSD separates the measured activity into three distinct sets, namely, the set of all sources, the set of all sinks and the zero set.
+The separation is not quantitative but qualitative: even when the signs of electric charge are arbitrary, the distinction between them is not.
+Therefore, CSD represents on each of the two active sets (sinks and sources) exactly what we need, a unit of concentration per unit space, in this case, per area, and the sign can be ignored, as long as we perform the operations in the sets separately.
+Also, the separation by the zero value set of the CSD is in contrast to the zero value of the LFP, where it has no precise meaning as it only represents a practical ground value.
+This is acknowledged in most of the literature, where only the scales of the LFP recording are shown, without reference to the sign of the values.
+Here the so called zero set (all the points in our space which have a recorded CSD of zero)  has a precise meaning: it is an instantaneous lack of activity and a border between sinks and sources.
+This shall be exploited further in the section below.
 
-Once the sepparation has been made, we would have two densities, or more preciselly, two different density functions defined over two sepparate sets.
-Trying to use this for obtaining vector averages would still be to crude.
+Once the separation into sinks and sources has been made, we would have two densities, or more precisely, two different density functions defined over two separate sets.
+Trying to use this for obtaining vector averages would still be crude and inexact.
 If these sets spread out over non-convex geometries, the vector averages could lie outside the sets, having little or none physiological interpretation.
-Therefore we decided to introduce another concept from geometricall set theory in order to make the analysis more local in character.
+Therefore we decided to introduce another concept from geometrical sets in order to make the analysis more local in character.
 
@@
 A first visual inspection of the CSD color map at a given instant in time shows that both sources and sinks appear in rather large and significant patches.
-These appear to correspond to the expected double and triple poles of firing neurons in structured tissue \cite{} and inspired us to separate the sink and source  sets into their \emph{disjoint components}.
-This concept is used in mathematics to denote subsets of a given geometric set that do not touch each other.
+These appear to correspond to the expected double and triple poles of firing neurons in structured tissue \cite{Buzsaki2012} and inspired us to separate the sink and source  sets into their \emph{disjoint components}.
+This concept is used in mathematics to denote subsets of a given geometric set that do not touch each other or share no elements in common \cite{Halmos}.
 Their rough correspondence to poles of firing neurons would be most useful in tracking the successive locus of the activity.
@@
 Said measurements produce a function over a discrete point set, but we tend to think of them as continuous functions.
-Also, noise between patches corresponding to a single set could make a ``bridge'' between components, making identification of poles inexact.
+Noise between patches corresponding to a single set often makes a ``bridges'' between components, making identification of poles inexact.
 In order to avoid this we create a ``thick zero'' set of all the recording points whose CSD values are inside a small error interval around zero.
-This separates even more the components of the set, represents our ignorance of precise borders and helps us to simplify the following analysis.
+This separates more sharply the components of the set, represents our ignorance of precise borders and helps us to simplify the following analysis.
 
@@
 If each one of these patches could be interpreted as an active unit, we may asign to it a ``center of activity'' in order to follow their displacements.
-A Center of Mass for each nstantaneous disjoint sink/source would provide us a very sensitive measurement of ``active locus''.
+A Center of Mass for each instantaneous disjoint sink/source would provide us a very sensitive measurement of ``active locus''.
 We can then calculate at each time frame the CM for each disjoint component before tracking displacements of activity.
@@
 An instantaneous snapshot of such information would depict the centers of activity of an ordered structure of neurons and their relative intensities.
-But it is in the successive depiction of a series of such snapshots that this analysis can show power.
+But it is in the successive depiction of a series of such snapshots that this analysis can show interpretative power.
 The appearance , displacement, and disappearance of such centers could reveal very subtle details of the dynamics of neural activity.
-Very small, fast displacements of the Centers of mass, may be occur due to slight asynchronism between neurons which could belong to the same active group.
-Larger and  slower than a few firing periods displacements could indicate physiological connections between various groups.
+Very small, fast displacements of the Centers of mass, may be occur due to slight asynchronism between neurons which belong to the same active group.
+Larger and  longer than a few firing periods displacements could indicate physiological connections between different groups.
 
-The next step in our analysis would then a procedure to systematically associate to each CM at time $t$ a ``successor'' at time $t+\Delta t$, where $\Delta t$ is the sampling time step.
+The next step in our analysis is a procedure to systematically associate to each CM at time $t$ a ``successor'' at time $t+\Delta t$, where $\Delta t$ is the sampling time step.
 We have used a pseudo-distance function which takes into account the geometrical distance from the putativa succesors for a CM, and also the intensity.
@@
 We use the distance defined in eq.
-\ref{pseudodist}, a sort of Euclidian-like distance.
+\ref{pseudodist}, a pseudo-Euclidean distance.
 \item ``Connect the dots'', that is, apply succesively the step above for each CM, until it stops having a succesor or its intensity falls below the error criteria.
 Then plot the succesive CMs as one trayectory.
-  \item Plot all the trayectories and sort visually those that may have physiological interpretation from those who may artifacts of the method. 
+  \item Plot all the trayectories and sort visually those that may have physiological interpretation from those who may be artifacts of the method. 
 \end{enumerate}
@@
 More sophisticated methods such as iCSDA \cite{Leski2011} and kCSDA \cite{Potworowski2011} prove little advantage in estimating the sources under the ohmic, isotropic and homogeneous assumptions at the scales involved in our data, and turn out to be computer-costly for high density MEAs and dense temporal sampling.
+The complexity order exponents of the algorithms are high, in integrating and inverse matrix solving operations, and thus, become unusable for large enough matrices.
 

### 2016-02-10 ArticuloCSDyCM/Paper03.tex
mas figuras
Figura 1 choncha

@@
 As Micro Electrode Arrays (MEAs) become denser, this representation of the data becomes saturated and the discovery of patterns turns into a difficult task.
-Novel information is also cluttered or obfuscated by the same high density acquisition that allows us to discover it.
-Therefore, representing  the recorded data in ways which can unveil previously unobtainable information.
+Novel information is obfuscated by the same high density acquisition that allows us to discover it.
+Therefore, representing  the recorded data in ways which can unveil previously unobtainable information is a necessity.
 Our task in this paper is to present a method for tracing out ``functional connectivity trajectories'' of activity in structured neural tissue, and we shall exemplify the method by application to the data obtained from recordings of the rats hippocampus.
 
-Succesive  of distinct units on structured tissue would trace a ``trajectory'' of the activity.
+Succesive activation  of distinct units on structured tissue would trace a ``trajectory'' of the activity.
 A tracing of such trajectories would help to indicate causal or correlated ``functional conections'' over the tissue.
-Neighbouring neurons that fire in succesion, integrated subthreshold activity and ephaptic coupling could all be interpreted parts of the paths where the activity gets transmitted.
+Neighbouring neurons that fire in succesion, integrated subthreshold activity and ephaptic coupling could all be interpreted as paths for transmission of activity.
 Nevertheless, how to define and acquire the corresponging traces is an open question.
 
-In previous works, some form of vector averaging has been proposed and shown to reveal some trajectrories traced out by the activity.
+In previous works some form of vector averaging has been proposed and shown to reveal some trajectrories traced out by the activity.
 We shall call such methods Center of Mass Analyses (CMA).
@@
 This representation not only provides a mathematical fit measure for the weights in the vector averaging process, but helps to separate the data into disjoint components.
-Based on this, a further refinement of the technique, given by the polar structure of firing neurons, allowed us to obtain accurate, well defined trajectories of the activity in the studied tissue.
+Based on this, a further refinement of the technique, given by the polar structure of firing neurons, allowed us to obtain accurate, well defined trajectories of the activity in non convex structured tissues.
 
@@
 We have tested then the method on slices of rat's hippocampus, concentrating in the CA region.
-We exemplify the usage of the method with data that has been obtained by an experiment \textbf{bla bla bla describr experimentos}.
+We exemplify the usage of the method with data that has been obtained by an experiment with 4AP ...\textbf{bla bla bla describr experimentos}.
 
-\section{CSD, CM and Disjoint Components}
+\section{Materials and Methods}
+
+\subsection{CSD for high density MEAS}
+
+The concepts presented here require first obtaining the CSD from the recorded LFP.
+Any method would present the same utility.
+Thanks to the density and scales of the BioCAM 4096, in our case a ``classical'' numerical derivation of the CSD is the most adecuate route to take.
+A numerical Laplacian Operator which reduces the cross effect of rectangular grids is the convolution of the data with the following matrix \cite{Lindberg90}:
+\begin{equation}
+\nabla^2_{1/3}=(2/3)
+\begin{pmatrix}
+  0 & 1 & 0 \\
+  1 & -4 & 1 \\
+  0 & 1 & 0
+\end{pmatrix}
++1/3
+\begin{pmatrix}
+  0.5 & 0 & 0.5 \\
+  0 & -2 & 0 \\
+  0.5 & 0 & 0.5
+\end{pmatrix}  
+\end{equation}
+This sort of operators are extremely sensitive to edges.
+The noise of the data can be perceived in the spatial domain as rapidly varying edges.
+A Gaussian spatial denoising can be performed before or after the Laplacian operator to the data in order to reduce excessive borders.
+Our Gaussian blur filter has a $\sigma$ value of three electrodes or $126 \mu m$.
+Intuitively this means that it smooths over each structure smaller than the soma of a typical pyramidal cell.
+This is  in accordance to our mean field approach, where we expect smooth differentiable fields over the scale of the neurons, but not in finer scales (for details see \cite{Bedard11}, further work on this subject will be presented elsewhere \cite{IsabelYo}).
+
+On previous works it has been argued that numerical difference operators sacrifice all data on the borders of the array.
+For electrode arrays of lesser density, this could posse a problem, but in our case we only renounce to 27 out of 4095 channels (one is grounded).
+More sophisticated methods such as iCSDA \cite{Leski2011} and kCSDA \cite{Potworowski2011} prove little advantage in estimating the sources under the ohmic, isotropic and homogeneous assumptions at the scales involved in our data, and turn out to be computer-costly for high density MEAs and dense temporal sampling.
+The complexity order exponents of the algorithms are high, in integrating and inverse matrix solving operations, and thus, become unusable for large enough matrices.
 
@@
 This might have occurred because  samplings at lesser spatial resolution did not permit estimations of this quantity with enough precision to make reliable spatial implications, even with the use of sophisticated CSD estimation methods.
-In our case, we want to use the density as a mathematician would: a function which assigns weights to points in the space of interest, and permits us to carry linear operations over them.
+In our case, we want to use the density as a mathematician would: a function which assigns weights to points in the space of interest and permits us to carry linear operations over them.
 The space here is the sampling points of the electrodes at a given instant in time, and the weight, the CSD, is amount interchange of charged ions from the inside to the outside of cellular membranes, per unit space.
@@
 This is acknowledged in most of the literature, where only the scales of the LFP recording are shown, without reference to the sign of the values.
-Here the so called zero set (all the points in our space which have a recorded CSD of zero)  has a precise meaning: it is an instantaneous lack of activity and a border between sinks and sources.
+In the CSD representation,  the so called zero set (all the points in our space which have a recorded CSD of zero)  has a precise meaning: it is an instantaneous lack of activity and a border between sinks and sources.
 This shall be exploited further in the section below.
@@
 Trying to use this for obtaining vector averages would still be crude and inexact.
-If these sets spread out over non-convex geometries, the vector averages could lie outside the sets, having little or none physiological interpretation.
+If these sets spread out over large non-convex geometries, the vector averages could lie outside the sets, having little or none physiological interpretation.
 Therefore we decided to introduce another concept from geometrical sets in order to make the analysis more local in character.
@@
 Due to the finite precision of the measurement, the validity of the mean field approach, and the noise in the measurements, the sinks and sources sets can make contact with each other.
-Said measurements produce a function over a discrete point set, but we tend to think of them as continuous functions.
+Said measurements produce a function over a discrete point set.
 Noise between patches corresponding to a single set often makes a ``bridges'' between components, making identification of poles inexact.
@@
 
- \section{The Procedure}
-\subsection{Recipe in a nutshell}
+ \subsection{The Procedure}
 
@@
 We recommend simpler, difference based approaches.
-Our approach is detailed below.
 \item Separate the CSD data into three sets: sources, sinks, and border or indiscernible (below error) data.
@@
 
-\subsection{CSD for high density MEAS}
-
-The concepts presented here require first obtaining the CSD from the recorded LFP.
-Any method would present the same utility.
-Thanks to the density and scales of the BioCAM 4096, in our case a ``classical'' numerical derivation of the CSD is the most adecuate route to take.
-A numerical Laplacian Operator which reduces the cross effect of rectangular grids is the convolution of the data with the following matrix \cite{Lindberg90}:
-\begin{equation}
-\nabla^2_{1/3}=(2/3)
-\begin{pmatrix}
-  0 & 1 & 0 \\
-  1 & -4 & 1 \\
-  0 & 1 & 0
-\end{pmatrix}
-+1/3
-\begin{pmatrix}
-  0.5 & 0 & 0.5 \\
-  0 & -2 & 0 \\
-  0.5 & 0 & 0.5
-\end{pmatrix}  
-\end{equation}
-This sort of operators are extremely sensitive to edges.
-The noise of the data can be perceived in the spatial domain as rapidly varying edges.
-A Gaussian spatial denoising can be performed before or after the Laplacian operator to the data in order to reduce excessive borders.
-Our Gaussian blur filter has a $\sigma$ value of three electrodes or $126 \mu m$.
-Intuitively this means that it smooths over each structure smaller than the soma of a typical pyramidal cell.
-This is  in accordance to our mean field approach, where we expect smooth differentiable fields over the scale of the neurons, but not in finer scales (for details see \cite{Bedard11}, further work on this subject will be presented elsewhere \cite{IsabelYo}).
-
-On previous works it has been argued that numerical difference operators sacrifice all data on the borders of the array.
-For electrode arrays of lesser density, this could posse a problem, but in our case we only renounce to 27 out of 4095 channels (one is grounded).
-More sophisticated methods such as iCSDA \cite{Leski2011} and kCSDA \cite{Potworowski2011} prove little advantage in estimating the sources under the ohmic, isotropic and homogeneous assumptions at the scales involved in our data, and turn out to be computer-costly for high density MEAs and dense temporal sampling.
-The complexity order exponents of the algorithms are high, in integrating and inverse matrix solving operations, and thus, become unusable for large enough matrices.
-
-\section{Experimental examples}
+\section{Results}
 
@@
 So the most violent epileptic burst provided us with the data necessary to prove the robustness of the method.
+a
 
-\section{Application to epileptic-like activity in the hippocampus}
+\subsection{Application to epileptic-like activity in the hippocampus}
 Chronologically speaking, we performed first the analysis to the most complicated data (numerically and qualitatively), which where a series of epileptic burst that occurred after 15 minutes of application of 4AP to living slices of hippocampus.
@@
 The necessity of having good error value for the ``thick zero'' set was more apparent here than anywhere else.
-Also the posibilty to create new structures of numerical data was seen here.
+Also the posibilty to create new structures of numerical data was first seen here.
 But also the most strickig features of the propagation of activity, coded now as trajectories of the CM, could be seen here, as the epileptic burst covered the whole CA structure.
+A sample of the raw data in $\mu V$ can be seen in the figure \textbf{Tururu}.
+Three key moments in the epileptic burst have been selected, one before the onset of the attack, one at the peak and one at the middle of the subsequent waves.
+Traces from selected electrodes along the line labeled A-F are presented in the subfigure D.
 
-\section{Application to evoked activity in idem.}
+In figure \textbf{Tururu2} we present the CSD for the same selected moments.
+
+In the figure \textbf{Tururu3} we present our method using the peak moment as an example.
+
+\subsection{Application to evoked activity in idem.}
 
@@
 
-\section{Application to spontaneous activity in idem.}
+\subsection{Application to spontaneous activity in idem.}
 

### 2017-09-25 ArticuloCSDyCM/Paper03.tex
avances paper

@@
 \usepackage{courier}
-%\usepackage{lipsum}% just to generate filler text
+
 \usepackage{setspace}
@@
 
-\author{Laboratorio 19 Crew}
+\DeclareMathOperator{\arcsinh}{arcsinh}
+\newcommand{\rd}{\mathrm{d}}
+
+\author{WPK Zapfe and the Laboratorio 19 Crew}
 \title{CSD, Disjoint components and trayectory detection}
@@
 Therefore, representing  the recorded data in ways which can unveil previously unobtainable information is a necessity.
-Our task in this paper is to present a method for tracing out ``functional connectivity trajectories'' of activity in structured neural tissue, and we shall exemplify the method by application to the data obtained from recordings of the rats hippocampus.
+Various techniques for sepparation of functional generators are avaible now, most come from the application of blind source sepparation analysis, such as ICA.
+Our task in this paper is to present a method for tracing out ``functional connectivity'', which would apport independent information that is complementary to those analysis.
+Our  analysis would be applicable mostly to activity in structured neural tissue.
+We shall exemplify the method by application to the data obtained from recordings of the rats hippocampus.
 
-Succesive activation  of distinct units on structured tissue would trace a ``trajectory'' of the activity.
+Succesive activation  of distinct nearby units on structured tissue would trace a ``trajectory'' of the activity.
 A tracing of such trajectories would help to indicate causal or correlated ``functional conections'' over the tissue.
-Neighbouring neurons that fire in succesion, integrated subthreshold activity and ephaptic coupling could all be interpreted as paths for transmission of activity.
+Neighbouring neurons that fire in succesion, integrated subthreshold activity and ephaptic coupling could all be represented as paths for transmission of activity.
 Nevertheless, how to define and acquire the corresponging traces is an open question.
@@
 The \emph{wheights} that have been chosen for obtaining the average have been selected on heuristic grounds.
-In the works of Chao \emph{et al.} \cite{Chao05, Chao07}, the density of spikes in a certain interval was used, a sort of ``density of action'', while Manjarrez \emph{et al.} \cite{Manjarrez07, Manjarrez09} used the positive half of the LFP recording as their weight.
+In the works of Chao \emph{et al.} \cite{Chao05, Chao07}, the density of spikes in a certain time interval was used, a sort of ``density of action'', while Manjarrez \emph{et al.} \cite{Manjarrez07, Manjarrez09} used the positive half of the LFP recording as their weight.
 Here we advance on such methods by incorporating mathematically sound definitions and concepts which allow us to rigorously separate spatially distinct traces.
@@
 
-The application of the method requires electro-physiological data acquired through bi-dimensional MEAs, although the extension to three dimensional arrays is possible in principle.
+The application of the method requires electro-physiological data acquired through bidimensional MEAs, although the extension to three dimensional arrays is possible in principle.
 The MEA should have dense enough spatial and temporal samplings, so as to obtain good estimates of the CSD and their changes in time.
@@
 
-\section{Materials and Methods}
+\section{Method and Calculations}
 
@@
 The concepts presented here require first obtaining the CSD from the recorded LFP.
-Any method would present the same utility.
-Thanks to the density and scales of the BioCAM 4096, in our case a ``classical'' numerical derivation of the CSD is the most adecuate route to take.
+Any correct method would present the same utility.
+We present to methods and show the robustness of the result for the Center of Mass tracking.
+The first method is a classical finite diference operator, while the second is the kCSDA presented in \cite{Potworowski2011}, which is an inverse problem approach.
+
+Thanks to the density and scales of the BioCAM 4096, a numerical finite difference method
+can give a good aproximation to the CSD, if the number of failing channels is low.
+For experiments that were made without electrical stimulation this was the case, where the order of saturated or otherwise failing electrodes was around 10.
 A  numerical Laplacian Operator which reduces the cross effect of rectangular grids is the convolution of the data with the following matrix \cite{Lindberg90}:
@@
 Intuitively this means that it smooths over each structure smaller than the soma of a typical pyramidal cell.
-This is  in accordance to our mean field approach, where we expect smooth differentiable fields over the scale of the neurons, but not in finer scales (for details see \cite{Bedard11}, further work on this subject will be presented elsewhere \cite{IsabelYo}).
+This is  in accordance to our mean field approach, where we expect smooth differentiable fields over the scale of the neurons, but not in finer scales (for details see \cite{Bedard11}.
 
@@
 For electrode arrays of lesser density, this could posse a problem, but in our case we only renounce to 27 out of 4095 channels (one is grounded).
-More sophisticated methods such as iCSDA \cite{Leski2011} and kCSDA \cite{Potworowski2011} prove little advantage in estimating the sources under the ohmic, isotropic and homogeneous assumptions at the scales involved in our data, and turn out to be computer-costly for high density MEAs and dense temporal sampling.
+If very few electrodes fail, more sophisticated methods such as iCSDA \cite{Leski2011} and kCSDA \cite{Potworowski2011} prove little advantage in estimating the sources under the ohmic, isotropic and homogeneous assumptions.
+In case that around 30 or more electrodes
+fail, specially if they are on the region of intereset, the use of indirect methods
+becomes necessary. The Kernel CSDA is then the tool to use. This methods relies on the
+assumption that a finite family of reasonable functions can span (with acceptable accuracy) the possible CSD at each instant in time. The idea behind the method is simple:
+presenting one model function of this family and obtaining the LFP that such a function
+would generate, we calculate the coefficients for the entire family by proyecting the
+experimental data over this ideal LFP. A general presentation is in the reference, here
+we shall put only the implementation that we used.
+
+We shall borrow the notation in \ref{Potworowski2011}. Our model function for the CSD is a hard disk function, given by:
+\begin{equation}
+  \tilde{b}_j(x,y)=\begin{cases}
+  1, \text{ for } (x-x_j)^2+(y-y_j)^2 \leq R^2 \\
+  0 \text{ otherwise.}
+  \end{cases}
+\end{equation}
+The $R$ parameter is the effective radius of the charge density, which for our
+calculations was taken as half the interelectrode distance, that is $R=21 \mu m$.
+Each pair $(x_j, y_j)$ is the center of one such model function, which we take as
+the position of each  functioning electrode in an adecuate coordinate system.
+The LFP generated by this function would then be:
+\begin{equation}
+  \tilde{b}_j(x,y)=\frac{1}{2\pi\sigma}
+  \iint \! \rd x \rd y \arcsinh \biggr(
+  \frac{2 h } {\sqrt{(x-x_j)^2+(y-y_j)^2}}
+  \biggl)
+  \tilde{b}_j(x,y)
+\end{equation}.
+Then the kCSDA consist of proyecting the experimental data (LFP), frame by frame,
+over these model functions and obtaining the  coefficients that would spawn an idealized
+LFP approximation. These coefficients are the same that would produce the
+corresponing CSD. The first operator is the projection operator $K$ of the experimental
+LFP (from now on $V(x,y)$) on the space of the idealized LFP (which we call $V^* (x,y)$).
+\begin{equation}
+  K_{jk}=\sum_{l=1}^{M}b_l(x_j,y_j)b_l(x_k,y_k).
+\end{equation}
+The sum is done over the set of $M$ electrodes that we want to use.
+The other operator is the one that maps the space of $V^*(x,y)$ into the idealized
+CSD functions, now called $C^*(x,y)$. 
+\begin{equation}
+  \tilde{K}_{jk}=\sum_{l=1}^Mb_l(x_j,y_j)\tilde{b}_l(x_k,y_k).
+\end{equation}
+Then $C^*(x,y)$ can be obtained on the region of interest by:
+\begin{equation}
+  C^*(x_j,y_j)=\sum_{k,l=1}^M \tilde{K}_{kj} K^{-1}_{ml} V(x_l,y_l)
+\end{equation}
+Notice how the method incorporates a denoizing mechanism, so
+no smoothing or other denoizing is needed.
+
+These methods turn out to be computer-costly for high density MEAs and dense temporal sampling, but paralelization, specially on
+GPUs, solves this issue.
+
 The complexity order exponents of the algorithms are high, in integrating and inverse matrix solving operations, and thus, become unusable for large enough matrices.

### 2017-09-25 ArticuloCSDyCM/Paper03.tex
avances paper

@@
   \biggl)
-  \tilde{b}_j(x,y)
-\end{equation}.
+  \tilde{b}_j(x,y),
+\end{equation}
+  with $h$ being the effective height of the function support, in our case,
+  the half width of the hippocampus slice ($h=150 \mu m$).
+
 Then the kCSDA consist of proyecting the experimental data (LFP), frame by frame,
@@
 Notice how the method incorporates a denoizing mechanism, so
-no smoothing or other denoizing is needed.
+no smoothing procedures are used.
+The great advantage of this method is that is possible to obtain a reasonable
+approximation to CSD even when the prescene of unusable channels is near or inside
+the region of interest. The difference method, on the other hand, affects all surrounding
+channels to a failing channel, up to 8 of them, so it becomes unusuable if enough
+saturated electrodes are over the r.o.i.
 
-These methods turn out to be computer-costly for high density MEAs and dense temporal sampling, but paralelization, specially on
+It is important to mention that our data has noice up to 0.1 of the magnitude of the
+signal, which means that regularization of the matrix $K$ is unavoidable (see section 4.1 in \cite{Potworowski2011} and references therein).
+We used a regularization parameter $\lambda=0.001 |e_{max}|$, where $e_{max}$ is the principal eigenvalue of $K$.
+
+The obtention of $K$ and $\tilde{K}$ turn out to be computer-costly for high density MEAs and dense temporal sampling, but paralelization, specially on
 GPUs, solves this issue.
 
-The complexity order exponents of the algorithms are high, in integrating and inverse matrix solving operations, and thus, become unusable for large enough matrices.
+We have performed comparition of both methods over experimental data that has
+no failing channels in order to adjust the parameters of the second method. We
+have obtained very precise results in both methods.
 
@@
 The Current Source Density fits the mathematical and interpretative requirements.
-We shall denote it in the following as $I$.
+We shall denote it in the following as $C$ (following \cite{Potworowski2011}).
 This measure is usually presented in arbitrary units, which permits us to interpret it two-fold.
@@
 \begin{equation}
-  I:=\nabla \cdot J
+  C:=\nabla \cdot J
 \end{equation}
@@
 \begin{equation}
-  I=\frac{\partial \rho}{\partial t}
+  C=\frac{\partial \rho}{\partial t}
 \end{equation}
-Current Source Density (CSD) as a mathematical density (a quasi-probability density function)  allows  to rigorously apply the concepts of vector averaging.
+Current Source Density as a mathematical density (a quasi-probability density function)  allows  to rigorously apply the concepts of vector averaging.
 This use has been overlooked.

### 2017-09-26 ArticuloCSDyCM/Paper03.tex
mas paper

@@
 Various techniques for sepparation of functional generators are avaible now, most come from the application of blind source sepparation analysis, such as ICA.
-Our task in this paper is to present a method for tracing out ``functional connectivity'', which would apport independent information that is complementary to those analysis.
-Our  analysis would be applicable mostly to activity in structured neural tissue.
-We shall exemplify the method by application to the data obtained from recordings of the rats hippocampus.
+Our task in this paper is to present a method for tracing out ``functional connectivity'', which would bring independent information that is complementary to those analysis.
+Our  method would be applicable mostly to activity in structured neural tissue.
+We shall exemplify it by application to the data obtained from recordings of the rats hippocampus.
 
@@
 A tracing of such trajectories would help to indicate causal or correlated ``functional conections'' over the tissue.
-Neighbouring neurons that fire in succesion, integrated subthreshold activity and ephaptic coupling could all be represented as paths for transmission of activity.
-Nevertheless, how to define and acquire the corresponging traces is an open question.
+Neighbouring neurons that fire in succesion, integrated subthreshold activity and ephaptic coupling could all be represented as pathways for transmission of activity.
+Nevertheless, how to define and acquire the putative paths is an open question.
 
-In previous works some form of vector averaging has been proposed and shown to reveal some trajectrories traced out by the activity.
+In previous works some form of vector averaging has been proposed and shown to reveal some trajectrories drawn by the activity.
 We shall call such methods Center of Mass Analyses (CMA).
@@
 This representation not only provides a mathematical fit measure for the weights in the vector averaging process, but helps to separate the data into disjoint components.
-Based on this, a further refinement of the technique, given by the polar structure of firing neurons, allowed us to obtain accurate, well defined trajectories of the activity in non convex structured tissues.
+Based on this, a further refinement of the technique, given by the multipolar signature of firing neurons, allowed us to obtain accurate, well defined trajectories of the activity in non convex structured tissues.
 
-The application of the method requires electro-physiological data acquired through bidimensional MEAs, although the extension to three dimensional arrays is possible in principle.
+The application of the method requires electro-physiological data acquired through bidimensional MEAs, although the extension to three dimensional arrays would, in principle, be possible.
 The MEA should have dense enough spatial and temporal samplings, so as to obtain good estimates of the CSD and their changes in time.
@@
 We have tested then the method on slices of rat's hippocampus, concentrating in the CA region.
-We exemplify the usage of the method with data that has been obtained by an experiment with 4AP ...\textbf{bla bla bla describr experimentos}.
+We exemplify the usage of the method with data that has been obtained by an experiment with ...\textbf{bla bla bla describr experimentos}.
 
@@
 Any correct method would present the same utility.
+The first method is a  finite difference operator, while the second is the kCSDA presented in \cite{Potworowski2011}, which is an inverse problem approach.
+Where both methods are usuable, we prefer the former, as is faster and requires less supositions.
 We present to methods and show the robustness of the result for the Center of Mass tracking.
-The first method is a classical finite diference operator, while the second is the kCSDA presented in \cite{Potworowski2011}, which is an inverse problem approach.
 
@@
 \begin{equation}
-\nabla^2_{1/3}=(2/3)
+\nabla^2_{1/3}=2/3Es
 \begin{pmatrix}
@@
 \end{equation}
-This sort of operators are extremely sensitive to edges.
+This sort of operators are extremely sensitive to abrupt changes in the values.
 The noise of the data can be perceived in the spatial domain as rapidly varying edges.
@@
 Intuitively this means that it smooths over each structure smaller than the soma of a typical pyramidal cell.
-This is  in accordance to our mean field approach, where we expect smooth differentiable fields over the scale of the neurons, but not in finer scales (for details see \cite{Bedard11}.
+This is  in accordance to our mean field approach, where we expect smooth differentiable fields over the scale of the neurons, but not in finer scales (for details see \cite{Bedard11}) .
 
@@
 In case that around 30 or more electrodes
-fail, specially if they are on the region of intereset, the use of indirect methods
-becomes necessary. The Kernel CSDA is then the tool to use. This methods relies on the
+fail, specially if they are on the region of intereset, the use of such methods
+becomes necessary. The kCSDA method relies on the
 assumption that a finite family of reasonable functions can span (with acceptable accuracy) the possible CSD at each instant in time. The idea behind the method is simple:
@@
 would generate, we calculate the coefficients for the entire family by proyecting the
-experimental data over this ideal LFP. A general presentation is in the reference, here
-we shall put only the implementation that we used.
+experimental data over this ideal LFP model function family. A general presentation is in the reference, here
+we shall recount only the implementation that we used.
 
@@
 no failing channels in order to adjust the parameters of the second method. We
-have obtained very precise results in both methods.
+have obtained consistent results between methods.
 

### 2017-10-13 ArticuloCSDyCM/Paper03.tex
paper 03 avanzado

@@
 Therefore, representing  the recorded data in ways which can unveil previously unobtainable information is a necessity.
-Various techniques for sepparation of functional generators are avaible now, most come from the application of blind source sepparation analysis, such as ICA.
+Various techniques for separation of functional generators are available now, most come from the application of blind source separation analysis, such as ICA.
 Our task in this paper is to present a method for tracing out ``functional connectivity'', which would bring independent information that is complementary to those analysis.
@@
 Our weight is actually the natural density function provided by the data: the Current Source Density.
-This representation not only provides a mathematical fit measure for the weights in the vector averaging process, but helps to separate the data into disjoint components.
-Based on this, a further refinement of the technique, given by the multipolar signature of firing neurons, allowed us to obtain accurate, well defined trajectories of the activity in non convex structured tissues.
+This representation not only provides a mathematical rigorous  measure for the weights in the vector averaging process, but helps to separate the data into disjoint components.
+This allows us to track separate process in the same region of interest.
 
-The application of the method requires electro-physiological data acquired through bidimensional MEAs, although the extension to three dimensional arrays would, in principle, be possible.
+The application of the method requires electro-physiological data acquired through bi-dimensional MEAs, although the extension to three dimensional arrays would, in principle, be possible.
 The MEA should have dense enough spatial and temporal samplings, so as to obtain good estimates of the CSD and their changes in time.
@@
 Any correct method would present the same utility.
-The first method is a  finite difference operator, while the second is the kCSDA presented in \cite{Potworowski2011}, which is an inverse problem approach.
-Where both methods are usuable, we prefer the former, as is faster and requires less supositions.
+The first method is a  finite difference operator, while the second is the kCSDA presented by Potworowski \emph{et all} \cite{Potworowski2011}, which is an inverse problem approach.
+Where both methods are usable, we prefer the former, as is faster and requires less assumptions.
 We present to methods and show the robustness of the result for the Center of Mass tracking.
@@
 Thanks to the density and scales of the BioCAM 4096, a numerical finite difference method
-can give a good aproximation to the CSD, if the number of failing channels is low.
+can give a good approximation to the CSD, if the number of failing channels is low.
 For experiments that were made without electrical stimulation this was the case, where the order of saturated or otherwise failing electrodes was around 10.
@@
 \begin{equation}
-\nabla^2_{1/3}=2/3Es
+\nabla^2_{1/3}=2/3
 \begin{pmatrix}
@@
 On previous works it has been argued that numerical difference operators sacrifice all data on the borders of the array.
-For electrode arrays of lesser density, this could posse a problem, but in our case we only renounce to 27 out of 4095 channels (one is grounded).
+For electrode arrays of lesser density, this could posse a problem, but in our case we only renounce to 27 out of 4095 channels (one is grounded), and those are outside the region of interest.
 If very few electrodes fail, more sophisticated methods such as iCSDA \cite{Leski2011} and kCSDA \cite{Potworowski2011} prove little advantage in estimating the sources under the ohmic, isotropic and homogeneous assumptions.
 In case that around 30 or more electrodes
-fail, specially if they are on the region of intereset, the use of such methods
+fail, specially if they are on the region of interest, the use of such methods
 becomes necessary. The kCSDA method relies on the
@@
 presenting one model function of this family and obtaining the LFP that such a function
-would generate, we calculate the coefficients for the entire family by proyecting the
+would generate, we calculate the coefficients for the entire family by projecting the
 experimental data over this ideal LFP model function family. A general presentation is in the reference, here
@@
 
-We shall borrow the notation in \ref{Potworowski2011}. Our model function for the CSD is a hard disk function, given by:
+We shall borrow the notation in \cite{Potworowski2011}. Our model function for the CSD is a hard disk function, given by:
 \begin{equation}
@@
 The $R$ parameter is the effective radius of the charge density, which for our
-calculations was taken as half the interelectrode distance, that is $R=21 \mu m$.
+calculations was taken as half the inter-electrode distance, that is $R=21 \mu m$.
 Each pair $(x_j, y_j)$ is the center of one such model function, which we take as
-the position of each  functioning electrode in an adecuate coordinate system.
+the position of each  functioning electrode in an adequate coordinate system.
 The LFP generated by this function would then be:
@@
 
-Then the kCSDA consist of proyecting the experimental data (LFP), frame by frame,
+Then the kCSDA consist of projecting the experimental data (LFP), frame by frame,
 over these model functions and obtaining the  coefficients that would spawn an idealized
 LFP approximation. These coefficients are the same that would produce the
-corresponing CSD. The first operator is the projection operator $K$ of the experimental
+corresponding CSD. The first operator is the projection operator $K$ of the experimental
 LFP (from now on $V(x,y)$) on the space of the idealized LFP (which we call $V^* (x,y)$).
@@
 \end{equation}
-Notice how the method incorporates a denoizing mechanism, so
+Notice how the method incorporates a denoising mechanism, so
 no smoothing procedures are used.
 The great advantage of this method is that is possible to obtain a reasonable
-approximation to CSD even when the prescene of unusable channels is near or inside
+approximation to CSD even when the presence of unusable channels is near or inside
 the region of interest. The difference method, on the other hand, affects all surrounding
-channels to a failing channel, up to 8 of them, so it becomes unusuable if enough
+channels to a failing channel, up to 8 of them, so it becomes unusable if enough
 saturated electrodes are over the r.o.i.
 
-It is important to mention that our data has noice up to 0.1 of the magnitude of the
+It is important to mention that our data has noise up to 0.1 of the magnitude of the
 signal, which means that regularization of the matrix $K$ is unavoidable (see section 4.1 in \cite{Potworowski2011} and references therein).
@@
 
-The obtention of $K$ and $\tilde{K}$ turn out to be computer-costly for high density MEAs and dense temporal sampling, but paralelization, specially on
+The attainment of $K$ and $\tilde{K}$ turn out to be computer-costly for high density MEAs and dense temporal sampling, but paraletization, specially on
 GPUs, solves this issue.
 
-We have performed comparition of both methods over experimental data that has
+We have performed comparison of both methods over experimental data that has
 no failing channels in order to adjust the parameters of the second method. We
-have obtained consistent results between methods.
+have obtained consistent results between them.
 
@@
 Therefore, CSD represents on each of the two active sets (sinks and sources) exactly what we need, a unit of concentration per unit space, in this case, per area, and the sign can be ignored, as long as we perform the operations in the sets separately.
-Also, the separation by the zero value set of the CSD is in contrast to the zero value of the LFP, where it has no precise meaning as it only represents a practical ground value.
+Also, the separation by the zero value set of the CSD is in contrast to the zero value of the LFP, where it has no precise meaning as it only represents a practical reference value.
 This is acknowledged in most of the literature, where only the scales of the LFP recording are shown, without reference to the sign of the values.
@@
 To clarify the concept and how we are we using it an example is in order.
-In figure \ref{ejemplodisjuntos} we show geometrical sets enclosed by simple curves as color patches, where every color indicates a Set.
-The set A consists of a single component, that means that one can make  path between any two points in the set that consist entirely of points of the same set.
-In contrast, the set B is made of two disjoint components, labeled B1 and B2.
+In figure \ref{Disjuntos} A we show geometrical sets enclosed by simple curves as color patches, where every color indicates a Set.
+The set $A$ consists of a single component, that means that one can make  path between any two points in the set that consist entirely of points of the same set.
+In contrast, the set $B$ is made of two disjoint components, labeled $B1$ and $B2$.
 If one chooses a point in each of the components, one cannot draw a path joining them without leaving the set.
 But each one of those components is connected, meaning that if we choose the pair of points in one component, then again we can trace a path joining them in the set.
-So the set B consists of two disjoint components, but each component can be regarded as a connected set on their own.
-Likewise the set C is made of three disjoint components.
+So the set $B$ consists of two disjoint components, but each component can be regarded as a connected set on their own.
+Likewise the set $C$ is made of three disjoint components.
+In the subfigre \ref{Disjuntos} Aiii we show the center of mass of the sets.
+The set $A$ is concave, so, it can have a center of mass outside the set.
+The
+set $B$ has also a center of mass outside the set. It has two disjoint components that are aproximatelly
+equall and separated by a long distance (this is not an implication: a set can be disjoint and have
+a CM inside the set). For the set $C$ our aproach has been used: we treat every disjoint component of the
+set as an independent entity and calculate the center of mass for each one.
+Notice that this doesn't preclude the posibility of having a concave subset with a CM that lies outside.
+But, as we shall see,
+as the electrode resolution is finite, the posibility of having concave shapes is unlikely, and they
+tend to have a CM near their borders. In the figure \ref{Disjuntos} B we show this with experimental
+data (we are using evoked activity data, as explained in section \ref{sec:evocada}). In graph Bi we
+sepparate the CSD data into the sink and sources set. We take sinks as example. In the grabh Bii
+we label each disjoint component, obtained by a single pass algorhythm. There are 25 components, and
+the biggest two are not strictly convex but they have tolerable deviations from it. In the
+subfigure Biii we indicate the CM of only those components that are above certain threshold of
+integrated intensity (the dot is proportional in area to such intensity, there are five CM to be
+seen). Each CM gets listed as a probable step in a trajectory.
+
+\begin{figure}
+  \includegraphics[width=0.9\textwidth]{Figuras/Disjuntos01.pdf}
+  \caption{ Explanation of the geometrical concepts used on our analysis. In subfigure A we show
+    with an artificial example the concepts of disjointness and convexivity. The subfigure Ai shows
+    three sets, $A$, $B$, and $C$. $A$ fails to be convex because it is concave, and $C$ fails to be
+    be convex because it is made of two disjoint components. In each of them is possible to draw a line
+    connecting two points inside the set that is not entirely inside the set.
+In $B$ that is not possible, the set $B$ is convex in a mathematical sense.
+In subfigure Aii we made some modification to whos that
+    each set can be made of many disjoint components, and each one of them can be convex or not.
+In Aiii we show the center of mass for $A$ and $B$ taken as a wholle, while for $C$ we break the set into each of the disjoint components and calculate each CM separately.
+There is no guarantee that each disjoint component would be convex, but as the resolution of the data is finite, the smaller components are more likely to
+    be convex. In figure B we shoe everything with a frame of experimental data. In Bi we sepparate the
+    data into three sets, Sources, Sinks, and Zeros. We discard the Zeros and keep the interesting ones.
+    In Figure Bii we take the Sink Set and separate it into disjoint components, that are labeled from one to 25 in this particular example.
+From this subsets we calculate the total mass of each one and only
+    considerate for the next step those who are above certain threshold of integrated intensity. The
+    next step is shown in Biii, where we plot the Center of Mass for each of such subsets. Deviations
+    from convexity seem to be unimportant, as most of the mass is concentrated in the inner electrodes
+    of each set.
+  }
+  \label{disjuntos}
+\end{figure}
 
@@
 The appearance , displacement, and disappearance of such centers could reveal very subtle details of the dynamics of neural activity.
-Very small, fast displacements of the Centers of mass, may be occur due to slight asynchronism between neurons which belong to the same active group.
+Very small, fast displacements of the Centers of mass, may be occur due to slight asynchronous between neurons which belong to the same active group.
 Larger and  longer than a few firing periods displacements could indicate physiological connections between different groups.
@@
 The next step in our analysis is a procedure to systematically associate to each CM at time $t$ a ``successor'' at time $t+\Delta t$, where $\Delta t$ is the sampling time step.
-We have used a pseudo-distance function which takes into account the geometrical distance from the putativa succesors for a CM, and also the intensity.
-For two CM inside our measurement space we define our pseudo-distance as follows.
-Let $q_1=(x_1, y_1)$ be the coordinates of the first point and $q_2=(x_2,y_2)$, the corresponding coordinates of the second, and let $I_1, I_2$ be their corresponding integrated intensities.
-Then the ``distance'' that separates them shall be
-\begin{equation}\label{pseudodist}
-d(q_1,q_2):=\sqrt{(x_1-x_2)^2+(y_1+y_2)^2+C(I_1-I_2)},
+We used an euclididan distance.
+Let $q_1=(x_1, y_1)$ be the coordinates of the first point and $q_2=(x_2,y_2)$, the corresponding coordinates of the second.
+Then the distance is:
+d(q_1,q_2):=\sqrt{(x_1-x_2)^2+(y_1+y_2)^2}.
 \end{equation}
-where $C$ is a conveniently chosen constant, so as to give a similar importance to the intensities as to the positions.
 The integrated intensity of a CM is simply the denominator in the eq. \ref{cmparadisj}, that is, the total ``mass'' obtained by integrating the CSD in a specific disjoint component:
 \begin{equation}
-I_k(t)=\sum_j I (q_{j,k}, t)
+  I_k(t)=\sum_j I (q_{j,k}, t).
 \end{equation}
@@
 By continuously applying such procedure, we can begin to trace trajectories of the CMs, as they appear, wander, and fade.
+
+When a specific trajectory cannot find a successor point that is above the
+threshold intensity, we assume that it is finished, so we do not concatenate
+points that are separated in time by more than $\Delta t$. In order to paste different
+trajectories, more rigorous analysis of their origin and interpretation would be needed.
 
@@
  \item Acquire high density electro-physiological data.
-The data must be two or three dimensional in space.
-Distance between the recording sites must be of the order of the size of the typical neurons involved, and such sites must be on a dense enough grid to perform numerical difference operations over them.
+The data must be two dimensional in space.
+Distance between the recording sites must be of the order of the size of the typical neurons involved or less, and such sites must be on a dense enough grid to perform numerical difference operations over them.
 The sampling frequency should be enough to detect sub-threshold activity, of at least 5kHz.
@@
 Unordered, highly homogeneous tissue, will not yield significant results.
-\item Obtain, frame by frame, the CSD from the data.
-Any method may be used, but computer expensive techniques are discouraged if the data is dense enough.
+ \item Calculate the CSD from the data.
+Any method may be used, but computer expensive techniques are discouraged if the data too dense.
 Common sense and tests are guidelines here.
-We recommend simpler, difference based approaches.
-\item Separate the CSD data into three sets: sources, sinks, and border or indiscernible (below error) data.
+We recommend simpler, difference based approaches wherever possible.
+If many electrodes fail near the r.o.i.
+kCSD is a good alternative if paraletization is available.
+If many electrodes are failing over the r.o.i., it is better to discard the data.
+\item Separate the CSD data into three sets: sources, sinks, and border or indiscernible (below error magnitude) data.
+Again, sensible criteria have to be defined.
+A rule of thumb would be to use the noise level of the recording as the amplitude for the ``thick zero'' range of values.
 \item Perform disjoint component analysis in the sources and sinks sets, frame by frame.
   A single pass algorithm could be used here \cite{Vincent91, Abubaker07}.
-\item  Obtain the CM for each disjoint component, using eq. \ref{cmparadisj}.
+\item  Obtain the CM for each disjoint component, using eq. \ref{cmparadisj}. Discard
+  all CM that are below some significant value. 
 \item For each frame and its successor, and for each CM, perform a ``most probable successor'' detection.
@@
 
-\section{Results}
+\section{Examples from Electro physiological data}
 
 The necessity to create such analysis came originally from the study of electro-physiological data obtained from the rat's hippocampus.
-The highly structured CA region of the hippocampus provides strong, discernible patterns of Sources and Sinks.
-Following the pattern of activity using Center of Mass analysis directly from the data, though, proves unfruitful, due to the non-convex nature of the layered tissue.
-Using CSD and disjoint component tracking over them provided the right conceptual framework to follow the displacement of active areas during highly active events, such as epileptic-like activity that was pharmacologically induced.
+The highly layered CA structure of the hippocampus provides strong, discernible patterns of Sources and Sinks.
+Following the pattern of activity using Center of Mass analysis directly from the data, though, proves unfruitful, due to the non-convex nature of the somatic layer.
+Using CSD and disjoint component tracking over them provided the right conceptual framework to follow the displacement of active areas during highly active events, such as epileptic-like activity that was pharmacological induced.
 The proposed analysis would not produce vector averages over different simultaneously active areas, on the contrary, it would separate them and track them independently.
 So the most violent epileptic burst provided us with the data necessary to prove the robustness of the method.
-a
+
+\subsection{Application to evoked activity in idem.}\label{sec:evocada}
+
+Evoked activity by electrical stimulus yields simpler and cleaner patterns in the LFP
+measurements. This facilitates the exemplification of our analysis, so we shall
+start the exposition of the usage with these kind of data.
+The data was obtained by stimulating certain specific points on the dentate gyrus
+of the slice of hippocampus with different voltage intensities. Details of the experiment
+will be published elsewhere \cite{FrancoGut2017}. We shall borrow some of
+the best experimental data for our purposes.
+
+A simple example of the procedure and the treatment of the data is shown in figure
+\ref{CSD_Evocada01}.  The subfigure A  shows a diagram of the hippocampus, with
+the somatic layers (CA and DG) as thick black lines and the stimulus point
+shown as a blue arrow. The $a,b,c,d,e$ dots are selected electrodes for
+showing LFP and CSD traces. Those are shown in the subfigure B. In subfigure
+C the same data is presented as a spatiotemporal color map that
+indicates shows the expected poles of firing neurons in transverse to the
+pyramidal stratum.  In the subfigure D we show the CSD for a series of time frames
+over the square delimited in the subfigure A. The idea for treating the different
+patches separately becomes evident here. The sources and sinks form patches
+with an apparent anterograde movement from CA3 to CA1. The patches change polarity
+as especteed, with the somatic layer becoming a strong sink between 3 to 4 ms
+after the stimulus, and then, at the 6 ms, inverting its polarity in a way that suggest
+successive activation of groups that can be presented as movement of the patch.
+
+\begin{figure}[h]
+  \includegraphics[width=0.9\textwidth]{Figuras/CSD_Evocada01.pdf}
+  \caption{ CSDA for evoked activity and visual inspection of disjoint
+    components. In A we present a diagram of the hippocampus slice over
+    the acquisition device. The somatic layers of CA and DG are represented as
+    thick black lines, and the stimulus site is indicated by a blue arrow. A selection
+    of six contiguous electrodes labeled $a,b,c,d,e$ in the r.o.i. are presented as blue
+    dots. In B we show the LFP traces and their CSD representation for those
+    selected electrodes. In C we present the same information as a color map,
+    with blue for sinks and red for sources, and a shorter time interval. The
+    map covers the 8ms after the stimulus is made (shown in the temporal axis by a
+    green doted line). The change of polarity around and over the str. pyr. (between b and
+    d) is clearly visible. In D we show the CSDA for various timestamps in the sub-region
+    indicated in A as the r.o.i, which covers a square of 24 by 24 electrodes over CA3.
+    in the first plot, 1ms before the stimulus, we show the grid of electrodes and a
+    shaded strip indicating the approximate location of str. pyr. A model neuron is
+    drawn for scale. At 4ms after the stimulus, the saturation of the electric shock
+    recedes and we can began to appreciate real neuronal activity on the responsive
+    sector of CA3. A great sink appears over the somatic layer, surrounded by two
+    sources and some smaller patches of activity. At the following frame the patch
+    appears to have moved anterogradly, towards CA1. Next frame, the sink appears
+    to have inverted its polarity as do their surrounding patches, and the movement
+    continues until approximately 9ms after stimulus, where the activity ceases to be
+    discernible from noise.}
+  \label{CSD_Evocada01}
+\end{figure}
+
+The evoked activity seems to yield much cleaner results, as we can now compare the trajectories of successive, very similar events.
+Our CSD and CM analysis can be used to detect consistent features in these experiments, such as the direction the waxing and waning of activity.
+Small deviations from the trajectory in successive experiments that are above the numerical error could indicate inconsistencies in the network below.
 
@@
 The evoked activity seems to yield much cleaner results, as we can now compare the trajectories of successive, very similar events.
-Our CSD and CM analysis can be used to detect consistent features in these experiments, such as the direction the waxing and wanning of activity.
+Our CSD and CM analysis can be used to detect consistent features in these experiments, such as the direction the waxing and waning of activity.
 Small deviations from the trajectory in successive experiments that are above the numerical error could indicate inconsistencies in the network below.

### 2017-11-01 ArticuloCSDyCM/Paper03.tex
cambios menores

@@
 Spatially tracking the activity of electro-physiological data in a concise, rigorous and interpretable manner poses a challenging problem.
-Measurement recordings are usually presented first as a series of traces that correspond each to a sampling point in the studied structure.
+Measurement recordings are usually presented first as a series of traces that correspond each to a sampling point on the tissue.
 As Micro Electrode Arrays (MEAs) become denser, this representation of the data becomes saturated and the discovery of patterns turns into a difficult task.
 Novel information is obfuscated by the same high density acquisition that allows us to discover it.
-Therefore, representing  the recorded data in ways which can unveil previously unobtainable information is a necessity.
+Representing  the recorded data in ways which can unveil previously obfuscated information is a necessity.
 Various techniques for separation of functional generators are available now, most come from the application of blind source separation analysis, such as ICA.

### 2017-11-08 ArticuloCSDyCM/Paper03.tex
cambios menores

@@
 Our task in this paper is to present a method for tracing out ``functional connectivity'', which would bring independent information that is complementary to those analysis.
-Our  method would be applicable mostly to activity in structured neural tissue.
-We shall exemplify it by application to the data obtained from recordings of the rats hippocampus.
+Our  method would be applicable mostly to activity in highly structured neural tissue.
+We shall exemplify it by application to the data obtained from recordings of the rats hippocampus, taking advantage of its lamminar
+orderer structure.
 
@@
 This representation not only provides a mathematical rigorous  measure for the weights in the vector averaging process, but helps to separate the data into disjoint components.
-This allows us to track separate process in the same region of interest.
+This allows us to track separate process in the same region of interest (r.o.i.) by tracking sepparately the poles of action: influx and eflux of charged ions.
 
-The application of the method requires electro-physiological data acquired through bi-dimensional MEAs, although the extension to three dimensional arrays would, in principle, be possible.
+The application of the method requires electro-physiological data acquired through bi-dimensional MEAs, although the extension to three dimensional arrays would be possible in principle.
 The MEA should have dense enough spatial and temporal samplings, so as to obtain good estimates of the CSD and their changes in time.
@@
 The concepts presented here require first obtaining the CSD from the recorded LFP.
-Any correct method would present the same utility.
+Any method would present the same utility.
 The first method is a  finite difference operator, while the second is the kCSDA presented by Potworowski \emph{et all} \cite{Potworowski2011}, which is an inverse problem approach.
 Where both methods are usable, we prefer the former, as is faster and requires less assumptions.
-We present to methods and show the robustness of the result for the Center of Mass tracking.
+We present both methods and show the robustness of the result for the Center of Mass tracking.
 
@@
 becomes necessary. The kCSDA method relies on the
-assumption that a finite family of reasonable functions can span (with acceptable accuracy) the possible CSD at each instant in time. The idea behind the method is simple:
-presenting one model function of this family and obtaining the LFP that such a function
-would generate, we calculate the coefficients for the entire family by projecting the
-experimental data over this ideal LFP model function family. A general presentation is in the reference, here
-we shall recount only the implementation that we used.
+assumption that a finite family of convenient functions can span, with acceptable accuracy, the possible CSD at each instant in time. The idea behind the method is simple:
+One presents a model function of this family and obtains the LFP that such a function
+would generate. Then one calculates the coefficients for the entire family by projecting the
+experimental data over this ideal LFP model function family. A general presentation is in the reference \cite{Potworowski2011},we shall recount only the basic implementation that we used.
 
-We shall borrow the notation in \cite{Potworowski2011}. Our model function for the CSD is a hard disk function, given by:
+We shall borrow the notation in the original paper \cite{Potworowski2011}. Our model function for the CSD is a hard disk function, described by:
 \begin{equation}
@@
 
-It is important to mention that our data has noise up to 0.1 of the magnitude of the
+It is important to mention that our data has noise up to 0.1 of the mean magnitude of the
 signal, which means that regularization of the matrix $K$ is unavoidable (see section 4.1 in \cite{Potworowski2011} and references therein).
@@
 
-The attainment of $K$ and $\tilde{K}$ turn out to be computer-costly for high density MEAs and dense temporal sampling, but paraletization, specially on
+The attainment of $K$ and $\tilde{K}$ turns out to be computer-costly for high density MEAs and dense temporal sampling, but paraletization, specially on
 GPUs, solves this issue.
-
 We have performed comparison of both methods over experimental data that has
@@
 In our case, we want to use the density as a mathematician would: a function which assigns weights to points in the space of interest and permits us to carry linear operations over them.
-The space here is the sampling points of the electrodes at a given instant in time, and the weight, the CSD, is amount interchange of charged ions from the inside to the outside of cellular membranes, per unit space.
+The space here is the sampling points of the electrodes at a given instant in time, and the weight, the CSD, is amount of interchange of charged ions from the inside to the outside of cellular membranes, per unit space.
 This is an indicator of activity in the neurons, therefore the CSD is also, indirectly, a \emph{``density of local activity''}.
@@
 The separation is not quantitative but qualitative: even when the signs of electric charge are arbitrary, the distinction between them is not.
-Therefore, CSD represents on each of the two active sets (sinks and sources) exactly what we need, a unit of concentration per unit space, in this case, per area, and the sign can be ignored, as long as we perform the operations in the sets separately.
+Therefore, CSD represents on each of the two active sets (sinks and sources) exactly what we need, a unit of concentration per unit space, in this case, per area, and the sign can be ignored, as long as we perform the operations in each set separately.
 Also, the separation by the zero value set of the CSD is in contrast to the zero value of the LFP, where it has no precise meaning as it only represents a practical reference value.
@@
 Likewise the set $C$ is made of three disjoint components.
-In the subfigre \ref{Disjuntos} Aiii we show the center of mass of the sets.
+In the subfigure \ref{Disjuntos} Aiii we show the center of mass of the sets.
 The set $A$ is concave, so, it can have a center of mass outside the set.
@@
 But, as we shall see,
-as the electrode resolution is finite, the posibility of having concave shapes is unlikely, and they
-tend to have a CM near their borders. In the figure \ref{Disjuntos} B we show this with experimental
+as the electrode resolution is finite, the posibility of having highly non compact shapes is unlikely, and they
+tend to have a CM inside or near their borders. In the figure \ref{Disjuntos} B we show this with experimental
 data (we are using evoked activity data, as explained in section \ref{sec:evocada}). In graph Bi we
-sepparate the CSD data into the sink and sources set. We take sinks as example. In the grabh Bii
+sepparate the CSD data into the sink and sources set. We take sinks as example. In the graph Bii
 we label each disjoint component, obtained by a single pass algorhythm. There are 25 components, and
@@
 In $B$ that is not possible, the set $B$ is convex in a mathematical sense.
-In subfigure Aii we made some modification to whos that
+In subfigure Aii we made some modification to show that
     each set can be made of many disjoint components, and each one of them can be convex or not.
-In Aiii we show the center of mass for $A$ and $B$ taken as a wholle, while for $C$ we break the set into each of the disjoint components and calculate each CM separately.
+In Aiii we show the center of mass for $A$ and $B$ taken as a whole, while for $C$ we break the set into each of the disjoint components and calculate each CM separately.
 There is no guarantee that each disjoint component would be convex, but as the resolution of the data is finite, the smaller components are more likely to
@@
 Very small, fast displacements of the Centers of mass, may be occur due to slight asynchronous between neurons which belong to the same active group.
-Larger and  longer than a few firing periods displacements could indicate physiological connections between different groups.
+Larger displacements and  longer intervals (more than two or so firing periods) could indicate physiological connections between different groups.
 

### 2017-11-23 ArticuloCSDyCM/Paper03.tex
Centros de Masa Nuevos

@@
 If these sets spread out over large non-convex geometries, the vector averages could lie outside the sets, having little or none physiological interpretation.
-Therefore we decided to introduce another concept from geometrical sets in order to make the analysis more local in character.
+Therefore we decided to introduce another concept from geometrical sets in order to reduce the analysis to local effects.
 
@@
 If each one of these patches could be interpreted as an active unit, we may asign to it a ``center of activity'' in order to follow their displacements.
-A Center of Mass for each instantaneous disjoint sink/source would provide us a very sensitive measurement of ``active locus''.
+A Center of Mass for each instantaneous disjoint sink/source would provide us a pragmatic measurement of an ``active locus''.
 We can then calculate at each time frame the CM for each disjoint component before tracking displacements of activity.
@@
 We must not take out of sight the phenomena which is producing the CSD distribution: neurons have a finite size, and the effect of their action a delimited extent.
-Every patch that we can identify as a disjoint component of the Sources/Sink sets indicates a coherent, joint active group of units.
+Every patch that we can identify as a disjoint component of the Sources/Sink sets indicates a  joint active set of units.
 A center of mass for each disjoint component would then indicate an average locus for this active group, a putative center of activity to which we can give a position, intensity and time.
@@
 Then the distance is:
+\begin{equation}
 d(q_1,q_2):=\sqrt{(x_1-x_2)^2+(y_1+y_2)^2}.
@@
  \item Calculate the CSD from the data.
-Any method may be used, but computer expensive techniques are discouraged if the data too dense.
+Any method may be used, but computer expensive techniques are discouraged if the data is too dense.
 Common sense and tests are guidelines here.
 We recommend simpler, difference based approaches wherever possible.
-If many electrodes fail near the r.o.i.
-kCSD is a good alternative if paraletization is available.
+If many electrodes fail near the r.o.i., kCSD is a good alternative if paraletization is available.
 If many electrodes are failing over the r.o.i., it is better to discard the data.
@@
 \item  Obtain the CM for each disjoint component, using eq. \ref{cmparadisj}. Discard
-  all CM that are below some significant value. 
+  all CM that are below some significant threshold value. A rule of thumb is one tenth of the average value for all CM. 
 \item For each frame and its successor, and for each CM, perform a ``most probable successor'' detection.
@@
 how far away and how different in intensity should be allowed a CM to be at time $t+1$ to be considered succesor in time of another at time $t$.
-We use the distance defined in eq.
-\ref{pseudodist}, a pseudo-Euclidean distance.
 \item ``Connect the dots'', that is, apply succesively the step above for each CM, until it stops having a succesor or its intensity falls below the error criteria.
 Then plot the succesive CMs as one trayectory.
-\item Plot all the trayectories and sort visually those that may have physiological interpretation from those who may be artifacts of the method. 
+\item Plot all the trayectories and sort visually those that may have physiological interpretation from those who may be artifacts of the method. This may need some form of expert validation.
 \end{enumerate}
@@
 So the most violent epileptic burst provided us with the data necessary to prove the robustness of the method.
+All the data comes from a series of experiments performed on slices of the rats hippocampus.
+
+\subsection{Application to facilitated activity}
 

### 2017-11-23 CenterOfMass/SoloCentrosdeMasa01.jl

@@ -8,16 +8,6 @@ esten en cero.
 =#
 
 
-nombre=ARGS[1]
-
-using JLD
-
-arx=load(nombre)
-#Checa como diablos se llama el CSD en cada archivo
-csd=arx["kCSDCorrecta"]
-(ancho,alto,tmax)=size(csd)
-
-
 # Funciones
 
 function vecindad8(punto::Array)
@@ -56,8 +46,8 @@ function ComponentesSP(DatosSignados::Array)
                     
                     
                     deleteat!(lista, indexin(Any[v], lista))
-                    push!(listaprofundeza, v)
-                    profundidad+=1
+                    push!(listaprofundeza, v) 
+                   profundidad+=1
                     push!(componentecurlab, v)
                 end
             end
@@ -111,8 +101,8 @@ function ObtenComponentesyCM(Datos::Array, tini=1,tfini=tmax, epsilon=1.0)
             #=
         componentesneg/pos son
             conjuntos con las listas de elemenentos de los
-        componentes en un instante dado. Se tiene que "cerar" siempre.
-        =#
+        componentes en un instante dado. Se tiene que "cerar" siempre. 
+       =#
         
         for p in componentesneg
             mu=length(p)        
@@ -135,7 +125,7 @@ function ObtenComponentesyCM(Datos::Array, tini=1,tfini=tmax, epsilon=1.0)
             end
         end
             centrosdemasaneg=centrosdemasaneg[2:end,:]
-            println(t,"hola")
+      #      println(t,"hola")
             CMNegativo[t]=centrosdemasaneg
  
         

### 2018-02-21 Reportes/ReporteCM02108.tex
Reporte 2018

karel@mezcalina Reportes $more Reporte01.tex 
\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}
\usepackage{graphicx}

\title{Reporte: Centro de Masa para potenciales evocados}

\author{W. P. Karel Zapfe}

\begin{document}

\maketitle

\section{Los Datos}

Para llevar a cabo está parte del trabajo, se me proporcionaron
los siguientes datos experimentales.
Dos experimentos largos de estimulación ortodrómica y uno
antidrómica. Cada experimento largo fue llevado a cabo
en dos etapas, una etapa control y una etapa con biliculina.
Cada etapa tiene estimulos de varios voltajes, en orden creciente,
de 1V a 9V.
El experimento antidrómico tiene una etapa control, una etapa con
mefluoquina y una etapa con nbqx-apv. La sucesión de estímulos es la misma.

El segundo experimento ortodrómico resultó inútil para los propósitos
de este análisis, dado que en todas las estimulaciones el número de
canales saturados sobre la región de interés (de ahora en adelante ROI)
era excesivo, y obfuscaban el cálculo densidad de fuentes de corriente
(de ahora en adelante CSD), así que nos concentramos en el primero experimento.
De ahí las estimulaciónes con 5V y 6V resultarón tener el registro más
limpio en condiciones de control, y la de 5V y 9V para la etapa
con biliculina.

El experimento antidrómico tuvo los mejores registros en 5V para control,
y en 7V para ambas drogas.

\section{Análisis exploratorio}

El primer paso para llevar a cabo el análisis de CSD-CM (Centro de Masa
para Densidades de Fuentes de Corriente) es conocer la posición de
los cuerpos anatómicos de interés en el registro, y simultaneamente,
conocer cuantos y cuales electrodos muestran respuestas fisiológicas y
cuales están saturados y resultan inútiles para nuestro propósito.
Una umbralización de Otsu sobre la variancia de cada canal 1 ms después del
estímulo nos da una indicación de posibles sedes para los somas de las neuronas.
complementado con la lista de los electrodos que registraron actividad fisiológica, podemos establecer la posición de CA y GD sobre el MEA. Una foto
podría ayudar de estar disponible, pero no cuento con ella, así que
tendremos que descanzar sobre los trazos fisiológicos y nuestro
conocimiento de la forma de la rebanada a priori. Muestro
los resultados así obtenidos para los casos analisados.

\section{ḱCSD y Regularización de Tikhonov}

Este tipo de experimento suele tener varios electrodos saturados, y pueden
estar sobre o muy cerca de la ROI. Un cálculo de diferencia finita para
obtener la CSD como hemos llevado a cabo anteriormente resulta muy inconveniente. Un solo electrodo ruidoso arruina a sus 8 vecinos en general, haciendo
que perdamos una parte sustancial del registro. En cambio, un análisis
de tipo inverso, tal como el kCSD \cite{Potworowski2011} nos da un resultado
mucho más suave y claro, con el costo de mayor tiempo computacional y unas
suposiciones extra. He decido implementer el proceso de kCSD para
registros masivos tales como los que proporciona el BIOCAM, paralelizando
sobre CPU el algorítmo, dado que es un problema trivialmente paralelizable.
Contamos asimismo con una versión parelela en GPU mucho más veloz
(desarrollada por Néstor Castillo para su tésis de licenciatura), pero
que requiere tarjetas gráficas nVidia de buena calidad para ser usada.

El kCSD requiere que hagamos una suposición razonable sobre
la forma de la función fuente-pozo cerca de un electrodo. La forma especifica
de la función no es importante mientras cumpla ciertas características
de regularidad y vaya acorde con nuestro conocimiento de como se puede
generar un espacio de funciones. He tomado como función modelo
una función característica circular con un radio de
media distancia inter electrodo (21 \mu m).

\section{}

\end{document}

### 2018-02-21 Graficando/MapaSaturadosParaEvodados.jl

@@ -1,0 +1,59 @@
+#=
+Para graficar los puntos de evocacion y asi de experimentos
+evocados
+=#
+
+using JLD, PyPlot
+
+nombre=ARGS[1]
+datos=load(nombre)
+
+
+lfp=datos["LFPSaturados"]
+saturados=datos["CanalesSaturados"]
+respuestas=datos["Canalesrespuesta"]
+retrazo=datos["retrazo"]
+frecuencia=datos["freq"]
+
+tantossaturados=length(saturados)
+tantosrespuesta=length(respuestas);
+
+## Esta es la latencia que recomendo Gis.
+latencia=round(Int, ceil(1.5*frecuencia))
+tejemplo=136
+ejemplo=lfp[:,:,tejemplo];
+
+xxsresp=zeros(tantosrespuesta)
+yysresp=zeros(tantosrespuesta)
+j=1
+for q in respuestas
+    xxsresp[j]=q[2]
+    yysresp[j]=q[1]
+    j+=1
+end
+xxssatu=zeros(tantossaturados)
+yyssatu=zeros(tantossaturados)
+j=1
+for q in saturados
+    xxssatu[j]=q[2]
+    yyssatu[j]=q[1]
+    j+=1
+end
+
+
+
+
+figure(figsize=(4.25,4))
+xlim(0,64)
+ylim(0,64)
+grid()
+limites=200
+imagen=imshow(ejemplo, interpolation="nearest", cmap="Spectral_r", 
+vmin=-limites,vmax=limites, extent=[0.5,64.5,0.5,64.5])
+cb=colorbar(imagen, fraction=0.046)
+scatter(xxsresp,yysresp, alpha=0.9, c="red", s=2, edgecolor="none", label="Respuestas")   
+scatter(xxssatu,yyssatu, alpha=0.9, c="black",s=2,  edgecolor="none", marker="x", label="Saturados") 
+legend()
+tight_layout()
+savefig("MapaSaturados.png",dpi=90)
+#scatter(46,31)

### 2018-02-23 Reportes/ReporteCM02108.tex
idi

@@
-karel@mezcalina Reportes $more Reporte01.tex 
 \documentclass{article}
@@
 \usepackage{graphicx}
+\usepackage{subcaption}
+\usepackage{textcomp}
 
@@
 en dos etapas, una etapa control y una etapa con biliculina.
-Cada etapa tiene estimulos de varios voltajes, en orden creciente,
+Cada etapa tiene estímulos de varios voltajes, en orden creciente,
 de 1V a 9V.
@@
 mefluoquina y una etapa con nbqx-apv. La sucesión de estímulos es la misma.
+Cada estímulo fue repetido 3 veces. He promediado los datos
+centrados en el pico del estímulo.
 
-El segundo experimento ortodrómico resultó inútil para los propósitos
+El segundo experimento ortodrómica resultó inútil para los propósitos
 de este análisis, dado que en todas las estimulaciones el número de
 canales saturados sobre la región de interés (de ahora en adelante ROI)
-era excesivo, y obfuscaban el cálculo densidad de fuentes de corriente
-(de ahora en adelante CSD), así que nos concentramos en el primero experimento.
-De ahí las estimulaciónes con 5V y 6V resultarón tener el registro más
+era excesivo, y ofuscaban el cálculo densidad de fuentes de corriente
+(de ahora en adelante CSD), así que nos concentramos en el primer experimento.
+De ahí las estimulaciones con 5V y 6V resultaron tener el registro más
 limpio en condiciones de control, y la de 5V y 9V para la etapa
@@
 para Densidades de Fuentes de Corriente) es conocer la posición de
-los cuerpos anatómicos de interés en el registro, y simultaneamente,
+los cuerpos anatómicos de interés en el registro, y simultáneamente,
 conocer cuantos y cuales electrodos muestran respuestas fisiológicas y
 cuales están saturados y resultan inútiles para nuestro propósito.
-Una umbralización de Otsu sobre la variancia de cada canal 1 ms después del
+Una umbralización de Otsu sobre la varianza de cada canal 1 ms después del
 estímulo nos da una indicación de posibles sedes para los somas de las neuronas.
@@
 podría ayudar de estar disponible, pero no cuento con ella, así que
-tendremos que descanzar sobre los trazos fisiológicos y nuestro
+tendremos que descansar sobre los trazos fisiológicos y nuestro
 conocimiento de la forma de la rebanada a priori. Muestro
-los resultados así obtenidos para los casos analisados.
+los resultados así obtenidos para los casos analizados en la figura \ref{exploratorio},
+y una muestra de los trazos en canales activos en la figura \ref{trazosejemplo1}.
+Los datos encontrados son poco esperanzadores. La estimulación ortodrómica parece
+no haber causado mucho registro en GD, a diferencia de la antidrómica, en donde
+podemos apreciar actividad en ambas estructuras.
 
-\section{ḱCSD y Regularización de Tikhonov}
+\begin{figure}[h]
+    \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{LFP_Promedio_est_con5-mapa.png}
+        \caption{Orto, Control, 5V}
+        \label{ocon5}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{LFP_Promedio_est_con6-mapa.png}
+        \caption{Orto, Control, 6V}
+        \label{ocon6}
+    \end{subfigure} \\
+ \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{LFP_Promedio_est_bic5-mapa.png}
+        \caption{Orto, Bic, 5V}
+        \label{ocon5}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{LFP_Promedio_est_bic9-mapa.png}
+        \caption{Orto, Bic,  9V}
+        \label{ocon6}
+    \end{subfigure} \\
+ \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{LFP_Promedio_310816_stim_antidromic5-mapa.png}
+        \caption{Anti, Control, 5V}
+        \label{ocon5}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{LFP_Promedio_310816_stim_antidromic_nbqx_apv7-mapa.png}
+        \caption{Anti, nbqxapv, 7V}
+        \label{ocon6}
+    \end{subfigure} \\
+    \caption{Análisis Exploratorio para los datos obtenidos,
+      se muestra un cuadro correspondiente a 2.58 ms después del estimulo,
+      en varios casos. Podemos ver que la orientación del experimento Antidrómico
+      es diferente (una reflexión horizontal, prácticamente) respecto al Ortodrómica.
+    }\label{exploratorio}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.95\textwidth]{TrazosEjemplares_bic5.png}
+  \caption{Trazos ejemplares del conjunto de canales detectados como
+    activos automáticamente, usando el experimento de Biliculina a 5V, ortodrómica.}
+  \label{trazosejemplo1}
+\end{figure}
+
+\section{kCSD y Regularización de Tikhonov}
 
@@
 mucho más suave y claro, con el costo de mayor tiempo computacional y unas
-suposiciones extra. He decido implementer el proceso de kCSD para
+suposiciones extra. He decido implementar el proceso de kCSD para
 registros masivos tales como los que proporciona el BIOCAM, paralelizando
-sobre CPU el algorítmo, dado que es un problema trivialmente paralelizable.
-Contamos asimismo con una versión parelela en GPU mucho más veloz
+sobre CPU el algoritmo, dado que es un problema trivialmente paralelizable.
+Contamos asimismo con una versión paralela en GPU mucho más veloz
 (desarrollada por Néstor Castillo para su tésis de licenciatura), pero
@@
 una función característica circular con un radio de
-media distancia inter electrodo (21 \mu m).
+media distancia inter electrodo (21 \textmu m). En los casos que he tratado,
+hasta ahora no parece haber diferencia considerable si cambiamos la
+función característica por una gaussiana o disminuimos o aumentamos el radio
+por un factor de dos, lo cual es una buena noticia, dado que quiere decir
+que nuestros resultados son estables frente a los supuestos que estamos haciendo.
 
-\section{}
+Los operadores del kCSD son dos: un operador normalmente designado por $K$ que representa
+la transformación del espacio de Fuentes al espacio de Voltajes, y un operador
+$\tilde{K}$ que representa la proyección de las fuentes sobre un espacio de funciones ideales del mismo espacio de fuentes (la base del espacio). El cálculo de estos operadores
+es la parte computacionalmente intensiva de la rutina, pero sólo tiene que hacer una vez
+por caso (i.e. por estímulo), dado que el cambio depende de cuales sean los electrodos
+inservibles en cada conjunto de datos.
+
+La concatenación directa de estos operadores presupone una señal ideal, sin ruido.
+En el mundo real la obtención del LFP es un proceso ruidos, en nuestro caso
+tenemos una razón de ruido a señal de aproximadamente 1/10.  Para disminuir el efecto
+del ruido en la aplicación del kCSDA se sobreestima la contribución de los canales
+vecinos ligeramente, de forma que el ruido del canal en cuestión quede
+aplacado por la suma de varias señales ligeramente diferentes. Encontrar
+la conveniente de aplacamiento requiere probar un parámetro sobre valores
+razonables y llevar a cabo validación cruzada.
+Valores muy bajos dan señales muy ruidosas, mientras que señales demasiado grandes
+suavizan en exceso la señal.
+Es bueno hacer la comparación contra un ejemplo donde podamos obtener la CSD
+por el método de diferencia finita, así que tengamos un ejemplo contra
+el cual comparar primero. He encontrado que un valor adecuado para el
+parámetro de regularización es 0.05. Los resultados para un ejemplo
+tomado de un pequeño recuadro de 24x24 electrodos se
+muestran el la figura \ref{regultik}.
+
+La regularización de Tikhinov basta hacerla una vez por cada nivel de
+ruido, es decir, si confiamos en que todos los experimentos evocados
+mantienen aproximadamente la misma razón ruido/señal, una vez
+bastará. Mas adelante podemos hacer una prueba más limpia.
+
+Una vez obtenido el parámetro adecuado, podemos obtener la CSD
+de cada conjunto de datos si ya contamos con los operadores
+$K$ y $\tilde{K}$.
+
+\begin{figure}[h]
+  \includegraphics[width=0.95\textwidth]{VarioskCSD03.png}
+  \caption{Prueba de regularización de Tikhonov}\label{regultik}
+\end{figure}
+
+\section{Componentes Disjuntos.}
+
+Obtenida la CSD del kCSDA, podemos empezar a separar los componentes disjuntos
+de pozos y fuentes. La rutina que se había utilizado anteriormente requirió
+un poco de ajuste, dado que las unidades arbitrarias del kCSDA reultan diferentes
+después de la regularización. He creado una función que toma como argumento
+un ``cero gordo'' alrededor del valor neutro de la CSD, y este debe
+ser ajustado respecto a la escala que tengamos. He encontrado que un valor
+de 0.01 veces el valor máximo de una fuente o un pozo es un buen criterio
+para separar los componentes. En las figuras \ref{cmcontrol} y \ref{cmbic}
+muestro
+ejemplos de componentes disjuntos para varios tiempos después del estimulo,
+con y sin biculina. Existe una regularidad en los componentes que permite
+discernir la punta de CA3.
+
+\begin{figure}
+  \centering
+    \begin{subfigure}[b]{0.95\textwidth}
+        \includegraphics[width=\textwidth]{DisjointComponents_Evento_control-5-120.png}
+    \end{subfigure}
+    \\
+     \begin{subfigure}[b]{0.95\textwidth}
+        \includegraphics[width=\textwidth]{DisjointComponents_Evento_control-5-127.png}
+    \end{subfigure}
+     \\
+      \begin{subfigure}[b]{0.95\textwidth}
+        \includegraphics[width=\textwidth]{DisjointComponents_Evento_control-5-134.png}
+    \end{subfigure}
+      \\
+       \begin{subfigure}[b]{0.95\textwidth}
+        \includegraphics[width=\textwidth]{DisjointComponents_Evento_control-5-141.png}
+    \end{subfigure}
+       \caption{Componentes Disjuntos para varios momentos después del estímulo, condiciones
+         de control, ortodrómica, 5V}\label{cmcontrol}
+\end{figure}
+
+\begin{figure}
+  \centering
+    \begin{subfigure}[b]{0.95\textwidth}
+        \includegraphics[width=\textwidth]{DisjointComponents_Evento_bic-5-120.png}
+    \end{subfigure}
+    \\
+     \begin{subfigure}[b]{0.95\textwidth}
+        \includegraphics[width=\textwidth]{DisjointComponents_Evento_bic-5-127.png}
+    \end{subfigure}
+     \\
+      \begin{subfigure}[b]{0.95\textwidth}
+        \includegraphics[width=\textwidth]{DisjointComponents_Evento_bic-5-134.png}
+    \end{subfigure}
+      \\
+       \begin{subfigure}[b]{0.95\textwidth}
+        \includegraphics[width=\textwidth]{DisjointComponents_Evento_bic-5-141.png}
+    \end{subfigure}
+       \caption{Componentes Disjuntos para varios momentos después del estímulo, condiciones
+         con biculina, ortodrómica, 5V}\label{cmbic}
+\end{figure}
+
+\section{Trayectorias}
+
+Finalmente comparemos las trayectorias de Centros de Masa para el experimento ortodrómico
+a 5V en condiciones control y biculina. Esta representación puede develar mejor los cambios
+de acción grupales en las estructuras fisiológicas \ref{traymedreqs}.
+
+\begin{figure}
+  \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+      \includegraphics[width=\textwidth]{LFP_Promedio_est_con5-Tray-t-134-medreqs.png}
+        \caption{control}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+       \includegraphics[width=\textwidth]{LFP_Promedio_est_bic5-Tray-t-134-medreqs.png}
+       \caption{biculina}
+     \end{subfigure}
+     \caption{Trayectorias con requisitos de aceptación intermedios, esto es,
+       cuya longitud sea mayor a 8 cuadros (correspondiente a poco mas de 1ms) y
+       8 u.a. de densidad.}
+\end{figure}
 

### 2018-02-23 Graficando/MapaSaturadosParaEvodados.jl

@@ -6,6 +6,8 @@ evocados
 using JLD, PyPlot
 
 nombre=ARGS[1]
+ng=nombre[1:end-4]
+nout=ng*"-mapa.png"
 datos=load(nombre)
 
 
@@ -13,13 +15,13 @@ lfp=datos["LFPSaturados"]
 saturados=datos["CanalesSaturados"]
 respuestas=datos["Canalesrespuesta"]
 retrazo=datos["retrazo"]
-frecuencia=datos["freq"]
+freq=datos["freq"]
 
 tantossaturados=length(saturados)
 tantosrespuesta=length(respuestas);
 
 ## Esta es la latencia que recomendo Gis.
-latencia=round(Int, ceil(1.5*frecuencia))
+latencia=round(Int, ceil(1.5*freq))
 tejemplo=136
 ejemplo=lfp[:,:,tejemplo];
 
@@ -47,13 +49,19 @@ figure(figsize=(4.25,4))
 xlim(0,64)
 ylim(0,64)
 grid()
+
+tiempo=round((tejemplo-retrazo)/freq, 2)
+
 limites=200
-imagen=imshow(ejemplo, interpolation="nearest", cmap="Spectral_r", 
+imagen=imshow(ejemplo, origin="lower", interpolation="nearest", cmap="Spectral_r", 
 vmin=-limites,vmax=limites, extent=[0.5,64.5,0.5,64.5])
 cb=colorbar(imagen, fraction=0.046)
-scatter(xxsresp,yysresp, alpha=0.9, c="red", s=2, edgecolor="none", label="Respuestas")   
-scatter(xxssatu,yyssatu, alpha=0.9, c="black",s=2,  edgecolor="none", marker="x", label="Saturados") 
+cb[:set_label]("Voltaje")
+scatter(xxsresp,yysresp, alpha=0.9, c="red", s=3, edgecolor="none", label="Respuestas")   
+scatter(xxssatu,yyssatu, alpha=0.9, c="black",s=3,  edgecolor="none", marker="x", label="Saturados") 
 legend()
 tight_layout()
-savefig("MapaSaturados.png",dpi=90)
+title("t= $tiempo ms")
+
+savefig(nout,dpi=90)
 #scatter(46,31)

### 2018-11-23 Reportes/SeminarioTraub2018.tex
SemTraub

\documentclass[serif]{beamer}

%\usepackage[numberedbib]{apacite}
\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}
\usepackage{natbib}
\usepackage{bibentry}
\bibliographystyle{apalike}

\begin{document}

\begin{frame}
    \frametitle{The track of activity and MEAs}
    Given a record of LFP is it possible to discover
    some track that the neural activity traces in a meaningful
    way?
    \begin{center}   
    \includegraphics[width=0.55\textwidth]{ImagenManjarrez01.jpg}
    \end{center}

    ZC Chao et all, \emph{Effects of Random External Background Stimulation
      on Network Synaptic Stability After Tetanization}, Neuroinformatics, 3, 2007.\\

    Manjarrez et all, \emph{Computing the Center of mass for traveling alpha waves in
  the human brain}, Brain Research 1145, 2007.   
  \end{frame}

   \begin{frame}
     \frametitle{CM is good, LFP a mass is not}
     \begin{equation}
       \langle x_k(t) \rangle =\frac{\sum_j x_{j,k} (t) m_{j,k} (t)} {\sum_j m_{j,k}(t)}
     \end{equation}   
     \begin{itemize}
       \item LFP (as most Potential measures) has no defined zero.
       \item The Spike Density is a true density, but has an heuristic parameter.
     \end{itemize}

     \begin{tabular}{ c c }
      \includegraphics[width=0.4\textwidth]{ComoManjarrezHaceCM01.png} &
       \includegraphics[width=0.4\textwidth]{ComoChaoHaceCM01.png} 
     \end{tabular}

    \end{frame}

  \begin{frame}
    \frametitle{A Natural density is at hand}
    Assuming constant, isotropic propiertes of the extracellular medium:
      \begin{equation}
       I_m/\sigma=-\nabla^2 V
      \end{equation}
      \begin{equation}
       \frac{\partial \rho}{\partial t} + \nabla \cdot \mathbf{J}=0 
      \end{equation}
      What about the sign?
      The sign help us to recognize two disjoint sets of activity:
      the Sources, and the Sinks.
      We have two different non overlapping measures: the
      Sink Density and the Source Density.
  \end{frame}

  \begin{frame}
     \frametitle{Implementation: naïve search}
     A direct approach proves useless, as the activity occurs at separated
     sites along CA1 and CA3. The concave geometry of our slice
     produces a centre of mass outside the CA, and without physiological
     interpretation.

       \begin{tabular}{ c c }
      \includegraphics[width=0.5\textwidth]{CSDSeminarioTraub.png} &
       \includegraphics[width=0.35\textwidth]{DiagCSDTraub.pdf} 
     \end{tabular}

  \end{frame}

  \begin{frame}
     \frametitle{A Natural Decomposition}
     Both Sink and Sources Set form patches over the interest region.
     There is a noice level that forms a ``thick'' zero. This allows us
     to sepparate each set into Disconnected components.
     \begin{tabular}{ l r }
      \includegraphics[width=0.35\textwidth]{DisjuntosUmbral.png} &
       \includegraphics[width=0.5\textwidth]{CSDCMSemTraub.png} 
     \end{tabular}
  \end{frame}

  \begin{frame}
    \frametitle{CM dynamics: putative trajectories for activity}
    Joining CM to CM at each time frame with a proximity threshold gives us
    putative trajectories for the activity, allready separated by type (either
    relasing or absorving positive ions of the enviroment) and
    site.
    \begin{center}
      \includegraphics[width=0.55\textwidth]{TrayectoriasLimpiasNumeros01.png} 
    \end{center}
  \end{frame}

  \begin{frame}
    There is still work to do here, and the Devil hides in the details.
    \begin{center}
      \includegraphics[width=0.55\textwidth]{TrayLimpSemTraub.pdf}
      \end{center}
  \end{frame}

  \begin{frame}
    \frametitle{And now for something completely different \ldots}
    There is other density avaible directly from the data:
    Power density.\\
    It also admits a decomposition into different interpretable sets:
    a decomposition by frequency bands.
    \begin{center}
      \includegraphics[width=0.55\textwidth]{DisconnectedByBands01.png}
      \end{center}
  \end{frame}

%  \begin{frame}{Bibliography}
   % \bibliographystyle{apacite}
  %  \bibliography{BiblioReportes01}
 % \end{frame}
\end{document}

### 2019-03-12 ArticuloCSDyCM/Paper03.tex
cambios paper

@@
 
-Spatially tracking the activity of electro-physiological data in a concise, rigorous and interpretable manner poses a challenging problem.
-Measurement recordings are usually presented first as a series of traces that correspond each to a sampling point on the tissue.
-As Micro Electrode Arrays (MEAs) become denser, this representation of the data becomes saturated and the discovery of patterns turns into a difficult task.
-Novel information is obfuscated by the same high density acquisition that allows us to discover it.
-Representing  the recorded data in ways which can unveil previously obfuscated information is a necessity.
-Various techniques for separation of functional generators are available now, most come from the application of blind source separation analysis, such as ICA.
-Our task in this paper is to present a method for tracing out ``functional connectivity'', which would bring independent information that is complementary to those analysis.
-Our  method would be applicable mostly to activity in highly structured neural tissue.
-We shall exemplify it by application to the data obtained from recordings of the rats hippocampus, taking advantage of its lamminar
-orderer structure.
+Spatially tracking succesive activity of electro-physiological data in a concise,
+rigorous and interpretable manner poses a challenging problem. Measurement recordings
+are usually presented first as a series of traces that correspond each to a sampling
+point on the tissue. As Micro Electrode Arrays (MEAs) become denser, this representation
+of the data becomes saturated and the discovery of patterns turns into a difficult task.
+Novel information is obfuscated by the same high density acquisition that allows us to
+discover it. Representing  the recorded data in ways which can unveil previously
+invisible patterns is a necessity. Various techniques for separation of functional
+generators are available now, most derived from the application of blind source
+separation analysis, such as ICA. 
+Our task in this paper is to present a method for tracing out
+``functional local connectivity'', which would bring independent information
+that is complementary to those analysis.  Our  method would be applicable mostly to
+activity in highly structured neural tissue. We shall exemplify it by application to
+the data obtained from recordings of the rats hippocampus, taking advantage of its lammiar orderer structure.
 
-Succesive activation  of distinct nearby units on structured tissue would trace a ``trajectory'' of the activity.
-A tracing of such trajectories would help to indicate causal or correlated ``functional conections'' over the tissue.
-Neighbouring neurons that fire in succesion, integrated subthreshold activity and ephaptic coupling could all be represented as pathways for transmission of activity.
-Nevertheless, how to define and acquire the putative paths is an open question.
+Local field potential can be regarded as produced by local multiunitary activity.
+A patch of simultaneous, similar, nearby traces can be regarded as one of those
+multiunitary elements. 
+Succesive activation  of distinct neighbouring elements on structured
+tissue would trace a
+``trajectory'' of the activity. A tracing of such trajectories would help to indicate
+causal or correlated ``functional local conections'' over the tissue. Neighbouring
+neurons that fire in succesion, integrated subthreshold activity and ephaptic coupling
+could all be represented as pathways for transmission of activity. Nevertheless,
+how to define and acquire the putative paths is an open question.
 
@@
 We shall call such methods Center of Mass Analyses (CMA).
-The \emph{wheights} that have been chosen for obtaining the average have been selected on heuristic grounds.
-In the works of Chao \emph{et al.} \cite{Chao05, Chao07}, the density of spikes in a certain time interval was used, a sort of ``density of action'', while Manjarrez \emph{et al.} \cite{Manjarrez07, Manjarrez09} used the positive half of the LFP recording as their weight.
+The ``wheights'' that have been chosen for obtaining the average have been selected on heuristic grounds.
+In the works of Chao \emph{et al.} \cite{Chao05, Chao07}, the density of spikes in a certain time interval was used, a sort of ``averaged density of action'', while Manjarrez \emph{et al.} \cite{Manjarrez07, Manjarrez09} used the positive half of the LFP recording as their weight.
 Here we advance on such methods by incorporating mathematically sound definitions and concepts which allow us to rigorously separate spatially distinct traces.
@@
 Our method permits us to map possible trajectories of activity inside structured neuronal tissue, without requiring ad hoc geometric constraints.
-As a prime example, non-convex structures which may show simultaneous activity in separate parts could test the robustness of our method.
+As a prime example, non-convex structures which may show simultaneous activity in separate parts could show the robustness of our method.
 We have tested then the method on slices of rat's hippocampus, concentrating in the CA region.
@@
 
-\section{Method and Calculations}
+\section{Methods and Calculations}
+
+\subsection{Preprocessing}
+
+After aquiring data from experiment, we pass the data through some preprocessing
+routines. For the most part, these serve to detect and discard unusable channels,
+such as saturated or failling electrodes. These vary from data set to data set,
+as they can be produced by bubbles or by saturation after electrical stimulation.
+We call this step the elimination of bad channels. We use a simple criterion for this:
+if the absolute value of the signal is above certain threshold (wich must be
+adjusted for evocated vs spontaneous acttivity) or if the standar deviation is
+larger than certain heuristically determined value, the channel is discarded.
+A similar but orthogonal criterium can be used for detection of channels with
+spikes.
 
@@
 Where both methods are usable, we prefer the former, as is faster and requires less assumptions.
-We present both methods and show the robustness of the result for the Center of Mass tracking.
+We have used both methods, kCSDA where
+the difference method is not applicable. The results are
+independent of the method save for multiplicative factor, which
+are irrelevant in the Center of Mass tracking.
 
 Thanks to the density and scales of the BioCAM 4096, a numerical finite difference method
-can give a good approximation to the CSD, if the number of failing channels is low.
-For experiments that were made without electrical stimulation this was the case, where the order of saturated or otherwise failing electrodes was around 10.
+can give a good approximation to the CSD, if the number of failing channels is low or
+are outside the r.o.i.
+For experiments that were made without electrical stimulation this turned out to be true, with the order of saturated or otherwise failing electrodes in the tens.
 A  numerical Laplacian Operator which reduces the cross effect of rectangular grids is the convolution of the data with the following matrix \cite{Lindberg90}:

### 2019-03-12 Preprocesamiento/PreprocTools.jl
cambios menores

@@ -4,7 +4,7 @@ using Statistics
 using HDF5
 using JLD
 
-export AbreyCheca, EncuentraTrancazosRaw,ActivAlrededorTrancazo, ActividadFueraTrancazo, FormaMatrizDatosCentrados, BuscaSaturados, BuscaSaturadosStd, BuscaCanalRespActPot, desviacionventanas
+export AbreyCheca, EncuentraTrancazosRaw,ActivAlrededorTrancazo, ActividadFueraTrancazo, FormaMatrizDatosCentrados, BuscaSaturados, BuscaSaturadosStd, BuscaRuidosos, BuscaCanalRespActPot, desviacionventanas
 
 #= Muchas funciones aqui presentes son para limpiar,
 manipular, cortar y suavizar datos. Ellas dependen
@@ -270,7 +270,8 @@ function suavegauss(trazo::Array, ms=0.5, freq=deffreq)
     # la sigma NO tiene porque ser entera, pero los cuadros si
     sigma=ms*freq
     cuatronv=round(Int, ceil(4*sigma)) # ¿que tan precisa la cola?
-    pesos=pesosgauss(simg    pesoT=sum(pesos)
+    pesos=pesosgauss(sigma, cuatronv)
+    pesoT=sum(pesos)
     l=length(trazo)
     cabeza=repeat([trazo[1]],cuatronv)
     cola=repeat([trazo[end]],cuatronv)

### 2019-03-13 Preprocesamiento/ArraySetTools.jl
RevisaHDF5 ahora sirve para detectar canales malos

@@ -1,0 +1,33 @@
+module ArraySetTools
+
+export rowstoset, elemtorow
+
+
+## ******************************************
+## funciones para manipular arrays y sets
+
+
+
+function rowstoset(xxs::Array)
+    # pasa los renglones de un arreglo a un conjunto como si jueran elementos
+    result=Set{Array{Int64,2}}()
+    (nrow,ncol)=size(xxs)
+    for j in 1:nrow
+        push!(result, reshape(xxs[j,:], 1, ncol))
+    end
+    return result
+end
+
+function elemtorow(xxs::Set)
+    #pasa los elementos de un conjunto a rengoles de un array
+    result=[0 0] 
+    for k in xxs
+        result=vcat(result, k)
+    end
+    return result[2:end,:]
+end
+
+
+
+
+end #module

### 2019-03-13 Preprocesamiento/PreprocTools.jl

@@ -2,9 +2,8 @@ module PreprocTools
 
 using Statistics
 using HDF5
-using JLD
 
-export AbreyCheca, EncuentraTrancazosRaw,ActivAlrededorTrancazo, ActividadFueraTrancazo, FormaMatrizDatosCentrados, BuscaSaturados, BuscaSaturadosStd, BuscaRuidosos, BuscaCanalRespActPot, desviacionventanas
+export AbreyCheca, EncuentraTrancazosRaw,ActivAlrededorTrancazo, ActividadFueraTrancazo, FormaMatrizDatosCentrados, BuscaSaturados, BuscaSaturadosStd, BuscaRuidosos, BuscaCanalRespActPot, desviacionventanas, tari, mediamov, gauss, pesosgauss, suavegauss
 
 #= Muchas funciones aqui presentes son para limpiar,
 manipular, cortar y suavizar datos. Ellas dependen
@@ -200,12 +199,8 @@ function BuscaCanalRespActPot(datos::Array,freq=deffreq, tini=0.5,
     return result[2:end, :]
 end
 
-
-function quitasatura(buenos::Array, malos::Array)
-    # quitemos del array de pares bueno los malos #
-    return 0
-end
-
+## Funciones mas numericas, que tienen que ver con
+## suavizar o resumir datos
 
 function desviacionventanas(data, ventmiliseg=100, freq=deffreq)
     ventana=round(Int, ventmiliseg*freq)
@@ -287,6 +282,7 @@ function suavegauss(trazo::Array, ms=0.5, freq=deffreq)
 
     
 end
+
 
 
 end #module

### 2019-03-26 ArticuloCSDyCM/Paper03.tex
mas paiper

@@
 are usually presented first as a series of traces that correspond each to a sampling
-point on the tissue. As Micro Electrode Arrays (MEAs) become denser, this representation
-of the data becomes saturated and the discovery of patterns turns into a difficult task.
+electrode on the tissue. Then a series of causality tests are made on the
+signal to search for possible efective conections. 
+As Micro Electrode Arrays (MEAs) become denser, that representation
+of the data saturates itself  and the discovery of patterns becomes a difficult task.
 Novel information is obfuscated by the same high density acquisition that allows us to
@@
 activity in highly structured neural tissue. We shall exemplify it by application to
-the data obtained from recordings of the rats hippocampus, taking advantage of its lammiar orderer structure.
+the data obtained from recordings of the rats hippocampus,
+taking advantage of its lammiar orderer structure.
 
@@
 We have tested then the method on slices of rat's hippocampus, concentrating in the CA region.
-We exemplify the usage of the method with data that has been obtained by an experiment with ...\textbf{bla bla bla describr experimentos}.
+We exemplify the usage of the method with data that has been obtained on experiments on rat's hippocampal slices that are centered on the GD and CA
+regions. The slices are put over an array of 64 by 64 electrodes
+(BioCAM 4096 by 3Brain \cite{BioCam}),
+and data is recorded at sampling rates between 7 and 14 kHz. The richness of the
+data permits us to obtain a prime use case for our method. Downscaling is
+possible in principle, but the array of electrodes should be at least 2D and
+a interelectrode distance of the order of magnitude of a typical pyramidal soma
+(around 100 $\mu$m).
 
@@
 After aquiring data from experiment, we pass the data through some preprocessing
-routines. For the most part, these serve to detect and discard unusable channels,
+routines. These serve to detect and discard unusable channels,
 such as saturated or failling electrodes. These vary from data set to data set,
-as they can be produced by bubbles or by saturation after electrical stimulation.
+as they can be produced by bubbles or by saturation after electrical stimulation
+or other random causes. 
 We call this step the elimination of bad channels. We use a simple criterion for this:
@@
 larger than certain heuristically determined value, the channel is discarded.
+
+%% esto puede no ser necesario. 
 A similar but orthogonal criterium can be used for detection of channels with
@@
 
-The concepts presented here require first obtaining the CSD from the recorded LFP.
-Any method would present the same utility.
+The concepts presented here require first obtaining the CSD from the recorded LFP. Any method would present the same utility, we have used two different
+approaches.
 The first method is a  finite difference operator, while the second is the kCSDA presented by Potworowski \emph{et all} \cite{Potworowski2011}, which is an inverse problem approach.
-Where both methods are usable, we prefer the former, as is faster and requires less assumptions.
+Where both methods are usable, we prefer the former,
+as it is faster and requires less assumptions.
 We have used both methods, kCSDA where
 the difference method is not applicable. The results are
-independent of the method save for multiplicative factor, which
-are irrelevant in the Center of Mass tracking.
+independent of the method save a for multiplicative factor, which
+is irrelevant in the Center of Mass tracking.
 
 Thanks to the density and scales of the BioCAM 4096, a numerical finite difference method
-can give a good approximation to the CSD, if the number of failing channels is low or
+produces an excelent approximation to the CSD,
+if the number of failing channels is low or
 are outside the r.o.i.
-For experiments that were made without electrical stimulation this turned out to be true, with the order of saturated or otherwise failing electrodes in the tens.
+For experiments that were made without electrical stimulation is the case, with the order of mangitude of unusable channels in the tens.
 A  numerical Laplacian Operator which reduces the cross effect of rectangular grids is the convolution of the data with the following matrix \cite{Lindberg90}:
@@
 A Gaussian spatial denoising can be performed before or after the Laplacian operator to the data in order to reduce excessive borders.
-Our Gaussian blur filter has a $\sigma$ value of three electrodes or $126 \mu m$.
-Intuitively this means that it smooths over each structure smaller than the soma of a typical pyramidal cell.
-This is  in accordance to our mean field approach, where we expect smooth differentiable fields over the scale of the neurons, but not in finer scales (for details see \cite{Bedard11}) .
+Our Gaussian blur filter has a $\sigma$ value of $126 \mu m$, which corresponds to three times the interelectrode distance.
+Intuitively this means that it smooths over each structures smaller than the soma of a typical pyramidal cell. 
+This is  in accordance to our mean field approach, where we expect smooth differentiable fields over the scale of the neurons, but not in finer scales (for details see \cite{Bedard11}).
 
@@
 becomes necessary. The kCSDA method relies on the
-assumption that a finite family of convenient functions can span, with acceptable accuracy, the possible CSD at each instant in time. The idea behind the method is simple:
-One presents a model function of this family and obtains the LFP that such a function
+assumption that a finite family of convenient functions can span, with acceptable accuracy, a very good aproximation to the  CSD at each instant in time.
+ A general presentation is in the reference \cite{Potworowski2011},we shall recount only the basic implementation that we used.
+One presents a family of model functions for the CSD and
+obtains the LFP that such functions
 would generate. Then one calculates the coefficients for the entire family by projecting the
-experimental data over this ideal LFP model function family. A general presentation is in the reference \cite{Potworowski2011},we shall recount only the basic implementation that we used.
-
+experimental data over this ideal LFP model function family.
 We shall borrow the notation in the original paper \cite{Potworowski2011}. Our model function for the CSD is a hard disk function, described by:
@@
 Notice how the method incorporates a denoising mechanism, so
-no smoothing procedures are used.
+no smoothing procedures are needed.
 The great advantage of this method is that is possible to obtain a reasonable
@@
 
-The attainment of $K$ and $\tilde{K}$ turns out to be computer-costly for high density MEAs and dense temporal sampling, but paraletization, specially on
+The attainment of $K$ and $\tilde{K}$ turns out to be computer-costly for high density MEAs, but paraletization, specially on
 GPUs, solves this issue.
+
+%% posiblemente quitar esto 
 We have performed comparison of both methods over experimental data that has
@@
 This is an indicator of activity in the neurons, therefore the CSD is also, indirectly, a \emph{``density of local activity''}.
+Having both
+the mathematician and electrophysiologicist interpretations presents we can
+perform the following analysis with clear readings.
 
-Apparently, CSD would not fit the requirements of non-negativity for its use as a mathematical density, but this turns out to be not a problem.
+Apparently, CSD would not fit the requirements of non-negativity for its use as a mathematical density, but this is not a real problem.
 CSD separates the measured activity into three distinct sets, namely, the set of all sources, the set of all sinks and the zero set.
 The separation is not quantitative but qualitative: even when the signs of electric charge are arbitrary, the distinction between them is not.
-Therefore, CSD represents on each of the two active sets (sinks and sources) exactly what we need, a unit of concentration per unit space, in this case, per area, and the sign can be ignored, as long as we perform the operations in each set separately.
+Therefore, CSD represents on each of the two active sets (sinks and sources) exactly what we need, a unit of concentration of changing charges per unit of space (in this case, per area), and the sign can be ignored, as long as we perform the operations in each set separately.
 Also, the separation by the zero value set of the CSD is in contrast to the zero value of the LFP, where it has no precise meaning as it only represents a practical reference value.
 This is acknowledged in most of the literature, where only the scales of the LFP recording are shown, without reference to the sign of the values.
-In the CSD representation,  the so called zero set (all the points in our space which have a recorded CSD of zero)  has a precise meaning: it is an instantaneous lack of activity and a border between sinks and sources.
+In the CSD representation,  the so called zero set (all the points in our space which have a recorded CSD of zero)  has a precise meaning: it is an instantaneous lack of activity and the border between sinks and sources.
 This shall be exploited further in the section below.
@@
 Trying to use this for obtaining vector averages would still be crude and inexact.
-If these sets spread out over large non-convex geometries, the vector averages could lie outside the sets, having little or none physiological interpretation.
+If these sets spread out over large non-convex areas, the vector averages could lie outside the sets, having little or none physiological interpretation.
 Therefore we decided to introduce another concept from geometrical sets in order to reduce the analysis to local effects.
 
-\subsection{CSD and Disjoint Components}
+\subsection{CSD and Disconnected Components}
 
 A first visual inspection of the CSD color map at a given instant in time shows that both sources and sinks appear in rather large and significant patches.
-These appear to correspond to the expected double and triple poles of firing neurons in structured tissue \cite{Buzsaki2012} and inspired us to separate the sink and source  sets into their \emph{disjoint components}.
-This concept is used in mathematics to denote subsets of a given geometric set that do not touch each other or share no elements in common \cite{Halmos}.
+These appear to correspond to the expected double and triple poles of firing neurons in structured tissue \cite{Buzsaki2012} and inspired us to separate the sink and source  sets into their \emph{disconnected components}.
+This concept is used in mathematics to denote subsets of a given geometric set that do not touch each other  \cite{Halmos}.
 Their rough correspondence to poles of firing neurons would be most useful in tracking the successive locus of the activity.
@@
 To clarify the concept and how we are we using it an example is in order.
-In figure \ref{Disjuntos} A we show geometrical sets enclosed by simple curves as color patches, where every color indicates a Set.
-The set $A$ consists of a single component, that means that one can make  path between any two points in the set that consist entirely of points of the same set.
-In contrast, the set $B$ is made of two disjoint components, labeled $B1$ and $B2$.
+In figure \ref{disconnectedsets} A we show geometrical sets enclosed by simple curves as color patches, where every color indicates a Set.
+The set $A$ consists of a single connected component, that means that one can make  path between any two points in the set that consist entirely of points of the same set.
+In contrast, the set $B$ is made of two disconnected components, labeled $B1$ and $B2$.
 If one chooses a point in each of the components, one cannot draw a path joining them without leaving the set.
 But each one of those components is connected, meaning that if we choose the pair of points in one component, then again we can trace a path joining them in the set.
-So the set $B$ consists of two disjoint components, but each component can be regarded as a connected set on their own.
-Likewise the set $C$ is made of three disjoint components.
-In the subfigure \ref{Disjuntos} Aiii we show the center of mass of the sets.
+So the set $B$ consists of two disconnected components, but each component can be regarded as a connected set on their own.
+Likewise the set $C$ is made of three disconnected components.
+In the subfigure \ref{disconnectedsets} Aiii we show the center of mass of the sets.
 The set $A$ is concave, so, it can have a center of mass outside the set.
 The
-set $B$ has also a center of mass outside the set. It has two disjoint components that are aproximatelly
-equall and separated by a long distance (this is not an implication: a set can be disjoint and have
+set $B$ has also a center of mass outside the set. It has two disconnected components that are aproximatelly
+equal and separated by a long distance (this is not an implication: a set can be disjoint and have
 a CM inside the set). For the set $C$ our aproach has been used: we treat every disjoint component of the
@@
 as the electrode resolution is finite, the posibility of having highly non compact shapes is unlikely, and they
-tend to have a CM inside or near their borders. In the figure \ref{Disjuntos} B we show this with experimental
+tend to have a CM inside or near their borders. In the figure \ref{disconnectedsets} B we show this with experimental
 data (we are using evoked activity data, as explained in section \ref{sec:evocada}). In graph Bi we
@@
   \caption{ Explanation of the geometrical concepts used on our analysis. In subfigure A we show
-    with an artificial example the concepts of disjointness and convexivity. The subfigure Ai shows
+    with an artificial example the concepts of connectedness and convexivity. The subfigure Ai shows
     three sets, $A$, $B$, and $C$. $A$ fails to be convex because it is concave, and $C$ fails to be
-    be convex because it is made of two disjoint components. In each of them is possible to draw a line
+    be convex because it is made of two disconnected components. In each of them is possible to draw a line
     connecting two points inside the set that is not entirely inside the set.
@@
 In subfigure Aii we made some modification to show that
-    each set can be made of many disjoint components, and each one of them can be convex or not.
+    each set can be made of many disconnected components, and each one of them can be convex or not.
 In Aiii we show the center of mass for $A$ and $B$ taken as a whole, while for $C$ we break the set into each of the disjoint components and calculate each CM separately.
-There is no guarantee that each disjoint component would be convex, but as the resolution of the data is finite, the smaller components are more likely to
-    be convex. In figure B we shoe everything with a frame of experimental data. In Bi we sepparate the
-    data into three sets, Sources, Sinks, and Zeros. We discard the Zeros and keep the interesting ones.
-    In Figure Bii we take the Sink Set and separate it into disjoint components, that are labeled from one to 25 in this particular example.
+There is no guarantee that each disconnected component would be convex, but as the resolution of the data is finite,  small components are more likely to
+    be convex. In figure B we show everything with a frame of experimental data. In Bi we sepparate the
+    data into three sets, Sources, Sinks, and Zeros. We discard the Zeros and keep the active ones.
+    In Figure Bii we take the Sink Set and separate it into disconnected components,
+    they are labeled from one to 25 in this particular example.
     From this subsets we calculate the total mass of each one and only
@@
   }
-  \label{disjuntos}
+  \label{disconnectedsets}
 \end{figure}
@@
 
-The ``thick zero'' would help us to sepparate the sinks and sources sets into its disjoint components, which would correspond then to clear patches of definite activity.
+The ``thick zero'' would help us to sepparate the sinks and sources sets into its disconnected components, which would correspond then to clear patches of definite activity.
 If each one of these patches could be interpreted as an active unit, we may asign to it a ``center of activity'' in order to follow their displacements.
-A Center of Mass for each instantaneous disjoint sink/source would provide us a pragmatic measurement of an ``active locus''.
-We can then calculate at each time frame the CM for each disjoint component before tracking displacements of activity.
+A Center of Mass for each instantaneous disconnected sink/source would provide us a pragmatic measurement of an ``active locus''.
+We can then calculate at each time frame the CM for each one of those components before tracking displacements of activity.
 
-At each disjoint component at a precise time $t$, indexed $k$, we have a number of points with CSD of the same sign, indexed $j$. Then, the CM of the $k$ component would have the usual definition:
+At each component at a precise time $t$, indexed $k$, we have a number of points with CSD of the same sign, indexed $j$. Then, the CM of the $k$ component would have the usual definition:
 \begin{equation}\label{cmparadisj}
@@
 \end{equation}
-where $I(q_{j,k},t)$ is the CSD at point $q_{j,k}=(x_{j,k}, y_{j,k})$ at time $t$.
-These disjoint components, while not rigorously convex, are less spread over the space of the measurements and tend to have their CM over the actual active points, or at the border of the active units.
+where $I(q_{j,k},t)$ is the CSD at point $q_{j,k}=(x_{j,k}, y_{j,k})$ at time $t$, and the
+sign of our ``mass'' eliminates itself graciously.
+These disconnected components, while not rigorously convex, are less spread over the space of the measurements
+and tend to have their CM over the actual active points, or at the border of the active channels.
 
@@
 We must not take out of sight the phenomena which is producing the CSD distribution: neurons have a finite size, and the effect of their action a delimited extent.
-Every patch that we can identify as a disjoint component of the Sources/Sink sets indicates a  joint active set of units.
+Every patch that we can identify as a disconnected component of the Sources/Sink sets indicates a  joint active set of units.
 A center of mass for each disjoint component would then indicate an average locus for this active group, a putative center of activity to which we can give a position, intensity and time.
@@
 Very small, fast displacements of the Centers of mass, may be occur due to slight asynchronous between neurons which belong to the same active group.
-Larger displacements and  longer intervals (more than two or so firing periods) could indicate physiological connections between different groups.
+Larger displacements and  longer intervals (more than two or so firing periods) could indicate physiological connections between different groups or units.
 
 The next step in our analysis is a procedure to systematically associate to each CM at time $t$ a ``successor'' at time $t+\Delta t$, where $\Delta t$ is the sampling time step.
-We used an euclididan distance.
 Let $q_1=(x_1, y_1)$ be the coordinates of the first point and $q_2=(x_2,y_2)$, the corresponding coordinates of the second.
-Then the distance is:
+Remember that these coordinates can be now rational numbers
+respect to the grid of electrodes, as CM are not restricted to be  at the grid points.
+The distance is:
 \begin{equation}
@@
 \end{equation}
-The integrated intensity of a CM is simply the denominator in the eq. \ref{cmparadisj}, that is, the total ``mass'' obtained by integrating the CSD in a specific disjoint component:
+The integrated intensity of a CM is simply the denominator in the eq. \ref{cmparadisj}, that is, the total ``mass'' obtained by integrating the CSD in a specific component:
 \begin{equation}
@@
 When a specific trajectory cannot find a successor point that is above the
-threshold intensity, we assume that it is finished, so we do not concatenate
+threshold intensity, we assume that it has ended, so we do not concatenate
 points that are separated in time by more than $\Delta t$. In order to paste different
-trajectories, more rigorous analysis of their origin and interpretation would be needed.
+trajectories, more rigorous analysis of their origin and interpretation should be used and is
+beyond the scope of this work.
 
@@
 The data must be two dimensional in space.
-Distance between the recording sites must be of the order of the size of the typical neurons involved or less, and such sites must be on a dense enough grid to perform numerical difference operations over them.
+Distance between the recording sites must be of the order of the typical size of the neurons involved or less, and such sites must be on a dense enough grid to perform numerical difference operations over them.
 The sampling frequency should be enough to detect sub-threshold activity, of at least 5kHz.
@@
 Unordered, highly homogeneous tissue, will not yield significant results.
- \item Calculate the CSD from the data.
-Any method may be used, but computer expensive techniques are discouraged if the data is too dense.
-Common sense and tests are guidelines here.
-We recommend simpler, difference based approaches wherever possible.
-If many electrodes fail near the r.o.i., kCSD is a good alternative if paraletization is available.
-If many electrodes are failing over the r.o.i., it is better to discard the data.
-\item Separate the CSD data into three sets: sources, sinks, and border or indiscernible (below error magnitude) data.
+ \item Calculate the CSD from the data. Any method may be used, but computer expensive techniques are discouraged if the
+   number of channels is over the thousands.
+   Common sense and tests are guidelines here. We recommend simpler, difference based approaches wherever possible. If many electrodes fail near the r.o.i., kCSD is a good alternative if done with
+   parallel routines.  If too many electrodes are failing over the r.o.i., it is better to discard the data. 
+ \item Separate the CSD data into three sets: sources, sinks, and border or indiscernible (below error magnitude) data
+   which we call the zero set.
 Again, sensible criteria have to be defined.
-A rule of thumb would be to use the noise level of the recording as the amplitude for the ``thick zero'' range of values.
-\item Perform disjoint component analysis in the sources and sinks sets, frame by frame.
-  A single pass algorithm could be used here \cite{Vincent91, Abubaker07}.
-\item  Obtain the CM for each disjoint component, using eq. \ref{cmparadisj}. Discard
-  all CM that are below some significant threshold value. A rule of thumb is one tenth of the average value for all CM. 
+A rule of thumb would be to use the mean of the peaks of the noise  of the recording as the amplitude for the ``thick zero'' range of values.
+\item Perform disconnected component analysis in the sources and sinks sets, frame by frame.
+  A single pass algorithm is adequate \cite{Vincent91, Abubaker07}.
+\item  Obtain the CM for each component, using eq. \ref{cmparadisj}. Discard
+  all CM that are below some significant threshold value. A rule of thumb is one tenth of the average value for all CM,
+  or that are in the lesser decil. 
 \item For each frame and its successor, and for each CM, perform a ``most probable successor'' detection.
@@
 Following the pattern of activity using Center of Mass analysis directly from the data, though, proves unfruitful, due to the non-convex nature of the somatic layer.
-Using CSD and disjoint component tracking over them provided the right conceptual framework to follow the displacement of active areas during highly active events, such as epileptic-like activity that was pharmacological induced.
+Using CSD and disconnected component tracking over the data provided the right conceptual framework to follow the displacement of active areas during highly active events, such as epileptic-like activity that was pharmacological induced.
 The proposed analysis would not produce vector averages over different simultaneously active areas, on the contrary, it would separate them and track them independently.
 So the most violent epileptic burst provided us with the data necessary to prove the robustness of the method.
-All the data comes from a series of experiments performed on slices of the rats hippocampus.
+All the experimental data comes from a series of experiments
+performed on slices of the rats hippocampus.
 
@@
 of the slice of hippocampus with different voltage intensities. Details of the experiment
-will be published elsewhere \cite{FrancoGut2017}. We shall borrow some of
+have been published elsewhere \cite{FrancoGut2017}. We shall borrow some of
 the best experimental data for our purposes.
 
-A simple example of the procedure and the treatment of the data is shown in figure
+The LFP is averaged over three events, produced by three succesive
+electrical stimuli in the same point at constant voltage. The
+response of the cels is allmost identical in each event, which
+allows us to use the average as a denoizing mechanism. After averaging
+the syncronized data we perform kCSDA, as in this case we end up
+with many saturated electrodes and dCDSA is not viable.
+Once that we have the CSD representation, we select noisy channels outside
+the r.o.i. and use them to obtain the threshold for sepparating
+the Sink and Sources sets. Then we procede as stated in the previous section.
+
+An example of the procedure and the treatment of the data is shown in figure
 \ref{CSD_Evocada01}.  The subfigure A  shows a diagram of the hippocampus, with
@@
 C the same data is presented as a spatiotemporal color map that
-indicates shows the expected poles of firing neurons in transverse to the
+shows the expected poles of firing neurons transverse to the
 pyramidal stratum.  In the subfigure D we show the CSD for a series of time frames
@@
 after the stimulus, and then, at the 6 ms, inverting its polarity in a way that suggest
-successive activation of groups that can be presented as movement of the patch.
+successive activation of groups that are represented as movement of the patch.
 
@@
 This turned out to be a test of the applicability and robustness of our method.
-The necessity of having good error value for the ``thick zero'' set was more apparent here than anywhere else.
+The necessity of having good error value for the zero set was more apparent here than anywhere else.
 Also the posibilty to create new structures of numerical data was first seen here.
@@
 
-\subsection{Application to evoked activity in idem.}
-
-The evoked activity seems to yield much cleaner results, as we can now compare the trajectories of successive, very similar events.
-Our CSD and CM analysis can be used to detect consistent features in these experiments, such as the direction the waxing and waning of activity.
-Small deviations from the trajectory in successive experiments that are above the numerical error could indicate inconsistencies in the network below.
-
 \subsection{Application to spontaneous activity in idem.}
@@
 We had one experiment in which we could trace very nicely the data on DG and CA simultaneously without adding excitatory drugs.
+
+\subsection{Counterexample: the stratial neuronal tissue}
+
+bla bla.
 

### 2019-03-27 Preprocesamiento/PreprocTools.jl
FFTW en linea, posible ataque

@@ -31,8 +31,9 @@ function AbreyCheca(x::String)
     maxvolt=archivo["/3BRecInfo/3BRecVars/MaxVolt"][1][1]
     minvolt=archivo["/3BRecInfo/3BRecVars/MinVolt"][1][1]
     bitdepth=archivo["/3BRecInfo/3BRecVars/BitDepth"][1][1]
+    inversion=archivo["SignalInversion"][1][1]
     duracionexperimento=numcuadros/frecuencia
-    factordeescala=(maxvolt-minvolt)/2^bitdepth
+    factordeescala=(maxvolt-minvolt)/2^bitdepth*inversion
     DatosCrudos=read(archivo["/3BData/Raw"])
     result=Dict("numcuadros" => numcuadros,
                 "frecuencia"=> frecuencia,

### 2019-03-28 Preprocesamiento/PreprocTools.jl
FFTW Raw listo

@@ -31,7 +31,7 @@ function AbreyCheca(x::String)
     maxvolt=archivo["/3BRecInfo/3BRecVars/MaxVolt"][1][1]
     minvolt=archivo["/3BRecInfo/3BRecVars/MinVolt"][1][1]
     bitdepth=archivo["/3BRecInfo/3BRecVars/BitDepth"][1][1]
-    inversion=archivo["SignalInversion"][1][1]
+    inversion=archivo["/3BRecInfo/3BRecVars/SignalInversion"][1][1]
     duracionexperimento=numcuadros/frecuencia
     factordeescala=(maxvolt-minvolt)/2^bitdepth*inversion
     DatosCrudos=read(archivo["/3BData/Raw"])

### 2019-04-01 ArticuloCSDyCM/Paper03.tex
paper paper paper

@@
 
-A first visual inspection of the CSD color map at a given instant in time shows that both sources and sinks appear in rather large and significant patches.
-These appear to correspond to the expected double and triple poles of firing neurons in structured tissue \cite{Buzsaki2012} and inspired us to separate the sink and source  sets into their \emph{disconnected components}.
+A first visual inspection of the CSD color map at a given instant in time shows that both sources and sinks appear in rather large and significant patches, with typical
+cross lengths around somo 150 $\mu m$.
+
+These appear to correspond to the expected poles of firing neurons in structured tissue \cite{Buzsaki2012} and inspired us to separate the sink and source  sets into their \emph{disconnected components}.
 This concept is used in mathematics to denote subsets of a given geometric set that do not touch each other  \cite{Halmos}.
-Their rough correspondence to poles of firing neurons would be most useful in tracking the successive locus of the activity.
+Their rough correspondence to poles of firing neurons is most useful in tracking the successive locus of the activity.
 
@@
 set $B$ has also a center of mass outside the set. It has two disconnected components that are aproximatelly
-equal and separated by a long distance (this is not an implication: a set can be disjoint and have
-a CM inside the set). For the set $C$ our aproach has been used: we treat every disjoint component of the
-set as an independent entity and calculate the center of mass for each one.
-Notice that this doesn't preclude the posibility of having a concave subset with a CM that lies outside.
-But, as we shall see,
+equal and separated by a long distance (this is not an implication: a set can be disconnected and have
+a CM inside the set). For the set $C$ our aproach has been used: we treat every disconnected component of the
+set as an independent entity and calculate the center of mass for each one. Notice that this doesn't preclude t
+he posibility of having a concave subset with a CM that lies outside. But, as we shall see,
 as the electrode resolution is finite, the posibility of having highly non compact shapes is unlikely, and they
@@
 seen). Each CM gets listed as a probable step in a trajectory.
+
+%%%% *********** Vas aqui
 

### 2019-04-09 ArticuloCSDyCM/Paper03.tex
todo el protopaper revisado

@@
-\documentclass{article}
+\documentclass[12pt]{article}
 
@@
 \usepackage{setspace}
+\usepackage{lineno} %ambos paquetes para borrador
+
 \usepackage{graphicx}
 \usepackage{longtable}
+
+\doublespacing
+\linenumbers
 
@@
 
-%%%% *********** Vas aqui
-
 \begin{figure}
   \includegraphics[width=0.9\textwidth]{Figuras/Disjuntos01.pdf}
-  \caption{ Explanation of the geometrical concepts used on our analysis. In subfigure A we show
-    with an artificial example the concepts of connectedness and convexivity. The subfigure Ai shows
-    three sets, $A$, $B$, and $C$. $A$ fails to be convex because it is concave, and $C$ fails to be
-    be convex because it is made of two disconnected components. In each of them is possible to draw a line
+  \caption{ Explanation of the geometrical concepts used on our analysis.
+    In subfigure A we show
+    with an artificial example the concepts of connectedness and convexivity.
+    The subfigure Ai shows
+    three sets, $A$, $B$, and $C$. $A$ fails to be convex because it is concave,
+    and $C$ fails to be
+    be convex because it is made of two disconnected components.
+    In each of them is possible to draw a line
     connecting two points inside the set that is not entirely inside the set.
@@
     In subfigure Aii we made some modification to show that
-    each set can be made of many disconnected components, and each one of them can be convex or not.
-In Aiii we show the center of mass for $A$ and $B$ taken as a whole, while for $C$ we break the set into each of the disjoint components and calculate each CM separately.
-There is no guarantee that each disconnected component would be convex, but as the resolution of the data is finite,  small components are more likely to
-    be convex. In figure B we show everything with a frame of experimental data. In Bi we sepparate the
-    data into three sets, Sources, Sinks, and Zeros. We discard the Zeros and keep the active ones.
+    each set can be made of many disconnected components,
+    and each one of them can be convex or not.  In Aiii we show the center of
+    mass for $A$ and $B$
+    taken as a whole, while for $C$ we break the set into each of the disconnected
+    components and calculate each CM separately. There is no guarantee that each
+    disconnected component would be convex, but as the resolution of the data is finite,
+    small components are more likely to
+    be convex. In figure B we show everything with a frame of experimental data.
+    In Bi we sepparate the
+    data into three sets, Sources, Sinks, and Zeros.
+    We discard the Zeros and keep the active ones.
     In Figure Bii we take the Sink Set and separate it into disconnected components,
@@
     From this subsets we calculate the total mass of each one and only
-    considerate for the next step those who are above certain threshold of integrated intensity. The
-    next step is shown in Biii, where we plot the Center of Mass for each of such subsets. Deviations
-    from convexity seem to be unimportant, as most of the mass is concentrated in the inner electrodes
+    considerate for the next step those who are above certain threshold. The
+    next step is shown in Biii, where we plot the Center of Mass for each of
+    such subsets. Deviations
+    from convexity are tolerable,
+    as most of the mass is concentrated in the inner electrodes
     of each set.
@@
 
-Theoretically, CSD would divide the set of measuring points into three subsets: the set of all sources, the set of all sinks, and the border between them i.e, the set of all measuring points that have a CSD positive value, the set of those who have a negative CSD value, and the set of those points having exactly the zero value in CSD space.
-Due to the finite precision of the measurement, the validity of the mean field approach, and the noise in the measurements, the sinks and sources sets can make contact with each other.
-Said measurements produce a function over a discrete point set.
-Noise between patches corresponding to a single set often makes a ``bridges'' between components, making identification of poles inexact.
-In order to avoid this we create a ``thick zero'' set of all the recording points whose CSD values are inside a small error interval around zero.
-This separates more sharply the components of the set, represents our ignorance of precise borders and helps us to simplify the following analysis.
+Theoretically, CSD would divide the set of measuring points into three subsets: the set of all sources, the set of all sinks, and the border between them i.e, the set of all measuring points that have a positive CSD  value, the set of those who have a negative CSD value, and the set of those points having exactly the zero value in CSD space.
+Due to the finite precision of the measurement,
+the sinks and sources sets can make contact with each other.
+Said measurements produce a function over the grid elements (the electrodes).
+Noise between patches corresponding to a single set often makes a ``bridge''
+between components, making identification of poles inexact.
+In order to avoid this we create a ``thick zero'' set of all the recording
+points whose CSD values are inside an error interval around zero.
+This separates more sharply the components of the set,
+represents our ignorance of precise borders and helps
+us to simplify the following analysis.
 
-The ``thick zero'' would help us to sepparate the sinks and sources sets into its disconnected components, which would correspond then to clear patches of definite activity.
-If each one of these patches could be interpreted as an active unit, we may asign to it a ``center of activity'' in order to follow their displacements.
-A Center of Mass for each instantaneous disconnected sink/source would provide us a pragmatic measurement of an ``active locus''.
-We can then calculate at each time frame the CM for each one of those components before tracking displacements of activity.
+The ``thick zero'' would help us to sepparate the sinks and sources sets into
+its disconnected components, which would correspond then to clear patches of
+definite activity. If each one of these patches could be interpreted as an active unit,
+we may asign to it a ``center of activity'' in order to follow their displacements.
+A Center of Mass for each instantaneous disconnected sink/source would provide us
+a pragmatic measurement of an ``active locus''. We can then calculate at each time
+frame the CM for each one of those components before tracking displacements of activity.
 
-At each component at a precise time $t$, indexed $k$, we have a number of points with CSD of the same sign, indexed $j$. Then, the CM of the $k$ component would have the usual definition:
+In each component, indexed $k$, at a given time $t$,, we have a number of points
+with CSD of the same sign, indexed $j$. Then, the CM of the $k$-th
+component would have the usual definition:
 \begin{equation}\label{cmparadisj}
@@
 sign of our ``mass'' eliminates itself graciously.
-These disconnected components, while not rigorously convex, are less spread over the space of the measurements
-and tend to have their CM over the actual active points, or at the border of the active channels.
+These disconnected components, while not rigorously convex,
+are less spread over the space of the measurements
+and tend to have their CM inside the are covered by the active elements
+of the grid.
 
@@
 
-We must not take out of sight the phenomena which is producing the CSD distribution: neurons have a finite size, and the effect of their action a delimited extent.
-Every patch that we can identify as a disconnected component of the Sources/Sink sets indicates a  joint active set of units.
-A center of mass for each disjoint component would then indicate an average locus for this active group, a putative center of activity to which we can give a position, intensity and time.
+We must not take out of sight the phenomena which is producing the CSD distribution:
+neurons have a finite size, and the effect of their action a delimited extent.
+Every patch that we can identify as a disconnected component of the Sources/Sink
+sets indicates a  joint active set of units. A center of mass for each disjoint
+component would then indicate an average locus for this active group, a putative center of activity to which we can give a position, intensity and time.
 An instantaneous snapshot of such information would depict the centers of activity of an ordered structure of neurons and their relative intensities.
@@
 The appearance , displacement, and disappearance of such centers could reveal very subtle details of the dynamics of neural activity.
-Very small, fast displacements of the Centers of mass, may be occur due to slight asynchronous between neurons which belong to the same active group.
-Larger displacements and  longer intervals (more than two or so firing periods) could indicate physiological connections between different groups or units.
+Very small, fast displacements of the Centers of mass, may be occur due to slight asynchronous action between neurons which belong to the same active group.
+Larger displacements and  longer intervals (more than two or so firing periods)
+could indicate physiological connections between different groups or units.
 
-The next step in our analysis is a procedure to systematically associate to each CM at time $t$ a ``successor'' at time $t+\Delta t$, where $\Delta t$ is the sampling time step.
-Let $q_1=(x_1, y_1)$ be the coordinates of the first point and $q_2=(x_2,y_2)$, the corresponding coordinates of the second.
-Remember that these coordinates can be now rational numbers
+The next step in our analysis is a procedure to systematically associate to each CM
+at time $t$ a ``successor'' at time $t+\Delta t$, where $\Delta t$ is the sampling
+time step. Let $q_1=(x_1, y_1)$ be the coordinates of the first point and $q_2=(x_2,y_2)$,
+the corresponding coordinates of the second. Remember that these coordinates can be now
+rational numbers
 respect to the grid of electrodes, as CM are not restricted to be  at the grid points.
@@
 \end{equation}
-The integrated intensity of a CM is simply the denominator in the eq. \ref{cmparadisj}, that is, the total ``mass'' obtained by integrating the CSD in a specific component:
+The integrated intensity of a CM is simply the denominator in the eq. \ref{cmparadisj},
+that is, the total ``mass'' obtained by integrating the CSD in a specific component:
 \begin{equation}
@@
 threshold intensity, we assume that it has ended, so we do not concatenate
-points that are separated in time by more than $\Delta t$. In order to paste different
-trajectories, more rigorous analysis of their origin and interpretation should be used and is
+points that are separated in time by more than $\Delta t$.
+In order to concatenate different
+trajectories, more rigorous analysis of their origin and interpretation
+should be used and is
 beyond the scope of this work.
@@
  \begin{enumerate}
- \item Acquire high density electro-physiological data.
-The data must be two dimensional in space.
-Distance between the recording sites must be of the order of the typical size of the neurons involved or less, and such sites must be on a dense enough grid to perform numerical difference operations over them.
-The sampling frequency should be enough to detect sub-threshold activity, of at least 5kHz.
-Also, the data should come from structured tissue.
-Unordered, highly homogeneous tissue, will not yield significant results.
- \item Calculate the CSD from the data. Any method may be used, but computer expensive techniques are discouraged if the
+ \item Acquire high density electro-physiological data. The data must be two dimensional
+   in space. Distance between the recording sites must be of the order of the typical
+   size of the neurons involved or less, and such sites must be on a dense enough grid to perform numerical difference operations over them.
+   The sampling frequency should be enough to detect sub-threshold activity,
+   of at least 3kHz. Also, the data should come from structured tissue.
+   Unordered, highly homogeneous tissue, will not yield iterpretable results. 
+ \item Calculate the CSD from the data. Any method may be used,
+   but computer expensive techniques are discouraged if the
    number of channels is over the thousands.
-   Common sense and tests are guidelines here. We recommend simpler, difference based approaches wherever possible. If many electrodes fail near the r.o.i., kCSD is a good alternative if done with
-   parallel routines.  If too many electrodes are failing over the r.o.i., it is better to discard the data. 
+   Common sense and tests are guidelines here. We recommend simpler, difference based approaches wherever possible. If many electrodes fail near the r.o.i.,
+   kCSD is a good alternative if done with
+   parallel routines.
+   If too many electrodes are failing over the r.o.i., it is better to discard the data. 
  \item Separate the CSD data into three sets: sources, sinks, and border or indiscernible (below error magnitude) data
-   which we call the zero set.
-Again, sensible criteria have to be defined.
+   which we call the zeros set. Again, sensible criteria have to be defined.
    A rule of thumb would be to use the mean of the peaks of the noise  of the recording as the amplitude for the ``thick zero'' range of values. 
 \item Perform disconnected component analysis in the sources and sinks sets, frame by frame.
-  A single pass algorithm is adequate \cite{Vincent91, Abubaker07}.
+  A single pass algorithm here is  adequate \cite{Vincent91, Abubaker07}.
 \item  Obtain the CM for each component, using eq. \ref{cmparadisj}. Discard
   all CM that are below some significant threshold value. A rule of thumb is one tenth of the average value for all CM,
-  or that are in the lesser decil. 
+  or those who are in the lesser decil. Anything that is around the peak
+  values of the noise integrated over around 5 electrodes would have to be discarded.
 \item For each frame and its successor, and for each CM, perform a ``most probable successor'' detection.
@@
 
-The necessity to create such analysis came originally from the study of electro-physiological data obtained from the rat's hippocampus.
+The necessity to create such analysis came originally from the study of
+electro-physiological data obtained from the rat's hippocampus.
 The highly layered CA structure of the hippocampus provides strong, discernible patterns of Sources and Sinks.
@@
 
-\subsection{Application to facilitated activity}
-
 \subsection{Application to evoked activity in idem.}\label{sec:evocada}
@@
 of the slice of hippocampus with different voltage intensities. Details of the experiment
-have been published elsewhere \cite{FrancoGut2017}. We shall borrow some of
-the best experimental data for our purposes.
+have been published elsewhere \cite{Franco2018}. We shall borrow some of
+the experimental data for our purposes.
 
@@
 electrical stimuli in the same point at constant voltage. The
-response of the cels is allmost identical in each event, which
-allows us to use the average as a denoizing mechanism. After averaging
+response of the cells is allmost identical in each event, which
+allows us to use averages as a denoizing mechanism. After averaging
 the syncronized data we perform kCSDA, as in this case we end up
@@
 with an apparent anterograde movement from CA3 to CA1. The patches change polarity
-as especteed, with the somatic layer becoming a strong sink between 3 to 4 ms
+as espected, with the somatic layer becoming a strong sink between 3 to 4 ms
 after the stimulus, and then, at the 6 ms, inverting its polarity in a way that suggest
@@
     map covers the 8ms after the stimulus is made (shown in the temporal axis by a
-    green doted line). The change of polarity around and over the str. pyr. (between b and
+    green doted line). The change of polarity around and over the
+    str. pyr. (between b and
     d) is clearly visible. In D we show the CSDA for various timestamps in the sub-region
@@
 Small deviations from the trajectory in successive experiments that are above the numerical error could indicate inconsistencies in the network below.
+
+\subsection{Application to facilitated activity}
+
+%%%% *********** Vas aqui
 

### 2019-04-24 ArticuloCSDyCM/Paper03.tex
nuevo jupyter para figuras

@@
 routines. These serve to detect and discard unusable channels,
-such as saturated or failling electrodes. These vary from data set to data set,
+such as saturated or failling electrodes. The
+set of failling electrodes varies from data set to data set,
 as they can be produced by bubbles or by saturation after electrical stimulation
-or other random causes. 
+or other causes. 
 We call this step the elimination of bad channels. We use a simple criterion for this:
-if the absolute value of the signal is above certain threshold (wich must be
-adjusted for evocated vs spontaneous acttivity) or if the standar deviation is
+if the absolute value of the signal is above certain threshold (which must be
+adjusted for evocated vs spontaneous acttivity) or if the standard deviation is
 larger than certain heuristically determined value, the channel is discarded.
@@
 
-Apparently, CSD would not fit the requirements of non-negativity for its use as a mathematical density, but this is not a real problem.
+Apparently, CSD would not fit the requirements of non-negativity for its use as a mathematical density, but this is not a problem.
 CSD separates the measured activity into three distinct sets, namely, the set of all sources, the set of all sinks and the zero set.
@@
 A first visual inspection of the CSD color map at a given instant in time shows that both sources and sinks appear in rather large and significant patches, with typical
-cross lengths around somo 150 $\mu m$.
+cross lengths around 150 $\mu m$.
 

### 2019-04-29 ArticuloCSDyCM/Paper03.tex
figuras tres columnas salieron ya

@@
 
+The figure \ref{lfpcsdcm}  compares the representation of the activity in
+the LFP, CSD and CM spaces. It has been exhaustively proven that the CSD
+representation provides better spatial localization of active sites, but the
+CM can serve as better dinamical tracker of action. The figure ilustrates
+that point by choosing the time interval of response in CA3 after the stimulus.
+CM representation permits us to track the activity with less
+interference from the saturated region. The small, clearly localized centers of
+mass separate clearly from the big patch over the saturated region.
+In the figure \ref{evotracks} we concatenate the succesive centers of mass
+for both sinks and sources and show how the anterograde reaction is
+very well localized over the Pyramidal Stratum at precise moments after
+stimulus.
+
 \begin{figure}[h]
@@
   \label{CSD_Evocada01}
+\end{figure}
+
+\begin{figure}[h]
+  \includegraphics[width=0.9\textwidth]{Figuras/CSD_Evocada01.pdf}
+  \caption{ A comparision of the representations of the displacement of the
+    activity. We show several timeframes after stimulus in three representations.
+    the leftmost column shows the averaged LFP as measured by the BioCAM.
+    In the center column we have the CSD representation, as obtained by
+    the kCSD method. In the right column we superimpose over the CSD representation
+    the Centers of Mass of each disconnected component, for both sinks and
+    sources. In the three columns the saturation of electrodes near
+    the point of stimulus is quite evident. Nevertheless, the CSD representation
+    provides much better contrast and the patches are better localized as espected.
+    The Centers of Mass provide a clear locus of the active subsets of units.
+    The area of the circular dots represent the integrated density. The
+    bigger ones at the right edge cover the saturated electrodes.
+    The anterograde movement of activity follows a high ``sink center'' in the
+    first three rows, and then it is followed by a ``source center'', of aproximately
+    the same integrated ``mass'', that ends spliting in three smaller components.    
+  }\label{lfpcsdcm}
 \end{figure}

### 2019-05-02 ArticuloCSDyCM/Paper03.tex
cambios menores

@@
 
+\usepackage{siunitx} % para unidades chidas como micrometros.
+
 \doublespacing
@@
 \DeclareMathOperator{\arcsinh}{arcsinh}
+\newcommand{\mum}[1]{\SI{#1}{\micro\metre}}
 \newcommand{\rd}{\mathrm{d}}
@@
 
-Spatially tracking succesive activity of electro-physiological data in a concise,
-rigorous and interpretable manner poses a challenging problem. Measurement recordings
-are usually presented first as a series of traces that correspond each to a sampling
+The spatial tracking activity in neural tissue poses chalenging problems,
+if one requires rigour, precision and clear interpretability. 
+Measurement recordings
+are usually presented  as a series of traces that correspond each to a sampling
 electrode on the tissue. Then a series of causality tests are made on the
@@
 separation analysis, such as ICA. 
-Our task in this paper is to present a method for tracing out
+We present here  a method for tracing out
 ``functional local connectivity'', which would bring independent information
-that is complementary to those analysis.  Our  method would be applicable mostly to
-activity in highly structured neural tissue. We shall exemplify it by application to
+that is complementary to blind source separation.
+Our  method would be applicable mostly to
+activity in highly structured neural tissue. We shall exemplify it by
 the data obtained from recordings of the rats hippocampus,
-taking advantage of its lammiar orderer structure.
+taking advantage of its highly stratified  structure.
 
-Local field potential can be regarded as produced by local multiunitary activity.
-A patch of simultaneous, similar, nearby traces can be regarded as one of those
-multiunitary elements. 
+Local field potential (LFP)
+can be regarded as generated by local multiunitary activity.
+A patch of simultaneous, highly correlated, nearby traces
+can be regarded as one of those multiunitary elements. 
 Succesive activation  of distinct neighbouring elements on structured
@@
 neurons that fire in succesion, integrated subthreshold activity and ephaptic coupling
-could all be represented as pathways for transmission of activity. Nevertheless,
+could all be encompased as pathways for transmission of activity. Nevertheless,
 how to define and acquire the putative paths is an open question.
 
-In previous works some form of vector averaging has been proposed and shown to reveal some trajectrories drawn by the activity.
+In previous works some form of vector averaging has been proposed and shown to reveal some trajectories drawn by the activity.
 We shall call such methods Center of Mass Analyses (CMA).
@@
 Our weight is actually the natural density function provided by the data: the Current Source Density.
-This representation not only provides a mathematical rigorous  measure for the weights in the vector averaging process, but helps to separate the data into disjoint components.
-This allows us to track separate process in the same region of interest (r.o.i.) by tracking sepparately the poles of action: influx and eflux of charged ions.
+This representation not only provides a mathematical rigorous  measure for the weights in the vector averaging process, but helps to separate the data into disjoint sets and disconnected components.
+This allows us to track separate process in the same region of interest
+(r.o.i.) by following sepparately the distinct poles of action: influx and
+eflux of charged ions.
 
-The application of the method requires electro-physiological data acquired through bi-dimensional MEAs, although the extension to three dimensional arrays would be possible in principle.
+The application of the method requires electro-physiological data acquired
+through bi-dimensional MEAs, although the extension to three dimensional arrays would be possible in principle.
 The MEA should have dense enough spatial and temporal samplings, so as to obtain good estimates of the CSD and their changes in time.
@@
 As a prime example, non-convex structures which may show simultaneous activity in separate parts could show the robustness of our method.
-We have tested then the method on slices of rat's hippocampus, concentrating in the CA region.
+We have tested then the method on slices of rat's hippocampus, concentrating in the CA3 region.
 We exemplify the usage of the method with data that has been obtained on experiments on rat's hippocampal slices that are centered on the GD and CA
@@
 and data is recorded at sampling rates between 7 and 14 kHz. The richness of the
-data permits us to obtain a prime use case for our method. Downscaling is
-possible in principle, but the array of electrodes should be at least 2D and
+data permits us to obtain a prime use case for our method. A
+ certain amount of downscaling would be 
+possible, but the array of electrodes should be at least 2 dimensional and
 a interelectrode distance of the order of magnitude of a typical pyramidal soma
-(around 100 $\mu$m).
+(around \mum{100}).
 
@@
 
-%% esto puede no ser necesario. 
-A similar but orthogonal criterium can be used for detection of channels with
-spikes.
-
 \subsection{CSD for high density MEAS}
@@
 as it is faster and requires less assumptions.
-We have used both methods, kCSDA where
+We have used kCSDA where
 the difference method is not applicable. The results are
-independent of the method save a for multiplicative factor, which
-is irrelevant in the Center of Mass tracking.
+independent of the method except for a multiplicative factor which
+can be normalized.
 
-Thanks to the density and scales of the BioCAM 4096, a numerical finite difference method
+Thanks to the density and scales of the BioCAM 4096,
+the numerical finite difference method
 produces an excelent approximation to the CSD,
 if the number of failing channels is low or
-are outside the r.o.i.
-For experiments that were made without electrical stimulation is the case, with the order of mangitude of unusable channels in the tens.
+they are outside the r.o.i.
+In the experiments that were made without electrical stimulation that was the case,
+with the order of magnitude of unusable channels in the tens.
 A  numerical Laplacian Operator which reduces the cross effect of rectangular grids is the convolution of the data with the following matrix \cite{Lindberg90}:
+
 \begin{equation}
@@
 \end{equation}
-This sort of operators are extremely sensitive to abrupt changes in the values.
-The noise of the data can be perceived in the spatial domain as rapidly varying edges.
-A Gaussian spatial denoising can be performed before or after the Laplacian operator to the data in order to reduce excessive borders.
-Our Gaussian blur filter has a $\sigma$ value of $126 \mu m$, which corresponds to three times the interelectrode distance.
+This sort of operators are very sensitive to abrupt changes in the values.
+The noise of the data can be perceived in the spatial domain as rapidly varying edges. A Gaussian spatial smoothing
+can be performed before or after the Laplacian operator to the data in order to reduce excessive borders.
+Our Gaussian blur filter has a $ \sigma$ value of \mum{63}, which corresponds to one and a half times the interelectrode distance.
 Intuitively this means that it smooths over each structures smaller than the soma of a typical pyramidal cell. 
@@
 assumption that a finite family of convenient functions can span, with acceptable accuracy, a very good aproximation to the  CSD at each instant in time.
- A general presentation is in the reference \cite{Potworowski2011},we shall recount only the basic implementation that we used.
-One presents a family of model functions for the CSD and
+A general presentation is in the reference \cite{Potworowski2011},we shall recount only the basic implementation as we used it. The
+method can be summarized into two steps:
+\begin{enumerate}
+\item One presents a family of model functions for the CSD and
 obtains the LFP that such functions
-would generate. Then one calculates the coefficients for the entire family by projecting the
+would generate.
+\item Then one procedess to calculate the coefficients for the entire family by projecting the
   experimental data over this ideal LFP model function family.
+\end{enumarate}
+
 We shall borrow the notation in the original paper \cite{Potworowski2011}. Our model function for the CSD is a hard disk function, described by:
@@
 The $R$ parameter is the effective radius of the charge density, which for our
-calculations was taken as half the inter-electrode distance, that is $R=21 \mu m$.
+calculations was taken as half the inter-electrode distance, that is $R=$\mum{21}$.
 Each pair $(x_j, y_j)$ is the center of one such model function, which we take as
@@
   with $h$ being the effective height of the function support, in our case,
-  the half width of the hippocampus slice ($h=150 \mu m$).
+  the half width of the hippocampus slice ($h=$\mum{150}).
 
@@
 
-The attainment of $K$ and $\tilde{K}$ turns out to be computer-costly for high density MEAs, but paraletization, specially on
-GPUs, solves this issue.
+The attainment of $K$ and $\tilde{K}$ turns out to be computer-costly for high density MEAs, but paraletization, specially on GPUs, solves this issue.
 
@@
 A first visual inspection of the CSD color map at a given instant in time shows that both sources and sinks appear in rather large and significant patches, with typical
-cross lengths around 150 $\mu m$.
+cross lengths around \mum{150}. .
 
@@
 we label each disjoint component, obtained by a single pass algorhythm. There are 25 components, and
-the biggest two are not strictly convex but they have tolerable deviations from it. In the
+the bigger two are non convex. In the
 subfigure Biii we indicate the CM of only those components that are above certain threshold of
-integrated intensity (the dot is proportional in area to such intensity, there are five CM to be
+integrated intensity (the circle size  is proportional in area to such intensity, there are five CM to be
 seen). Each CM gets listed as a probable step in a trajectory.
@@
 Due to the finite precision of the measurement,
-the sinks and sources sets can make contact with each other.
-Said measurements produce a function over the grid elements (the electrodes).
-Noise between patches corresponding to a single set often makes a ``bridge''
+it is imposible to define a precise zero value.
+The measurements produce a function over the grid elements (the electrodes).
+Noise between patches corresponding to a single set often makes bridges
 between components, making identification of poles inexact.
-In order to avoid this we create a ``thick zero'' set of all the recording
+In order to avoid this we create a ``thick zero set'' of all the recording
 points whose CSD values are inside an error interval around zero.
@@
 
-The ``thick zero'' would help us to sepparate the sinks and sources sets into
+The ``thick zero set'' would help us to sepparate the sinks and sources sets into
 its disconnected components, which would correspond then to clear patches of
-definite activity. If each one of these patches could be interpreted as an active unit,
+definite activity. If each one of these patches could be
+interpreted as an active unit,
 we may asign to it a ``center of activity'' in order to follow their displacements.
@@
 In each component, indexed $k$, at a given time $t$,, we have a number of points
+(representing electrode centers)
 with CSD of the same sign, indexed $j$. Then, the CM of the $k$-th
@@
 \end{equation}
-where $I(q_{j,k},t)$ is the CSD at point $q_{j,k}=(x_{j,k}, y_{j,k})$ at time $t$, and the
-sign of our ``mass'' eliminates itself graciously.
+where $I(q_{j,k},t)$ is the CSD at point $q_{j,k}=(x_{j,k}, y_{j,k})$ at time $t$,
+and the
+sign of our ``density mass function'' eliminates itself graciously.
 These disconnected components, while not rigorously convex,
-are less spread over the space of the measurements
-and tend to have their CM inside the are covered by the active elements
-of the grid.
+are not largerly spread over the space of the measurements.
 
@@
 time step. Let $q_1=(x_1, y_1)$ be the coordinates of the first point and $q_2=(x_2,y_2)$,
-the corresponding coordinates of the second. Remember that these coordinates can be now
+the corresponding coordinates of the second.
+Remember that these coordinates can be now
 rational numbers
-respect to the grid of electrodes, as CM are not restricted to be  at the grid points.
+respect to the grid of electrodes,
+as CM are not restricted to be  at the grid points.
 The distance is:
@@
 \end{equation}
-The integrated intensity of a CM is simply the denominator in the eq. \ref{cmparadisj},
+The integrated intensity of a CM is simply the denominator in the
+eq. \ref{cmparadisj},
 that is, the total ``mass'' obtained by integrating the CSD in a specific component:
@@
 In order to concatenate different
-trajectories, more rigorous analysis of their origin and interpretation
-should be used and is
-beyond the scope of this work.
+trajectories amore rigorous analysis of their origin and interpretation
+should is due and is beyond the scope of this work.
 
@@
  \begin{enumerate}
- \item Acquire high density electro-physiological data. The data must be two dimensional
+ \item Acquire high density electro-physiological data. The data must be two
+   dimensional
    in space. Distance between the recording sites must be of the order of the typical
@@
    but computer expensive techniques are discouraged if the
-   number of channels is over the thousands.
-   Common sense and tests are guidelines here. We recommend simpler, difference based approaches wherever possible. If many electrodes fail near the r.o.i.,
+   number of channels is above the thousands.
+   Common sense and tests are guidelines here.
+   We recommend simpler, difference based approaches wherever possible.
+   If many electrodes fail near the r.o.i.,
    kCSD is a good alternative if done with
    parallel routines.
-   If too many electrodes are failing over the r.o.i., it is better to discard the data. 
+   If too many electrodes are failing over the r.o.i.,
+   it is better to discard the data. 
  \item Separate the CSD data into three sets: sources, sinks, and border or indiscernible (below error magnitude) data
@@
 The necessity to create such analysis came originally from the study of
-electro-physiological data obtained from the rat's hippocampus.
+electro-physiological data obtained from the rat's hippocampus, using
+a very dense MEA, the BioCAM of 3Brain \cite{BioCAM}. 
 The highly layered CA structure of the hippocampus provides strong, discernible patterns of Sources and Sinks.
 Following the pattern of activity using Center of Mass analysis directly from the data, though, proves unfruitful, due to the non-convex nature of the somatic layer.
-Using CSD and disconnected component tracking over the data provided the right conceptual framework to follow the displacement of active areas during highly active events, such as epileptic-like activity that was pharmacological induced.
-The proposed analysis would not produce vector averages over different simultaneously active areas, on the contrary, it would separate them and track them independently.
+Using CSD and disconnected component tracking over the data provided the right conceptual framework to follow the displacement of active areas during highly active events, such as epileptic-like activity that was pharmacological
+or traumatically induced.
+The proposed analysis would not produce vector averages over different simultaneously active areas.
+On the contrary, it would separate them and track them independently.
 So the most violent epileptic burst provided us with the data necessary to prove the robustness of the method.
@@
 have been published elsewhere \cite{Franco2018}. We shall borrow some of
-the experimental data for our purposes.
+the experimental data for illustrating our purposes.
 
@@
 with many saturated electrodes and dCDSA is not viable.
-Once that we have the CSD representation, we select noisy channels outside
+In the figure \ref{lfpevocada} we show some frames of the post-stimulus
+acticity as a heat map, along
+with a diagram of the somatic layers and the
+stimulus point. The saturated area is quite large, and covers
+a large section of the GD structure. In this case we shall
+restrict the CM track analysis to the respondent area in CA3.
+
+Once that we have the CSD representation, we select channels outside
 the r.o.i. and use them to obtain the threshold for sepparating
 the Sink and Sources sets. Then we procede as stated in the previous section.
-
-An example of the procedure and the treatment of the data is shown in figure
-\ref{CSD_Evocada01}.  The subfigure A  shows a diagram of the hippocampus, with
-the somatic layers (CA and DG) as thick black lines and the stimulus point
-shown as a blue arrow. The $a,b,c,d,e$ dots are selected electrodes for
-showing LFP and CSD traces. Those are shown in the subfigure B. In subfigure
-C the same data is presented as a spatiotemporal color map that
-shows the expected poles of firing neurons transverse to the
-pyramidal stratum.  In the subfigure D we show the CSD for a series of time frames
-over the square delimited in the subfigure A. The idea for treating the different
-patches separately becomes evident here. The sources and sinks form patches
-with an apparent anterograde movement from CA3 to CA1. The patches change polarity
-as espected, with the somatic layer becoming a strong sink between 3 to 4 ms
-after the stimulus, and then, at the 6 ms, inverting its polarity in a way that suggest
-successive activation of groups that are represented as movement of the patch.
 
@@
 stimulus.
+
+The treatment of the data is shown in figure
+\ref{CSD_Evocada01}.  The subfigure A  shows a diagram of the hippocampus, with
+the somatic layers (CA and DG) as thick black lines and the stimulus point
+shown as a blue arrow. The $a,b,c,d,e$ dots are selected electrodes for
+showing LFP and CSD traces. Those are shown in the subfigure B. In subfigure
+C the same data is presented as a spatiotemporal color map that
+shows the expected poles of firing neurons transverse to the
+pyramidal stratum.  In the subfigure D we show the CSD for a series of time frames
+over the square delimited in the subfigure A. The idea for treating the different
+patches separately becomes evident here. The sources and sinks form patches
+with an apparent anterograde movement from CA3 to CA1.
+The patches change polarity
+as espected, with the somatic layer becoming a strong sink between 3 to 4 ms
+after the stimulus, and then, at the 6 ms, inverting its polarity in a way that suggest
+successive activation of groups that are represented as movement of the patch.
 

### 2019-06-25 Graficando/ploteaunchingocsd.jl
plotalot

@@ -1,47 +1,84 @@
 using HDF5
 using PyPlot
+using Statistics
 
 
 nombre=ARGS[1]
 
 arxivo=h5open(nombre)
 elementos=names(arxivo)
-
 csdlind=read(arxivo["CSDALindenberg"])
+
+sigma=std(csdlind)
+mu=mean(csdlind)
+
+function subruido(x,u)
+    result=0
+    if -u <x < u
+        result=0
+     else
+        result=x
+    end
+    return result
+ end
+
 (alto,ancho,nmax)=size(csdlind)
+intervalo=1:5:nmax
 #(xmin, xmax)=read(arxivo["cols"])
 #(ymin, ymax)=read(arxivo["rengs"])
 freq=read(arxivo["freq"])
 
-(mincsd, maxcsd)=extrema(csdlind)
-limites=min(abs(mincsd),abs(maxcsd))/1.5
 
+if length(ARGS) == 1
+    nomout="csdplot_"
+else
+    nomout=ARGS[2]
+end
+
+
+ini=floor(Int, 400*freq)
+fin=floor(Int, 500*freq)
+
+
+(mincsd, maxcsd)=extrema(csdlind)
+#limites=min(abs(mincsd),abs(maxcsd))/1.25
+limites=5*sigma
+
+close(arxivo)
+
+println(limites, " es la marca de color csd maximo/minimo")
+
+csdadulterada=map(x->subruido(x,2.5*sigma), csdlind) 
+
+
+csdlind=0
 #for interpol in lista
 
 
-for n =1:nmax
+#for n =1:5:nmax
+for n=ini:fin
 
     xlim(0,ancho+1)
     ylim(0, alto+1)
     
     fafa=figure(figsize=(4,4))
-    ejemplo=csdlind[:,:,n]
+    ejemplo=csdadulterada[:,:,n]
     tiempo=round(Int,n/freq)
 
     title("t = $tiempo ms")
     
    
     imshow(ejemplo, origin="lower",
-           cmap="coolwarm", interpolation="nearest",
+           cmap="coolwarm", interpolation="bicubic",
            vmin=-limites, vmax=limites,extent=[0.5,ancho+0.5,0.5,alto+0.5])
 
     nstring=lpad(n, 8, "0")
     
-    nomine="csd_eduardo_193005_control_01_$nstring.png"
+    nomine=nomout*"$nstring.png"
 
     savefig(nomine, dpi=92)
     close(fafa)
[…]

### 2019-06-27 ArticuloCSDyCM/Paper03.tex
papeleos

@@
 On previous works it has been argued that numerical difference operators sacrifice all data on the borders of the array.
-For electrode arrays of lesser density, this could posse a problem, but in our case we only renounce to 27 out of 4095 channels (one is grounded), and those are outside the region of interest.
+For electrode arrays of lesser density, this could posse a problem, but in our case we renounce to 253 out of 4095 channels (one is grounded), and those are outside the region of interest.
 If very few electrodes fail, more sophisticated methods such as iCSDA \cite{Leski2011} and kCSDA \cite{Potworowski2011} prove little advantage in estimating the sources under the ohmic, isotropic and homogeneous assumptions.
@@
 would generate.
-\item Then one procedess to calculate the coefficients for the entire family by projecting the
+\item Then one procedess to calculate
+  the coefficients for the entire family by projecting the
   experimental data over this ideal LFP model function family.
@@
 The $R$ parameter is the effective radius of the charge density, which for our
-calculations was taken as half the inter-electrode distance, that is $R=$\mum{21}$.
+calculations was taken as half the inter-electrode distance, that is $R=$\mum{21}.
 Each pair $(x_j, y_j)$ is the center of one such model function, which we take as
@@
 \begin{equation}
-  C=\frac{\partial \rho}{\partial t}
+  C=-\frac{\partial \rho}{\partial t}
 \end{equation}

### 2019-07-02 ArticuloCSDyCM/Paper03.tex
versiones del paper merged

@@
   experimental data over this ideal LFP model function family.
-\end{enumarate}
+\end{enumerate}
 
@@
 \begin{figure}
-  \includegraphics[width=0.9\textwidth]{Figuras/Disjuntos01.pdf}
+  \includegraphics[width=0.9\textwidth]{Figuras/ExplicaDisconexos01.pdf}
   \caption{ Explanation of the geometrical concepts used on our analysis.
-    In subfigure A we show
+    In the Row A of fiugres
     with an artificial example the concepts of connectedness and convexivity.
@@
     In Figure Bii we take the Sink Set and separate it into disconnected components,
-    they are labeled from one to 25 in this particular example.
-    From this subsets we calculate the total mass of each one and only
+    they are labeled from one to 25 in this particular example,
+    indicated by color
+    For each subset we sum over its mass, and 
     considerate for the next step those who are above certain threshold. The
@@
 \begin{figure}[h]
-  \includegraphics[width=0.9\textwidth]{Figuras/CSD_Evocada01.pdf}
+  \includegraphics[width=0.9\textwidth]{Figuras/CSD_EvocadaYTrazos.pdf}
   \caption{ CSDA for evoked activity and visual inspection of disjoint
@@
     thick black lines, and the stimulus site is indicated by a blue arrow. A selection
-    of six contiguous electrodes labeled $a,b,c,d,e$ in the r.o.i. are presented as blue
+    of six contiguous electrodes labeled $A-F$ in the r.o.i. are presented as blue
     dots. In B we show the LFP traces and their CSD representation for those
-    selected electrodes. In C we present the same information as a color map,
-    with blue for sinks and red for sources, and a shorter time interval. The
-    map covers the 8ms after the stimulus is made (shown in the temporal axis by a
-    green doted line). The change of polarity around and over the
-    str. pyr. (between b and
-    d) is clearly visible. In D we show the CSDA for various timestamps in the sub-region
+    selected electrodes.
+    In C we shoe the CSDA for various timestamps in te sub-region
     indicated in A as the r.o.i, which covers a square of 24 by 24 electrodes over CA3.
-    in the first plot, 1ms before the stimulus, we show the grid of electrodes and a
+    in the first image,  4ms after the stimulus,  a
     shaded strip indicating the approximate location of str. pyr. A model neuron is
@@
 
-\begin{figure}[h]
-  \includegraphics[width=0.9\textwidth]{Figuras/CSD_Evocada01.pdf}
-  \caption{ A comparision of the representations of the displacement of the
-    activity. We show several timeframes after stimulus in three representations.
-    the leftmost column shows the averaged LFP as measured by the BioCAM.
-    In the center column we have the CSD representation, as obtained by
-    the kCSD method. In the right column we superimpose over the CSD representation
-    the Centers of Mass of each disconnected component, for both sinks and
-    sources. In the three columns the saturation of electrodes near
-    the point of stimulus is quite evident. Nevertheless, the CSD representation
-    provides much better contrast and the patches are better localized as espected.
-    The Centers of Mass provide a clear locus of the active subsets of units.
-    The area of the circular dots represent the integrated density. The
-    bigger ones at the right edge cover the saturated electrodes.
-    The anterograde movement of activity follows a high ``sink center'' in the
-    first three rows, and then it is followed by a ``source center'', of aproximately
-    the same integrated ``mass'', that ends spliting in three smaller components.    
-  }\label{lfpcsdcm}
-\end{figure}
+ \includegraphics[width=0.9\textwidth]{Figuras/CMVariousTimes01.pdf}
++  \caption{ The CM analysis over the same images as figure \ref{CSD_Evocada01} C. Some trajectories
++    are calculated before the 4ms time stamp. The triangular mark shows the later point of the
++    trajectory, either the end or the last calculated position until that moment. The circular
++    mark is the start of the trajectory. We can see that activity moves form CA3 distal to
++    proximal in alternating waves of action (hiperpolarization, firing and repolarization).
++  }\label{cmevocada}
 

### 2019-07-05 CenterOfMass/CentrosdeMasa.jl
grafs trays

@@ -1,3 +1,6 @@
+module CentrosdeMasa
+
+
 #=
 Solo el calculo del centro de masa a partir
 de un JLD que contenga un arreglo 3D que
@@ -9,6 +12,8 @@ esten en cero.
 
 
 # Funciones
+
+export vecindad8, ComponentesSP, ObtenComponentesyCM
 
 function vecindad8(punto::Array)
     j=punto[1]
@@ -163,3 +168,4 @@ function ObtenComponentesyCM(Datos::Array, tini=1,tfini=tmax, epsilon=1.0)
     
        
 
+end # modulo

### 2019-07-09 ArticuloCSDyCM/Paper03.tex
grafos trays

@@
 \newcommand{\mum}[1]{\SI{#1}{\micro\metre}}
+\newcommand{\muV}[1]{\SI{#1}{\micro\volt}}
 \newcommand{\rd}{\mathrm{d}}
@@
 \subsection{Application to epileptic-like activity in the hippocampus}
-Chronologically speaking, we performed first the analysis to the most complicated data (numerically and qualitatively), which where a series of epileptic burst that occurred after 15 minutes of application of 4AP to living slices of hippocampus.
-The burst occurred with activity spreading over many sites simultaneously.
+We performed first the analysis to the most complicated data, numerically and qualitatively, which comes from a series of epileptic burst that occurred after 15 minutes of application of 4AP to living slices of hippocampus.
+Each burst occurrs with activity spreading over many sites allmost simultaneously.
 This turned out to be a test of the applicability and robustness of our method.
-The necessity of having good error value for the zero set was more apparent here than anywhere else.
-Also the posibilty to create new structures of numerical data was first seen here.
-But also the most strickig features of the propagation of activity, coded now as trajectories of the CM, could be seen here, as the epileptic burst covered the whole CA structure.
-A sample of the raw data in $\mu V$ can be seen in the figure \textbf{Tururu}.
-Three key moments in the epileptic burst have been selected, one before the onset of the attack, one at the peak and one at the middle of the subsequent waves.
-Traces from selected electrodes along the line labeled A-F are presented in the subfigure D.
+The necessity of having a good error value for the inactive set was more apparent here than anywhere else.
+Aulso the most strickig features of the propagation of activity, coded now as trajectories of the CM, could be seen here, as the epileptic burst covered the whole CA structure.
+A sample of the raw data in $\mu V$ can be seen in the figure \ref{lfpycsd4ap}.
+Five key moments in the epileptic burst have been selected, one just before the onset of the attack, the next one at the peak and one  following it, and the last at the subsiding moments.
+The figure compares the Raw data
+in $\muV{}$ in the first row, then the crude CSD representation in the second row, and
+in the third raw a CSD after a Gaussian Blur Filter has been applied,
+with a $\sigma$ value of $ \mum{42}$. After the Gaussian Blur acts a sort of
+spatial denoizing mechanism, the disjoint components are apparent to the eye.
+In the figure \ref{csdfrontera} we apply the border criterion for the disconnected
+components. Again, in the first row we have the CSD without the Gaussian Blur.
+Even though well localizated patches of activity are to be seen over the
+stratum pyramidale, large parts of the activity remain hidden by the noice.
+After the Gaussian Blur has been applied the joint sets of activity are evident.
 
-In figure \textbf{Tururu2} we present the CSD for the same selected moments.
+\begin{figure}
+  \includegraphics[width=0.9\textwidth]{./Figuras/LFPvsCSD_4AP.pdf}
+  \caption{Comparison of LFP and CSD representations
+    at the moments of the epileptic burst. In the first row
+    we present the data as meassured, no denoizing or filtering has
+    been applied. The units of measurement are in $\muV{}$. In
+    the middle row a direct application of a difference operator
+    gives us the CSD. This sort of operators are very sensitive
+    to noice, in this case, spatial hard edges, which become
+    greatly exagerated. In the third row we present the same
+    data after a Gaussian spatial filter ( Gaussian Blur) has
+    been applied. }\label{lfpycsd4ap}
+\end{figure}
 
-In the figure \textbf{Tururu3} we present our method using the peak moment as an example.
+\begin{figure}
+  \includegraphics[width=0.9\textwidth]{./Figuras/Fronteras2sigma.png}
+  \caption{The detection of disconnected components for the Sink and Source
+    Sets. We take the the threshold value as two times the
+    standard deviation of the whole data in CSD representation. In
+    the upper row the data without the Gaussian Blur filtering is shown.
+    Altough the location of some of the active sets is clear, a lot of the
+    joint activity gets hidden under the spatial noice. In the lower
+    row we repeat the detection with the smoothed data. Large, coherent
+    patches of activity are clearly shown, and the structure of the
+    stratum pyramidale is apparent. 
+  }\label{csdfrontera}
+\end{figure}
 

### 2019-07-09 CenterOfMass/TrayectoriasAux.jl

@@ -4,6 +4,10 @@ de trayectorias) y que quitan puntos "antes que" o "despues que"
 cierto tiempo
 =#
 
+module TrayectoriasAux
+
+export reducecatenario, AntesQue, DespuesQue, distprom
+
 dist(x1,y1,x2,y2)=sqrt((x1-x2)^2+(y1-y2)^2)
 
 function reducecadetenario(Catenario::Dict, gordmin=25,minlong=7,maxlong=100, distmin=0)
@@ -64,3 +68,5 @@ function distprom(Catenario::Dict)
     end
     return result
 end
+
+end # module

### 2019-07-10 CenterOfMass/TrayectoriasAux.jl
catenarios

@@ -10,19 +10,24 @@ export reducecatenario, AntesQue, DespuesQue, distprom
 
 dist(x1,y1,x2,y2)=sqrt((x1-x2)^2+(y1-y2)^2)
 
-function reducecadetenario(Catenario::Dict, gordmin=25,minlong=7,maxlong=100, distmin=0)
+function reducecatenario(Catenario::Dict;
+                         gordmin=25,minlong=7,maxlong=100, distmin=0)
+    
     result=Dict{Integer, Array{Any}}()
     
-for (k,p) in Catenario
+    for (k,p) in Catenario
     
         distrec=dist(p[1,1],p[1,2],p[end,1],p[end,2])
         
-    gordura=abs.(p[:,3])
-    longus,gordus=size(p)
-    if (mean(gordura)>gordmin) && (longus>minlong) && longus < maxlong && distrec>distmin
+        gordura=abs.(p[:,3])
+        longus,gordus=size(p)
+        if (mean(gordura)>gordmin) &&
+            (longus>minlong) &&
+            longus < maxlong && distrec>distmin
             result[k]=p
+        end
     end
-end
+
     return result
 end
 

### 2019-07-17 ArticuloCSDyCM/Paper03.tex
articulo con mas carnita

@@
 
-The spatial tracking activity in neural tissue poses chalenging problems,
-if one requires rigour, precision and clear interpretability. 
+The spatial tracking activity in neural tissue poses challenging problems,
+if one requires rigor, precision and clear interpretability. 
 Measurement recordings
@@
 electrode on the tissue. Then a series of causality tests are made on the
-signal to search for possible efective conections. 
+signal to search for possible effective connections. 
 As Micro Electrode Arrays (MEAs) become denser, that representation
@@
 Local field potential (LFP)
-can be regarded as generated by local multiunitary activity.
+can be regarded as generated by local multi-unitary activity.
 A patch of simultaneous, highly correlated, nearby traces
-can be regarded as one of those multiunitary elements. 
-Succesive activation  of distinct neighbouring elements on structured
+can be regarded as one of those multi-unitary elements. 
+Successive activation  of distinct neighboring elements on structured
 tissue would trace a
 ``trajectory'' of the activity. A tracing of such trajectories would help to indicate
-causal or correlated ``functional local conections'' over the tissue. Neighbouring
-neurons that fire in succesion, integrated subthreshold activity and ephaptic coupling
-could all be encompased as pathways for transmission of activity. Nevertheless,
+causal or correlated ``functional local connections'' over the tissue. Neighboring
+neurons that fire in succession, integrated sub-threshold activity and ephaptic coupling
+could all be encompassed as pathways for transmission of activity. Nevertheless,
 how to define and acquire the putative paths is an open question.
@@
 We shall call such methods Center of Mass Analyses (CMA).
-The ``wheights'' that have been chosen for obtaining the average have been selected on heuristic grounds.
+The ``weights'' that have been chosen for obtaining the average have been selected on heuristic grounds.
 In the works of Chao \emph{et al.} \cite{Chao05, Chao07}, the density of spikes in a certain time interval was used, a sort of ``averaged density of action'', while Manjarrez \emph{et al.} \cite{Manjarrez07, Manjarrez09} used the positive half of the LFP recording as their weight.
@@
 This allows us to track separate process in the same region of interest
-(r.o.i.) by following sepparately the distinct poles of action: influx and
-eflux of charged ions.
+(r.o.i.) by following separately the distinct poles of action: influx and
+efflux of charged ions.
 
@@
 data permits us to obtain a prime use case for our method. A
- certain amount of downscaling would be 
+ certain amount of down-scaling would be 
 possible, but the array of electrodes should be at least 2 dimensional and
-a interelectrode distance of the order of magnitude of a typical pyramidal soma
+a inter-electrode distance of the order of magnitude of a typical pyramidal soma
 (around \mum{100}).
@@
 
-\subsection{Preprocessing}
+\subsection{Prepossessing}
 
-After aquiring data from experiment, we pass the data through some preprocessing
+After acquiring data from experiment, we pass the data through some Prepossessing
 routines. These serve to detect and discard unusable channels,
-such as saturated or failling electrodes. The
-set of failling electrodes varies from data set to data set,
+such as saturated or failing electrodes. The
+set of fail-ling electrodes varies from data set to data set,
 as they can be produced by bubbles or by saturation after electrical stimulation
@@
 if the absolute value of the signal is above certain threshold (which must be
-adjusted for evocated vs spontaneous acttivity) or if the standard deviation is
+adjusted for evoked vs spontaneous activity) or if the standard deviation is
 larger than certain heuristically determined value, the channel is discarded.
@@
 the numerical finite difference method
-produces an excelent approximation to the CSD,
+produces an excellent approximation to the CSD,
 if the number of failing channels is low or
@@
 can be performed before or after the Laplacian operator to the data in order to reduce excessive borders.
-Our Gaussian blur filter has a $ \sigma$ value of \mum{63}, which corresponds to one and a half times the interelectrode distance.
+Our Gaussian blur filter has a $ \sigma$ value of \mum{63}, which corresponds to one and a half times the inter-electrode distance.
 Intuitively this means that it smooths over each structures smaller than the soma of a typical pyramidal cell. 
@@
 becomes necessary. The kCSDA method relies on the
-assumption that a finite family of convenient functions can span, with acceptable accuracy, a very good aproximation to the  CSD at each instant in time.
+assumption that a finite family of convenient functions can span, with acceptable accuracy, a very good approximation to the  CSD at each instant in time.
 A general presentation is in the reference \cite{Potworowski2011},we shall recount only the basic implementation as we used it. The
@@
 would generate.
-\item Then one procedess to calculate
+\item Then one proceeds to calculate
   the coefficients for the entire family by projecting the
@@
 Center of Mass Analysis (CMA) has been used previously in order to track putative  ``trajectories'' of the activity or information across neural tissue or even the whole brain \cite{Chao05, Chao07, Manjarrez07, Manjarrez09}.
-In such works the chosen ``mass'' was an heuristic measure adecuate for representing the corse-grain picture of displacement of activity.
-These present the problem of justifing the choice made on practical grounds.
-For calculating a vectorial average, the adecuate wheight per vector is a density, a positive quantity that measures the concentration of something per space unit, so it is only adequate that we search for a natural density in the electrophysiological experiments.
+In such works the chosen ``mass'' was an heuristic measure adequate for representing the coarse-grain picture of displacement of activity.
+These present the problem of justifying the choice made on practical grounds.
+For calculating a vectorial average, the adequate weight per vector is a density, a positive quantity that measures the concentration of something per space unit, so it is only adequate that we search for a natural density in the electrophysiological experiments.
 The Current Source Density fits the mathematical and interpretative requirements.
@@
 equal and separated by a long distance (this is not an implication: a set can be disconnected and have
-a CM inside the set). For the set $C$ our aproach has been used: we treat every disconnected component of the
+a CM inside the set). For the set $C$ our approach has been used: we treat every disconnected component of the
 set as an independent entity and calculate the center of mass for each one. Notice that this doesn't preclude t
-he posibility of having a concave subset with a CM that lies outside. But, as we shall see,
-as the electrode resolution is finite, the posibility of having highly non compact shapes is unlikely, and they
+he possibility of having a concave subset with a CM that lies outside. But, as we shall see,
+as the electrode resolution is finite, the possibility of having highly non compact shapes is unlikely, and they
 tend to have a CM inside or near their borders. In the figure \ref{disconnectedsets} B we show this with experimental
 data (we are using evoked activity data, as explained in section \ref{sec:evocada}). In graph Bi we
-sepparate the CSD data into the sink and sources set. We take sinks as example. In the graph Bii
-we label each disjoint component, obtained by a single pass algorhythm. There are 25 components, and
+separate the CSD data into the sink and sources set. We take sinks as example. In the graph Bii
+we label each disjoint component, obtained by a single pass algorithm. There are 25 components, and
 the bigger two are non convex. In the
@@
   \caption{ Explanation of the geometrical concepts used on our analysis.
-    In the Row A of fiugres
-    with an artificial example the concepts of connectedness and convexivity.
+    In the Row A of figures
+    with an artificial example the concepts of connectedness and convexity.
     The subfigure Ai shows
@@
 Due to the finite precision of the measurement,
-it is imposible to define a precise zero value.
+it is impossible to define a precise zero value.
 The measurements produce a function over the grid elements (the electrodes).
@@
 
-The ``thick zero set'' would help us to sepparate the sinks and sources sets into
+The ``thick zero set'' would help us to separate the sinks and sources sets into
 its disconnected components, which would correspond then to clear patches of
@@
 interpreted as an active unit,
-we may asign to it a ``center of activity'' in order to follow their displacements.
+we may assign to it a ``center of activity'' in order to follow their displacements.
 A Center of Mass for each instantaneous disconnected sink/source would provide us
@@
 These disconnected components, while not rigorously convex,
-are not largerly spread over the space of the measurements.
+are not largely spread over the space of the measurements.
 
@@
 In order to concatenate different
-trajectories amore rigorous analysis of their origin and interpretation
+trajectories more rigorous analysis of their origin and interpretation
 should is due and is beyond the scope of this work.
@@
    of at least 3kHz. Also, the data should come from structured tissue.
-   Unordered, highly homogeneous tissue, will not yield iterpretable results. 
+   Unordered, highly homogeneous tissue, will not yield interpretable results. 
  \item Calculate the CSD from the data. Any method may be used,
@@
 \item For each frame and its successor, and for each CM, perform a ``most probable successor'' detection.
-Here we must put numerical criteria according to the data avaible, i.e.
-how far away and how different in intensity should be allowed a CM to be at time $t+1$ to be considered succesor in time of another at time $t$.
-\item ``Connect the dots'', that is, apply succesively the step above for each CM, until it stops having a succesor or its intensity falls below the error criteria.
-Then plot the succesive CMs as one trayectory.
-\item Plot all the trayectories and sort visually those that may have physiological interpretation from those who may be artifacts of the method. This may need some form of expert validation.
+Here we must put numerical criteria according to the data available, i.e.
+how far away and how different in intensity should be allowed a CM to be at time $t+1$ to be considered successor in time of another at time $t$.
+\item ``Connect the dots'', that is, apply successively the step above for each CM, until it stops having a successor or its intensity falls below the error criteria.
+Then plot the successive CMs as one trajectory.
+\item Plot all the trajectories and sort visually those that may have physiological interpretation from those who may be artifacts of the method. This may need some form of expert validation.
  \end{enumerate}
@@
 
-The LFP is averaged over three events, produced by three succesive
+The LFP is averaged over three events, produced by three successive
 electrical stimuli in the same point at constant voltage. The
-response of the cells is allmost identical in each event, which
-allows us to use averages as a denoizing mechanism. After averaging
-the syncronized data we perform kCSDA, as in this case we end up
+response of the cells is almost identical in each event, which
+allows us to use averages as a denoising mechanism. After averaging
+the synchronized data we perform kCSDA, as in this case we end up
 with many saturated electrodes and dCDSA is not viable.
 In the figure \ref{lfpevocada} we show some frames of the post-stimulus
-acticity as a heat map, along
+activity as a heat map, along
 with a diagram of the somatic layers and the
@@
 Once that we have the CSD representation, we select channels outside
-the r.o.i. and use them to obtain the threshold for sepparating
+the r.o.i. and use them to obtain the threshold for separating
 the Sink and Sources sets. Then we procede as stated in the previous section.
@@
 representation provides better spatial localization of active sites, but the
-CM can serve as better dinamical tracker of action. The figure ilustrates
+CM can serve as better dynamical tracker of action. The figure illustrates
 that point by choosing the time interval of response in CA3 after the stimulus.
@@
 mass separate clearly from the big patch over the saturated region.
-In the figure \ref{evotracks} we concatenate the succesive centers of mass
+In the figure \ref{evotracks} we concatenate the successive centers of mass
 for both sinks and sources and show how the anterograde reaction is
@@
 The patches change polarity
-as espected, with the somatic layer becoming a strong sink between 3 to 4 ms
+as expected, with the somatic layer becoming a strong sink between 3 to 4 ms
 after the stimulus, and then, at the 6 ms, inverting its polarity in a way that suggest
@@
     selected electrodes.
-    In C we shoe the CSDA for various timestamps in te sub-region
+    In C we shoe the CSDA for various timestamps in the sub-region
     indicated in A as the r.o.i, which covers a square of 24 by 24 electrodes over CA3.
@@
 We performed first the analysis to the most complicated data, numerically and qualitatively, which comes from a series of epileptic burst that occurred after 15 minutes of application of 4AP to living slices of hippocampus.
-Each burst occurrs with activity spreading over many sites allmost simultaneously.
+Each burst occurs with activity spreading over many sites almost simultaneously.
 This turned out to be a test of the applicability and robustness of our method.
 The necessity of having a good error value for the inactive set was more apparent here than anywhere else.
-Aulso the most strickig features of the propagation of activity, coded now as trajectories of the CM, could be seen here, as the epileptic burst covered the whole CA structure.
+Also the most striking features of the propagation of activity, coded now as trajectories of the CM, could be seen here, as the epileptic burst covered the whole CA structure.
 A sample of the raw data in $\mu V$ can be seen in the figure \ref{lfpycsd4ap}.
@@
 with a $\sigma$ value of $ \mum{42}$. After the Gaussian Blur acts a sort of
-spatial denoizing mechanism, the disjoint components are apparent to the eye.
+spatial denoising mechanism, the disjoint components are apparent to the eye.
 In the figure \ref{csdfrontera} we apply the border criterion for the disconnected
 components. Again, in the first row we have the CSD without the Gaussian Blur.
-Even though well localizated patches of activity are to be seen over the
-stratum pyramidale, large parts of the activity remain hidden by the noice.
+Even though well localized patches of activity are to be seen over the
+stratum pyramidale, large parts of the activity remain hidden by the noise.
 After the Gaussian Blur has been applied the joint sets of activity are evident.
+
+In the figure \ref{tray4ap} we show the attack on the trajectories representation. Fine lines
+connect the blue (sink) and red (source) dots, that indicate the trajectories along various activation
+sites around the whole CA structure, and part of GD. Very small trajectories precede
+the onset of the epileptiform attack in CA1. At around 60ms after the 4AP dose a burst
+of localized trajectories appear near CA2. By the 70ms mark, the activity has spread
+retrograde through most of the CA structure. Between the 80 and 90 ms marks
+GD shows knot-like trajectories. By 140ms the remaining activity is in the proximal
+area of CA1.
 
@@
     at the moments of the epileptic burst. In the first row
-    we present the data as meassured, no denoizing or filtering has
+    we present the data as measured, no denoizing or filtering has
     been applied. The units of measurement are in $\muV{}$. In
@@
     gives us the CSD. This sort of operators are very sensitive
-    to noice, in this case, spatial hard edges, which become
-    greatly exagerated. In the third row we present the same
+    to noise, in this case, spatial hard edges, which become
+    greatly exaggerated. In the third row we present the same
     data after a Gaussian spatial filter ( Gaussian Blur) has
@@
 \begin{figure}
-  \includegraphics[width=0.9\textwidth]{./Figuras/Fronteras2sigma.png}
+  \includegraphics[width=0.9\textwidth]{./Figuras/Fronteras2sigma.pdf}
   \caption{The detection of disconnected components for the Sink and Source
@@
     the upper row the data without the Gaussian Blur filtering is shown.
-    Altough the location of some of the active sets is clear, a lot of the
-    joint activity gets hidden under the spatial noice. In the lower
+    Although the location of some of the active sets is clear, a lot of the
+    joint activity gets hidden under the spatial noise. In the lower
     row we repeat the detection with the smoothed data. Large, coherent
@@
   }\label{csdfrontera}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=0.9\textwidth]{./Figuras/SeqTray4AP01.pdf}
+  \caption{ The activity in the Center of Mass and Trajectories representation.
+    A single epileptic attack is shown from slightly before the onset until the
+    last wave of the attack. Sink and Source trajectories are shown in blue
+    and red, respectively. In order to not saturate the latter subfigures
+    we take a persistence value of 2 ms, meaning that we do not shown
+    trajectories 2 ms after they have ended.}\label{tray4ap}.
  \end{figure}

### 2019-07-17 FFTWandsuch/FFTWtoolkit.jl

@@ -8,7 +8,7 @@ export filtrobanda, wfft
 function indicecercano(a::Number, freqs::Array)
     #encuentra en índice que tiene el número  más cercano a "a" en el Array freqs
     # si a está más lejos de alguno de ellos que su paso
-    tol=abs((max(freqs)-min(freqs))/length(freqs))
+    tol=abs((maximum(freqs)-minimum(freqs))/length(freqs))
     result=findfirst(x->isapprox(a,x, atol=tol), freqs)
     return result
 end

### 2019-07-17 Preprocesamiento/ActividadSutil.jl

@@ -33,7 +33,7 @@ function buscasigmaactiva(datos::Array; malos=Set(), venms=50, freq=deffreq)
     return pur
 end
 
-end.
+end
 
 
 

### 2019-07-17 Preprocesamiento/PreprocTools.jl

@@ -115,7 +115,7 @@ function FormaMatrizDatosCentrados(xxs::Array, factor::Number)
     irrrelevante,largo=size(xxs)
     #Los datos originales son UInt16. No podemos
     # tener mas bits que eso. No sirve de nada.
-    aux=Array{Int16}(undef, 64,64, largo);
+    aux=Array{UInt16}(undef, 64,64, largo);
     result=Array{Float32}(undef, 64,64, largo);
     for j=1:64,k=1:64
         aux[k,j,:]=xxs[j+(k-1)*64,:]

### 2019-07-17 Preprocesamiento/rebanador.jl

@@ -28,7 +28,7 @@ factor=Datos["factor"] #Factor de conversion de numeros enteros a microVolts
 
 if autorebana
     tamax=400*1024*1024  #maximo tamaño tolerable de archivo en tu lap
-    tam=filesize(arbestring)
+    tam=filesize(abrestring)
     cachos=div(tam,tamax)+1
 else
     #tiempos interesantes es una lista de intervalos en segundos. 

### 2019-07-18 ArticuloCSDyCM/Paper03.tex
mas figuras

@@
 Having both
-the mathematician and electrophysiologicist interpretations presents we can
+the mathematician and electrophysiologist interpretations presents we can
 perform the following analysis with clear readings.
@@
 The
-set $B$ has also a center of mass outside the set. It has two disconnected components that are aproximatelly
+set $B$ has also a center of mass outside the set. It has two disconnected components that are approximately
 equal and separated by a long distance (this is not an implication: a set can be disconnected and have
@@
 
+%The treatment of the data is shown in figure
+%\ref{CSD_Evocada01}.  The subfigure A  shows a diagram of the hippocampus, with
+%the somatic layers (CA and DG) as thick black lines and the stimulus point
+%shown as a blue arrow. The $a,b,c,d,e$ dots are selected electrodes for
+%showing LFP and CSD traces. Those are shown in the subfigure B. In subfigure
+%C the same data is presented as a spatiotemporal color map that
+%shows the expected poles of firing neurons transverse to the
+%pyramidal stratum.  In the subfigure D we show the CSD for a series of time frames
+%over the square delimited in the subfigure A. The idea for treating the different
+%patches separately becomes evident here. The sources and sinks form patches
+%with an apparent anterograde movement from CA3 to CA1.
+%The patches change polarity
+%as expected, with the somatic layer becoming a strong sink between 3 to 4 ms
+%after the stimulus, and then, at the 6 ms, inverting its polarity in a way that suggest
+%successive activation of groups that are represented as movement of the patch.
+
 The LFP is averaged over three events, produced by three successive
@@
 the r.o.i. and use them to obtain the threshold for separating
-the Sink and Sources sets. Then we procede as stated in the previous section.
+the Sink and Sources sets. Then we proceed as stated in the previous section.
 
 The figure \ref{lfpcsdcm}  compares the representation of the activity in
-the LFP, CSD and CM spaces. It has been exhaustively proven that the CSD
+the LFP, CSD and CM spaces. The CSD
 representation provides better spatial localization of active sites, but the
@@
 mass separate clearly from the big patch over the saturated region.
+In the figure \ref{lfpcsdcmsub} we repeat the procedure but focusing in
+the r.o.i., which includes CA3 str. pyr. We take advantage of the
+detailed frame to show the borders of the disconnected sets that
+we have chosen for the CM analysis in the right column.
+
 In the figure \ref{evotracks} we concatenate the successive centers of mass
@@
 very well localized over the Pyramidal Stratum at precise moments after
-stimulus.
+stimulus. This dataset provides us with a very clean example of the
+trajectory analysis. The very well localized activity and coherent
+action gives us trajectories that are in accordance with the canonical
+source and sink distribution, but adding the temporal dynamical displacement
+to it. We have been very thorough with this example as it will help us
+to understand the more complicated data that shall be presented in
+the next examples.
 
-The treatment of the data is shown in figure
-\ref{CSD_Evocada01}.  The subfigure A  shows a diagram of the hippocampus, with
-the somatic layers (CA and DG) as thick black lines and the stimulus point
-shown as a blue arrow. The $a,b,c,d,e$ dots are selected electrodes for
-showing LFP and CSD traces. Those are shown in the subfigure B. In subfigure
-C the same data is presented as a spatiotemporal color map that
-shows the expected poles of firing neurons transverse to the
-pyramidal stratum.  In the subfigure D we show the CSD for a series of time frames
-over the square delimited in the subfigure A. The idea for treating the different
-patches separately becomes evident here. The sources and sinks form patches
-with an apparent anterograde movement from CA3 to CA1.
-The patches change polarity
-as expected, with the somatic layer becoming a strong sink between 3 to 4 ms
-after the stimulus, and then, at the 6 ms, inverting its polarity in a way that suggest
-successive activation of groups that are represented as movement of the patch.
+\begin{figure}
+  \includegraphics[width=0.9\textwidth]{Figuras/lfpevocado01.pdf}
+  \caption{A set of frames showing the stimulus and posterior activity.
+    The frame titled ``0 ms'' is just before the stimulus was
+    made.  A diagram of the Granular and Pyramidal Layers is shown
+    against the noice background. The black rectangle delimites the
+    r.o.i. and the black arrow shows the site of stimulation in GD.
+  }\label{lfpevocada}
+\end{figure}
 
 \begin{figure}[h]
-  \includegraphics[width=0.9\textwidth]{Figuras/CSD_EvocadaYTrazos.pdf}
-  \caption{ CSDA for evoked activity and visual inspection of disjoint
-    components. In A we present a diagram of the hippocampus slice over
-    the acquisition device. The somatic layers of CA and DG are represented as
-    thick black lines, and the stimulus site is indicated by a blue arrow. A selection
-    of six contiguous electrodes labeled $A-F$ in the r.o.i. are presented as blue
-    dots. In B we show the LFP traces and their CSD representation for those
-    selected electrodes.
-    In C we shoe the CSDA for various timestamps in the sub-region
-    indicated in A as the r.o.i, which covers a square of 24 by 24 electrodes over CA3.
-    in the first image,  4ms after the stimulus,  a
-    shaded strip indicating the approximate location of str. pyr. A model neuron is
-    drawn for scale. At 4ms after the stimulus, the saturation of the electric shock
-    recedes and we can began to appreciate real neuronal activity on the responsive
-    sector of CA3. A great sink appears over the somatic layer, surrounded by two
-    sources and some smaller patches of activity. At the following frame the patch
-    appears to have moved anterogradly, towards CA1. Next frame, the sink appears
-    to have inverted its polarity as do their surrounding patches, and the movement
-    continues until approximately 9ms after stimulus, where the activity ceases to be
-    discernible from noise.}
-  \label{CSD_Evocada01}
+\includegraphics[width=0.9\textwidth]{Figuras/lfpcsdcm01.pdf}
+\caption{ Comparison of the activity represented as LFP, CSD and CM.
+  The time after
+  stimulation in ms is shown in the central column for each row.
+  In this series of frames a relatively big patch of the electrodes
+  still appear saturated, but even so the CSD shows good localization
+  of the relevant deflections from the neutral or noise base signal.
+  The CM representation concentrates the locus of action even more.
+  The circular dots in the third column represent the center of mass
+  of coherent active subsets.
+}\label{lfpcsdcm}
 \end{figure}
 
+\begin{figure}[h]
+\includegraphics[width=0.9\textwidth]{Figuras/lfpcsdcm02.pdf}
+\caption{ The same as previous figure, but focusing on the
+  region of interest, including the stratum pyramidale of CA3.
+}\label{lfpcsdcmsub}
+\end{figure}
+
+%\begin{figure}[h]
+%  \includegraphics[width=0.9\textwidth]{Figuras/CSD_EvocadaYTrazos.pdf}
+%  \caption{ CSDA for evoked activity and visual inspection of disjoint
+%    components. In A we present a diagram of the hippocampus slice over
+%    the acquisition device. The somatic layers of CA and DG are represented as
+%    thick black lines, and the stimulus site is indicated by a blue arrow. A selection
+%    of six contiguous electrodes labeled $A-F$ in the r.o.i. are presented as blue
+%    dots. In B we show the LFP traces and their CSD representation for those
+%    selected electrodes.
+%    In C we shoe the CSDA for various timestamps in the sub-region
+%    indicated in A as the r.o.i, which covers a square of 24 by 24 electrodes over CA3.
+%    in the first image,  4ms after the stimulus,  a
+%    shaded strip indicating the approximate location of str. pyr. A model neuron is
+%    drawn for scale. At 4ms after the stimulus, the saturation of the electric shock
+%    recedes and we can began to appreciate real neuronal activity on the responsive
+%    sector of CA3. A great sink appears over the somatic layer, surrounded by two
+%    sources and some smaller patches of activity. At the following frame the patch
+%    appears to have moved anterogradly, towards CA1. Next frame, the sink appears
+%    to have inverted its polarity as do their surrounding patches, and the movement
+%    continues until approximately 9ms after stimulus, where the activity ceases to be
+%    discernible from noise.}
+%  \label{CSD_Evocada01}
+% \end{figure}
+
+\begin{figure}
  \includegraphics[width=0.9\textwidth]{Figuras/CMVariousTimes01.pdf}
-+  \caption{ The CM analysis over the same images as figure \ref{CSD_Evocada01} C. Some trajectories
-+    are calculated before the 4ms time stamp. The triangular mark shows the later point of the
-+    trajectory, either the end or the last calculated position until that moment. The circular
-+    mark is the start of the trajectory. We can see that activity moves form CA3 distal to
-+    proximal in alternating waves of action (hiperpolarization, firing and repolarization).
-+  }\label{cmevocada}
+  \caption{ The CM analysis over the same images as figure \ref{lfpcsdcmsub} C. Some trajectories
+    are calculated before the 4ms time stamp. The triangular mark shows the later point of the
+    trajectory, either the end or the last calculated position until that moment. The circular
+    mark is the start of the trajectory. We can see that activity moves form CA3 distal to
+    proximal in alternating waves of action (hiperpolarization, firing and repolarization).
+  }\label{cmevocada}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=0.9\textwidth]{./Figuras/trayevocadasub01.pdf}
+  \caption{The evoked trajectories in the r.o.i. We begin to track them
+    at the 0.9 ms mark, after letting the saturation of the stimulus recede.
+    At 3ms a very clear sink trajectory is moving retrogradly towards CA2,
+    sheepherd by two shorter source trayectories. At 5.6 a depolarization
+    trajectory starts, following almost the same path and presenting also its
+    two companions, ending around the 7.7ms timestamp. No significant
+  trajectories are observed after the 10ms mark. }\label{trayevo}
+\end{figure}
 
@@
     at the moments of the epileptic burst. In the first row
-    we present the data as measured, no denoizing or filtering has
+    we present the data as measured, no denoising or filtering has
     been applied. The units of measurement are in $\muV{}$. In

### 2019-07-19 ArticuloCSDyCM/Paper03.tex
tabla de trayectorias

@@
 The highly layered CA structure of the hippocampus provides strong, discernible patterns of Sources and Sinks.
-Following the pattern of activity using Center of Mass analysis directly from the data, though, proves unfruitful, due to the non-convex nature of the somatic layer.
+Following the paths of activity using Center of Mass analysis directly from the data, though, proves unfruitful, due to the non-convex nature of the somatic layer.
 Using CSD and disconnected component tracking over the data provided the right conceptual framework to follow the displacement of active areas during highly active events, such as epileptic-like activity that was pharmacological
@@
 the LFP, CSD and CM spaces. The CSD
-representation provides better spatial localization of active sites, but the
-CM can serve as better dynamical tracker of action. The figure illustrates
+representation provides better spatial localization of active sites, and the
+CM serves as better dynamical tracker. The figure illustrates
 that point by choosing the time interval of response in CA3 after the stimulus.
@@
 
-\begin{figure}
+\begin{figure}[h]
  \includegraphics[width=0.9\textwidth]{Figuras/CMVariousTimes01.pdf}
@@
 
-\begin{figure}
+\begin{figure}[h]
   \includegraphics[width=0.9\textwidth]{./Figuras/trayevocadasub01.pdf}
@@
 \subsection{Application to epileptic-like activity in the hippocampus}
-We performed first the analysis to the most complicated data, numerically and qualitatively, which comes from a series of epileptic burst that occurred after 15 minutes of application of 4AP to living slices of hippocampus.
-Each burst occurs with activity spreading over many sites almost simultaneously.
-This turned out to be a test of the applicability and robustness of our method.
+We performed first the analysis to the most complicated data,
+numerically and qualitatively, which comes from a series of epileptic burst
+that occurred after 15 minutes of application of 4AP to living slices of
+hippocampus. Each burst occurs with activity spreading over many sites almost
+simultaneously. This turned out to be a test of the applicability and
+robustness of our method.
 The necessity of having a good error value for the inactive set was more apparent here than anywhere else.

### 2019-07-19 CenterOfMass/TrayectoriasAux.jl

@@ -6,6 +6,8 @@ cierto tiempo
 
 module TrayectoriasAux
 
+using Statistics
+
 export reducecatenario, AntesQue, DespuesQue, distprom
 
 dist(x1,y1,x2,y2)=sqrt((x1-x2)^2+(y1-y2)^2)

### 2019-08-16 Preprocesamiento/ArraySetTools.jl
tuabuela

@@ -8,9 +8,13 @@ export rowstoset, elemtorow
 
 
 
-function rowstoset(xxs::Array)
+function rowstoset(xxs::Array; enteros=true)
     # pasa los renglones de un arreglo a un conjunto como si jueran elementos
-    result=Set{Array{Int64,2}}()
+    if enteros
+        result=Set{Array{Int64,2}}()
+    else
+        result=Set{Array{Float64,2}}()
+    end
     (nrow,ncol)=size(xxs)
     for j in 1:nrow
         push!(result, reshape(xxs[j,:], 1, ncol))

### 2019-09-02 Reportes/clusteringkarel.tex
nuevas notas

As an extension to the trajectory concept in the CSD-CM we
shall propose a more general concept: spatio-temporal clustering of the CM.
This extends the notion of trajectory to more general grouping of the
CM data.

In view ot the necesity of clasifing the CM by anatomical location,
the need of an automatic, unsupervised clasification of the various CM
over the underlying physiological structure was possed.  The CM data
consists of 4 continous numerical dimenisons and a binary categorical
one. These are the 2 dimensions of of spatial location, which are constrained
to the dimensions of the MEA, the temporal variable (which can be
considered discrete if we use sampling frames as time unit), and the intensity of the CM.
The sign of the last one can be considered a separate categorical variable which
has the values ``Sink'' or negative, and ``Source'' or positive. 
Our problem was then to use the numerical variables to group the CM data in
a way that could sheed some light over the anatomical structure that
generates it. In the figure \ref{puntostodos} we present the CM data
in the MEA coordinates obtained from an experiment with facilitated activity
during a burst of spikes ( around 285 ms of recording ). A visual inspection
seems to suggest a clustering of the points in layers and segments over and near the
Cornu Ammonis Structure, with fewer points over other structures. Most interesting
is the three layered clustering that seems to be around CA3 and the two layeered
structure in CA1.

In the figure \ref{puntosconcolor} we also colorize the dost by temporal emergence.
This also reveals that the layering has also a reccurent order of appearance. Intensity
layering was also explored, but heuristic reasons based on previous trajectory
analysis put doubts on the validity of this clasification. Namely, we expect
that the genaralization of trajectories to also have a rising and fading behaviour,
and thus the intensity should not be used to sepparate different cluster, but to
find regular activity inside each cluster. It could be argued that intensity also
classifies multiunitary activity by type and number of putative units, but then,
this classification would require using intervals of intensity instead of mere
points and would complicate the analysis considerably.

We decided to  use only the spatiotemporal variables for classification. 
The temporal variable has different units, so a prudent normalization is due
in order to not to give to it a preponderant or negligible weight in relation to the
spatial ones. In absence of any other information, the sensible thing to do is to
scale it so that its maximum span is of the same order of magnitude to the
spatial length of the data, namely 64 arbitrary units, which is the convenient unit
of measurement of the MEA ( 64 electrodes correspond to 2.27 mm).
With this normalizazion, the data of the CM exist in a 3 dimensioanl cube of 64 a.u.
of length on each dise. We can then procede to cluster the data acording to
certain criteria.

Clustering algorhytms vary not only in implementation, but in design criteria.
Thus they produce different results. We had to test a few of them. Our choice at the
end was the so called DBSCAN (density based spatial clustering of applications with noice).
This algorhytm is conceptually clean and can produce heteregenous clusters, in contrast
to others such as HDC ( Hierarquical Data clustering). The DBSCAN clustering depends
on two parameters: a radio of neighboorhood and a minimal number of neighbours. Each
point that is demed to be part of a certain cluster has to have that number of neighbours
that are also part of the same set inside an sphere of the choosen radius. These
parameters can be choosen on physioligical grounds altough we can device a more
abstract way of optimizing them.

We decided to use a certain ``esthetic'' criterium for choosing the optimal parameters
for the DBSCAN algorythm.  It is its accordance with Zipf's Law.  This could post out
the accordance with the allometric inverse algebraic laws that are observed in
anatomical studies, but it is, in the end, a somewhat arbitrary criterium. Even thou, a good
fit with Zipfs Law could be an indictor of a natural clustering of data points.
A problem that appeard using this criterium is that it is such a pervarsive
behaviour, that it gives a good fit to allmost any set of reazonable parameters
that one encounters. So we had to come up with criteria for what these ``reasonable
parameters'' may be.  We settled that we shouldn't have to many unasigned points, and
that no cluster should be too big, or that there were to many clusters.

### 2019-09-03 Reportes/clusteringkarel.tex
mas figuras en reporte

@@
+\documentclass{article}
+
+\usepackage{graphicx}
+\usepackage{subcaption}
+
+\begin{document}
+
 As an extension to the trajectory concept in the CSD-CM we
-shall propose a more general concept: spatio-temporal clustering of the CM.
+shall propose another,  more general, concept: spatial-temporal
+clustering of the CM.
 This extends the notion of trajectory to more general grouping of the
@@
 
-In view ot the necesity of clasifing the CM by anatomical location,
-the need of an automatic, unsupervised clasification of the various CM
-over the underlying physiological structure was possed.  The CM data
-consists of 4 continous numerical dimenisons and a binary categorical
+In view of the need of classifying the CM by anatomical location,
+the need of an automatic, unsupervised classification of the various CM
+over the underlying physiological structure was seen.  The CM data
+consists of 4 continuous numerical dimensions and a binary categorical
 one. These are the 2 dimensions of of spatial location, which are constrained
@@
 Cornu Ammonis Structure, with fewer points over other structures. Most interesting
-is the three layered clustering that seems to be around CA3 and the two layeered
+is the three layered clustering that seems to be around CA3 and the two layered
 structure in CA1.
 
-In the figure \ref{puntosconcolor} we also colorize the dost by temporal emergence.
-This also reveals that the layering has also a reccurent order of appearance. Intensity
+\begin{figure}
+\centering
+\begin{subfigure}{0.40\textwidth}
+\includegraphics[width=\textwidth]{puntoscmfacneg.png}
+\caption{Sinks}
+\end{subfigure}
+\begin{subfigure}{0.40\textwidth}
+\includegraphics[width=\textwidth]{puntoscmfacpos.png}
+\caption{Sources}
+\end{subfigure}
+
+\caption{The distribution of the CM of 285ms of data of facilitated activity
+during a burst. We have of the order of 9000 points in each
+set. Visual inspection seems to reveal a multilayered structure in the
+CA3 and CA1 region, with more ambigous clustering in the putative
+CA2 location.}\label{puntostodos}
+
+\end{figure}
+
+In the figure \ref{puntosconcolor} we also colorize the dots by temporal emergence.
+This also reveals that the layering has also a recurrent order of appearance. Intensity
 layering was also explored, but heuristic reasons based on previous trajectory
-analysis put doubts on the validity of this clasification. Namely, we expect
-that the genaralization of trajectories to also have a rising and fading behaviour,
-and thus the intensity should not be used to sepparate different cluster, but to
+analysis put doubts on the validity of this classification. Namely, we expect
+that the generalization of trajectories to also have a rising and fading behavior,
+and thus the intensity should not be used to separate different cluster, but to
 find regular activity inside each cluster. It could be argued that intensity also
-classifies multiunitary activity by type and number of putative units, but then,
+classifies multi-unitary activity by type and number of putative units, but then,
 this classification would require using intervals of intensity instead of mere
 points and would complicate the analysis considerably.
+
+\begin{figure}
+\centering
+\begin{subfigure}{0.40\textwidth}
+\includegraphics[width=\textwidth]{puntoscmfacnegtcolor.png}
+\caption{Sinks}
+\end{subfigure}
+\begin{subfigure}{0.40\textwidth}
+\includegraphics[width=\textwidth]{puntoscmfacpostcolor.png}
+\caption{Sources}
+\end{subfigure}
+
+\caption{Same data as in fig. \ref{puntostodos} but colorized using
+the temporal appearence. The colorbar is in miliseconds.
+}\label{puntosconcolor}
+
+\end{figure}
 
@@
 of measurement of the MEA ( 64 electrodes correspond to 2.27 mm).
-With this normalizazion, the data of the CM exist in a 3 dimensioanl cube of 64 a.u.
-of length on each dise. We can then procede to cluster the data acording to
+With this normalization, the data of the CM exist in a 3 dimensional cube of 64 a.u.
+of length on each side. We can then proceed to cluster the data according to
 certain criteria.
 
-Clustering algorhytms vary not only in implementation, but in design criteria.
+Clustering algorithms vary not only in implementation, but in design criteria.
 Thus they produce different results. We had to test a few of them. Our choice at the
-end was the so called DBSCAN (density based spatial clustering of applications with noice).
-This algorhytm is conceptually clean and can produce heteregenous clusters, in contrast
-to others such as HDC ( Hierarquical Data clustering). The DBSCAN clustering depends
-on two parameters: a radio of neighboorhood and a minimal number of neighbours. Each
-point that is demed to be part of a certain cluster has to have that number of neighbours
-that are also part of the same set inside an sphere of the choosen radius. These
-parameters can be choosen on physioligical grounds altough we can device a more
+end was the so called DBSCAN (density based spatial clustering of applications with noise).
+This algorithm is conceptually clean and can produce heterogeneous clusters, in contrast
+to others such as HDC ( Hierarchical Data clustering). The DBSCAN clustering depends
+on two parameters: a radio of neighborhood and a minimal number of neighbors. Each
+point that is deemed to be part of a certain cluster has to have that number of neighbors
+that are also part of the same set inside an sphere of the chosen radius. These
+parameters can be chosen on physiological grounds although we can device a more
 abstract way of optimizing them.
 
-We decided to use a certain ``esthetic'' criterium for choosing the optimal parameters
+We decided to use a certain ``aesthetic'' criterium for choosing the optimal parameters
 for the DBSCAN algorythm.  It is its accordance with Zipf's Law.  This could post out
@@
 that no cluster should be too big, or that there were to many clusters.
+
+\end{document}

### 2019-09-04 Reportes/clusteringkarel.tex
mas figuras en reporte
y ahora tablas

@@
 
+\author{K. Zapfe}
+\title{Spatiotemporal clusters in the CSD-CM representation}
+
 \begin{document}
+
+\maketitle
 
@@
 \caption{Same data as in fig. \ref{puntostodos} but colorized using
-the temporal appearence. The colorbar is in miliseconds.
+the temporal appearence. The colorbar is in miliseconds. We can observe
+that there is near simultaneus activation of the ``Sources Layers'' in CA3,
+between other visual groupings. 
 }\label{puntosconcolor}
@@
 Thus they produce different results. We had to test a few of them. Our choice at the
-end was the so called DBSCAN (density based spatial clustering of applications with noise).
+end was the so called DBSCAN (Density Based Spatial Clustering of Applications with Noise).
 This algorithm is conceptually clean and can produce heterogeneous clusters, in contrast
@@
 parameters can be chosen on physiological grounds although we can device a more
-abstract way of optimizing them.
+abstract way of optimizing them. An example of the results of this
+algorythm is shown on fig. \ref{puntoscluster}.
+A strict choice of parameters (small radius and many neighbors) produces
+very dense clusters and a lot of unasigned points. A laxer choice (a big radius
+with one to three neighboors) produces few clusters, with sparcely populated
+areas,  and fewer unasigned points. A quick testing of the algorythm 
+shows that we have to explore the validity of clusters in the range of
+$r \in [0.75, 2 ]$ ( electrode lengths) and $ n \in \lbrace 2,3,4,5 \rbrace$.
+
+\begin{figure}[h]
+\centering
+\begin{subfigure}{0.40\textwidth}
+\includegraphics[width=\textwidth]{dbscannegejem01.png}
+\caption{Sinks}
+\end{subfigure}
+\begin{subfigure}{0.40\textwidth}
+\includegraphics[width=\textwidth]{dbscanposejem01.png}
+\caption{Sources}
+\end{subfigure}
+
+\caption{Same data as in previous figures, but the colorization is
+  from a realization of DBSCAN clustering. The parameters used were
+  $r=1.75$ a.u. for the neighborhood radius and 3 neighbors for each point.
+  Those parameters give around 300 clusters, many of them with less than
+  10 points. A good choice of parameters requires balancing
+  the number of clusters and the number of unasigned points. 
+}\label{puntoscluster}
+
+\end{figure}
 
@@
 the accordance with the allometric inverse algebraic laws that are observed in
-anatomical studies, but it is, in the end, a somewhat arbitrary criterium. Even thou, a good
+anatomical studies, but it is, in the end, a somewhat arbitrary criterium.
+Nevertheless, a good
 fit with Zipfs Law could be an indictor of a natural clustering of data points.
 A problem that appeard using this criterium is that it is such a pervarsive
-behaviour, that it gives a good fit to allmost any set of reazonable parameters
+behaviour of ranked data
+that it gives a good fit to allmost any set of reazonable parameters
 that one encounters. So we had to come up with criteria for what these ``reasonable
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\textwidth]{zipfexample.png}
+  \caption{An instance of Zipf's Law, with DBSCAN parameters
+    $r=1.5$ a.u. and $n=4$. The agreement with Zipf's
+    gives an exponent of $-1.007$ against the ideal $-1$ and
+    an $R^2$ of $0.986$.}
+\end{figure}
+
+\begin{tabular}{|r|r|r|r|r|r|}
+ radius &  n  & $R^2$ & u & m & ng  \\ \hline
+  1 &  2 &  0.978 & 606 & 606 & 1285 \\
+  1 &  3 &  0.984 & 1346 & 1346 & 915 \\
+  1 &  4 &  0.974 & 2300 & 2300 & 658 \\
+  1 &  5 &  0.951 & 3193 & 3193 & 493 \\
+  1 &  6 &  0.918 & 4023 & 4023 & 375 \\
+  1.5 &  2 &  0.987 & 276 & 1007 & 691 \\
+  1.5 &  3 &  0.987 & 660 & 1007 & 499 \\
+  1.5 &  4 &  0.985 & 1060 & 1060 & 391 \\
+  1.5 &  5 &  0.986 & 1523 & 1523 & 315 \\
+  1.5 &  6 &  0.99 & 2040 & 2040 & 258 \\
+  2 &  2 &  0.95 & 156 & 5355 & 343 \\
+  2 &  3 &  0.928 & 382 & 5355 & 230 \\
+  2 &  4 &  0.912 & 566 & 5349 & 174 \\
+  2 &  5 &  0.906 & 777 & 5308 & 132 \\
+  2 &  6 &  0.915 & 976 & 3734 & 112 \\
+  2.5 &  2 &  0.911 & 113 & 6144 & 225 \\
+  2.5 &  3 &  0.875 & 243 & 6144 & 160 \\
+  2.5 &  4 &  0.849 & 349 & 6142 & 127 \\
+  2.5 &  5 &  0.828 & 475 & 6127 & 100 \\
+  2.5 &  6 &  0.849 & 602 & 6008 & 81 \\
+  3 &  2 &  0.909 & 89 & 7408 & 163 \\
+  3 &  3 &  0.874 & 187 & 7408 & 114 \\
+  3 &  4 &  0.847 & 267 & 7404 & 89 \\
+  3 &  5 &  0.829 & 343 & 7402 & 71 \\
+  3 &  6 &  0.817 & 418 & 7400 & 58 
+\end{tabular}
+
 \end{document}

### 2019-09-04 Reportes/clusteringkarel.tex
hay que ponerle el exp a la tabla

@@
 parameters'' may be.  We settled that we shouldn't have to many unasigned points, and
-that no cluster should be too big, or that there were to many clusters.
+that no cluster should be too big, or that there were to many clusters. In the
+table \ref{tablarquad} we show the fitting results for some values of
+$r$ and $n$. The model that was fitted did not specify the exponent, it
+is remarkable how the various models gave values for this parameter so close
+to the ideal $-1$.
 
@@
 
+\begin{center}
 \begin{tabular}{|r|r|r|r|r|r|}
- radius &  n  & $R^2$ & u & m & ng  \\ \hline
+ r &  n  & $R^2$ & u & m & ng  \\ \hline
   1 &  2 &  0.978 & 606 & 606 & 1285 \\
@@
 \end{tabular}
+\caption{Table of $R^2$ values fitting Zipf's Law to the clustering
+  rank and size for various parameters. Here $r$ is the neighborhood
+  radius, $n$ the number of required neighboors including itself,
+  $R^2$ is the coefficient of determination, $u$ is the number of
+  unasigned points, $m$ the size in points of the largest
+  cluster, and $ng$ the number of clusters. }
+\label{tablarquad}
+\end{center}
 

### 2019-09-04 Reportes/clusteringkarel.tex
mas tabl

--- 2019-09-04 Reportes/clusteringkarel.tex
@@
 \usepackage{subcaption}
+%\usepackage{natbib}
+\usepackage{caption}
 
@@
 \maketitle
+
+\section{Problem Statement}
 
@@
 the need of an automatic, unsupervised classification of the various CM
-over the underlying physiological structure was seen.  The CM data
-consists of 4 continuous numerical dimensions and a binary categorical
-one. These are the 2 dimensions of of spatial location, which are constrained
-to the dimensions of the MEA, the temporal variable (which can be
-considered discrete if we use sampling frames as time unit), and the intensity of the CM.
-The sign of the last one can be considered a separate categorical variable which
-has the values ``Sink'' or negative, and ``Source'' or positive. 
+over the underlying physiological structure began to increase.  The CM data
+consists of 4  numerical dimensions and a binary categorical
+one, detailed as follows:
+\begin{itemize}
+\item Two spatial dimensions, namely the location of the CM over the
+  area covered by the MEA. In our case this is a 64 by 64 microelectrode
+  array that spans over a square with a side of 2.27 mm. A convenient unit
+  of length is to use the grid interspacing, so we have a 64 by 64
+  cartesian grid. The CM data can have in principle continous values in
+  these coordinates, as it obtained by vector averaging over
+  CSD data.
+\item The temporal coordinate. This one is actually discrete, as it
+  is constrained to the sampling rate of the MEA, which for experimental
+  apparatus that we have, can vary between 7kHz and 18kHz. A timestep unit
+  then is is a fraction of milisecond. Proper normalization shall be done so
+  that it appears comparable to the spatial coordinates.
+\item The intensity of the CM. This can be also treated as a continous
+  variable, whose resolution depends on the noise level of the apparatus
+  and experiment in question.
+\item The sign of the mentioned intensity. Rather than incluiding this as
+  part of the numerical value, we use it as a categorical clasifcator that
+   separates ``Sink'' and ``Source'' data.
+\end{itemize}
+
 Our problem was then to use the numerical variables to group the CM data in
@@
 seems to suggest a clustering of the points in layers and segments over and near the
-Cornu Ammonis Structure, with fewer points over other structures. Most interesting
+Cornu Ammonis Structure, with fewer and sparser
+points over other structures. Most interesting
 is the three layered clustering that seems to be around CA3 and the two layered
@@
 that the generalization of trajectories to also have a rising and fading behavior,
-and thus the intensity should not be used to separate different cluster, but to
+and thus the intensity should not be used to separate different clusters, but to
 find regular activity inside each cluster. It could be argued that intensity also
@@
 
+\section{The Approach}
+
 We decided to  use only the spatiotemporal variables for classification. 
@@
 of measurement of the MEA ( 64 electrodes correspond to 2.27 mm).
-With this normalization, the data of the CM exist in a 3 dimensional cube of 64 a.u.
-of length on each side. We can then proceed to cluster the data according to
-certain criteria.
+With this normalization, the data of the CM exist in a 3 dimensional cube of 64
+arbitrary units at each  side. We can then proceed to test clustering methods.
 
@@
 Thus they produce different results. We had to test a few of them. Our choice at the
-end was the so called DBSCAN (Density Based Spatial Clustering of Applications with Noise).
+end was the so called DBSCAN
+(Density Based Spatial Clustering of Applications with Noise \cite{DBSCAN}).
 This algorithm is conceptually clean and can produce heterogeneous clusters, in contrast
-to others such as HDC ( Hierarchical Data clustering). The DBSCAN clustering depends
+to others such as HDC ( Hierarchical Data clustering \cite{ClustHB}).
+The DBSCAN clustering depends
 on two parameters: a radio of neighborhood and a minimal number of neighbors. Each
@@
 shows that we have to explore the validity of clusters in the range of
-$r \in [0.75, 2 ]$ ( electrode lengths) and $ n \in \lbrace 2,3,4,5 \rbrace$.
+$r \in [0.75, 3 ]$ ( electrode lengths) and $ n \in \lbrace 2,3,4,5 \rbrace$.
+
+In order to give a sensible interpretation we must not forget the
+fact that
+our neighborhoods are three dimensional balls with one temporal dimension.
+We are allowing for many CMs to be predecesors of a single one and
+the other way around, to have many succesors of each point. Strict causality
+is not tested at this stage of the analysis.
 
@@
 We decided to use a certain ``aesthetic'' criterium for choosing the optimal parameters
-for the DBSCAN algorythm.  It is its accordance with Zipf's Law.  This could post out
+for the DBSCAN algorythm.  It is its accordance with Zipf's Law. We expected
+a general allometric scalling law of algebraic decay as expected
+from other biological systems \cite{WestBrown2004, West99}, with unspecified
+exponent, but the universality of Zipf's Law imposed itself. This may be
+due to the fact that Zipfian distribution appears when there is some
+sort of hierarchy between the sets.
+
+We started by proposing a
+relation between rank and size of given by the general expression
+\begin{equation}
+s(k)=c \cdot k^a 
+\end{equation}
+where $s$ is size of the cluster, $k$ is its rank ( taking the bigger as the
+first) and $a, c$ constants to be determined. We used a least squares fitting
+over the logarithmic values and the linearized model, so as not to give
+preponderance to the bigger sets in the fitting, and we tested fits by
+comparing the $a$ exponent and the $R^2$ coefficent of deerminantion (see table
+\ref{tablarquad}).
+
+This could post out
 the accordance with the allometric inverse algebraic laws that are observed in
@@
 parameters'' may be.  We settled that we shouldn't have to many unasigned points, and
-that no cluster should be too big, or that there were to many clusters.
+that no cluster should be too big, or that there were to many clusters. In the
+table \ref{tablarquad} we show the fitting results for some values of
+$r$ and $n$. The model that was fitted did not specify the exponent, it
+is remarkable how the various models gave values for this parameter so close
+to the ideal $-1$. The unasigned points form also a set that is taken into
+account.
 
@@
   \caption{An instance of Zipf's Law, with DBSCAN parameters
-    $r=1.5$ a.u. and $n=4$. The agreement with Zipf's
-    gives an exponent of $-1.007$ against the ideal $-1$ and
+    $r=1.5$ a.u. and 4 neighbors ($n=5$). The agreement with Zipf's
+    gives an exponent of $-1.008$ against the ideal $-1$ and
     an $R^2$ of $0.986$.}
@@
 
-\begin{tabular}{|r|r|r|r|r|r|}
- radius &  n  & $R^2$ & u & m & ng  \\ \hline
-  1 &  2 &  0.978 & 606 & 606 & 1285 \\
-  1 &  3 &  0.984 & 1346 & 1346 & 915 \\
-  1 &  4 &  0.974 & 2300 & 2300 & 658 \\
-  1 &  5 &  0.951 & 3193 & 3193 & 493 \\
-  1 &  6 &  0.918 & 4023 & 4023 & 375 \\
-  1.5 &  2 &  0.987 & 276 & 1007 & 691 \\
-  1.5 &  3 &  0.987 & 660 & 1007 & 499 \\
-  1.5 &  4 &  0.985 & 1060 & 1060 & 391 \\
-  1.5 &  5 &  0.986 & 1523 & 1523 & 315 \\
-  1.5 &  6 &  0.99 & 2040 & 2040 & 258 \\
-  2 &  2 &  0.95 & 156 & 5355 & 343 \\
-  2 &  3 &  0.928 & 382 & 5355 & 230 \\
-  2 &  4 &  0.912 & 566 & 5349 & 174 \\
-  2 &  5 &  0.906 & 777 & 5308 & 132 \\
-  2 &  6 &  0.915 & 976 & 3734 & 112 \\
-  2.5 &  2 &  0.911 & 113 & 6144 & 225 \\
-  2.5 &  3 &  0.875 & 243 & 6144 & 160 \\
-  2.5 &  4 &  0.849 & 349 & 6142 & 127 \\
-  2.5 &  5 &  0.828 & 475 & 6127 & 100 \\
-  2.5 &  6 &  0.849 & 602 & 6008 & 81 \\
-  3 &  2 &  0.909 & 89 & 7408 & 163 \\
-  3 &  3 &  0.874 & 187 & 7408 & 114 \\
-  3 &  4 &  0.847 & 267 & 7404 & 89 \\
-  3 &  5 &  0.829 & 343 & 7402 & 71 \\
-  3 &  6 &  0.817 & 418 & 7400 & 58 
+\begin{table}
+\begin{center}
+  \begin{tabular}{|r|r||r|r|r|r|r|}
+ $r$ &  $n$  & $a$ & $ R^2$ & $U$ & $M$ & $N$  \\ \hline 
+  1 &  2 & -0.755 &  0.978 & 606 & 606 & 1285 \\
+  1 &  3 & -0.71 &  0.984 & 1346 & 1346 & 915 \\
+  1 &  4 & -0.683 &  0.974 & 2300 & 2300 & 658 \\
+  1 &  5 & -0.655 &  0.951 & 3193 & 3193 & 493 \\
+  1 &  6 & -0.647 &  0.918 & 4023 & 4023 & 375 \\
+  1.5 &  2 & -0.99 &  0.987 & 276 & 1007 & 691 \\
+  1.5 &  3 & -0.974 &  0.987 & 660 & 1007 & 499 \\
+  1.5 &  4 & -0.994 &  0.985 & 1060 & 1060 & 391 \\
+  1.5 &  5 & -1.008 &  0.986 & 1523 & 1523 & 315 \\
+  1.5 &  6 & -1.006 &  0.99 & 2040 & 2040 & 258 \\
+  2 &  2 & -0.962 &  0.95 & 156 & 5355 & 343 \\
+  2 &  3 & -0.976 &  0.928 & 382 & 5355 & 230 \\
+  2 &  4 & -1.013 &  0.912 & 566 & 5349 & 174 \\
+  2 &  5 & -1.081 &  0.906 & 777 & 5308 & 132 \\
+  2 &  6 & -1.203 &  0.915 & 976 & 3734 & 112 \\
+  2.5 &  2 & -1.002 &  0.911 & 113 & 6144 & 225 \\
+  2.5 &  3 & -1.006 &  0.875 & 243 & 6144 & 160 \\
+  2.5 &  4 & -1.034 &  0.849 & 349 & 6142 & 127 \\
+  2.5 &  5 & -1.092 &  0.828 & 475 & 6127 & 100 \\
+  2.5 &  6 & -1.213 &  0.849 & 602 & 6008 & 81 \\
+  3 &  2 & -1.043 &  0.909 & 89 & 7408 & 163 \\
+  3 &  3 & -1.05 &  0.874 & 187 & 7408 & 114 \\
+  3 &  4 & -1.083 &  0.847 & 267 & 7404 & 89 \\
+  3 &  5 & -1.138 &  0.829 & 343 & 7402 & 71 \\
+  3 &  6 & -1.206 &  0.817 & 418 & 7400 & 58 \\
 \end{tabular}
+\caption{Table of resulting values fitting algebraic laws to the clustering
+  rank and size for various parameters. The inputs are $r$, the neighborhood 
+  radius and $n$, the number of required neighboors including itself.
+  The resultants are the following.
+  $a$ is the exponent,
+  $R^2$ is the coefficient of determination, 
+  $U$ is the number of
+  unasigned points, $M$ the size in points of the largest
+  cluster, and $N$ the number of clusters. }
+\label{tablarquad}
+\end{center}
+\end{table}
+
+%\bibliographystyle{te}
+
+\bibliographystyle{abstract}
+\bibliography{BiblioReportes01}
 

### 2019-10-15 ArticuloCSDyCM/Paper03.tex
talacha auxiliar
bla bla

@@
-\documentclass[12pt]{article}
+\documentclass[12pt, letterpaper]{article}
 
@@
 
-\doublespacing
-\linenumbers
+% \doublespacing
+% \linenumbers
 
@@
 Evoked activity by electrical stimulus yields simpler and cleaner patterns in the LFP
-measurements. This facilitates the exemplification of our analysis, so we shall
-start the exposition of the usage with these kind of data.
-The data was obtained by stimulating certain specific points on the dentate gyrus
-of the slice of hippocampus with different voltage intensities. Details of the experiment
+measurements. This facilitates the exemplification of our analysis, and gives us
+a good starting case. It comes at a cost, nonetheless, as many of the electrodes
+remain saturated after the stimulus and well beyond the measurement interval.
+The data was obtained by stimulating  specific points on the dentate gyrus
+in the slice of hippocampus with different voltage intensities. Details of the experiment
 have been published elsewhere \cite{Franco2018}. We shall borrow some of
-the experimental data for illustrating our purposes.
+the experimental data for our purposes.
 
@@
 allows us to use averages as a denoising mechanism. After averaging
-the synchronized data we perform kCSDA, as in this case we end up
-with many saturated electrodes and dCDSA is not viable.
+we perform kCSDA. In this case there are
+many saturated electrodes and dCDSA is not viable.
 In the figure \ref{lfpevocada} we show some frames of the post-stimulus
-activity as a heat map, along
+activity as a color map, along
 with a diagram of the somatic layers and the
@@
 the LFP, CSD and CM spaces. The CSD
-representation provides better spatial localization of active sites, and the
+representation provides better spatial localization of active sites, but the
 CM serves as better dynamical tracker. The figure illustrates
@@
 for both sinks and sources and show how the anterograde reaction is
-very well localized over the Pyramidal Stratum at precise moments after
+very well localized over the Pyramidal Stratum at a precise moment after
 stimulus. This dataset provides us with a very clean example of the
-trajectory analysis. The very well localized activity and coherent
-action gives us trajectories that are in accordance with the canonical
+trajectory analysis. The localized activity and coherent
+action gives us trajectories that are in accordance with the expected
 source and sink distribution, but adding the temporal dynamical displacement
@@
 the next examples.
+
+The evoked activity seems to yield much cleaner results, as we can now compare the trajectories of successive, very similar events.
+Our CSD and CM analysis can be used to detect consistent features in these experiments, such as the direction the waxing and waning of activity.
+Small deviations from the trajectory in successive experiments that are above the numerical error could indicate inconsistencies in the network below.
 
@@
 
-The evoked activity seems to yield much cleaner results, as we can now compare the trajectories of successive, very similar events.
-Our CSD and CM analysis can be used to detect consistent features in these experiments, such as the direction the waxing and waning of activity.
-Small deviations from the trajectory in successive experiments that are above the numerical error could indicate inconsistencies in the network below.
-
 \subsection{Application to facilitated activity}
@@
 \subsection{Application to epileptic-like activity in the hippocampus}
-We performed first the analysis to the most complicated data,
+This is probably the most complicated data, 
 numerically and qualitatively, which comes from a series of epileptic burst
@@
 robustness of our method.
-The necessity of having a good error value for the inactive set was more apparent here than anywhere else.
-Also the most striking features of the propagation of activity, coded now as trajectories of the CM, could be seen here, as the epileptic burst covered the whole CA structure.
-A sample of the raw data in $\mu V$ can be seen in the figure \ref{lfpycsd4ap}.
+The necessity of having a good error value for the inactive set was more apparent here
+than in the other cases. Also, the most striking features of the propagation of activity,
+coded now as trajectories of the CM, could be seen here, as the epileptic burst covered the whole
+CA structure. A sample of the raw data in $\muV{}$ can be seen in the figure \ref{lfpycsd4ap}.
 Five key moments in the epileptic burst have been selected, one just before the onset of the attack, the next one at the peak and one  following it, and the last at the subsiding moments.
 The figure compares the Raw data
-in $\muV{}$ in the first row, then the crude CSD representation in the second row, and
+in $\muV{}$ in the first row, then the raw CSD representation in the second row, and
 in the third raw a CSD after a Gaussian Blur Filter has been applied,
@@
 
-In the figure \ref{tray4ap} we show the attack on the trajectories representation. Fine lines
-connect the blue (sink) and red (source) dots, that indicate the trajectories along various activation
+In the figure \ref{tray4ap} we show the epileptic burst on the trajectories representation.
+Fine lines
+connect the blue (sink) and red (source) dots, that indicate the trajectories
+along various activation
 sites around the whole CA structure, and part of GD. Very small trajectories precede
@@
 
-bla bla.
-
 \bibliographystyle{abstract}

### 2019-11-06 CenterOfMass/TrayectoriasAux.jl
Figuras facilitadas

@@ -33,6 +33,26 @@ function reducecatenario(Catenario::Dict;
     return result
 end
 
+
+function cortacatenarios(Catenario::Dict; nini,nfin)
+    # reduce las trayectorias por cuadro de inicio y final
+
+    result=Dict{Integer, Array{Any}}()
+    
+    for (k,p) in Catenario
+
+        paux=AntesQue(p,nfin)
+        paux=DespuesQue(paux,nini)
+        if length(paux) != 0
+            result[k]=paux
+        end
+        
+    end
+
+    return result
+end
+
+
 function AntesQue(Datos::Array, tiempo)
     Cadena=[0 0 0 0]
  

### 2019-11-06 Preprocesamiento/ArraySetTools.jl

@@ -13,7 +13,7 @@ function rowstoset(xxs::Array; enteros=true)
     if enteros
         result=Set{Array{Int64,2}}()
     else
-        result=Set{Array{Float64,2}}()
+        result=Set{Array{Float64,2}}()sp
     end
     (nrow,ncol)=size(xxs)
     for j in 1:nrow

### 2019-11-07 ArticuloCSDyCM/Paper03.tex
figuras facilitadas incorporadas

@@
 \usepackage{longtable}
+\usepackage{float}
 
@@
 
-\begin{figure}
+\begin{figure}[H]
   \includegraphics[width=0.9\textwidth]{Figuras/lfpevocado01.pdf}
@@
 
-\begin{figure}[h]
-\includegraphics[width=0.9\textwidth]{Figuras/lfpcsdcm01.pdf}
+\begin{figure}[H]
+\includegraphics[width=0.7\textwidth]{Figuras/lfpcsdcm_completa_evocada3R7.pdf}
 \caption{ Comparison of the activity represented as LFP, CSD and CM.
@@
 
-\begin{figure}[h]
-\includegraphics[width=0.9\textwidth]{Figuras/lfpcsdcm02.pdf}
+\begin{figure}[H]
+\includegraphics[width=0.7\textwidth]{Figuras/lfpcsdcm_sub_evocada3R7.pdf}
 \caption{ The same as previous figure, but focusing on the
-  region of interest, including the stratum pyramidale of CA3.
+  region of interest, including the stratum pyramidale of CA3. A different
+  color scheme was choosen on the last column, so as to provide borders
+  that show how the disconnected componentes where choosen. 
 }\label{lfpcsdcmsub}
@@
 
-\begin{figure}[h]
- \includegraphics[width=0.9\textwidth]{Figuras/CMVariousTimes01.pdf}
-  \caption{ The CM analysis over the same images as figure \ref{lfpcsdcmsub} C. Some trajectories
-    are calculated before the 4ms time stamp. The triangular mark shows the later point of the
-    trajectory, either the end or the last calculated position until that moment. The circular
-    mark is the start of the trajectory. We can see that activity moves form CA3 distal to
-    proximal in alternating waves of action (hiperpolarization, firing and repolarization).
-  }\label{cmevocada}
-\end{figure}
+%\begin{figure}[h]
+% \includegraphics[width=0.9\textwidth]{Figuras/CMVariousTimes01.pdf}
+%  \caption{ The CM analysis over the same images as figure \ref{lfpcsdcmsub}. Some trajectories
+%    are calculated before the 4ms time stamp. The triangular mark shows the later point of the
+%    trajectory, either the end or the last calculated position until that moment. The circular
+%    mark is the start of the trajectory. We can see that activity moves form CA3 distal to
+%    proximal in alternating waves of action (hiperpolarization, firing and repolarization).
+%  }\label{cmevocada}
+%\end{figure}
 
-\begin{figure}[h]
-  \includegraphics[width=0.9\textwidth]{./Figuras/trayevocadasub01.pdf}
+\begin{figure}[H]
+ \includegraphics[width=0.7\textwidth]{./Figuras/trayevocadasub.pdf}
   \caption{The evoked trajectories in the r.o.i. We begin to track them
@@
     two companions, ending around the 7.7ms timestamp. No significant
-  trajectories are observed after the 10ms mark. }\label{trayevo}
+  trajectories are observed after the 10ms mark. }\label{evotracks}
 \end{figure}
 
-\subsection{Application to facilitated activity}
+\subsection{Application to facilitated activity through injury}
+
+The facilitated activity through a small injury produces very clean
+and ordered activity in the CA layers, as measured by the CM analysis.
+The results can be seen in the figures
+\ref{diagfaci}, \ref{lfpcsdcmfaci}, \ref{facitracks} and
+\ref{facitracktodas}.
+A previous diagram is on the figure \ref{diagfaci}.
+
+\begin{figure}[H]
+  \includegraphics[width=0.7\textwidth]{Figuras/diagfacilitada.pdf}
+  \caption{A diagram of the approximate location of the pyramidal and granular strata on the hippocampus slice, during a peak of activity as measured in \muV{}.
+  }
+  \label{diagfaci}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[width=0.7\textwidth]{Figuras/lfpcsdcm_facilitada.pdf}
+\caption{ Comparison of the activity represented as LFP, CSD and CM,
+  in the same style as fig. \ref{lfpcsdcm}, but for facilitated activity.
+  In this case the time counter starts at some moment around 80 ms. before the
+  onset of the burst of activity.
+}\label{lfpcsdcmfaci}
+\end{figure}
+
+\begin{figure}[H]
+ \includegraphics[width=0.7\textwidth]{./Figuras/trayfaciljuntas.pdf}
+ \caption{The trajectories for  the facilitated activity, in
+ sequence of aparition around the burst of action.}\label{facitracks}
+\end{figure}
+
+\begin{figure}[H]
+ \includegraphics[width=0.7\textwidth]{./Figuras/traytodasfacilitada.png}
+ \caption{The trajectories for  the facilitated activity, shown
+ all together.}\label{facitracktodas}
+\end{figure}
 
@@
 
-\begin{figure}
+\begin{figure}[H]
   \includegraphics[width=0.9\textwidth]{./Figuras/LFPvsCSD_4AP.pdf}
@@
 
-\begin{figure}
+\begin{figure}[H]
   \includegraphics[width=0.9\textwidth]{./Figuras/Fronteras2sigma.pdf}
@@
 
-\begin{figure}
+\begin{figure}[H]
   \includegraphics[width=0.9\textwidth]{./Figuras/SeqTray4AP01.pdf}
@@
 
-\subsection{Application to spontaneous activity in idem.}
-
-We had one experiment in which we could trace very nicely the data on DG and CA simultaneously without adding excitatory drugs.
-
 \subsection{Counterexample: the stratial neuronal tissue}

### 2019-11-08 ArticuloCSDyCM/Paper03.tex
Estriado 1004AP es el bueno

@@
 
+ Larger scale recordings (macroscopic
+structures) may gain something of this analysis, but we shall focus
+our result to the neurons scale ( distances of the order of \mum{10}).
+
+The stratial tissue is an example of where this analysis cannot be used.
+The limitations of this analysis are that it works  in
+ordered, stratified neuronal tissue, and highly homogeneous tissue
+without clear structures makes the distintion of both the CSD
+disconnected sets unfeasible.
+
 \bibliographystyle{abstract}

### 2019-11-15 ArticuloCSDyCM/Paper03.tex
mas trazos evocado

@@
-\documentclass[12pt, letterpaper]{article}
+\documentclass[11pt, letterpaper]{article}
 
@@
 \usepackage{setspace}
-\usepackage{lineno} %ambos paquetes para borrador
+\usepackage{lineno} %ambos paquetes para borrado
 
@@
 \usepackage{siunitx} % para unidades chidas como micrometros.
+
+\usepackage{caption}
 
@@
 \newcommand{\rd}{\mathrm{d}}
+
+\newenvironment{intension}{\fontfamily{phv}\selectfont}{\par}
 
@@
 These present the problem of justifying the choice made on practical grounds.
-For calculating a vectorial average, the adequate weight per vector is a density, a positive quantity that measures the concentration of something per space unit, so it is only adequate that we search for a natural density in the electrophysiological experiments.
+For calculating a vectorial average, the adequate weight per vector is a density, a positive quantity that measures the concentration of something per space unit, so it is only adequate that we search for a natural density in the electro-physiological experiments.
 The Current Source Density fits the mathematical and interpretative requirements.
@@
 Having both
-the mathematician and electrophysiologist interpretations presents we can
+the mathematician and electro-physiologist interpretations presents we can
 perform the following analysis with clear readings.
@@
 
-\begin{figure}
-  \includegraphics[width=0.9\textwidth]{Figuras/ExplicaDisconexos01.pdf}
-  \caption{ Explanation of the geometrical concepts used on our analysis.
+\begin{figure}[H]
+  \captionsetup{singlelinecheck=off} 
+  \includegraphics[width=0.9\textwidth]{Figuras/explicadisconected.pdf}
+  \caption[tu abuela]{
+   %& \begingroup
+    %  \fontfamily{ppl}\selectfont
+    \begin{intension}
+      El propósito de esta figura es aclarar los conceptos usados
+      para obtener los centros de masa en componentes disconexos.
+      En la parte A
+      se detalla con ejemplos artificiales los conceptos de
+      convexo, disjunto y disconexo, y se muestran sus
+      efectos sobre el cálculo de centros de masa. En la parte B
+      se encuentra una aplicación a un recuadro de datos experimentales
+      para que quede claro como se ven los pasos del análisis
+      en un caso real.
+      La primera parte indica a grosso modo la localización
+      del str. pyr. y como en un cuadro específico
+      se separan los datos en los conjuntos disjuntos de Pozos y Fuentes.
+      En las partes ii y iii se muestra como se separan los componentes
+      disconexos y se obtienen para cada uno su centro de masa.
+      \end{intension}
+   %   \endgroup
+
+    Explanation of the geometrical concepts used on our analysis.
     In the Row A of figures
@@
     be convex. In figure B we show everything with a frame of experimental data.
-    In Bi we sepparate the
+    In Bi we separate the
     data into three sets, Sources, Sinks, and Zeros.
@@
 electro-physiological data obtained from the rat's hippocampus, using
-a very dense MEA, the BioCAM of 3Brain \cite{BioCAM}. 
+a very dense MEA, the BioCAM of 3Brain \cite{BioCam}. 
 The highly layered CA structure of the hippocampus provides strong, discernible patterns of Sources and Sinks.
@@
 \begin{figure}[H]
+  \captionsetup{singlelinecheck=off} 
   \includegraphics[width=0.9\textwidth]{Figuras/lfpevocado01.pdf}
-  \caption{A set of frames showing the stimulus and posterior activity.
+  \caption[tu abuela]{
+    \begin{intension}
+      Aquí mostramos como se ve el registro crudo del potencial
+      en un evento evocado. Tenmos los primeros recuadros super
+      saturados, y a partir del milisegundo 4 podemos
+      empezar a discernir las estructuras. Los electrodos
+      más cercanos al sitio de estimulación quedan inutilizados.
+      Es un ejemplo de las medidas en crudo en un caso dificil de usar.
+     \end{intension}
+
+    A set of frames showing the stimulus and posterior activity.
     The frame titled ``0 ms'' is just before the stimulus was
     made.  A diagram of the Granular and Pyramidal Layers is shown
-    against the noice background. The black rectangle delimites the
+    against the noise background. The black rectangle delimits the
     r.o.i. and the black arrow shows the site of stimulation in GD.
@@
 \begin{figure}[H]
+  \captionsetup{singlelinecheck=off} 
 \includegraphics[width=0.7\textwidth]{Figuras/lfpcsdcm_completa_evocada3R7.pdf}
-\caption{ Comparison of the activity represented as LFP, CSD and CM.
+\caption[tu abueala]{
+  \begin{intension}
+    En esta imagen y la siguiente se compara la representación de
+    LFP, CSD y CM. La idea es que se pueda apreciar la ventaja
+    en limpieza y localización que ofrece CSD, y que se entienda
+    como manejar los centros de masa a partir de aquella.
+    También muestra como es posible separar la acción de un grupo
+    de electrodos inservible porque su Centro de Masa no se mueve.
+  \end{intension}
+
+  Comparison of the activity represented as LFP, CSD and CM.
   The time after
@@
 \begin{figure}[H]
+   \captionsetup{singlelinecheck=off} 
    \includegraphics[width=0.7\textwidth]{Figuras/lfpcsdcm_sub_evocada3R7.pdf}
-\caption{ The same as previous figure, but focusing on the
+   \caption[tu abuela]{
+     \begin{intension}
+       Misma intensión general que la figura anterior pero en un
+       recuadro específico. La idea es mostrar claramente los componentes
+       disjuntos en CA3 y su Centro de Masa a detalle, en la tercera columna.
+       Un punto muy claro es ver como los componentes se pueden disolver
+       en componentes más pequeños.
+     \end{intension}
+
+     The same as previous figure, but focusing on the
   region of interest, including the stratum pyramidale of CA3. A different
-  color scheme was choosen on the last column, so as to provide borders
-  that show how the disconnected componentes where choosen. 
+  color scheme was chosen on the last column, so as to provide borders
+  that show how the disconnected components where chosen. 
 }\label{lfpcsdcmsub}
@@
 \begin{figure}[H]
+   \captionsetup{singlelinecheck=off} 
  \includegraphics[width=0.7\textwidth]{./Figuras/trayevocadasub.pdf}
-  \caption{The evoked trajectories in the r.o.i. We begin to track them
+ \caption[tu abuela]{
+   \begin{intension}
+     Primera muestra de trayectorias. Escogemos sólo la parte
+     respondiente del potencial evocado, para claridad. Se debe
+     poder apreciar como las trayectorias corren sobre
+     los diferentes estratos y sus tiempos de desplazamiento
+     son esperados dentro de las escalas de distancia y sincronización.
+   \end{intension}
+   The evoked trajectories in the r.o.i. We begin to track them
     at the 0.9 ms mark, after letting the saturation of the stimulus recede.
     At 3ms a very clear sink trajectory is moving retrogradly towards CA2,
-    sheepherd by two shorter source trayectories. At 5.6 a depolarization
+    shepherd by two shorter source trajectories. At 5.6 a depolarization
     trajectory starts, following almost the same path and presenting also its
@@
 \begin{figure}[H]
+   \captionsetup{singlelinecheck=off} 
   \includegraphics[width=0.7\textwidth]{Figuras/diagfacilitada.pdf}
-  \caption{A diagram of the approximate location of the pyramidal and granular strata on the hippocampus slice, during a peak of activity as measured in \muV{}.
+  \caption[tu abuela]{
+  \begin{intension}
+    Mostrar un diagrama de los estratos importantes. Se puede incorporar
+    a otra figura.
+  \end{intension}
+  A diagram of the approximate location of the pyramidal and granular strata on the hippocampus slice, during a peak of activity as measured in \muV{}.
   }
@@
 \begin{figure}[H]
+   \captionsetup{singlelinecheck=off} 
 \includegraphics[width=0.7\textwidth]{Figuras/lfpcsdcm_facilitada.pdf}
-\caption{ Comparison of the activity represented as LFP, CSD and CM,
+\caption[tu abuela]{
+  \begin{intension}
+    Intensión similar a la figura \ref{lfpcsdcmsub}, pero
+    usando el área de registro completa. Se muestra la diversidad
+    de escalas que abarcan los posibles componentes, por ejemplo,
+    los muy grandes en las lineas centrales, en CA3, y los pequeños en
+    la cola de CA1 en los últimos renglones. 
+  \end{intension}
+
+  Comparison of the activity represented as LFP, CSD and CM,
   in the same style as fig. \ref{lfpcsdcm}, but for facilitated activity.
@@
 \begin{figure}[H]
+  \captionsetup{singlelinecheck=off}
   \includegraphics[width=0.7\textwidth]{./Figuras/trayfaciljuntas.pdf}
- \caption{The trajectories for  the facilitated activity, in
- sequence of aparition around the burst of action.}\label{facitracks}
+\caption[tu abuela]{
+  \begin{intension}
+    Imagen de área de registro completo mostranto las trayectorias.
+    Después de la figura \ref{evotracks}, esta figura muestra
+    la riqueza estructural en los datos generados con éste análisis,
+    y muestra la posibilidad ( intuitivamente ) de agrupar visualmente
+    los diferentes manchones de trayectorias a esta escala. 
+  \end{intension}
+ The trajectories for  the facilitated activity, in
+ sequence of apparition around the burst of action.}\label{facitracks}
 \end{figure}
@@
 \begin{figure}[H]
+   \captionsetup{singlelinecheck=off} 
  \includegraphics[width=0.7\textwidth]{./Figuras/traytodasfacilitada.png}
- \caption{The trajectories for  the facilitated activity, shown
+ \caption[tu abuela]{
+   \begin{intension}
+     Remate de la figura \ref{facitracks}. En la ampliación es mucho más fácil
+     discernir diferentes tipos de trayectorias. En CA3 corren a lo largo
+     de los estratos, mientras que en CA1 son primordialmente transversales.
+     También se puede apreciar un área de transición con menor densidad. 
+   \end{intension}
+
+ The trajectories for  the facilitated activity, shown
  all together.}\label{facitracktodas}
@@
 \begin{figure}[H]
+   \captionsetup{singlelinecheck=off} 
   \includegraphics[width=0.9\textwidth]{./Figuras/LFPvsCSD_4AP.pdf}
-  \caption{Comparison of LFP and CSD representations
+  \caption[tu abuela]
+  { \begin{intension}
+      En esta figura muestro LFP contra CSD con ruido espacial y
+      sin ruido espacial. Esta vez estoy probando ordenar por
+      columnas en lugar de por renglones, probando otra representación.
+      A diferencia de otras figuras, aquí muestro explicitamente
+      la ayuda de usar un filtro gaussiano espacial para
+      ``limar el ruido''. 
+    \end{intension}
+    Comparison of LFP and CSD representations
     at the moments of the epileptic burst. In the first row
@@
 \begin{figure}[H]
+   \captionsetup{singlelinecheck=off} 
   \includegraphics[width=0.9\textwidth]{./Figuras/Fronteras2sigma.pdf}
-  \caption{The detection of disconnected components for the Sink and Source
+  \caption[tu abuela]
+    {
+      \begin{intension}
+        Una vez más enfatizamos la importancia del filtro gaussiano.
+        En el renglón superior mostramos como los componentes disconexos
+        quedan demasiado pequeños, y se pierde significancia en un mar de
+        ruido, mientras que en el inferior aparecen subconjuntos coherentes.
+      \end{intension}
+
+            The detection of disconnected components for the Sink and Source
     Sets. We take the the threshold value as two times the
@@
 \begin{figure}[H]
+   \captionsetup{singlelinecheck=off} 
   \includegraphics[width=0.9\textwidth]{./Figuras/SeqTray4AP01.pdf}
-  \caption{ The activity in the Center of Mass and Trajectories representation.
+  \caption[tu abuela]{
+    \begin{intension}
+      Partiendo del renglón inferior de la figura \ref{csdfrontera}
+      mostramos la aparición y sucesión de trayectorias bien definidas
+      sobre str.pyr. y el giro dentado (primer ejemplo donde es claramente
+      discernible su aportación a las trayectorias). 
+    \end{intension}
+    The activity in the Center of Mass and Trajectories representation.
     A single epileptic attack is shown from slightly before the onset until the
@@
 The stratial tissue is an example of where this analysis cannot be used.
-The limitations of this analysis are that it works  in
+The limitation of this technique is that it works  in
 ordered, stratified neuronal tissue, and highly homogeneous tissue
-without clear structures makes the distintion of both the CSD
+without clear structures makes the distinction of both the CSD and
 disconnected sets unfeasible.
+
+\begin{figure}[H]
+ \captionsetup{singlelinecheck=off} 
+  \includegraphics[width=0.7\textwidth]{Figuras/lfpestriado.pdf}
+  \caption[tu abuela]{
+    \begin{intension}
+      Esta figura contrasta mucho con la figura \ref{lfpevocada}.
+      Bajo la misma representación, tenemos muy claro que no hay
+      estructuras discernibles. Sin embargo, requiere una serie de trazos
+      bien escogidos para saber si hay actividad relevante.
+    \end{intension}
+    LFP stratial Tissue, some samples. There are no discernible
+    structures. The apparent active set at the lower left corner
+  is an artifact due to some bubbles in the experiment.}
+\end{figure}
+
+\begin{figure}[H]
+   \captionsetup{singlelinecheck=off} 
+  \includegraphics[width=0.7\textwidth]{Figuras/traytodasestriado.pdf}
+  \caption[tu abuela]{
+    \begin{intension}
+      Una muestra clara de que el tejido estriado no
+      permite la búsqueda de trayectorias definidas. A pesar
+      de contar con un registro mucho más amplio, apenas alcanzamos a
+      ver un par de trayectorias ``fuente'' en lo que aparente ser
+      un artefacto del ruido y no cuentan con ``pozos'' que acompañen.
+    \end{intension}
+    After 30 seconds of recording, no visible trajectories can be found.
+The minimum displacement criteria is not full-filed for most of the putative centers of mass.
+Only a small set of trajectories is seen on the lower corner and these seem to correspond to an artifact due to a bubble in the recording.
+}
+
+\end{figure}
 

### 2019-11-19 ArticuloCSDyCM/Paper03.tex
trazos estriado listo

@@
 \usepackage{caption}
+\usepackage{subcaption}
 
@@
 
+In the figure \ref{trazselectevo} we show a sample of the LFP
+raw traces and how they appear in the CSD representation. A diagram
+over a frame of the LFP is shown delimiting the region of interest and
+some selected electrodes.
+
 Once that we have the CSD representation, we select channels outside
@@
   }\label{lfpevocada}
+\end{figure}
+
+\begin{figure}[H]
+   \centering
+    \begin{subfigure}[t]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{Figuras/lfpcuadrodiagramaevocado.pdf}
+        \caption{Select Electrodes}
+        \label{fig:sel}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[t]{0.2\textwidth}
+        \includegraphics[width=\textwidth]{Figuras/lfpevocselect_3R7.pdf}
+        \caption{LFP, Raw}
+        \label{fig:lfpraw}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+    %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[t]{0.2\textwidth}
+        \includegraphics[width=\textwidth]{Figuras/csdevocselect_3R7.pdf}
+        \caption{CSD}
+        \label{fig:csd}
+    \end{subfigure}
+    \caption{
+      \begin{intension}
+        Mostremos unos cuantos trazos de la medida cruda y su representación
+        en el espacio CSD, para establecer lo adecuado del experimento.
+      \end{intension}
+      Electrode signal across some selected electrodes over the CA3 region.
+      We show in subfigure \ref{fig:sel} a frame of the raw signal in the same color
+      scale as fig. \ref{lfpevocada}. The GD and CA3 structures appear as red stains,
+      their general shape recognizable. In subfigure \ref{fig:lfpraw}
+      the raw traces of the selected electrodes are shown, and in \ref{fig:csd} the
+      csd traces come along for comparision. In this last subfigure the
+      vertical scale is in arbitrary units. }
+  \label{trazselectevo}
 \end{figure}
@@
     is an artifact due to some bubbles in the experiment.}
+  \label{cuadestriado}
+\end{figure}
+
+\begin{figure}[H]
+   \centering
+    \begin{subfigure}[b]{0.55\textwidth}
+        \includegraphics[width=\textwidth]{Figuras/mapestriado_1004AP.pdf}
+        \caption{A Frame}
+        \label{maptraz:map}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.35\textwidth}
+        \includegraphics[width=\textwidth]{Figuras/trazosestriado_1004AP.pdf}
+        \caption{Traces}
+        \label{maptraz:traz}
+    \end{subfigure}
+    \caption{
+      \begin{intension}
+        Mostramos aquí unos trazos ejemplares para asegurarnos de la
+        calidad del experimento.
+      \end{intension}
+      Selected Traces of the Stratial Tissue. A frame such as the ones
+      in figure \ref{cuadestriado} shows the location of the selected electrodes,
+    which where algorhytmically chosen.}
+
 \end{figure}

### 2019-11-20 ArticuloCSDyCM/Paper03.tex
Cuadernos separados para trazos

@@
 
-\subsection{Application to evoked activity in idem.}\label{sec:evocada}
+\subsection{Application to evoked activity in the rat's hippocampus}\label{sec:evocada}
 
@@
 \ref{facitracktodas}.
-A previous diagram is on the figure \ref{diagfaci}.
+A previous diagram is on the figure \ref{diagfaci}, and some
+exemplary traces in the figure \ref{trazfaci}.
 
@@
   \label{diagfaci}
+\end{figure}
+
+\begin{figure}[H]
+   \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{Figuras/mapstddfaci.pdf}
+        \caption{Location of electrodes}
+        \label{maptrazfaci:map}
+    \end{subfigure}
+    \qquad
+    %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[t]{0.15\textwidth}
+        \includegraphics[width=\textwidth]{Figuras/trazoselectfaci.pdf}
+        \caption{Traces}
+        \label{maptrazfaci:traz}
+    \end{subfigure}
+    \caption{
+      \begin{intension}
+        Mostrar los trazos en variar regiones del hipocampo para
+        aclarar el suceso del ataque epiléptico.
+      \end{intension} 
+      A selection of traces of the recordings for various electrodes.
+      In the subfigure \ref{maptrazfaci:map}, the selected electrodes
+      are shown against the standard deviation of the recording of the
+      potential, which gives a good indication of the more active
+      sites, and therefore, of the structure of the hippocampus.
+      The electrode labeled ``a'' is over CA3 proximal, on the other
+      extreme ``h'' is in the middle of CA1.
+    }\label{trazfaci}
 \end{figure}
@@
 \subsection{Application to epileptic-like activity in the hippocampus}
+
 This is probably the most complicated data, 
@@
 coded now as trajectories of the CM, could be seen here, as the epileptic burst covered the whole
-CA structure. A sample of the raw data in $\muV{}$ can be seen in the figure \ref{lfpycsd4ap}.
+CA structure. A sample of the raw data in $\muV{}$ can be seen in the figure \ref{lfpycsd4ap}. In the figure \ref{traz4ap} some exemplarly traces
+can be seen.
 Five key moments in the epileptic burst have been selected, one just before the onset of the attack, the next one at the peak and one  following it, and the last at the subsiding moments.
@@
     been applied. }\label{lfpycsd4ap}
+\end{figure}
+
+\begin{figure}[H]
+   \centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[width=\textwidth]{Figuras/mapstdd4AP.pdf}
+        \caption{Location of electrodes}
+        \label{maptraz4ap:map}
+    \end{subfigure}
+    \qquad
+    %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[t]{0.15\textwidth}
+        \includegraphics[width=\textwidth]{Figuras/trazoselect4AP.pdf}
+        \caption{Traces}
+        \label{maptraz4ap:traz}
+    \end{subfigure}
+    \caption{
+      \begin{intension}
+        Mostrar los trazos en variar regiones del hipocampo para
+        aclarar el suceso del ataque epiléptico.
+      \end{intension} 
+      A selection of traces of the recordings for various electrodes.
+      In the subfigure \ref{maptraz4ap:map}, the selected electrodes
+      are shown against the standard deviation of the recording of the
+      potential, which gives a good indication of the more active
+      sites, and therefore, of the structure of the hippocampus.
+      The electrode labeled ``a'' is putatively over GD, while ``h''
+      is in the medial CA1.}\label{traz4ap}
 \end{figure}

### 2020-06-03 Reportes/Reporte20200601.tex
Reporte Engrish

\documentclass{article}

%\usepackage[utf8]{inputenc}
\usepackage{fontspec}
\usepackage{amsmath}
%\usepackage[spanish]{babel}

\usepackage{graphicx}

\usepackage{caption}
\usepackage{subcaption}

\usepackage{hyperref}

\setmainfont[Ligatures=TeX]{Droid Serif}

\newcommand{\Jd}{\mathbf{J}}
\newcommand{\EF}{\mathbf{E}}
\newcommand{\cond}{\boldsymbol{\sigma}}
\DeclareMathOperator{\diver}{div}
\DeclareMathOperator{\grad}{grad}

\title{Entropy, Symbolization and Mutual Information \\
Teorical-Numerical Quarentine Dissertation}
\author{ W. P. Karel Zapfe}

\begin{document}

\maketitle

\section{Introduction}

Electroneurophysiology is strongly based on the measurement and characterization of electrical signals given by live neurons, either in vitro or in vivo. This signal is often recorded as an array of voltage measurements, done at precise high frequencies. Along the years, the sampling and precision of these measurements have increased by several orders of magnitude. Moreover, the capacity to sample hundreds or thousands of sites simultaneously exists now, sometimes spread over macroscopic areas, such as the whole scalp of the head, or concentrated in very small areas, as in MMEAs (Massive Micro-electrode arrays). Such signals are very rich, and many efforts are done to characterize their complexity. A common approach is to estimate the Entropy of the signal, and from there to use other derived measures such as Mutual Information or Transfer Entropy to search for causality between different channels.

It turns out that the very first step is sometimes done routinely and without much thought to fundamentals, and, as we shall show here, can lead to inconsistent or plain wrong results. As we shall see, Entropy in the sense of Shannon is often wrongly used, either due to conceptual laxity or by error in the numerical algorithmic design. On the other hand, slightly different calculations of Entropy may point different "complexities" in the data. As we are most interested in causality tests, we may focus on the most consistent and robust definition of Entropy. It appears that the definition of Entropy in the sense of Shannon is a tricky thing, and Shannon himself made an erroneous extrapolation of his discrete definition that is wrong. Even though this has been pointed since 1957 by E.T. Jaynes \cite{Jaynes1957}, and adequate alternatives have been shown to exist
\cite{Kullback1951, Schroeder2004}, researchers routinely take the incorrect definition as granted, without realising that does not converge nor has the same interpretation \cite{Jaynes1968}. On the other side, the correct use of Shannon's Entropy involves
a process of symbolization of the continuous data, and this introduces further
questions. As we shall see in the following, some symbolization alter drastically
the Entropy estimation, thus forcing the researchers to take a sort of Occam's approach:
take the simplest one, even if it looks really crude, so that artifacts given
by the model do not play a large role. A lot of recent work in Neuroscience take this
path, at a very extreme level, encoding the neurophysiological signals in
only two signs, above or below a threshold. This can be done on the electric potential field, separating putative spikes from absence of spikes, or in a  more
focused manner, such as taking the sign of
the difference of the signals, as in \cite{Steuer2004}, or on
the time domain, if the inter-spike interval is larger or below a certain value,
as in \cite{Steuer2001}.

In this report we shall explore the difficulties of trying to establish the
Entropy of a continuous LFP signal, and we shall explore a symbolic approach based
on the decomposition of the signals on the CSD representation. We shall also explore
the notion of block entropy and apply it to this symbolization. Finally we shall
see if those measurements can be used to produce adequate estimations of the
Mutual Information and Transfer Entropy, so that we can explore their
usability as causality tests.

\section{Shannon Entropy}

Shannon Entropy was devised as a measure of the complexity of a message encoded in discrete symbols. This is very important because its strict definition works very well in the case that the symbolization is given, but it fails ( spectacularly one may say ) if one takes instead a possible continuum of symbols, or one imposes another encoding on the symbols. Claude Shannon himself wrongly stated that the continuous case was a limit form
of his original definition \cite{Shannon1948}, and although this was corrected
by subsequent researchers in the 50's and 60's, the error still persists in statements
that claim, as an example, that ``the entropy of the Normal distribution is equal
to the log of the standard deviation'', as for example the Wikipedia article on
Normal Distribution and the references therein \cite{WikiNormal}.

Let us remember Shannon's definition. Suppose that we have a set of discrete symbols, i.e. , a series of separated, well defined values, which could be letters, digits, or any other symbol of such a kind. For simplicity we represent them as numbers, so $n$ discrete symbols would be $\{1, 2, \ldots n \}$. We call such a set an alphabet. Suppose that we have a random process that picks one by one these symbols and gives us a string of them. Such a string would be called a "message". Now let us suppose that the process is ergodic, that means, a very long "message" would be a good representation of the probabilities of the process that picks the symbols. In such a case the process would eventually be able to produce all the possible messages that those symbols permit, and their frequency's would represent the probabilities of the random process. If we knew precisely the probabilities $P$ of every symbol $s_j$, the entropy would be:
\begin{equation}\label{eqDiffEnt}
  H=-\sum_{j=1}^n P(s_j) \log (P (s_j))
\end{equation}
The base of the logarithm is usually taken as two, in which case the Entropy has as units the bit. In that case it can be interpreted as the average number of binary option question one has to make to guess symbol correctly.  A change of base of the logarithm is simply a change of units. We shall call this definition and only this one Shannon's Entropy.
We may not have the knowledge of the $P(s_j)$ directly, so we must infer them from the messages that we sample, and by doing so we may say that we are measuring the Entropy of the messages and not of the process. This has to be taken into account if the messages are not "random enough".

Shannon, in his original paper, stated that the same idea could be done by extending it to continuous distributions:
\begin{equation}
H=-\int p(x) \log (p(x)) d x
\end{equation}
for a random continuous variable and a probability distribution $p(x)$. It turns out that the limit does not converge. One can make some algebraic tricks to make it converge, but that would affect the interpretation. The fact that this looks identical to the Thermodynamic Entropy definition gives a false sense of legitimacy. Edwin T. Jaynes \cite{Jaynes1957}  made a rigorous critique of this confusion and showed how a true relationship between Thermodynamic and Shannon's Entropy can be found. Sometimes this formula is
labeled ``Differential Entropy'', but it has a hidden assumption.
The Differential Entropy would correspond to the Kullback-Leibler divergence
with respect to a homogeneous probability, and that only makes sense if one
has a finite support!
Our way out of this incorrect usage shall be to honestly accept that real measurements
of LFP do indeed have finite support into our derivations. That actually is more sensible than taking the analytical formula without criticism and is according to the nature of real experimental data. This approach leads to the so called ``Limiting density
of Discrete Points'' \cite{Jaynes1968, Jaynes2003}, a special case of the
Kullback-Leibler Divergence \cite{Kullback1959}.

\section{Binarization of the Data and Symbolic Sequences}

Authors take at face value the equation \ref{eqDiffEnt} and try to make a numerical aproximation to it. This is achieved first by making an heuristic estimate to the distribution function of continous data. The usuall approach is first to make an histogram of the data. By so doing one is actually making a symbolization of the data ( called binarization, from bins, not from binary digits ), and if one applies Shannon's formula, then one is actually calculating the entropy of the distribution of symbols from that specific binarization. One could optimistically think that such a calculation could be a reasonable approach to a limiting form of the Differential Entropy, but that is not the case. A simple example sufices.
If one has a finite set of symbols ( we shall call it an alphabet from now on), such as $\{1,2\ldots,n\}$ and a random process that picks them with equal probabilies, the Shannon Entropy either of the process or of the distribution that dictates the process is 
$$
H=-\sum \frac{1}{n} \log \frac{1}{n} = \log n
$$
A simple proof shows that this has the highest entropy for a process with $n$ symbols, taking for example the entropy of one symbol to be higher or lower than the rest, and normalizing the others so that they are still equiprobable. Then It can be shown that this probability function has lower entropy.

If one takes a continous interval over the real numbers, say, from $a$ to $b$, and one puts a flat probabilty distribution function over it, the Differential Entropy will be:
$$
H=-\int \frac{1}{b-a} \log \frac{1}{b-a} d x = \log (b-a)
$$
This somehow paralels the discrete Shanon entropy for equiprobable process, but it hides a trap.

Supose that we want to obtain the above by means of sampling a process, and that we put the data in bins of
equal length. After enough samples we may have a reasonable approach to an even distribution, but we are carefull
and don't take that for granted, so we approach the integral by a numerical discrete sum. Let us suppose that we have $n$ bins, and that each one turned to be more or less equally filled with data. Say that each bin holds 1/n of the total amount of data. Then we would obtain Shanons entropy for $n$ symbols, $H=\log n$. If we would try instead to make a Rieman sum over the expression $ p(x) \log (p(x))$ we end with another result: The bins have lenght $\Delta x= (b-a)/n$, and $p(x)=1/(b-a)$. 
the Integral form by the next sum:
$$
\tilde{H}=-\sum p_n(x) \log p_n(x) \Delta x \\
=-\sum \frac{1}{b-a}\log (\frac{1}{b-a}) \frac{b-a}{n} \\
= \log (b-a)
$$
The Riemann sum in this case converges trivialy to the Integral form. But the numerical sampling does not! In fact, if we refine the partition of our data (and have enough data to spread them over the new bins), the Shannon Entropy begins to grow, as $\log n$. What is going on? Two things. The most overlooked is that $p(x)$ *is not* a probability: it needs to be multiplied by $dx$ to become one. The other problem is that  Shannon's entropy is the entropy of \emph{the partition itself}. The partition induces probabilites (discrete ones) and is a symbolization of the data. The Riemman Integral does not depend on the partition, as long as it converges uniformly. The Entropies do not match each other because they are measuring different things.

People familiar with Ergodic Theory shall recognize that these kind of measures, which depend on the partition, are usefull, but one cannot omit stating the partition when using them. Moreover, given that a partition is a symbolization, how can we chose the right partition? We cannot approach with experimental data the Riemman sum, except on very simple examples. 
In the following we shall use some experimental data
to drive the point.

\section{ First Experimental Numerical Example }

We shall show some calculations derived from using naively the mentioned approach. Our data comes from electrophysiological experiments, recorded with the BioCAM form BrainWave, a MMEA that has 4095 recording electrodes over an area of 2.27mm by 2.27 mm. An extra
electrode is used as ground value. In the data shown below, the data sampling rate was of 7.022 KHz.

\subsection{The binarization process: different criteria}

There are different rules to select adecuate bin sizes in one-dimensional data. The rules try to make the histogramm "smooth enough" and without gaps. We are compeled to do so, because we asume that the phenomena that produces our data is continous in nature, and therefore, it induces a continous probabilty density function. We are looking for a histogram that captures the general shape of this pdf. If the bins are to narrow, we may end up with many empty bins, or many bins with just one data point inside it. If they are too big they may not look like the shape of the underlying distribution.

Below we compare three  methods that produce different bin sizes for the same data: The Freedman and Diaconis rule \cite{FD1981}, the Scott's rule \cite{Scott1979} and the square root of the number of data rule. The first two use statistical characterization over the values of the sample, such as the interquantil range or the standard deviation. Those are measures of the spread of the center of the distribution. The last one only takes into account the number of samples. Also we set up an "heuristic" binsize, in which we do not take into account anything of the sample, but we simply pick a binsize, and set up a range. In our case the range is from $-2000 \mu V$ to $2000 \mu V$

\section{Block Entropy}

Another caveat, often overlooked, is that Entropy of an array of symbols is invariant to ordering. One could in principle regard a signal as a sampling of random variables, reorder them, and obtain the same distribution, and thus, the same entropy. As an example, we could take every letter, punctuation mark and space of this paragraph, garble them into nonsense, and claim that they have the same Entropy as the original paragraph, and therefore, the same Information, which is as much nonsense as the random letters. But actually this is done implicitly in a lot of numerical estimations of Entropy.

\bibliography{./BiblioReportes01}{}
\bibliographystyle{plain}

\end{document}

### 2020-06-03 Entropy/sonidoneuronas01.jl

@@ -1,0 +1,43 @@
+using HDF5, StatsBase, Statistics,PyPlot
+push!(LOAD_PATH, "../Preprocesamiento/")
+using ArraySetTools, OrdenaSets
+
+using WAV
+
+arxname="/home/karel/BRWFiles/Facilitada/Completo_19115s1cut_single_event_preproc.h5"
+arx=h5open(arxname)
+names(arx)
+
+canales=read(arx["CanalesBuenos"]);
+malos=read(arx["CanalesMalos"]);
+#lfp=read(arx["dset"]);
+lfp=read(arx["LFPSaturados"]);
+#freq=read(arx["SamplingRate"])/1000
+freq=read(arx["freq"])
+
+dd=size(lfp)
+if length(dd)==2
+    nmax=dd[2]
+    lfp=reshape(lfp, 64,64,nmax);
+end
+nmax=size(lfp,3)
+typeof(lfp)
+
+decibelear=-1
+fs=8000
+
+largito=repeat(lfpex, inner=12).*(10.0^decibelear);
+wavplay(largito, fs)
+
+lfpexmalo=lfp[62,4,:];
+
+largito2=repeat(lfpexmalo, inner=6).*(10.0^decibelear);
+wavplay(largito2, fs)
+
+lfpexrui=lfp[30,30,:];
+
+largito3=repeat(lfpexrui, inner=12).*(10.0^decibelear);
+wavplay(largito3, fs)
+
+
+

### 2020-06-04 Reportes/Reporte20200601.tex
reporte avance

@@
 \DeclareMathOperator{\grad}{grad}
+\DeclareMathOperator{\iqr}{iqr}
 
@@
 
-Authors take at face value the equation \ref{eqDiffEnt} and try to make a numerical aproximation to it.
-This is achieved first by making an heuristic estimate to the distribution function of continous data.
-The usuall approach is first to make an histogram of the data.
+Authors take at face value the equation \ref{eqDiffEnt} and try to make a numerical approximation to it.
+This is achieved first by making an heuristic estimate to the distribution function of continuous data.
+The usually approach is first to make an histogram of the data.
 By so doing one is actually making a symbolization of the data ( called binarization, from bins, not from binary digits ), and if one applies Shannon's formula, then one is actually calculating the entropy of the distribution of symbols from that specific binarization.
 One could optimistically think that such a calculation could be a reasonable approach to a limiting form of the Differential Entropy, but that is not the case.
-A simple example sufices.
-If one has a finite set of symbols ( we shall call it an alphabet from now on), such as $\{1,2\ldots,n\}$ and a random process that picks them with equal probabilies, the Shannon Entropy either of the process or of the distribution that dictates the process is
+A simple example suffices.
+If one has a finite set of symbols ( we shall call it an alphabet from now on), such as $\{1,2\ldots,n\}$ and a random process that picks them with equal probabilities, the Shannon Entropy either of the process or of the distribution that dictates the process is
 $$
@@
 
-If one takes a continous interval over the real numbers, say, from $a$ to $b$, and one puts a flat probabilty distribution function over it, the Differential Entropy will be:
+If one takes a continuous interval over the real numbers, say, from $a$ to $b$, and one puts a flat probability distribution function over it, the Differential Entropy will be:
 $$
@@
 $$
-This somehow paralels the discrete Shanon entropy for equiprobable process, but it hides a trap.
+This somehow parallels the discrete Shannon entropy for equiprobable process, but it hides a trap.
 
-Supose that we want to obtain the above by means of sampling a process, and that we put the data in bins of
-equal length. After enough samples we may have a reasonable approach to an even distribution, but we are carefull
+Suppose that we want to obtain the above by means of sampling a process, and that we put the data in bins of
+equal length. After enough samples we may have a reasonable approach to an even distribution, but we are careful
 and don't take that for granted, so we approach the integral by a numerical discrete sum.
@@
 Say that each bin holds 1/n of the total amount of data.
-Then we would obtain Shanons entropy for $n$ symbols, $H=\log n$.
-If we would try instead to make a Rieman sum over the expression $ p(x) \log (p(x))$ we end with another result: The bins have lenght $\Delta x= (b-a)/n$, and $p(x)=1/(b-a)$.
+Then we would obtain Shannon's entropy for $n$ symbols, $H=\log n$.
+If we would try instead to make a Riemann sum over the expression $ p(x) \log (p(x))$ we end with another result: The bins have length $\Delta x= (b-a)/n$, and $p(x)=1/(b-a)$.
 the Integral form by the next sum:
@@
 $$
-The Riemann sum in this case converges trivialy to the Integral form.
+The Riemann sum in this case converges trivially to the Integral form.
 But the numerical sampling does not!
@@
 The other problem is that  Shannon's entropy is the entropy of \emph{the partition itself}.
-The partition induces probabilites (discrete ones) and is a symbolization of the data.
-The Riemman Integral does not depend on the partition, as long as it converges uniformly.
+The partition induces probabilities (discrete ones) and is a symbolization of the data.
+The Riemann Integral does not depend on the partition, as long as it converges uniformly.
 The Entropies do not match each other because they are measuring different things.
 
-People familiar with Ergodic Theory shall recognize that these kind of measures, which depend on the partition, are usefull, but one cannot omit stating the partition when using them.
+People familiar with Ergodic Theory shall recognize that these kind of measures, which depend on the partition, are useful, but one cannot omit stating the partition when using them.
 Moreover, given that a partition is a symbolization, how can we chose the right partition?
-We cannot approach with experimental data the Riemman sum, except on very simple examples.
+We cannot approach with experimental data the Riemann sum, except on very simple examples.
 In the following we shall use some experimental data
@@
 
-There are different rules to select adecuate bin sizes in one-dimensional data.
-The rules try to make the histogramm "smooth enough" and without gaps.
-We are compeled to do so, because we asume that the phenomena that produces our data is continous in nature, and therefore, it induces a continous probabilty density function.
+There are different rules to select adequate bin sizes in one-dimensional data.
+The rules try to make the histogram "smooth enough" and without gaps.
+We are compelled to do so, because we assume that the phenomena that produces our data is continuous in nature, and therefore, it induces a continuous probability density function.
 We are looking for a histogram that captures the general shape of this pdf.
@@
 
-Below we compare three  methods that produce different bin sizes for the same data: The Freedman and Diaconis rule \cite{FD1981}, the Scott's rule \cite{Scott1979} and the square root of the number of data rule.
-The first two use statistical characterization over the values of the sample, such as the interquantil range or the standard deviation.
-Those are measures of the spread of the center of the distribution.
-The last one only takes into account the number of samples.
-Also we set up an "heuristic" binsize, in which we do not take into account anything of the sample, but we simply pick a binsize, and set up a range.
-In our case the range is from $-2000 \mu V$ to $2000 \mu V$
+Below we compare three  methods that produce different bin sizes for the same data.
+The Freedman and Diaconis rule \cite{FD1981}, takes as ruler the interquantile
+range. This makes the criterion stable with respect to outliers. The exact expression
+is
+\begin{equation}
+  l_{FD}=2 \iqr / \sqrt[3]{N},
+\end{equation}
+where $\iqr$ is the interquantile range and $N$ is the number of data entries.
+A very similar one is Scott's rule \cite{Scott1979}:
+\begin{equation}
+  l_{S}=3.49 \sigma / \sqrt[3]{N},
+\end{equation}
+where $\sigma$ is the standard deviation. If the data behaves more or less like a
+Gaussian, it gives a good bin size.
+We also use the ``square root'' rule, in which we use as many data bins as the
+square root of the number of data, a usual rule of thumb for obtaining histograms.
+Also we set up an "heuristic" bin size, in which we do not take into account anything of the sample, but we simply pick a bin size, and set up a range.
+In our case the range is from $-2000 \mu V$ to $2000 \mu V$, and bin size of $4.7 \mu V$.
+
+After obtaining histograms for each channel using each of the criteria above, we
+calculate Shannon's Entropy on each channel. In the figure \ref{entrocomp} we compare
+the results. 
+\begin{figure}
+  \includegraphics[width=0.9\textwidth]{ComparaEntropias01.png}
+  \caption{Comparison of Shannon's Entropy of the same data, but
+    different binning. The criteria produce qualitatively different results.}
+  \label{entrocomp}
+\end{figure}
+
+\subsection{Discussion of the Entropies with different bin sizes criteria}
+
+Inspecting the above figures, it shall be obvious that Shannon's Entropy depends highly on the partition of the range of data.
+The different color maps don't differ only in scale: they behave qualitatively in contrary ways.
+Both the Scott and the Square Root show the opposite behaviour as the Freed-Diaconis and the Heuristic fixed method.
+In the former higher entropy is on the noise channels, the later have higher entropy on the very active CA3 region and their neighbourhood (we discuss the count-map method at the end).
+Why is that and what does it mean?
+
+Let us recant some properties of Shannon Entropy.
+In a finite alphabet, higher entropies imply a flatter distribution.
+If all the symbols have the same probability, we get the maximum entropy.
+Also, a distribution with more symbols tends to also have a higher entropy, unless the symbols have very different probabilities.
+The different criteria for obtaining the optimal bin size take into account that there may be a continuous distribution underneath it, so we are optimizing in search of it.
+But every channel produces a different bin size.
+Scott and Square root methods are very sensible to outliers, (the second one is actually dependent solely on the extrema, that is, the more outliers of them all), so they tend to produce larger bins.
+If the channel is active, we end with some bins that are large enough to contain all noise contribution and others that contain the activity.
+On the other hand,the noisy channels produce smaller bins, and the distribution of noise tend to be more homogeneous, like a truncated Gaussian.
+This will produce a higher entropy, as we have more bins, with a relatively flatter distribution.
+
+In the Freed Diaconis method, we use the interquantile range as a ruler, so it is more stable in presence of outliers.
+That makes the bin size more consistent across channels, and therefore, the active channels tend to have a larger spread of probabilities over more or less the same alphabet, giving more Entropy to channels with more diverse signals.
+The heuristic criterion behaves more like this, as we use the exact same bin size and number of bins for all channels.
+
+It turns out to be that histograms that are optimized to find a continuous distribution are inconsistent to estimate the Shannon Entropy, and this is rightly so: the Differential Entropy is \emph{not} the limit of Shannon's entropy, so that any continuous distribution can not induce a consistent discrete entropies that converge to it.
+
+This gives us the reason to explore our true limiting partition.
+Our apparatus has an intrinsic discrete alphabet: the measurements are given in integer multiples of $2.014 \mu V$, so our symbols are actually those discrete values of the electric Voltage.
+The count-map function gives us an histogram of those discrete values.
+The entropy thus calculated is qualitatively similar to the Freed Diaconis Criterion, but with higher values, as it maximizes the symbols.
+Moreover, it also avoids misplacing measurements in the same basket, leaving other symbols empty- this is the correct alphabet for measuring the Entropy of
+the signal: is consistent across channels and it minimizes our assumptions. But it shall
+not generalize to more complex measures, such as Mutual Information or
+Transfer Entropy, as we shall see below.
 

### 2020-06-11 Reportes/Reporte20200601.tex
Extreme Symbolization

@@
-\documentclass{article}
+\documentclass[10pt]{article}
 
@@
 
+\subsection{Diggresion: using countmap to discover numerical artifacts on
+  the data}
+
+When using the countmap method, we obtain a frequency histogram of discrete values.
+This is so because the apparatus gives us discrete values, it is the nature
+of any experiment to have finite resolution. The BioCam in this data has a
+minimal resolution of exactly $\Delta V=2.01416015625 \mu V $.
+When plotting the frequency
+of each value we end up with a figure as seen in fig, \ref{histocountmap1}.
+Using a coarser binning to produce the histogram would have made us miss
+this efect, or at least made it more dificult to discover.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.65\textwidth]{histoconcorazones01.png}
+    \end{center}
+ \caption{Absolute frequencies of the discrete values of all active channels.
+   Some values clearly deviate from the statistic, which is apparently
+   Gaussian. These atypical values seem to be more or less
+   evenly spaced, indicated by red marks here.}\label{histocountmap1}
+\end{figure}
+
+The spacing of the suspicious values near the center seems to be
+around $12 \Delta V \approx 25 \mu V$, but other differences appear.
+There seems no sensible way of how to deal with these values, because
+substituting them randomly could alter the statistical propiertes of the
+signal. Nevertheless we can ignore them when calculating the Entropy
+by the countmap method, as the other values give a well defined shape
+of the Probability Function. Given that the weird behavior is less
+detectable on the edges of the Gaussian, we may only eliminate those
+near the center. A larger range graph is presented on the figure
+\ref{histocountmap2}.
+
+\begin{figure}
+  \begin{center}
+ \includegraphics[width=0.85\textwidth]{histoconcorazones02.png}   
+  \end{center}
+
+ \caption{Broader view of the same daa as figure \ref{histocountmap2}.
+   A very suspicios irregularity is seen, but is very dificult to
+ quantify it.}
+ \label{histocountmap2}   
+\end{figure}
+
+In contrast, a histogram would smooth those strange values, making it
+difficult to detect. In the figure \ref{histoscott01} we show the
+effect of using the Scott's Method for the histogram. As can be seen,
+the ``bad values'' are much more dificult to spot, and the
+histogram looks really weird. Any further method that we apply over such
+probabilistic calculation will have unreliable or false effects. This reveals
+the necesesity to be very very strict with the Histograms that we calculate.
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width=0.65\textwidth]{histoscottbuenos01.png}  
+  \end{center}
+    \caption{The same data as on previous figures but binned with Scott's method.}
+  \label{histoscott01}
+\end{figure}
+
+Let us inspect the data a bit more. How do those values look on the trace of
+the signal?
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.95\textwidth]{trazosospechos01.png}
+  \end{center}
+  \caption{An exemplary trace with the suspicious values marked. Note that
+  many occur on local maxima or minima.}
+  \label{trazomarcas01}
+\end{figure}
+
+In the figure \ref{trazomarcas01} we can see that the strange values occur often
+at turning points in the data, marking some characteriztic oscillation of the noise.
+This is clearly an artifact of the electrodes measuring the potential, as they
+seem more prone to round of to certain values at the moment in which the
+potential changes sign of its temporal derivative. Although a denoizing filter may
+help us to clean those values, I suspect that other artifacts may acumulate and
+alter the value of the Entropy.
+
+\section{Higher dimensional and causality measures}
+
+We have established that the method of counting the discrete values of
+the measured LFP is the one that gives a trustworthy estimation of the
+probabiliy function. In some sense, it is a good representative
+of the probabilites of the \emph{measured values} of the data.
+As a side effect, it also helps us to reveal numerical artifacts
+on the data. Without a relieable measure of the Probabilites
+we cannot  calculate causality functions, such as Mutual Information and
+Transfer Entropy, which are our true aim.
+Could we use the countmap method for such calculations?
+The answer is no. The problem is that the probabilites of joint events
+decrease as a power of the number of events that we are measuring: the data
+points get diluted in a higher dimensional space. We are seeing
+discrete events that are sampled from a process that has around 1000 symbols.
+They appear to be more or less normaly distributed, with a standard deviation
+of around 12 simbols. So around the central 50 simbols we have 95 percent of all
+events (the 68-95-99 rule of gaussian distributions).
+In a very rough Fermi approximation these symbols could be thought of being
+equiprobable. If we measure the joint probability of two channels,
+the appareance of a certain pair
+of symbols in those ranges is around 1/2500. We have around 2000 samples in
+the example experiment. That means that we are going to observe each pair
+at most once, which makes a very poor statistic indeed.
+If that were not bad enough,
+the central symbols account mostly for the background and thermal noise.
+Those are the more probable pairs of symbols that we are going to see: the
+ones corresponding to spikes are very rare, and to see
+two correlated spikes is going to be an extremely rare event with this
+resulution and number of samples.
+
+To ilustrate this, we shall select two exemplary channels and
+determine the joint distribution of their symbols. We select channels from
+the very active CA3 region. Channels labeled $A=[50,28]$ and $B=[60,35]$ have both
+around 40 putative spikes. We shall plot the joint distribution of events,
+$P(X_A, X_B)$ in the figure \ref{histopAB}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.8\textwidth]{histopAB.png}
+  \end{center}
+  \caption{Frequency of the order pair of values for Channels $A$ and $B$.
+    Both axis are in $\mu V$.  The color indicates the frequency of each pair
+  of values.}
+  \label{histopAB}
+\end{figure}
+
+Most of the values in the plot are zero. Some of them are 1, and the peak of
+the distribution is exactly at zero, with 10 ocurrences of the event $(0,0)$.
+This could be highly disappointing: the only probable correlated occurences are
+just noise. On closer inspection we find some valuers that show possible correlation
+beyond the noise center. The pair of values $(215.5 \mu V, 88.6 \mu V)$ occurs twice.
+This is still very poor but maybe it could point to something.
+
 \section{Block Entropy}
@@
 
+1
 \bibliography{./BiblioReportes01}{}

### 2020-06-18 Reportes/Reporte20200601.tex
reporte extendido

@@
 Transfer Entropy, which are our true aim.
+Let us recall the definition of the first measure. Let us say that
+$y$ is a result of the random process $Y$ and $x$ of the random
+process $X$ (for simplicity in the notation we shall identify
+the process with its encoding alphabet, and from now on we shall
+restrict ourselves to discrete process). Mutual Information
+between process $Y$ and $X$ is given by \cite{Cover1991}:
+\begin{equation}
+  I(X;Y)=\sum_{x \in X} \sum_{y \in Y} P(x,y) \log \frac{P(x,y)}{P(x)P(y)}
+\end{equation}
+where $P(y,x)$ is the probability of the joint event and the other are
+individual probabilities. Here is where our problem grows in dimensionality.
+These joint probabilities have as domain a cartesian product of the
+separated spaces $X$ and $Y$. If one has $n$ symbols in one, and
+$m$ symbols in the other, the joint space alphabet would have $ n \times m$ 
+symbols.
 Could we use the countmap method for such calculations?
@@
 
+\section{Symbolization strategies}
+
+The problem with the histograms is that we induce a certain partition of the
+data, and there it is a certain degree of arbitrairness. We have also seen
+that partitions that are too fine would be useless for the higher dimensional
+measures. We shall test some symbolization strategies that do not depend
+on the exact values of the data, as we have also tested that some of these
+values may be unreliable due to measurement artifacts.
+Steuer \emph{et al.} \cite{Steuer2004} used a two symbol system based on
+the finite time derivative of the signal. I shall use the same basic idea
+but with three symbols, using $1$ for increasing signal, $0$ for stationary, and
+$-1$ for decreasing signal.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=1.0\textwidth]{entropia2s01.png}
+  \end{subfigure}
+   \begin{subfigure}[b]{0.49\textwidth}
+    \includegraphics[width=1.0\textwidth]{histo2s01.png}
+  \end{subfigure}
+   \caption{The entropy for 3 symbols over all channels  and its histogram. The later was done using only the putative active channels. No discernible features are visible,
+     except the unusable channels, that have zero entropy. Compare with figure
+     \ref{entrocomp}. }
+   \label{entro2s01}
+\end{figure}
+
+In the figure \ref{entro2s01} we show the results of this symbolization procedure.
+As one can see, there is no range and therefore no contrast between the
+active channels and the inactive ones. This symbolization is not adecuate. The
+noise, which varies rapidly between increasing and decreasing signals, drowns
+out all features of neuronal activity when reduced to such a simple characteristic.
+Steuer \emph{et al.} do not provide clues of the timestep they used, but the caveat
+is there. I propose to extend the alphabet in the following manner.
+Instead of considerin one time step, which is a sampling step of the BrainWave
+BioCAM, we shall use time windows that are of the order of magnitud of physiological
+events, around 1ms. In each window we shall calculate the sums of the signs
+of the differences, that is, sum the signs of the former symbolization. The new symbols
+will be a measure of how much was moving the signal in the positive or negative
+direction, discretized in reasonable amounts. If a window contains mostly noice,
+the sum will be near zero. Otherwise it shall depart notably from zero
+in a window which encompases the polarization or depolarization of a neuron. A window
+may encompass $n$ frames, so the alphabet shall range from $-n$ to $n$, having
+$2n+1$ symbols. We shall choose the $n$ value as the ones which provides the
+higher contrast between entropies, that is, the largest range of entropies between
+active and inactive channels. The largest spread of entropy values ignoring the
+unusable channels appears at $n=8$, which at this sampling rate is a little longer
+than one milisecond. This could be a good symbolic alphabet for obtaining the
+Mutual Information, although it is still quite large with 17 symbols. 
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.85\textwidth]{comparasumaentropias01.png}
+    \end{center}
+  \caption{Comparission of the summed symbol entropies. Each image corresponds
+    to summing over the stated number of symbols on the header.}
+  \label{compsumentro}
+\end{figure}
+
+In the figure  \ref{h8sim} we
+show the distribution of symbols for the sum over 8 frames, ranging from
+$-16$ to $+16$. The distribution seems consistent across channels, having a
+gross normal appearence with richer structure therein.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=3.0in]{histogramas_8simbolos.png}
+  \caption{Histogram for two channels on the putative active Area, coordinates indicated in the legend. Compare with figure \ref{entro2s01}.}
+  \label{h8sim}
+\end{figure}
+
 \section{Block Entropy}
@@
 But actually this is done implicitly in a lot of numerical estimations of Entropy.
+In the last section
+I proposed adding the symbols of a certain time window to obtain a richer symbolization,
+and thus describe a larger set of messages. There is more detailed version
+of this approach. The idea is to regard finite succesion of symbols as
+higher order symbols, in a larger alphabet, a vocabulary in some sense. The entropy
+that one obtains calculating $n$ symbol words is called ``n-block Entropy''.
+This measure of entropy does not converge as we take larger words. As we have
+established, Shanon's Entropy grows as the logarythm of the size of the
+alphabet, or in this case, the vocabulary. If the signal is rich in pattens, and
+long enough, many words shall appear. What it is done here is to establish
+an average entropy increasse per new symbol. If Shannon's Entropy for the $n$ block
+vocabulary is denoted $H_n$, the average increase in entropy for each new symbol
+will be:
+\begin{equation}
+    h_n=(H_{n+1}-H_n)    
+\end{equation}
+This quantity converges under normal circumstances, and its limit is called
+the \emph{source entropy}. Unconveniently, it converges very slowly, so in principle
+we shall use very large words to obtain a good approximation. This becomes a
+problem, because we are constrained by the data set length to obtain adecuate
+probabilities in the estimation of this entropy. Let us
+think that the data length is $M$ symbols
+long, and we are trying to establish $n$ word probabilities. It the alphabet
+contians $N$ different symbols, there are $N^n$ possible words. If they were
+equiprobable, in order to detect on average once each word, t
+he next relationship should hold:
+\begin{equation}
+  1 \approx M/(N^n)
+\end{equation}
+that means that to be marginally able to detect an arbitrary $n$ length word,
+$n$ shall be at most:
+\begin{equation}
+  n < \log(M)/\log(N).
+\end{equation}
+or, equivalently, $ n < \log_N (M) $, that is, the logarythm of the length of
+the measured data in the base of the symbol alphabet size. In a concrete case,
+if we have $2000$ data points per channel, and I propose a three symbol
+alphabet, we cannot realiably measure probabilities beyond $n=7$. As a curious
+note, this is very close to the best case in the  ``summed'' symbol that
+I proposed on the last section. The similitude may derive in that
+the summed symbol is a ``compressed encoding'' of the word symbol, and that
+both may cointain similar information, altough on second looks,
+the summed symbol, especially if the range  of the sum is relative large, may
+encode as the same symbol very different signal segments, and the compression
+is not homgeneous. The quantity of words $W_3$ that add up exactly to $m$ using at most
+$n$ symbols choosen from the alphabet $S_3={-1,0,1}$ is
+\begin{equation}
+  W_3=\sum_{j=0}^\nu \frac{n!}{(m-n-2j)!(n+j)!j!}, 
+\end{equation}
+where $\nu$ is the integer part of $(n-)/2$ and represents the pairs of $+1$ and
+$-1$ that add up to zero and don't alter the sum. as an example we can calculate
+how many words are commpresed in the 8 symbols sum for different values of $m$.
+\begin{center}
+  \begin{tabular}{ l | r }
+    $m$ & $W_3$ \\
+    \hline
+    0 & 1107 \\
+    1 & 1016 \\
+    2 & 784 \\
+    3 & 504 \\
+    4 & 266 \\
+    5 & 112 \\
+    6 & 36 \\
+    7 & 8 \\
+    8 & 1 \\
+    \hline
+  \end{tabular}
+\end{center}
+The most extreme case is the $m=0$, where there could be such disimilar symbols
+compresed as the sequence $0,0,0,0,0,0,0,0$, a completely flat signal,
+and the sequence $-1,-1,-1,-1,+1,+1,+1,+1$, which could be a spike! It is interesting
+to note that even so, the entropy thus obtained is concentrated in the
+physiological
+expected areas, namely CA3. After some pondering, it turns out that
+the lossy compression is only effective if we perform the change of dictionary
+( the sum over windows) if we do not make a running window, but a time stepped
+window. The running window does not lose in general the ternary message, because
+after $n$ time steps, we have enough linear equations to recover the
+subyacent signs. Therefore, the succesive messages are not independent. They
+follow a very specific and strict succesion, and that produces higher entropies,
+revealing the structure of the original complex message that is subyacent to the
+encoding. Heuristically, that also could point out to the fact that
+the optimal summing window has 8 time frames, and the best block entropy
+that we can calculate is with 7 time frames.
 
-1
+We should not be to hasty to regard the expanded and detailed vocabulary
+as the best simbolization strategy. As before, excess symbols may actually be a
+hindrance to a good interpretation of the entropy measure. Let us see the
+entropies from 1 to 6 symbols in the figure \ref{entro6symbs}.
+
+\begin{figure}
+  \includegraphics[width=0.8\textwidth]{comparaentrobloc01.png}
+  \caption{Block Entropies for words with three symbols from the
+    set $\{ -1, 0, 1 \}$. 
+  }\label{entro6symbs}
+\end{figure}
+
+A bit to the surprise and dismay of the author, the Block Entropy for this
+data provides less contrast (and therefore less information about the channels)
+than the sum of the symbol entropies. A plausble explanation is that the
+more detailed symbolic set flattens the distribution, making noisy and
+active channels more alike in this encoding. We shall test this numerically
+on the next figure, \ref{histo6simbols}. The histogram looks terribly
+messy and without any ordering, much in contrast with the one in figure
+\ref{h8sim} where a natural ordering seems to be present. Worse yet,
+the comparision with a noisey channel doesn't seem to yield any qualitative
+differences, but a sort of grouping of the probabilites that seem similar
+in both cases, thus indicating that this symbolization is better at
+describing the noise than the physiological signal.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{\textwidth}
+  \includegraphics[width=0.8\textwidth]{histo6simbols.png}  
+  \end{subfigure} \\
+  \begin{subfigure}[b]{\textwidth}
+  \includegraphics[width=0.8\textwidth]{histo6simbruido.png}  
+  \end{subfigure} 
+  \caption{Distribution for the 6 symbol words in an active
+    Channel. The labels are ordered numerically from
+    $(-1,-1, -1, -1,-1,-1)$ to $(1,1,1,1,1,1)$. In the above
+    plot the distribution for an active channel is shown. Below we see
+    a graphic for a noisy channel. 
+  }\label{histo6simbols}
+\end{figure}
+
 \bibliography{./BiblioReportes01}{}

### 2020-09-01 ArticuloCSDyCM/Paper03.tex
nuevos statistics para no perder figuras

@@
 
-The spatial tracking activity in neural tissue poses challenging problems,
-if one requires rigor, precision and clear interpretability. 
-Measurement recordings
-are usually presented  as a series of traces that correspond each to a sampling
-electrode on the tissue. Then a series of causality tests are made on the
-signal to search for possible effective connections. 
-As Micro Electrode Arrays (MEAs) become denser, that representation
-of the data saturates itself  and the discovery of patterns becomes a difficult task.
+Tracking the activity in neural tissue poses challenging problems, if one
+wants a clear, spatial and physiologicall interpretation.
+
+Electro-physiological recordings usually consist of 
+ series of traces, each one corresponds to a sampling
+electrode on the tissue. Causality tests are made between the
+signals in search for effective connections. 
+As Micro Electrode Arrays (MEAs) become denser,the data saturates itself
+and the discovery of patterns becomes a difficult task.
+Testing for combination of signals in large numbers
+brings the so called ``combinatorial explosion'' into
+play. 
 Novel information is obfuscated by the same high density acquisition that allows us to
@@
 separation analysis, such as ICA.
-We present here  a method for tracing out
+In this paper we present a method that aims rather to integrate
+the signals into functional components that
+have physiologicall interpretation.
+This is a technique for tracing out
 ``functional local connectivity'', which would bring independent information
-that is complementary to blind source separation.
+that is complementary to Blind Source Separation.
 Our  method would be applicable mostly to
@@
 Local field potential (LFP)
-can be regarded as generated by local multi-unitary activity.
-A patch of simultaneous, highly correlated, nearby traces
-can be regarded as one of those multi-unitary elements. 
+is generated by local multi-unitary activity.
+A set of simultaneous and neighbouring traces
+can be regarded as representant
+of those multi-unitary elements. 
 Successive activation  of distinct neighboring elements on structured
@@
 The MEA should have dense enough spatial and temporal samplings, so as to obtain good estimates of the CSD and their changes in time.
-
+1
 Our method permits us to map possible trajectories of activity inside structured neuronal tissue, without requiring ad hoc geometric constraints.

### 2020-09-01 CenterOfMass/CentrosdeMasa.jl

@@ -48,11 +48,9 @@ function ComponentesSP(DatosSignados::Array)
             y=pop!(listaprofundeza)
             for v in vecindad8(y)
                 if in(v, lista)
-                    
-                    
                     deleteat!(lista, indexin(Any[v], lista))
                     push!(listaprofundeza, v) 
-                   profundidad+=1
+                    profundidad+=1
                     push!(componentecurlab, v)
                 end
             end

### 2020-09-01 CenterOfMass/TrayectoriasAux.jl

@@ -8,9 +8,20 @@ module TrayectoriasAux
 
 using Statistics
 
-export reducecatenario, AntesQue, DespuesQue, distprom
+export reducecatenario, AntesQue, DespuesQue, distprom,
+    dist, set2dict
 
 dist(x1,y1,x2,y2)=sqrt((x1-x2)^2+(y1-y2)^2)
+
+function set2dict(xx::Set)
+    j=1
+    result=Dict{Integer, Any}()
+    for x in xx
+        result[j]=x
+        j+=1
+    end
+    return result
+end
 
 function reducecatenario(Catenario::Dict;
                          gordmin=25,minlong=7,maxlong=100, distmin=0)
@@ -46,8 +57,8 @@ function cortacatenarios(Catenario::Dict; nini,nfin)
         if length(paux) != 0
             result[k]=paux
         end
-        
-    end
+         
+   end
 
     return result
 end
@@ -94,6 +105,11 @@ function distprom(Catenario::Dict)
         result[k]=d
     end
     return result
+end
+
+function distprom(Catenario::Set)
+    aux=set2dict(Catenario)
+    distprom(aux)
 end
 
 end # module

### 2020-09-01 Isabel/CortarMatrix.jl

@@ -10,7 +10,9 @@ using HDF5
 arxname=ARGS[1]
 
 #memoria usada
-pmem = filesize(arxname)/Sys.total_memory()
+ramsiz=Sys.total_memory()
+pmem = filesize(arxname)/ramsiz
+
 
 #000000000000 Funciones chidas
 
@@ -61,8 +63,9 @@ function cortayguarda(filenombre::String, pmem=0.125)
     nch=size(chans, 1)
     nframes=Int(M/nch)
     
-    
-    ncachos=max(ceil(Int, 1/(1-pmem)), 8) # al menos partela en 4
+
+    nrams=ceil(Int, filesize(filenombre)*8/Sys.total_memory())  
+    ncachos=max(nrams, 8) # al menos partela en 4
     fcacho=floor(Int, nframes/ncachos) # frames por cacho dara un poco menos que el total.
         
     paso=nch*fcacho #cuantos datos tiene cada cacho.

### 2020-09-01 Isabel/Z_auxiliaresBRW.jl

@@ -9,13 +9,20 @@ function brw_things( file_brw::String ) # depende de brw file
     Chs = read( brw[ "/3BRecInfo/3BMeaStreams/RawRanges" ] )[ "Chs" ]; 
     x = zeros( Int, size( Chs, 1 ), 3 );
     for i = 1:size( Chs, 1 )
-	x[ i, : ] = [ i, Chs[ i ].data[ 1 ], Chs[ i ].data[ 2 ] ]; # Channels, [number coord1 coord2]
+        ## Los genios de 3Brain le volvieron a cambiar los nombres a las variables
+        ## antes era data con dos entradas
+        ## ahora es Row y Col
+	##x[ i, : ] = [ i, Chs[ i ].data[ 1 ], Chs[ i ].data[ 2 ] ]; # Channels, [number coord1 coord2]
+        x[ i, : ] = [ i, Chs[ i ].Row, Chs[ i ].Col ]; # Channels, [number coord1 coord2]
     end  
     
     ValidChs = read( brw[ "/3BData/3BInfo/3BNoise/ValidChs" ] ); # Valid Channels (software desition)
     y = zeros( Int, size( ValidChs, 1 ), 3 );
     for i = 1:size( ValidChs, 1 )
-	y[i,:] = [ ( ( ValidChs[ i ].data[ 1 ] - 1)*64 + ValidChs[ i ].data[ 2 ] ), ValidChs[ i ].data[ 1 ], ValidChs[ i ].data[ 2 ] ];# Valid Channels, [number coord1 coord2]
+         ## Los genios de 3Brain le volvieron a cambiar los nombres a las variables
+        ## antes era data con dos entradas
+        ## ahora es Row y Col
+	y[i,:] = [ ( ( ValidChs[ i ].Row - 1)*64 + ValidChs[ i ].Col ), ValidChs[ i ].Row, ValidChs[ i ].Col ];# Valid Channels, [number coord1 coord2]
     end
     
     RawEncodedTOC = read( brw[ "/3BData/RawEncodedTOC" ] )[ 2 ]; # numero real de frames registrados

### 2020-11-14 ArticuloCSDyCM/Paper03.tex
simplifique oraciones

@@
 
-Tracking the activity in neural tissue poses challenging problems, if one
-wants a clear, spatial and physiologicall interpretation.
+Tracking the propagation of activity in neural tissue poses challenging
+problems.
 
 Electro-physiological recordings usually consist of 
- series of traces, each one corresponds to a sampling
+ series of traces, each one corresponding to a sampling
 electrode on the tissue. Causality tests are made between the
@@
 Testing for combination of signals in large numbers
-brings the so called ``combinatorial explosion'' into
-play. 
+brings ``combinatorial explosion'' into
+play. This makes the causality tests time consuming and
+prone to produce false positives.
 Novel information is obfuscated by the same high density acquisition that allows us to
 discover it. Representing  the recorded data in ways which can unveil previously
-invisible patterns is a necessity. Various techniques for separation of functional
-generators are available now, most derived from the application of blind source
+invisible patterns is a necessity. Various techniques for separation of signals
+are available now, most derived from the application of blind source
 separation analysis, such as ICA.
@@
 A set of simultaneous and neighbouring traces
-can be regarded as representant
+can be regarded as representant of the action
 of those multi-unitary elements. 
@@
 (r.o.i.) by following separately the distinct poles of action: influx and
-efflux of charged ions.
+efflux of charged ions, which correspond to two distinct phases of the
+neuronal activity: depolarization and repolarization.
 
 The application of the method requires electro-physiological data acquired
-through bi-dimensional MEAs, although the extension to three dimensional arrays would be possible in principle.
-The MEA should have dense enough spatial and temporal samplings, so as to obtain good estimates of the CSD and their changes in time.
-1
+through bi-dimensional MEAs, although the extension to three dimensional arrays
+is straightforward.
+The MEA should have dense enough spatial and temporal samplings,
+so as to obtain good estimates of the CSD and their changes in time.
+
 Our method permits us to map possible trajectories of activity inside structured neuronal tissue, without requiring ad hoc geometric constraints.
 As a prime example, non-convex structures which may show simultaneous activity in separate parts could show the robustness of our method.
-We have tested then the method on slices of rat's hippocampus, concentrating in the CA3 region.
+We have tested the method on slices of rat's hippocampus, concentrating in the CA3 region.
 We exemplify the usage of the method with data that has been obtained on experiments on rat's hippocampal slices that are centered on the GD and CA
@@
 and data is recorded at sampling rates between 7 and 14 kHz. The richness of the
-data permits us to obtain a prime use case for our method. A
- certain amount of down-scaling would be 
+data is an exemplary use case for our method. A certain amount of down-scaling would be 
 possible, but the array of electrodes should be at least 2 dimensional and
 a inter-electrode distance of the order of magnitude of a typical pyramidal soma
-(around \mum{100}).
+(around \mum{100} at most).
 
@@
 such as saturated or failing electrodes. The
-set of fail-ling electrodes varies from data set to data set,
+set of failling electrodes varies from data set to data set,
 as they can be produced by bubbles or by saturation after electrical stimulation
-or other causes. 
+or any other cause. 
 We call this step the elimination of bad channels. We use a simple criterion for this:
 if the absolute value of the signal is above certain threshold (which must be
-adjusted for evoked vs spontaneous activity) or if the standard deviation is
-larger than certain heuristically determined value, the channel is discarded.
+adjusted for evoked vs spontaneous activity, but
+is of the order of \muV{1000}) or if the standard deviation is
+larger than a two times the averague r.m.s.  of the normal noise,
+the channel is discarded.
 
@@
 
-The concepts presented here require first obtaining the CSD from the recorded LFP. Any method would present the same utility, we have used two different
-approaches.
-The first method is a  finite difference operator, while the second is the kCSDA presented by Potworowski \emph{et all} \cite{Potworowski2011}, which is an inverse problem approach.
+The first step is to obtain the CSD from the recorded LFP.
+Any reliable method would work.
+We have used two different
+approaches. The first method is a  finite difference operator,
+while the second is the kCSDA presented by
+Potworowski \emph{et all} \cite{Potworowski2011},
+which is an inverse problem approach.
 Where both methods are usable, we prefer the former,
@@
 the difference method is not applicable. The results are
-independent of the method except for a multiplicative factor which
-can be normalized.
+independent of the method except for a multiplicative factor.
+As there is still another unknown constant factor
+(the resistivity of the medium) this is irrelevant.
+Therefore, all of the CSD results are presented in arbitrary units.
 
@@
 they are outside the r.o.i.
-In the experiments that were made without electrical stimulation that was the case,
-with the order of magnitude of unusable channels in the tens.
-A  numerical Laplacian Operator which reduces the cross effect of rectangular grids is the convolution of the data with the following matrix \cite{Lindberg90}:
+In most examples the number  of unusable channels is in the tens.
+A  numerical Laplacian Operator which reduces the cross effect of
+rectangular grids is the convolution of the data with the following
+matrix \cite{Lindberg90}:
 
@@
 For electrode arrays of lesser density, this could posse a problem, but in our case we renounce to 253 out of 4095 channels (one is grounded), and those are outside the region of interest.
-If very few electrodes fail, more sophisticated methods such as iCSDA \cite{Leski2011} and kCSDA \cite{Potworowski2011} prove little advantage in estimating the sources under the ohmic, isotropic and homogeneous assumptions.
+If very few electrodes fail, more sophisticated methods such as iCSDA \cite{Leski2011} and kCSDA \cite{Potworowski2011} prove little advantage.
 In case that around 30 or more electrodes
-fail, specially if they are on the region of interest, the use of such methods
+fail, specially if they are on the region of interest or its borders,
+the use of such methods
 becomes necessary. The kCSDA method relies on the
@@
 Notice how the method incorporates a denoising mechanism, so
-no smoothing procedures are needed.
+no smoothing filters are needed.
 The great advantage of this method is that is possible to obtain a reasonable
@@
 The attainment of $K$ and $\tilde{K}$ turns out to be computer-costly for high density MEAs, but paraletization, specially on GPUs, solves this issue.
-
-%% posiblemente quitar esto 
-We have performed comparison of both methods over experimental data that has
-no failing channels in order to adjust the parameters of the second method. We
-have obtained consistent results between them.
 
@@
 Having both
-the mathematician and electro-physiologist interpretations presents we can
-perform the following analysis with clear readings.
-
-Apparently, CSD would not fit the requirements of non-negativity for its use as a mathematical density, but this is not a problem.
+the mathematician and electro-physiologist interpretations present we can
+perform the following analysis.
+
+Apparently, CSD would not fit the requirements of non-negativity for its use as a mathematical density.
+This is not a problem.
 CSD separates the measured activity into three distinct sets, namely, the set of all sources, the set of all sinks and the zero set.
@@
 cross lengths around \mum{150}. .
-
 These appear to correspond to the expected poles of firing neurons in structured tissue \cite{Buzsaki2012} and inspired us to separate the sink and source  sets into their \emph{disconnected components}.
@@
 To clarify the concept and how we are we using it an example is in order.
-In figure \ref{disconnectedsets} A we show geometrical sets enclosed by simple curves as color patches, where every color indicates a Set.
+In figure \ref{disconnectedsets} \textbf{A} we show geometrical sets enclosed by simple curves as color patches, where every color indicates a Set.
 The set $A$ consists of a single connected component, that means that one can make  path between any two points in the set that consist entirely of points of the same set.
@@
 Likewise the set $C$ is made of three disconnected components.
-In the subfigure \ref{disconnectedsets} Aiii we show the center of mass of the sets.
+In the subfigure \ref{disconnectedsets} \textbf{Aiii} we show the center of mass of the sets.
 The set $A$ is concave, so, it can have a center of mass outside the set.
@@
 as the electrode resolution is finite, the possibility of having highly non compact shapes is unlikely, and they
-tend to have a CM inside or near their borders. In the figure \ref{disconnectedsets} B we show this with experimental
-data (we are using evoked activity data, as explained in section \ref{sec:evocada}). In graph Bi we
-separate the CSD data into the sink and sources set. We take sinks as example. In the graph Bii
+tend to have a CM inside or near their borders. In the figure \ref{disconnectedsets}
+\texxtbf{B} we show this with experimental
+data (we are using evoked activity data, as explained in section \ref{sec:evocada}). In graph
+\textbf{Bi} we
+separate the CSD data into the sink and sources set. We take sinks as example. In the graph
+\textbf{Bii}
 we label each disjoint component, obtained by a single pass algorithm. There are 25 components, and
 the bigger two are non convex. In the
-subfigure Biii we indicate the CM of only those components that are above certain threshold of
+subfigure \textbf{Biii}
+we indicate the CM of only those components that are above certain threshold of
 integrated intensity (the circle size  is proportional in area to such intensity, there are five CM to be
@@
     Explanation of the geometrical concepts used on our analysis.
-    In the Row A of figures
+    In the Row \textbf{A} of figures we show
     with an artificial example the concepts of connectedness and convexity.
-    The subfigure Ai shows
+    The subfigure \textbf{Ai} shows
     three sets, $A$, $B$, and $C$. $A$ fails to be convex because it is concave,
@@
     In $B$ that is not possible, the set $B$ is convex in a mathematical sense.
-    In subfigure Aii we made some modification to show that
+    In subfigure \textbf{Aii} we made some modification to show that
     each set can be made of many disconnected components,
-    and each one of them can be convex or not.  In Aiii we show the center of
+    and each one of them can be convex or not.  In \textbf{Aiii} we show the center of
     mass for $A$ and $B$
@@
     small components are more likely to
-    be convex. In figure B we show everything with a frame of experimental data.
-    In Bi we separate the
+    be convex. In figure \textbf{B} we show everything with a frame of experimental data.
+    In \textbf{Bi} we separate the
     data into three sets, Sources, Sinks, and Zeros.
     We discard the Zeros and keep the active ones.
-    In Figure Bii we take the Sink Set and separate it into disconnected components,
+    In Figure \textbf{Bii}
+    we take the Sink Set and separate it into disconnected components,
     they are labeled from one to 25 in this particular example,
@@
     considerate for the next step those who are above certain threshold. The
-    next step is shown in Biii, where we plot the Center of Mass for each of
-    such subsets. Deviations
-    from convexity are tolerable,
+    next step is shown in \textbf{Biii},
+    where we plot the Center of Mass for each of
+    such subsets. Deviations from convexity are tolerable,
     as most of the mass is concentrated in the inner electrodes
@@
 
-Theoretically, CSD would divide the set of measuring points into three subsets: the set of all sources, the set of all sinks, and the border between them i.e, the set of all measuring points that have a positive CSD  value, the set of those who have a negative CSD value, and the set of those points having exactly the zero value in CSD space.
-Due to the finite precision of the measurement,
-it is impossible to define a precise zero value.
+Theoretically, CSD would divide the set of measuring points into three subsets: the set of all sources, the set of all sinks, and the border between them.
+It is impossible to define a precise zero value.
 The measurements produce a function over the grid elements (the electrodes).
@@
 between components, making identification of poles inexact.
-In order to avoid this we create a ``thick zero set'' of all the recording
+In order to avoid this we create a ``thick neutral set'' of all the recording
 points whose CSD values are inside an error interval around zero.
@@
 
-The ``thick zero set'' would help us to separate the sinks and sources sets into
+The ``thick neutral set'' would help us to separate the sinks and sources sets into
 its disconnected components, which would correspond then to clear patches of
@@
    If many electrodes fail near the r.o.i.,
-   kCSD is a good alternative if done with
-   parallel routines.
+   kCSD is a good alternative if performed using
+   parallel CPU or GPU computing.
    If too many electrodes are failing over the r.o.i.,
    it is better to discard the data. 
- \item Separate the CSD data into three sets: sources, sinks, and border or indiscernible (below error magnitude) data
-   which we call the zeros set. Again, sensible criteria have to be defined.
-   A rule of thumb would be to use the mean of the peaks of the noise  of the recording as the amplitude for the ``thick zero'' range of values. 
+ \item Separate the CSD data into three sets: sources, sinks, and neutral.
+   Again, sensible criteria have to be defined.
+   A rule of thumb would be to use the mean of the peaks of the noise  of the recording as the amplitude for the ``thick neutral'' range of values. 
 \item Perform disconnected component analysis in the sources and sinks sets, frame by frame.
@@
   all CM that are below some significant threshold value. A rule of thumb is one tenth of the average value for all CM,
-  or those who are in the lesser decil. Anything that is around the peak
-  values of the noise integrated over around 5 electrodes would have to be discarded.
+  or those who are in the lesser decil. 
 \item For each frame and its successor, and for each CM, perform a ``most probable successor'' detection.
